### 2019-05-13 poster/main.tex
add poster section

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Jacobs Landscape Poster
% LaTeX Template
% Version 1.1 (14/06/14)
%
% Created by:
% Computational Physics and Biophysics Group, Jacobs University
% https://teamwork.jacobs-university.de:8443/confluence/display/CoPandBiG/LaTeX+Poster
% 
% Further modified by:
% Nathaniel Johnston (nathaniel@njohnston.ca)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[final]{beamer}

\usepackage[scale=1.24]{beamerposter} % Use the beamerposter package for laying out the poster

\usetheme{confposter} % Use the confposter theme supplied with this template

\setbeamercolor{block title}{fg=ngreen,bg=white} % Colors of the block titles
\setbeamercolor{block body}{fg=black,bg=white} % Colors of the body of blocks
\setbeamercolor{block alerted title}{fg=white,bg=dblue!70} % Colors of the highlighted block titles
\setbeamercolor{block alerted body}{fg=black,bg=dblue!10} % Colors of the body of highlighted blocks
% Many more colors are available for use in beamerthemeconfposter.sty

%-----------------------------------------------------------
% Define the column widths and overall poster size
% To set effective sepwid, onecolwid and twocolwid values, first choose how many columns you want and how much separation you want between columns
% In this template, the separation width chosen is 0.024 of the paper width and a 4-column layout
% onecolwid should therefore be (1-(# of columns+1)*sepwid)/# of columns e.g. (1-(4+1)*0.024)/4 = 0.22
% Set twocolwid to be (2*onecolwid)+sepwid = 0.464
% Set threecolwid to be (3*onecolwid)+2*sepwid = 0.708

\newlength{\sepwid}
\newlength{\onecolwid}
\newlength{\twocolwid}
\newlength{\threecolwid}
\setlength{\paperwidth}{48in} % A0 width: 46.8in
\setlength{\paperheight}{36in} % A0 height: 33.1in
\setlength{\sepwid}{0.024\paperwidth} % Separation width (white space) between columns
\setlength{\onecolwid}{0.22\paperwidth} % Width of one column
\setlength{\twocolwid}{0.464\paperwidth} % Width of two columns
\setlength{\threecolwid}{0.708\paperwidth} % Width of three columns
\setlength{\topmargin}{-0.5in} % Reduce the top margin size
%-----------------------------------------------------------

\usepackage{graphicx}  % Required for including images

\usepackage{booktabs} % Top and bottom rules for tables

%----------------------------------------------------------------------------------------
%	TITLE SECTION 
%----------------------------------------------------------------------------------------

\title{Imputation Methods in Water Quality Modeling} % Poster title

\author{Carter Allen, Dr. Edsel Pena, Dr. John Grego} % Author(s)

\institute{Department of Statistics, University of South Carolina} % Institution(s)

%----------------------------------------------------------------------------------------

\begin{document}

\addtobeamertemplate{block end}{}{\vspace*{2ex}} % White space under blocks
\addtobeamertemplate{block alerted end}{}{\vspace*{2ex}} % White space under highlighted (alert) blocks

\setlength{\belowcaptionskip}{2ex} % White space under figures
\setlength\belowdisplayshortskip{2ex} % White space under equations

\begin{frame}[t] % The whole poster is enclosed in one beamer frame

\begin{columns}[t] % The whole poster consists of three major columns, the second of which is split into two columns twice - the [t] option aligns each column's content to the top

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\onecolwid} % The first column

%----------------------------------------------------------------------------------------
%	OBJECTIVES
%----------------------------------------------------------------------------------------

\begin{alertblock}{Objectives}

This project addresses the need for predictive tools to forecast expected concentrations of fecal bacteria Charleston, SC. \textit{Enterococcus faecalis} will be modeled in terms of precipitation in the past 72 hours, location, and tidal stage. Three distict missing/censored data imputation methods are developed and compared to acheive the desired model.

\end{alertblock}

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\begin{block}{Introduction}

The predictive model is built from data obtained through Charleston Waterkeeper's Recreational Water Quality Monitoring Program (RWQMP), a sampling program that is consistent with protocol stipulated by the South Carolina Department of Health and Environmental Control, as to preserve the compatibility of data collected by the two organizations [3]. The RWQMP measures the concentration of \textit{Enterococcus faecalis} bacteria in units of most probable number of individuals per 100 mL of sample (MPN/100 mL) [3]. Charleston Waterkeeper's records contain both missing and left-cencored (<10) observations. Of the 969 total observations from 2013-2015, 140 are censored and 28 are missing.

\end{block}

%------------------------------------------------

\begin{figure}
\includegraphics[width=0.9\linewidth]{SiteLocations.png}
\caption{Locations of Charleston Waterkeeper's Sample Sites}
\end{figure}

%----------------------------------------------------------------------------------------

\end{column} % End of the first column

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\twocolwid} % Begin a column which is two columns wide (column 2)

\begin{columns}[t,totalwidth=\twocolwid] % Split up the two columns wide column

\begin{column}{\onecolwid}\vspace{-.6in} % The first column within column 2 (column 2.1)

%----------------------------------------------------------------------------------------
%	MATERIALS
%----------------------------------------------------------------------------------------

\begin{block}{Exploratory Analysis}
\begin{figure}
\includegraphics[width=0.9\linewidth]{Ln(Ent)vSitevTide.png}
\caption{ln(Ent.) conditioned on tidal stage and sample site}
\end{figure}
\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of column 2.1

\begin{column}{\onecolwid}\vspace{-.6in} % The second column within column 2 (column 2.2)

%----------------------------------------------------------------------------------------
%	METHODS
%----------------------------------------------------------------------------------------

\begin{block}{Comparison Criteria}

Lorem ipsum dolor \textbf{sit amet}, consectetur adipiscing elit. Sed laoreet accumsan mattis. Integer sapien tellus, auctor ac blandit eget, sollicitudin vitae lorem. Praesent dictum tempor pulvinar. Suspendisse potenti. Sed tincidunt varius ipsum, et porta nulla suscipit et. Etiam congue bibendum felis, ac dictum augue cursus a. \textbf{Donec} magna eros, iaculis sit amet placerat quis, laoreet id est. In ut orci purus, interdum ornare nibh. Pellentesque pulvinar, nibh ac malesuada accumsan, urna nunc convallis tortor, ac vehicula nulla tellus eget nulla. Nullam lectus tortor, \textit{consequat tempor hendrerit} quis, vestibulum in diam. Maecenas sed diam augue.

\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of column 2.2

\end{columns} % End of the split of column 2 - any content after this will now take up 2 columns width

%----------------------------------------------------------------------------------------
%	IMPORTANT RESULT
%----------------------------------------------------------------------------------------

\begin{alertblock}{Important Results}

\textit{Enterococcus} counts are significantly influenced by precipication, tidal stage, and location. The best performing imputation method of the three tested employs sampling from a truncated normal distribution.

\end{alertblock}

%----------------------------------------------------------------------------------------

\begin{columns}[t,totalwidth=\twocolwid] % Split up the two columns wide column again

\begin{column}{\onecolwid} % The first column within column 2 (column 2.1)

%----------------------------------------------------------------------------------------
%	MATHEMATICAL SECTION
%----------------------------------------------------------------------------------------

\begin{block}{Imputation Methods}

\begin{enumerate}
\item \textit{A naive approach}: Censored values are sampled from a \( Uniform (0,10) \) and missing values are replaced with the overall mean \textit{Enterococcus} for 2013-2015.
\item \textit{Sampling from a Truncated Normal}: Censored values are sampled from a \(truncNorm()\) distribution, and missing values are predicted by the model made from all non-missing values.
\item \textit{Expected value of a Truncated Normal}: Let \( Y = \) \textit{Enterococcus} concentrations. If \( Y \sim logNormal(\mu,\sigma^2) \implies ln(Y) \sim Normal(\mu,\sigma^2) \), it can be shown that \( E[ln(Y)|a<ln(Y)<b] = \mu+\sigma{\frac{\phi(\frac{a-\mu}{\sigma})-\phi(\frac{b-\mu}{\sigma})}{\Phi(\frac{b-\mu}{\sigma})-\Phi(\frac{a-\mu}{\sigma})}} \). When $a = -\infty$ and $b = ln(10)$,

\end{enumerate}

\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of column 2.1

\begin{column}{\onecolwid} % The second column within column 2 (column 2.2)

%----------------------------------------------------------------------------------------
%	RESULTS
%----------------------------------------------------------------------------------------

\begin{block}{Results}

\begin{figure}
\includegraphics[width=0.8\linewidth]{placeholder.jpg}
\caption{Figure caption}
\end{figure}

Nunc tempus venenatis facilisis. Curabitur suscipit consequat eros non porttitor. Sed a massa dolor, id ornare enim:

\begin{table}
\vspace{2ex}
\begin{tabular}{l l l}
\toprule
\textbf{Treatments} & \textbf{Response 1} & \textbf{Response 2}\\
\midrule
Treatment 1 & 0.0003262 & 0.562 \\
Treatment 2 & 0.0015681 & 0.910 \\
Treatment 3 & 0.0009271 & 0.296 \\
\bottomrule
\end{tabular}
\caption{Table caption}
\end{table}

\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of column 2.2

\end{columns} % End of the split of column 2

\end{column} % End of the second column

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\onecolwid} % The third column

%----------------------------------------------------------------------------------------
%	CONCLUSION
%----------------------------------------------------------------------------------------

\begin{block}{Conclusion}

Nunc tempus venenatis facilisis. \textbf{Curabitur suscipit} consequat eros non porttitor. Sed a massa dolor, id ornare enim. Fusce quis massa dictum tortor \textbf{tincidunt mattis}. Donec quam est, lobortis quis pretium at, laoreet scelerisque lacus. Nam quis odio enim, in molestie libero. Vivamus cursus mi at \textit{nulla elementum sollicitudin}.

\end{block}

%----------------------------------------------------------------------------------------
%	ADDITIONAL INFORMATION
%----------------------------------------------------------------------------------------

\begin{block}{Additional Information}

Maecenas ultricies feugiat velit non mattis. Fusce tempus arcu id ligula varius dictum. 
\begin{itemize}
\item Curabitur pellentesque dignissim
\item Eu facilisis est tempus quis
\item Duis porta consequat lorem
\end{itemize}

\end{block}

%----------------------------------------------------------------------------------------
%	REFERENCES
%----------------------------------------------------------------------------------------

\begin{block}{References}

\nocite{*} % Insert publications even if they are not cited in the poster
\small{\bibliographystyle{unsrt}
\bibliography{sample}\vspace{0.75in}}

\end{block}

%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------

\setbeamercolor{block title}{fg=red,bg=white} % Change the block title color

\begin{block}{Acknowledgements}

\small{\rmfamily{Nam mollis tristique neque eu luctus. Suspendisse rutrum congue nisi sed convallis. Aenean id neque dolor. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.}} \\

\end{block}

%----------------------------------------------------------------------------------------
%	CONTACT INFORMATION
%----------------------------------------------------------------------------------------

\setbeamercolor{block alerted title}{fg=black,bg=norange} % Change the alert block title colors
\setbeamercolor{block alerted body}{fg=black,bg=white} % Change the alert block body colors

\begin{alertblock}{Contact Information}

\begin{itemize}
\item Web: \href{http://www.university.edu/smithlab}{http://www.university.edu/smithlab}
\item Email: \href{mailto:john@smith.com}{john@smith.com}
\item Phone: +1 (000) 111 1111
\end{itemize}

\end{alertblock}

\begin{center}
\begin{tabular}{ccc}
\includegraphics[width=0.4\linewidth]{logo.png} & \hfill & \includegraphics[width=0.4\linewidth]{logo.png}
\end{tabular}
\end{center}

%----------------------------------------------------------------------------------------

\end{column} % End of the third column

\end{columns} % End of all the columns in the poster

\end{frame} % End of the enclosing frame

\end{document}

### 2019-05-13 poster/poster.tex
poster almost done

--- 2019-05-13 poster/main.tex
@@
 
+\usepackage{amsmath}
+\usepackage{graphicx}
+\usepackage{blindtext}
 \usepackage[scale=1.24]{beamerposter} % Use the beamerposter package for laying out the poster
@@
 
-\title{Imputation Methods in Water Quality Modeling} % Poster title
+\title{Multivariate Skew-Normal Mixture Model for Infant Development Clustering} % Poster title
 
-\author{Carter Allen, Dr. Edsel Pena, Dr. John Grego} % Author(s)
+\author{Carter Allen$^1$; Brian Neelon, PhD$^1$; Sara E. Benjamin-Neelon, PhD, JD, MPH$^2$} % Author(s)
 
-\institute{Department of Statistics, University of South Carolina} % Institution(s)
+\institute{$^1$Department of Public Health Sciences, Medical University of South Carolina; $^2$Bloomberg School of Public Health, Johns Hopkins University}
 
@@
 
-\begin{alertblock}{Objectives}
+\begin{alertblock}{Abstract}
 
-This project addresses the need for predictive tools to forecast expected concentrations of fecal bacteria Charleston, SC.
-\textit{Enterococcus faecalis} will be modeled in terms of precipitation in the past 72 hours, location, and tidal stage.
-Three distict missing/censored data imputation methods are developed and compared to acheive the desired model.
+We propose a novel Bayesian model for infant development patterns that addresses primary research questions in this area while allowing for skewness and correlation of outcomes.
+Our model is based on finite mixtures of multivariate skew normal (MSN) distributions, where covariates are allowed on both the multivariate outcomes and probability of latent class membership.
+We also allow for missing outcome data by imputing missing outcomes from their conditional multivariate normal distributions.
+We demonstrate our method using data from the Nurture study.
 
@@
 
-The predictive model is built from data obtained through Charleston Waterkeeper's Recreational Water Quality Monitoring Program (RWQMP), a sampling program that is consistent with protocol stipulated by the South Carolina Department of Health and Environmental Control, as to preserve the compatibility of data collected by the two organizations [3].
-The RWQMP measures the concentration of \textit{Enterococcus faecalis} bacteria in units of most probable number of individuals per 100 mL of sample (MPN/100 mL) [3].
-Charleston Waterkeeper's records contain both missing and left-cencored (<10) observations.
-Of the 969 total observations from 2013-2015, 140 are censored and 28 are missing.
+A primary goal in infant development research is to identify \textbf{latent development classes} and explain class membership in relation to covariates of interest.
+It is often also of interest to relate covariates to mean longitudinal growth patterns.
+Infant development data are inherently correlated longitudinally, often skewed, and frequently missing due to longitudinal attrition and standard practices are ill-suited to addressing these research questions due to their ignorance of one or more of these features of development data.
 
@@
 
+\begin{block}{Motivation}
 \begin{figure}
-\includegraphics[width=0.9\linewidth]{SiteLocations.png}
-\caption{Locations of Charleston Waterkeeper's Sample Sites}
+\includegraphics[width=0.9\linewidth]{bayley_dens.jpg}
+\caption{Density of Bayley composite scores for Nurture infants at 3, 6, 9 and 12 months of age.}
 \end{figure}
+\end{block}
 
@@
 %----------------------------------------------------------------------------------------
-%	MATERIALS
+%	
 %----------------------------------------------------------------------------------------
 
-\begin{block}{Exploratory Analysis}
-\begin{figure}
-\includegraphics[width=0.9\linewidth]{Ln(Ent)vSitevTide.png}
-\caption{ln(Ent.) conditioned on tidal stage and sample site}
-\end{figure}
+\begin{block}{Clustering}
+
+A primary concern of our model is with identification of latent infant development clusters.
+We accomplish this via multinomial logit regression model on cluster membership, which utilizes P\'{o}lya-Gamma data-aumentation to allow for updating of all parameters using Gibbs sampling.
+
 \end{block}
@@
 
-\begin{block}{Comparison Criteria}
+\begin{block}{Conditional Imputation}
 
-Lorem ipsum dolor \textbf{sit amet}, consectetur adipiscing elit.
-Sed laoreet accumsan mattis.
-Integer sapien tellus, auctor ac blandit eget, sollicitudin vitae lorem.
-Praesent dictum tempor pulvinar.
-Suspendisse potenti.
-Sed tincidunt varius ipsum, et porta nulla suscipit et.
-Etiam congue bibendum felis, ac dictum augue cursus a.
-\textbf{Donec} magna eros, iaculis sit amet placerat quis, laoreet id est.
-In ut orci purus, interdum ornare nibh.
-Pellentesque pulvinar, nibh ac malesuada accumsan, urna nunc convallis tortor, ac vehicula nulla tellus eget nulla.
-Nullam lectus tortor, \textit{consequat tempor hendrerit} quis, vestibulum in diam.
-Maecenas sed diam augue.
+We allow for missingness of outcomes in the MSN mixture model by imputing missing values from their conditional multivariate normal distributions. We note that
+
+$$Y_i|X_i,t_i,\boldsymbol\beta,\psi \sim N_k(X_i \boldsymbol\beta + t_i \psi, \boldsymbol\Sigma)$$
+
+This allows us to appeal to standard conditional forms of the multivariate normal distribution. Let $Y_i = [Y^{miss}_{i_{q \times 1}} | Y^{obs}_{i_{k - q \times 1}}]^T$. We have
+
+$$Y_i^{miss}|Y_i^{obs},X_i,t_i,\boldsymbol\beta,\psi \sim N(\mu^{miss},\boldsymbol\Sigma^{miss})$$
+
+where $\mu^{miss}$ and $\Sigma^{miss}$ take standard forms.
+Each missing outcome is imputed "online", i.e.
+once per MCMC iteration.
+This provides more opportunities to explore the parameter space than multiple imputation and avoids multiplicative run-time scaling in $m$, the number of imputations.
 
@@
 
-\textit{Enterococcus} counts are significantly influenced by precipication, tidal stage, and location.
-The best performing imputation method of the three tested employs sampling from a truncated normal distribution.
+We developed a novel Bayesian MSN mixture model, and showed superior performance compared to standard approaches.
+We applied the MSN mixture model to data from the Nurture study and discovered two distinct development classes characterized by differences in development trajectories and demographics.
 
@@
 
-\begin{block}{Imputation Methods}
+\begin{block}{MSN Regression}
 
-\begin{enumerate}
-\item \textit{A naive approach}: Censored values are sampled from a \( Uniform (0,10) \) and missing values are replaced with the overall mean \textit{Enterococcus} for 2013-2015.
-\item \textit{Sampling from a Truncated Normal}: Censored values are sampled from a \(truncNorm()\) distribution, and missing values are predicted by the model made from all non-missing values.
-\item \textit{Expected value of a Truncated Normal}: Let \( Y = \) \textit{Enterococcus} concentrations.
-If \( Y \sim logNormal(\mu,\sigma^2) \implies ln(Y) \sim Normal(\mu,\sigma^2) \), it can be shown that \( E[ln(Y)|a<ln(Y)<b] = \mu+\sigma{\frac{\phi(\frac{a-\mu}{\sigma})-\phi(\frac{b-\mu}{\sigma})}{\Phi(\frac{b-\mu}{\sigma})-\Phi(\frac{a-\mu}{\sigma})}} \).
-When $a = -\infty$ and $b = ln(10)$,
+We model the effect of covariates on longitudinal development outcomes through the use of a MSN regression model.
+The MSN distribution can be represented as the superposition of a MN random variable with a latent truncated normal random effect.
+Let $\mathbf{Y}_{n \times k}$ be the observation matrix such that $Y_{ij}$ is the observation for subject $i$ at timepoint $j$.
 
-\end{enumerate}
+$$\mathbf{Y}_{n \times k} = \mathbf{X}_{n \times p}\boldsymbol\beta_{p \times k} + t_{n \times 1}\psi_{1 \times k} + \boldsymbol\epsilon_{n \times k}$$
+
+where $X_i$ is the $1 \times p$ vector of covariate values for subject $i$, $\beta_j$ is the $i \times k$ vector of fixed effects coefficients for timepoint $j$, $t_i \stackrel{iid}{\sim} N_{[0,\infty)}(0,1)$ is a truncated normal random effect, $\psi$ is the vector containing skewness parameters for each timepoint, and $\boldsymbol\epsilon_i \sim N_k(0,\boldsymbol\Sigma_{k \times k})$ is the correlated error term.
 
@@
 
-\begin{block}{Results}
+\begin{block}{Simulation Study}
 
-\begin{figure}
-\includegraphics[width=0.8\linewidth]{placeholder.jpg}
-\caption{Figure caption}
-\end{figure}
+\begin{table}[t]
 
-Nunc tempus venenatis facilisis. Curabitur suscipit consequat eros non porttitor. Sed a massa dolor, id ornare enim:
+\caption{\label{tab:unnamed-chunk-4}Parameter estimates for n = 1000, k = 4, p = 2, h = 3, v = 3 simulation.}
+\centering
+\fontsize{20}{20}\selectfont
+\begin{tabular}{llllllll}
+\toprule
+\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Class 1} & \multicolumn{2}{c}{Class 2} & \multicolumn{2}{c}{Class 3} \\
+\cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6} \cmidrule(l{3pt}r{3pt}){7-8}
+Component & Parm. & True & Est. (95\% CrI) & True & Est. (95\% CrI) & True & Est. (95\% CrI)\\
+\midrule
+\addlinespace[0.3em]
+\multicolumn{8}{l}{\textbf{ }}\\
+\hspace{1em}MSN & $\beta_{11}$ & -0.54 & -0.51 (-0.61, -0.41) & -0.19 & -0.22 (-0.35, -0.09) & 1.84 & 1.81 (1.65, 1.99)\\
+\hspace{1em}Reg. & $\beta_{12}$ & -1.06 & -1.07 (-1.15, -0.98) & -0.16 & -0.13 (-0.25, -0.02) & 2.03 & 2.09 (1.93, 2.23)\\
+\hspace{1em} & $\beta_{13}$ & -1.28 & -1.22 (-1.32, -1.11) & 0.59 & 0.58 (0.46, 0.69) & 1.8 & 1.74 (1.58, 1.9)\\
+\hspace{1em} & $\beta_{14}$ & -1.91 & -1.83 (-1.92, -1.74) & -0.2 & -0.12 (-0.25, -0.01) & 1.43 & 1.37 (1.22, 1.52)\\
 
-\begin{table}
-\vspace{2ex}
-\begin{tabular}{l l l}
-\toprule
-\textbf{Treatments} & \textbf{Response 1} & \textbf{Response 2}\\
-\midrule
-Treatment 1 & 0.0003262 & 0.562 \\
-Treatment 2 & 0.0015681 & 0.910 \\
-Treatment 3 & 0.0009271 & 0.296 \\
+\addlinespace[0.3em]
+\multicolumn{8}{l}{\textbf{ }}\\
+\hspace{1em} & $\sigma^2_{11}$ & 1 & 0.9 (0.78, 1.03) & 1 & 1.04 (0.88, 1.26) & 1 & 1.39 (1.08, 1.76)\\
+\hspace{1em} & $\sigma^2_{12}$ & -0.1 & -0.12 (-0.22, -0.02) & -0.32 & -0.31 (-0.44, -0.18) & 0.08 & 0.27 (0.04, 0.52)\\
+\hspace{1em} & $\sigma^2_{13}$ & -0.18 & -0.17 (-0.28, -0.08) & -0.52 & -0.5 (-0.66, -0.38) & 0.78 & 1.19 (0.87, 1.54)\\
+\hspace{1em} & $\sigma^2_{14}$ & 0.39 & 0.31 (0.22, 0.43) & 0.04 & 0.05 (-0.07, 0.22) & 0.41 & 0.72 (0.47, 0.99)\\
+\hspace{1em} & $\sigma^2_{22}$ & 1 & 0.97 (0.84, 1.12) & 1 & 0.94 (0.77, 1.14) & 1 & 1.18 (0.91, 1.5)\\
+
+\addlinespace[0.3em]
+\multicolumn{8}{l}{\textbf{ }}\\
+\hspace{1em} & $\psi_{1}$ & -0.33 & -0.28 (-0.4, -0.18) & 0.67 & 0.64 (0.5, 0.78) & -1 & -0.91 (-1.12, -0.68)\\
+\hspace{1em} & $\psi_{2}$ & -0.33 & -0.21 (-0.34, -0.09) & 0.67 & 0.7 (0.55, 0.84) & -1 & -0.96 (-1.16, -0.76)\\
+\hspace{1em} & $\psi_{3}$ & -0.33 & -0.27 (-0.39, -0.14) & 0.67 & 0.72 (0.58, 0.86) & -1 & -0.91 (-1.14, -0.69)\\
+\hspace{1em} & $\psi_{4}$ & -0.33 & -0.18 (-0.3, -0.07) & 0.67 & 0.73 (0.6, 0.87) & -1 & -0.93 (-1.13, -0.73)\\
+\addlinespace[0.3em]
+\multicolumn{8}{l}{\textbf{ }}\\
+\hspace{1em}Multi. & $\delta_{11}$ & -0.53 & -0.5 (-0.74, -0.24) & -0.53 & -0.5 (-0.74, -0.24) & -0.53 & -0.5 (-0.74, -0.24)\\
+\hspace{1em}Logit & $\delta_{12}$ & 0.44 & 0.28 (0.02, 0.56) & 0.44 & 0.28 (0.02, 0.56) & 0.44 & 0.28 (0.02, 0.56)\\
+\hspace{1em} & $\delta_{21}$ & 0.34 & 0.38 (0.09, 0.63) & 0.34 & 0.38 (0.09, 0.63) & 0.34 & 0.38 (0.09, 0.63)\\
+\addlinespace[0.3em]
+\multicolumn{8}{l}{\textbf{ }}\\
+\hspace{1em}Clust. & $\pi_l$ & 0.41 & 0.41 (0.4, 0.42) & 0.36 & 0.36 (0.35, 0.37) & 0.23 & 0.23 (0.22, 0.24)\\
 \bottomrule
 \end{tabular}
-\caption{Table caption}
 \end{table}
@@
 
-\begin{block}{Conclusion}
+\begin{figure}
+\includegraphics[height=1\linewidth,width=1\linewidth]{beta_trace.pdf}
+\caption{Trace plots of regression parameter estimates in simulation}
+\end{figure}
 
-Nunc tempus venenatis facilisis.
-\textbf{Curabitur suscipit} consequat eros non porttitor.
-Sed a massa dolor, id ornare enim.
-Fusce quis massa dictum tortor \textbf{tincidunt mattis}.
-Donec quam est, lobortis quis pretium at, laoreet scelerisque lacus.
-Nam quis odio enim, in molestie libero.
-Vivamus cursus mi at \textit{nulla elementum sollicitudin}.
+\begin{block}{Application Results}
 
-\end{block}
-
-%----------------------------------------------------------------------------------------
-%	ADDITIONAL INFORMATION
-%----------------------------------------------------------------------------------------
-
-\begin{block}{Additional Information}
-
-Maecenas ultricies feugiat velit non mattis. Fusce tempus arcu id ligula varius dictum. 
-\begin{itemize}
-\item Curabitur pellentesque dignissim
-\item Eu facilisis est tempus quis
-\item Duis porta consequat lorem
-\end{itemize}
+\begin{figure}
+\includegraphics[height=0.8\linewidth,width=1\linewidth]{clust_trends.pdf}
+\caption{Mean development patterns in each estimated class}
+\end{figure}
 
@@
 
-\begin{block}{References}
-
-\nocite{*} % Insert publications even if they are not cited in the poster
-\small{\bibliographystyle{unsrt}
-\bibliography{sample}\vspace{0.75in}}
-
-\end{block}
-
 %----------------------------------------------------------------------------------------
@@
 %----------------------------------------------------------------------------------------
-
-\setbeamercolor{block title}{fg=red,bg=white} % Change the block title color
-
-\begin{block}{Acknowledgements}
-
-\small{\rmfamily{Nam mollis tristique neque eu luctus.
-Suspendisse rutrum congue nisi sed convallis.
-Aenean id neque dolor.
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.}} \\
-
-\end{block}
 
@@
 
-\begin{alertblock}{Contact Information}
+\begin{alertblock}{Further Resources}
 
-\begin{itemize}
-\item Web: \href{http://www.university.edu/smithlab}{http://www.university.edu/smithlab}
-\item Email: \href{mailto:john@smith.com}{john@smith.com}
-\item Phone: +1 (000) 111 1111
-\end{itemize}
+\centering
+https://carter-allen.github.io/MVSN-FMM
 
 \end{alertblock}
-
-\begin{center}
-\begin{tabular}{ccc}
-\includegraphics[width=0.4\linewidth]{logo.png} & \hfill & \includegraphics[width=0.4\linewidth]{logo.png}
-\end{tabular}
-\end{center}
 

### 2019-05-13 poster/poster.tex
poster done

@@
 We accomplish this via multinomial logit regression model on cluster membership, which utilizes P\'{o}lya-Gamma data-aumentation to allow for updating of all parameters using Gibbs sampling.
+The multinomial logit model is as follows for $l = 1,...,h$.
+
+$$P(Z_i = l|\w_i) = \pi_{il} = \frac{e^{w_i^T \delta_l}}{\sum_{r = 1}^h e^{w_i^T \delta_r}}$$
+
+where $w_i$ is the vector of class probability covariates for subject $i$, $\delta_l$ contains the multinomial regression parameters for class $l$, and $h$ is the number of putative clusters specified \textit{a priori}.
+
+During our MCMC estimation procedure, the class labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as class assignments.
 
@@
 We developed a novel Bayesian MSN mixture model, and showed superior performance compared to standard approaches.
-We applied the MSN mixture model to data from the Nurture study and discovered two distinct development classes characterized by differences in development trajectories and demographics.
+We applied the MSN mixture model to data from the Nurture study and discovered three distinct development classes characterized by differences in development trajectories and demographics.
 
@@
 \begin{figure}
-\includegraphics[height=1\linewidth,width=1\linewidth]{beta_trace.pdf}
+\includegraphics[height=0.9\linewidth,width=1\linewidth]{beta_trace.pdf}
 \caption{Trace plots of regression parameter estimates in simulation}
@@
 
+\tiny  \textbf{(1) Fruhwirth-Schnatter, S and Pyne, S.
+(2010)}.
+Bayesian inference for finite mixtures of univariate ...
+Biostatistics; \textbf{(2) Benjamin-Neelon SE, Ostbye T, Bennett GG, et al.} Cohort profile for the Nurture Observational Study ...
+BMJ Open 2017.
+
+\
+
+\tiny \textbf{Funding:} \textit{This work is supported by a grant from the NIH (\textbf{R01DK094841})}.
+
 %----------------------------------------------------------------------------------------

### 2019-05-14 manuscript/manuscript.tex
manuscript section

%  ALWAYS USE THE referee OPTION WITH PAPERS SUBMITTED TO BIOMETRICS!!!
%  You can see what your paper would look like typeset by removing
%  the referee option.  Because the typeset version will be in two
%  columns, however, some of your equations may be too long. DO NOT
%  use the \longequation option discussed in the user guide!!!  This option
%  is reserved ONLY for equations that are impossible to split across 
%  multiple lines; e.g., a very wide matrix.  Instead, type your equations 
%  so that they stay in one column and are split across several lines, 
%  as are almost all equations in the journal.  Use a recent version of the
%  journal as a guide. 
%  
\documentclass[useAMS,referee]{biom}
\usepackage{amsmath}

\def\bSig\mathbf{\Sigma}
\newcommand{\VS}{V\&S}
\newcommand{\tr}{\mbox{tr}}

%  Here, place your title and author information.  Note that in 
%  use of the \author command, you create your own footnotes.  Follow
%  the examples below in creating your author and affiliation information.
%  Also consult a recent issue of the journal for examples of formatting.

\title[]{A multivariate skew-normal finite mixture model for analysis of infant development trajectories}

%  Here are examples of different configurations of author/affiliation
%  displays.  According to the Biometrics style, in some instances,
%  the convention is to have superscript *, **, etc footnotes to indicate 
%  which of multiple email addresses belong to which author.  In this case,
%  use the \email{ } command to produce the emails in the display.

%  In other cases, such as a single author or two authors from 
%  different institutions, there should be no footnoting.  Here, use
%  the \emailx{ } command instead.

%  The examples below corrspond to almost every possible configuration
%  of authors and may be used as a guide.  For other configurations, consult
%  a recent issue of the the journal.

%  Single author -- USE \emailx{ } here so that no asterisk footnoting
%  for the email address will be produced.

%\author{John Author\emailx{email@address.edu} \\
%Department of Statistics, University of Warwick, Coventry CV4 7AL, U.K.}

%  Two authors from the same institution, with both emails -- use
%  \email{ } here to produce the asterisk footnoting for each email address

%\author{John Author$^{*}$\email{author@address.edu} and
%Kathy Authoress$^{**}$\email{email2@address.edu} \\
%Department of Statistics, University of Warwick, Coventry CV4 7AL, U.K.}

%  Exactly two authors from different institutions, with both emails  
%  USE \emailx{ } here so that no asterisk footnoting for the email address
%  is produced.

\author
{Carter Allen\emailx{allecart@musc.edu} \\
Department of Public Health Sciences, Medical University of South Carolina, Charleston, SC, U.S.A
\and
Brian Neelon, PhD\emailx{neelon@musc.edu} \\
Department of Public Health Sciences, Medical University of South Carolina, Charleston, SC, U.S.A
\and
Sara Benjamin-Neelon, PhD, MPH, RD\emailx{sara.neelon@jhu.edu} \\
Department of Health, Behavior and Society, Johns Hopkins University, Baltimore, MD, U.S.S}

%  Three or more authors from same institution with all emails displayed
%  and footnoted using asterisks -- use \email{ }

%\author{John Author$^*$\email{author@address.edu}, 
%Jane Author$^{**}$\email{jane@address.edu}, and 
%Dick Author$^{***}$\email{dick@address.edu} \\
%Department of Statistics, University of Warwick, Coventry CV4 7AL, U.K}

%  Three or more authors from same institution with one corresponding email
%  displayed

%\author{John Author$^*$\email{author@address.edu}, 
%Jane Author, and Dick Author \\
%Department of Statistics, University of Warwick, Coventry CV4 7AL, U.K}

%  Three or more authors, with at least two different institutions,
%  more than one email displayed

%\author{John Author$^{1,*}$\email{author@address.edu}, 
%Kathy Author$^{2,**}$\email{anotherauthor@address.edu}, and 
%Wilma Flinstone$^{3,***}$\email{wilma@bedrock.edu} \\
%$^{1}$Department of Statistics, University of Warwick, Coventry CV4 7AL, U.K \\
%$^{2}$Department of Biostatistics, University of North Carolina at 
%Chapel Hill, Chapel Hill, North Carolina, U.S.A. \\
%$^{3}$Department of Geology, University of Bedrock, Bedrock, Kansas, U.S.A.}

%  Three or more authors with at least two different institutions and only
%  one email displayed

%\author{John Author$^{1,*}$\email{author@address.edu}, 
%Wilma Flinstone$^{2}$, and Barney Rubble$^{2}$ \\
%$^{1}$Department of Statistics, University of Warwick, Coventry CV4 7AL, U.K \\
%$^{2}$Department of Geology, University of Bedrock, Bedrock, Kansas, U.S.A.}

\begin{document}

%  This will produce the submission and review information that appears
%  right after the reference section.  Of course, it will be unknown when
%  you submit your paper, so you can either leave this out or put in 
%  sample dates (these will have no effect on the fate of your paper in the
%  review process!)

\date{{\it Received October} 2007. {\it Revised February} 2008.  {\it
Accepted March} 2008.}

%  These options will count the number of pages and provide volume
%  and date information in the upper left hand corner of the top of the 
%  first page as in published papers.  The \pagerange command will only
%  work if you place the command \label{firstpage} near the beginning
%  of the document and \label{lastpage} at the end of the document, as we
%  have done in this template.

%  Again, putting a volume number and date is for your own amusement and
%  has no bearing on what actually happens to your paper!

\pagerange{\pageref{firstpage}--\pageref{lastpage}} 
\volume{64}
\pubyear{2008}
\artmonth{December}

%  The \doi command is where the DOI for your paper would be placed should it
%  be published.  Again, if you make one up and stick it here, it means 
%  nothing!

\doi{10.1111/j.1541-0420.2005.00454.x}

%  This label and the label ``lastpage'' are used by the \pagerange
%  command above to give the page range for the article.  You may have 
%  to process the document twice to get this to match up with what you 
%  expect.  When using the referee option, this will not count the pages
%  with tables and figures.

\label{firstpage}

%  put the summary for your paper here

\begin{abstract}
In studies of infant motor development, a crucial research goal is to identify latent classes of infants that experience delayed development, as this is a known risk factor for adverse outcomes later in life. However, there are a number of statistical challenges in modeling infant development: the data are typically skewed, exhibit intermittent missingness, and are highly correlated across the repeated measurements collected during infancy. Using data from the Nurture study, a cohort of over 600 mother-infant pairs followed from pregnancy to 12 months postpartum, we develop a flexible Bayesian latent class model for the analysis infant motor development. Our model has a number of attractive features. First, we adopt the multivariate skew normal distribution with class-specific parameters that accommodate the inherent correlation and skewness in the data. Second, we model the class membership probabilities using a novel Pólya-Gamma data-augmentation scheme, thereby improving predictions of the class membership allocations. Lastly, we impute missing responses under missing at random assumption by drawing from appropriate conditional skew normal distributions. Bayesian inference is achieved through straightforward Gibbs sampling, and can be carried out in available software such as R.  Through simulation studies, we show that the proposed model yields improved inferences over models that ignore skewness. In addition, our imputation method yields improvements compared to conventional missing data methods, including multiple imputation and complete or available case analysis. When applied to Nurture data, we identified two distinct development classes: one characterized by delayed “U-shaped” development and a higher percentage of male infants and another characterized by more steady development and a lower percentage of males. The classes also differed in terms of key demographic variables, such as infant race and maternal pre-pregnancy body mass index. These findings can aid investigators in targeting interventions during this critical early-life developmental window.
\end{abstract}

%  Please place your key words in alphabetical order, separated
%  by semicolons, with the first letter of the first word capitalized,
%  and a period at the end of the list.
%

\begin{keywords}
A key word; But another key word; Still another key word; Yet another key word.
\end{keywords}

%  As usual, the \maketitle command creates the title and author/affiliations
%  display

\maketitle

%  If you are using the referee option, a new page, numbered page 1, will
%  start after the summary and keywords.  The page numbers thus count the
%  number of pages of your manuscript in the preferred submission style.
%  Remember, ``Normally, regular papers exceeding 25 pages and Reader Reaction 
%  papers exceeding 12 pages in (the preferred style) will be returned to 
%  the authors without review. The page limit includes acknowledgements, 
%  references, and appendices, but not tables and figures. The page count does 
%  not include the title page and abstract. A maximum of six (6) tables or 
%  figures combined is often required.''

%  You may now place the substance of your manuscript here.  Please use
%  the \section, \subsection, etc commands as described in the user guide.
%  Please use \label and \ref commands to cross-reference sections, equations,
%  tables, figures, etc.
%
%  Please DO NOT attempt to reformat the style of equation numbering!
%  For that matter, please do not attempt to redefine anything!

\section{Introduction}
\label{s:intro}

\section{Model}
\label{s:model}

\subsection{Multivariate Skew Normal Regression}

We model the effect of covariates on longitudinal development outcomes through the use of a MSN regression model. The MSN distribution can be represented as the superposition of a MN random variable with a latent truncated normal random effect. Let $\mathbf{Y}_{n \times k}$ be the observation matrix such that $Y_{ij}$ is the observation for subject $i$ at timepoint $j$.

$$\mathbf{Y}_{n \times k} = \mathbf{X}_{n \times p}\boldsymbol\beta_{p \times k} + t_{n \times 1}\psi_{1 \times k} + \boldsymbol\epsilon_{n \times k}$$

where $X_i$ is the $1 \times p$ vector of covariate values for subject $i$, $\beta_j$ is the $i \times k$ vector of fixed effects coefficients for timepoint $j$, $t_i \stackrel{iid}{\sim} N_{[0,\infty)}(0,1)$ is a truncated normal random effect, $\psi$ is the vector containing skewness parameters for each timepoint, and $\boldsymbol\epsilon_i \sim N_k(0,\boldsymbol\Sigma_{k \times k})$ is the correlated error term.

\subsection{Latent Class Finite Mixture}

\subsection{Multinomial Regression on Class Probabilities}

\subsection{Conditional MVSN Imputation}

We allow for missingness of outcomes in the MSN mixture model by imputing missing values from their conditional multivariate normal distributions. We note that

$$Y_i|X_i,t_i,\boldsymbol\beta,\psi \sim N_k(X_i \boldsymbol\beta + t_i \psi, \boldsymbol\Sigma)$$

This allows us to appeal to standard conditional forms of the multivariate normal distribution. Let $Y_i = [Y^{miss}_{i_{q \times 1}} | Y^{obs}_{i_{k - q \times 1}}]^T$. We have

$$Y_i^{miss}|Y_i^{obs},X_i,t_i,\boldsymbol\beta,\psi \sim N(\mu^{miss},\boldsymbol\Sigma^{miss})$$

where $\mu^{miss}$ and $\Sigma^{miss}$ take standard forms. Each missing outcome is imputed "online", i.e. once per MCMC iteration. This provides more opportunities to explore the parameter space than multiple imputation and avoids multiplicative run-time scaling in $m$, the number of imputations.

\section{Bayesian Inference}
\label{s:inf}

\section{Simulation Studies}
\label{s:sim}

\section{Application}
\label{s:app}

\section{Discussion}
\label{s:discuss}

Put your final comments here.

%  The \backmatter command formats the subsequent headings so that they
%  are in the journal style.  Please keep this command in your document
%  in this position, right after the final section of the main part of 
%  the paper and right before the Acknowledgements, Supplementary Materials,
%  and References sections.

\backmatter

%  This section is optional.  Here is where you will want to cite
%  grants, people who helped with the paper, etc.  But keep it short!

\section*{Acknowledgements}

%  If your paper refers to supplementary web material, then you MUST
%  include this section!!  See Instructions for Authors at the journal
%  website http://www.biometrics.tibs.org

\section*{Supplementary Materials}

%  Here, we create the bibliographic entries manually, following the
%  journal style.  If you use this method or use natbib, PLEASE PAY
%  CAREFUL ATTENTION TO THE BIBLIOGRAPHIC STYLE IN A RECENT ISSUE OF
%  THE JOURNAL AND FOLLOW IT!  Failure to follow stylistic conventions
%  just lengthens the time spend copyediting your paper and hence its
%  position in the publication queue should it be accepted.

%  We greatly prefer that you incorporate the references for your
%  article into the body of the article as we have done here 
%  (you can use natbib or not as you choose) than use BiBTeX,
%  so that your article is self-contained in one file.
%  If you do use BiBTeX, please use the .bst file that comes with 
%  the distribution.  In this case, replace the thebibliography
%  environment below by 
%
%  \bibliographystyle{biom} 
% \bibliography{mybibilo.bib}

\begin{thebibliography}{}

\bibitem{ } Cox, D. R. (1972). Regression models and life tables (with
discussion).  \textit{Journal of the Royal Statistical Society, Series B}
\textbf{34,} 187--200.

\bibitem{ }  Hastie, T., Tibshirani, R., and Friedman, J. (2001). \textit{The 
Elements of Statistical Learning: Data Mining, Inference, and Prediction}.
New York: Springer.

\end{thebibliography}

\appendix

%  To get the journal style of heading for an appendix, mimic the following.

\section{}
\subsection{Full Conditional Distributions}

\label{lastpage}

\end{document}

### 2019-05-14 manuscript/manuscript.tex
manuscript format

@@
 \usepackage{amsmath}
+\usepackage{graphicx}
+\usepackage{blindtext}
 
@@
 \and
-Brian Neelon, PhD\emailx{neelon@musc.edu} \\
+Brian Neelon, PhD \\
 Department of Public Health Sciences, Medical University of South Carolina, Charleston, SC, U.S.A
 \and
-Sara Benjamin-Neelon, PhD, MPH, RD\emailx{sara.neelon@jhu.edu} \\
+Sara Benjamin-Neelon, PhD, MPH, RD \\
 Department of Health, Behavior and Society, Johns Hopkins University, Baltimore, MD, U.S.S}
@@
 
+A primary concern of our model is with identification of latent infant development clusters.
+We accomplish this via multinomial logit regression model on cluster membership, which utilizes P\'{o}lya-Gamma data-aumentation to allow for updating of all parameters using Gibbs sampling.
+The multinomial logit model is as follows for $l = 1,...,h$.
+
+$$P(Z_i = l|w_i) = \pi_{il} = \frac{e^{w_i^T \delta_l}}{\sum_{r = 1}^h e^{w_i^T \delta_r}}$$
+
+where $w_i$ is the vector of class probability covariates for subject $i$, $\delta_l$ contains the multinomial regression parameters for class $l$, and $h$ is the number of putative clusters specified \textit{a priori}.
+
+During our MCMC estimation procedure, the class labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as class assignments.
+
 \subsection{Conditional MVSN Imputation}

### 2019-05-22 manuscript/manuscript.tex
outline
derivations

@@
 
+\setcounter{tocdepth}{3}
+\tableofcontents
+
+\newpage
+
 %  If you are using the referee option, a new page, numbered page 1, will
@@
 
+\section{Nurture Study}
+\label{s:nurt}
+\subsection{Baseline Demographics and Description of Variables}
+
+\subsection{Statistical Challenges}
+\subsubsection{Skewness of Bayley score residuals}
+\subsubsection{Attrition and Intermittent Missingness}
+
 \section{Model}
@@
 where $X_i$ is the $1 \times p$ vector of covariate values for subject $i$, $\beta_j$ is the $i \times k$ vector of fixed effects coefficients for timepoint $j$, $t_i \stackrel{iid}{\sim} N_{[0,\infty)}(0,1)$ is a truncated normal random effect, $\psi$ is the vector containing skewness parameters for each timepoint, and $\boldsymbol\epsilon_i \sim N_k(0,\boldsymbol\Sigma_{k \times k})$ is the correlated error term.
-
-\subsection{Latent Class Finite Mixture}
 
@@
 
-\section{Bayesian Inference}
-\label{s:inf}
+\subsection{Bayesian Inference}
+
+\subsubsection{MCMC Algorithm}
+
+\subsubsection{Assessment of MCMC Convergence}
+
+\subsubsection{Label Switching}
+
+\subsubsection{Prior Choice}
 
@@
 \label{s:sim}
+
+\subsection{Simulation to Compare to Multivariate Normal}
+
+\subsection{Simulation to Compare Imputation Methods}
+
+\subsection{Simulation to Assess Sensitivity to Mispecified H}
 
@@
 \label{s:discuss}
+
+\section{Appendix}
 
@@
 \section*{Supplementary Materials}
+
+\subsection{Glossary of Notation}
+
+\subsection{Derivation of Full Conditional Distributions}
+
+\subsubsection{Multivariate Skew-Normal Regression}
+
+Without loss of generality, we derive the full conditional distributions for the multivariate skew-normal regression model component under the assumption that all observations belong to a single cluster.
+To make the extension to the case where more than one cluster is specified, simply apply these distributional forms to cluster specific parameters and data.
+
+\subsubsection{Multinomial Logit Regression}
+
+\subsubsection{Multivariate Normal Conditional Imputation}
 
@@
 
-\appendix
+% \appendix
 
 %  To get the journal style of heading for an appendix, mimic the following.
-
-\section{}
-\subsection{Full Conditional Distributions}
 

### 2019-05-22 simulation/simulation.R

@@ -264,7 +264,7 @@ for(i in 1:nsim)
         delta.l <- delta[,l]
         delta.notl <- delta[,-l]
         u.l <- 1*(c == l+1)
-        c.l <- log(1 + exp(W %*% delta.notl))
+        c.l <- log(1 + exp(rowSums(W %*% delta.notl)))
         eta <- W %*% delta.l - c.l
         w <- rpg(n, 1, eta)
         z.l <- (u.l - 1/2)/w + c.l 

### 2019-05-22 manuscript/manuscript.tex
glossary

@@
 \usepackage{blindtext}
+\usepackage{booktabs}
 
@@
 
+\begin{itemize}
+
+    \item $\mathbf{Y}$: A $n \times p$ matrix containing all multivariate skew-normal regression outcomes such that $y_{ij}$ is the $j^{th}$ outcome observed for subject $i$, where $i = 1,...,n$ and $j = 1,...p$.
+
+    \item $\mathbf{X}$: A $n \times m$ matrix containing all multivariate skew-normal regression covariates such that $x_{ij}$ is the $j^{th}$ covariate value for subject $i$.
+
+    \item $\mathbf{B}$: A $m \times p$ matrix containing all multivariate skew-normal regression coefficients such that $\mathbf{B} = \left [ \boldsymbol\beta_1,...,\boldsymbol\beta_p \right ]$, where $\beta_{ij}$ is interpreted as the effect of covariate $i$ on outcome $j$ for $i = 1,...,m$ and $j = 1,...,p$.
+
+    \item $\mathbf{E}$: A $n \times p$ matrix of error terms in the multivariate skew-normal regression model component.
+$\mathbf{E}$ is made up of row vectors $\boldsymbol\epsilon_i = (\epsilon_{i1},...,\epsilon_{ip})$, where $ \boldsymbol\epsilon_i \stackrel{iid}{\sim} N_p(0, \boldsymbol\Sigma)$ for $i = 1,...,n$.
+
+    \item $\boldsymbol\Sigma$: A $p \times p$ covariance matrix that defines the correlation between the $p$ multivariate normal outcomes.
+
+    \item $\boldsymbol\Omega$: A $p \times p$ covariance scale matrix that defines the correlation between the $p$ multivariate skew-normal outcomes.
+
+    \item $\boldsymbol\psi$: A $p \times 1$ vector containing the skewness parameter for each outcome.
+
+    \item $\boldsymbol\alpha$: A $p \times 1$ vector containing the skewness parameter for each outcome.
+
+    \item $\mathbf{t}$: A $n \times 1$ vector of truncated normal random effects used in the stochastic representation of the multivariate skew-normal distribution.
+For $i = 1,...,n$, $t_i \stackrel{iid}{\sim}T_[0,\infty)(0,1)$
+
+    \item $\mathbf{X}^*$: A $n \times (m + 1)$ matrix constructed by column binding $\mathbf{t}$ to $\mathbf{X}$
+
+    \item $\mathbf{B}^*$: A $(m+1) \times p$ matrix constructed by row binding $\boldsymbol\psi^T$. to $\mathbf{B}$.
+
+\end{itemize}
+
 \subsection{Derivation of Full Conditional Distributions}

### 2019-05-22 manuscript/manuscript.tex
glossary

@@
 
-    \item $\mathbf{B}^*$: A $(m+1) \times p$ matrix constructed by row binding $\boldsymbol\psi^T$. to $\mathbf{B}$.
+    \item $\mathbf{B}^*$: A $(m+1) \times p$ matrix constructed by row binding $\boldsymbol\psi^T$ to $\mathbf{B}$.
 
@@
 To make the extension to the case where more than one cluster is specified, simply apply these distributional forms to cluster specific parameters and data.
+
+The multivariate skew-normal regression model can be written as follows.
+
+$$\mathbf{Y} = \mathbf{X}\mathbf{B} + \mathbf{t} \boldsymbol\psi^T + \mathbf{E} = \mathbf{X}^* \mathbf{B}^* + \mathbf{E}$$
+
+Under the assumption that $t_i \stackrel{iid}{\sim}T_[0,\infty)(0,1)$ and that $\mathbf{E}$ is made of row vectors $\boldsymbol\epsilon_i = (\epsilon_{i1},...,\epsilon_{ip})$ for $i = 1,...,n$,
 

### 2019-05-22 manuscript/manuscript.tex
appendix

@@
 
+\newpage
+
 \subsection{Glossary of Notation}
@@
     \item $\mathbf{t}$: A $n \times 1$ vector of truncated normal random effects used in the stochastic representation of the multivariate skew-normal distribution.
-For $i = 1,...,n$, $t_i \stackrel{iid}{\sim}T_[0,\infty)(0,1)$
+For $i = 1,...,n$, $t_i \stackrel{iid}{\sim}T_{[0,\infty)(0,1)}$
 
@@
 
+\newpage
 \subsection{Derivation of Full Conditional Distributions}
@@
 
-Under the assumption that $t_i \stackrel{iid}{\sim}T_[0,\infty)(0,1)$ and that $\mathbf{E}$ is made of row vectors $\boldsymbol\epsilon_i = (\epsilon_{i1},...,\epsilon_{ip})$ for $i = 1,...,n$,
+For convenience, we define $\mathbf{X}^*$ as a $n \times (m + 1)$ matrix constructed by column binding $\mathbf{t}$ to $\mathbf{X}$, and $\mathbf{B}^*$ as a $(m+1) \times p$ matrix constructed by row binding $\boldsymbol\psi^T$ to $\mathbf{B}$.
+We assume that $t_i \stackrel{iid}{\sim}T_{[0,\infty})(0,1)$ and that $\mathbf{E}$ is made of row vectors $\boldsymbol\epsilon_i = (\epsilon_{i1},...,\epsilon_{ip})$ for $i = 1,...,n$, where $ \boldsymbol\epsilon_i \stackrel{iid}{\sim} N_p(0, \boldsymbol\Sigma)$.
+
+The conditional likelihood for this model is given below.
+
+$$p(\mathbf{Y}|\mathbf{X}^*,\mathbf{B}^*,\boldsymbol\Sigma) \propto |\boldsymbol\Sigma|^{-n/2}\exp \left \{ -\frac{1}{2} \tr(\mathbf{Y} - \mathbf{X}^* \mathbf{B}^*)^T(\mathbf{Y} - \mathbf{X}^* \mathbf{B}^*)\boldsymbol\Sigma^{-1} \right \}$$
 
@@
 \subsubsection{Multivariate Normal Conditional Imputation}
+
+The multivariate normal conditional imputation derivations are given for a single cluster without loss of generality.
+In practive, the data and parameters in this section would be replaced by cluster specific estimates in the case of clustering.
+
+For a given observation vector $\mathbf{y} \sim N_p(\boldsymbol\mu, \boldsymbol\Sigma)$, we allow for missingness in at most $p - 1$ of the multivariate outcomes through the use of a conditional imputation step embedded within our Gibbs sampler.
+Suppose $\mathbf{y}$ contains $q$ missing observations and can be partitioned into two vectors $\mathbf{y_{1}}$ and $\mathbf{y_{2}}$ such that $\mathbf{y_{1}}$ is a $q \times 1$ vector of missing observations and $\mathbf{y_{2}}$ is a $(p-q) \times 1$ vector of complete observations.
+Similarly, partition $\boldsymbol\mu$ and $\boldsymbol\Sigma$ as follows.
+
+$$\boldsymbol\mu = \begin{bmatrix} \boldsymbol\mu_1 \\ \boldsymbol\mu_2 \end{bmatrix}$$
+
+$$\boldsymbol\Sigma = \begin{bmatrix} \boldsymbol\Sigma_{11} & \boldsymbol\Sigma_{12} \\ \boldsymbol\Sigma_{21} & \boldsymbol\Sigma_{22} \end{bmatrix}$$
+
+We will use these quantities to derive the conditional distribution $f(\mathbf{y_1}|\mathbf{y_2},\boldsymbol\mu,\boldsymbol\Sigma)$.
+
+\begin{align*} 
+f(\mathbf{y_1}|\mathbf{y_2},\boldsymbol\mu,\boldsymbol\Sigma) & \propto  f(\mathbf{y_1},\mathbf{y_2}|\boldsymbol\mu,\boldsymbol\Sigma) \\
+& \propto \exp \left \{ -\frac{1}{2}(\mathbf{y} - \boldsymbol\mu)^T \boldsymbol\Sigma^{-1} (\mathbf{y} - \boldsymbol\mu) \right \} \\
+& = \exp \left \{ -\frac{1}{2}\begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix}^T \boldsymbol\Sigma^{-1} \begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix} \right \} \\
+& =  \exp \left \{ -\frac{1}{2}\begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix}^T \begin{bmatrix} \boldsymbol\Sigma_{11} & \boldsymbol\Sigma_{12} \\ \boldsymbol\Sigma_{21} & \boldsymbol\Sigma_{22} \end{bmatrix}^{-1} \begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix} \right \} \\
+& = \exp \left \{ -\frac{1}{2}\begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix}^T \begin{bmatrix} \boldsymbol\Sigma_{11}^* & \boldsymbol\Sigma_{12}^* \\ \boldsymbol\Sigma_{21}^* & \boldsymbol\Sigma_{22}^* \end{bmatrix} \begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix} \right \}
+\end{align*}
+
+where the blockwise inversion formula is used to invert $\boldsymbol\Sigma$ as follows.
+
+\begin{align*}
+\boldsymbol\Sigma_{11}^* & = \boldsymbol\Sigma_{11}^{-1} + \boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12}(\boldsymbol\Sigma_{22} - \boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12})^{-1}\boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}
+\end{align*}
 

### 2019-05-23 manuscript/manuscript.tex
appendix

@@
 \usepackage{booktabs}
+\usepackage{breqn}
 
@@
 The multivariate skew-normal regression model can be written as follows. 
-
 $$\mathbf{Y} = \mathbf{X}\mathbf{B} + \mathbf{t} \boldsymbol\psi^T + \mathbf{E} = \mathbf{X}^* \mathbf{B}^* + \mathbf{E}$$
@@
 The conditional likelihood for this model is given below. 
+$$p(\mathbf{Y}|\mathbf{X}^*,\mathbf{B}^*,\boldsymbol\Sigma) \propto |\boldsymbol\Sigma|^{-n/2}\exp \left \{ -\frac{1}{2} \tr(\mathbf{Y} - \mathbf{X}^* \mathbf{B}^*)^T(\mathbf{Y} - \mathbf{X}^* \mathbf{B}^*)\boldsymbol\Sigma^{-1} \right \}$$
 
-$$p(\mathbf{Y}|\mathbf{X}^*,\mathbf{B}^*,\boldsymbol\Sigma) \propto |\boldsymbol\Sigma|^{-n/2}\exp \left \{ -\frac{1}{2} \tr(\mathbf{Y} - \mathbf{X}^* \mathbf{B}^*)^T(\mathbf{Y} - \mathbf{X}^* \mathbf{B}^*)\boldsymbol\Sigma^{-1} \right \}$$
+We choose conjugate priors for $\mathbf{B}^*$ and $\boldsymbol\Sigma$ as follows. 
+$$\boldsymbol\Sigma \sim \text{inverse-Wishart}(\mathbf{V}_0,\nu_0) $$
+$$\mathbf{B}^*|\boldsymbol\Sigma \sim MN_{(m+1)\times p}(\mathbf{B}_0^*,\mathbf{L}_0^{-1},\boldsymbol\Sigma)$$
+
+We now dervive the joint posterior distribution of the parameters $\mathbf{B}^*$ and $\boldsymbol\Sigma$.
+\begin{align*} 
+p(\mathbf{B}^*,\boldsymbol\Sigma|\mathbf{X}^*,\mathbf{Y}) & \propto  p(\mathbf{Y}|\mathbf{X}^*,\mathbf{B}^*,\boldsymbol\Sigma)p(\mathbf{B}^*|\boldsymbol\Sigma)p(\boldsymbol\Sigma)\\
+&  \propto |\boldsymbol\Sigma|^{-n/2}\exp \left \{ -\frac{1}{2} \tr((\mathbf{Y} - \mathbf{X^*}\mathbf{B}^*)^T(\mathbf{Y} - \mathbf{X^*}\mathbf{B}^*)\boldsymbol\Sigma^{-1})\right \} \\
+& \times |\boldsymbol\Sigma|^{-m/2} \exp \left \{ -\frac{1}{2} \tr((\mathbf{B}^* - \mathbf{B}^*_0)^T \mathbf{L}_0(\mathbf{B}^* - \mathbf{B}^*_0)\boldsymbol\Sigma^{-1})\right \} \\
+& \times |\boldsymbol\Sigma|^{(\nu_0 + p + 1)/2} \exp \left \{ -\frac{1}{2} \tr (\mathbf{V}_0 \boldsymbol\Sigma^{-1}) \right \}
+\end{align*}
 
@@
 Similarly, partition $\boldsymbol\mu$ and $\boldsymbol\Sigma$ as follows.
-
-$$\boldsymbol\mu = \begin{bmatrix} \boldsymbol\mu_1 \\ \boldsymbol\mu_2 \end{bmatrix}$$
-
-$$\boldsymbol\Sigma = \begin{bmatrix} \boldsymbol\Sigma_{11} & \boldsymbol\Sigma_{12} \\ \boldsymbol\Sigma_{21} & \boldsymbol\Sigma_{22} \end{bmatrix}$$
+$$\boldsymbol\mu = \begin{bmatrix} \boldsymbol\mu_1 \\ \boldsymbol\mu_2 \end{bmatrix} \ \ \ \ \ \boldsymbol\Sigma = \begin{bmatrix} \boldsymbol\Sigma_{11} & \boldsymbol\Sigma_{12} \\ \boldsymbol\Sigma_{21} & \boldsymbol\Sigma_{22} \end{bmatrix}$$
 
 We will use these quantities to derive the conditional distribution $f(\mathbf{y_1}|\mathbf{y_2},\boldsymbol\mu,\boldsymbol\Sigma)$. 
-
 \begin{align*} 
@@
 & =  \exp \left \{ -\frac{1}{2}\begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix}^T \begin{bmatrix} \boldsymbol\Sigma_{11} & \boldsymbol\Sigma_{12} \\ \boldsymbol\Sigma_{21} & \boldsymbol\Sigma_{22} \end{bmatrix}^{-1} \begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix} \right \} \\
-& = \exp \left \{ -\frac{1}{2}\begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix}^T \begin{bmatrix} \boldsymbol\Sigma_{11}^* & \boldsymbol\Sigma_{12}^* \\ \boldsymbol\Sigma_{21}^* & \boldsymbol\Sigma_{22}^* \end{bmatrix} \begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix} \right \}
+& = \exp \left \{ -\frac{1}{2}\begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix}^T \begin{bmatrix} \boldsymbol\Sigma_{11}^* & \boldsymbol\Sigma_{12}^* \\ \boldsymbol\Sigma_{21}^* & \boldsymbol\Sigma_{22}^* \end{bmatrix} \begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix} \right \} \\
+& = \exp \left \{ -\frac{1}{2} \left[ (\mathbf{y}_1 - \boldsymbol\mu_{cond})^T \boldsymbol\Sigma_{cond}^{-1}(\mathbf{y}_1 - \boldsymbol\mu_{cond})\right]\right \} 
 \end{align*}
+$$\Rightarrow \mathbf{y_1}|\mathbf{y_2},\boldsymbol\mu,\boldsymbol\Sigma \sim N_q(\boldsymbol\mu_{cond},\boldsymbol\Sigma_{cond})$$
+$$\boldsymbol\mu_{cond} = \boldsymbol\mu_1 + \boldsymbol\Sigma_{12}\boldsymbol\Sigma_{22}^{-1}(\mathbf{y}_2 - \boldsymbol\mu_2), \ \ \ \ \ \ \boldsymbol\Sigma_{cond} = \boldsymbol\Sigma_{11} - \boldsymbol\Sigma_{12}\boldsymbol\Sigma_{22}^{-1}\boldsymbol\Sigma_{21}$$
 
-where the blockwise inversion formula is used to invert $\boldsymbol\Sigma$ as follows.
-
+The blockwise inversion formula was used to invert $\boldsymbol\Sigma$ according to the following reparameterizations.
 \begin{align*}
-\boldsymbol\Sigma_{11}^* & = \boldsymbol\Sigma_{11}^{-1} + \boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12}(\boldsymbol\Sigma_{22} - \boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12})^{-1}\boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}
+\boldsymbol\Sigma_{11}^* & = \boldsymbol\Sigma_{11}^{-1} + \boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12}(\boldsymbol\Sigma_{22} - \boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12})^{-1}\boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\\
+\boldsymbol\Sigma_{12}^* & = -\boldsymbol\Sigma_{11} \boldsymbol\Sigma_{12}(\boldsymbol\Sigma_{22}-\boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12})^{-1}\\
+\boldsymbol\Sigma_{21}^* & = -(\boldsymbol\Sigma_{22} - \boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12})^{-1}\boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\\
+\boldsymbol\Sigma_{22}^* & = (\boldsymbol\Sigma_{22} - \boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12})^{-1}     
 \end{align*}
@@
 
-\begin{thebibliography}{}
-
-\bibitem{ } Cox, D. R. (1972). Regression models and life tables (with
-discussion).  \textit{Journal of the Royal Statistical Society, Series B}
-\textbf{34,} 187--200.
-
-\bibitem{ }  Hastie, T., Tibshirani, R., and Friedman, J. (2001). \textit{The 
-Elements of Statistical Learning: Data Mining, Inference, and Prediction}.
-New York: Springer.
-
-\end{thebibliography}
+\newpage
 

### 2019-05-23 manuscript/manuscript.tex
appendix

@@
 
+\newpage
+
 \section{Nurture Study}
@@
 \subsubsection{Attrition and Intermittent Missingness}
+
+\newpage
 
@@
 
+\begin{itemize}
+
+\item Emphasize that PG data augmentation for the multinomial model results in a PG mixure of experts model, which is a computationally efficient way to model edge weights.
+
+\end{itemize}
+
+\subsubsection{Prior Choice}
+
 \subsubsection{MCMC Algorithm}
@@
 
-\subsubsection{Prior Choice}
+\newpage
 
@@
 
-\subsection{Simulation to Assess Sensitivity to Mispecified H}
+\subsection{Simulation to Assess Sensitivity to Mispecified K}
+
+\newpage
 
@@
 
+\begin{itemize}
+
+\item Include both time varying and non-time varying covariates for the within cluster covariate set.
+
+\end{itemize}
+
+\newpage
+
 \section{Discussion}
 \label{s:discuss}
+
+\begin{itemize}
+
+\item Discuss how we handle non-ignorable missingness
+
+\end{itemize}
 
@@
 
-    \item $\mathbf{Y}$: A $n \times p$ matrix containing all multivariate skew-normal regression outcomes such that $y_{ij}$ is the $j^{th}$ outcome observed for subject $i$, where $i = 1,...,n$ and $j = 1,...p$.
+    \item $\mathbf{Y}$: A $N \times J$ matrix containing all multivariate skew-normal  outcomes such that $y_{ij}$ is the $j^{th}$ outcome observed for subject $i$, where $i = 1,...,n$ and $j = 1,...J$.
 
-    \item $\mathbf{X}$: A $n \times m$ matrix containing all multivariate skew-normal regression covariates such that $x_{ij}$ is the $j^{th}$ covariate value for subject $i$.
+    \item $\mathbf{X}$: A $n \times p$ matrix containing all multivariate skew-normal regression covariates such that $x_{ij}$ is the $j^{th}$ covariate value for subject $i$.
 
@@
 
-    \item $\mathbf{t}$: A $n \times 1$ vector of truncated normal random effects used in the stochastic representation of the multivariate skew-normal distribution.
-For $i = 1,...,n$, $t_i \stackrel{iid}{\sim}T_{[0,\infty)(0,1)}$
+    \item $\mathbf{t}$: An $n \times 1$ vector of truncated normal random effects used in the stochastic representation of the multivariate skew-normal distribution.
+For $i = 1,...,n$, $t_i \stackrel{iid}{\sim}T_{[0,\infty)}(0,1)$
 
@@
 To make the extension to the case where more than one cluster is specified, simply apply these distributional forms to cluster specific parameters and data.
+Finally, we assume for the moment that we have complete data for all outcomes for each subject.
+We extend consider the case of missing data in section (INSERT SECTION).
 
-The multivariate skew-normal regression model can be written as follows. 
+The multivariate skew-normal regression model can be written as follows in matrix form. 
 $$\mathbf{Y} = \mathbf{X}\mathbf{B} + \mathbf{t} \boldsymbol\psi^T + \mathbf{E} = \mathbf{X}^* \mathbf{B}^* + \mathbf{E}$$
-
+The matrix $\mathbf{Y}$ is of dimension $n \times p$.
 For convenience, we define $\mathbf{X}^*$ as a $n \times (m + 1)$ matrix constructed by column binding $\mathbf{t}$ to $\mathbf{X}$, and $\mathbf{B}^*$ as a $(m+1) \times p$ matrix constructed by row binding $\boldsymbol\psi^T$ to $\mathbf{B}$.
@@
 We choose conjugate priors for $\mathbf{B}^*$ and $\boldsymbol\Sigma$ as follows. 
-$$\boldsymbol\Sigma \sim \text{inverse-Wishart}(\mathbf{V}_0,\nu_0) $$
-$$\mathbf{B}^*|\boldsymbol\Sigma \sim MN_{(m+1)\times p}(\mathbf{B}_0^*,\mathbf{L}_0^{-1},\boldsymbol\Sigma)$$
+$$\boldsymbol\Sigma \sim \text{inverse-Wishart}(\mathbf{V}_0,\nu_0)$$
+
+$$\mathbf{B}^*|\boldsymbol\Sigma \sim MatNorm_{(m+1), p}(\mathbf{B}_0^*,\mathbf{L}_0^{-1},\boldsymbol\Sigma)$$
 
@@
 p(\mathbf{B}^*,\boldsymbol\Sigma|\mathbf{X}^*,\mathbf{Y}) & \propto  p(\mathbf{Y}|\mathbf{X}^*,\mathbf{B}^*,\boldsymbol\Sigma)p(\mathbf{B}^*|\boldsymbol\Sigma)p(\boldsymbol\Sigma)\\
-&  \propto |\boldsymbol\Sigma|^{-n/2}\exp \left \{ -\frac{1}{2} \tr((\mathbf{Y} - \mathbf{X^*}\mathbf{B}^*)^T(\mathbf{Y} - \mathbf{X^*}\mathbf{B}^*)\boldsymbol\Sigma^{-1})\right \} \\
-& \times |\boldsymbol\Sigma|^{-m/2} \exp \left \{ -\frac{1}{2} \tr((\mathbf{B}^* - \mathbf{B}^*_0)^T \mathbf{L}_0(\mathbf{B}^* - \mathbf{B}^*_0)\boldsymbol\Sigma^{-1})\right \} \\
+&  \propto |\boldsymbol\Sigma|^{-n/2}\exp \left \{ -\frac{1}{2} \tr \left [(\mathbf{Y} - \mathbf{X^*}\mathbf{B}^*)^T(\mathbf{Y} - \mathbf{X^*}\mathbf{B}^*)\boldsymbol\Sigma^{-1} \right ]\right \} \\
+& \times |\boldsymbol\Sigma|^{-(m+1)/2} \exp \left \{ -\frac{1}{2} \tr \left [(\mathbf{B}^* - \mathbf{B}^*_0)^T \mathbf{L}_0(\mathbf{B}^* - \mathbf{B}^*_0)\boldsymbol\Sigma^{-1} \right ]\right \} \\
 & \times |\boldsymbol\Sigma|^{(\nu_0 + p + 1)/2} \exp \left \{ -\frac{1}{2} \tr (\mathbf{V}_0 \boldsymbol\Sigma^{-1}) \right \}

### 2019-05-24 manuscript/manuscript.tex
references

@@
 
+\begin{thebibliography}{}
+
+\bibitem{ } Franczak BC, Tortora C, Browne RP, McNicholas PD.
+Unsupervised learning via mixtures of skewed distributions with hypercube contours.
+\textit{Pattern Recognition Letters}.
+2015 Jun 1;58:69-76.
+
+\bibitem{ } Ganjali M, Baghfalaki T.
+A Bayesian shared parameter model for analysing longitudinal skewed responses with nonignorable dropout.
+\textit{International Journal of Statistics in Medical Research}.
+2014 Apr 1;3(2):103.
+
+\bibitem{ } Cox, D. R. (1972). Regression models and life tables (with
+discussion).  \textit{Journal of the Royal Statistical Society, Series B}
+\textbf{34,} 187--200.
+
+\bibitem{ }  Hastie, T., Tibshirani, R., and Friedman, J. (2001). \textit{The 
+Elements of Statistical Learning: Data Mining, Inference, and Prediction}.
+New York: Springer.
+
+\end{thebibliography}
+
 % \appendix

### 2019-05-24 manuscript/manuscript.tex
references

@@
 
-\bibitem{ } Cox, D. R. (1972). Regression models and life tables (with
-discussion).  \textit{Journal of the Royal Statistical Society, Series B}
-\textbf{34,} 187--200.
+\bibitem{ } Lagona F, Picone M. Model-based clustering of multivariate skew data with circular components and missing values. \textit{Journal of Applied Statistics}. 2012 May 1;39(5):927-45.
 
-\bibitem{ }  Hastie, T., Tibshirani, R., and Friedman, J. (2001). \textit{The 
-Elements of Statistical Learning: Data Mining, Inference, and Prediction}.
-New York: Springer.
+\bibitem{ } Lee SX, McLachlan GJ. Model-based clustering and classification with non-normal mixture distributions. Statistical Methods \& Applications. 2013 Nov 1;22(4):427-54.
+
+\bibitem{ } Fr\"{u}hwirth-Schnatter S, Pyne S. Bayesian inference for finite mixtures of univariate and multivariate skew-normal and skew-t distributions. Biostatistics. 2010 Jan 27;11(2):317-36.
+
+\bibitem{ } Tiao GC, Zellner A. On the Bayesian estimation of multivariate regression. \textit{Journal of the Royal Statistical Society}: Series B (Methodological). 1964 Jul;26(2):277-85.
+
+\bibitem{ } Zhou X, Reiter JP. A note on Bayesian inference after multiple imputation. \textit{The American Statistician}. 2010 May 1;64(2):159-63.
+
+\bibitem{ } Vrbik I, Mcnicholas PD. Parsimonious skew mixture models for model-based clustering and classification. \textit{Computational Statistics \& Data Analysis}. 2014 Mar 1;71:196-210.
+
+\bibitem{ } Melnykov V, Maitra R. Finite mixture models and model-based clustering. \textit{Statistics Surveys}. 2010;4:80-116.
+
+\bibitem{ } Lee SX, Mclachlan GJ. On mixtures of skew normal and skew $$ t $$-distributions. \textit{Advances in Data Analysis and Classification}. 2013 Sep 1;7(3):241-66.
+
+\bibitem{ } Lin TI, Wang WL, McLachlan GJ, Lee SX.
+Robust mixtures of factor analysis models using the restricted multivariate skew-t distribution.
+\textit{Statistical Modelling}.
+2018 Feb;18(1):50-72.
 

### 2019-05-29 manuscript/manuscript.tex
references
poster

@@
 
-\title[]{A multivariate skew-normal finite mixture model for analysis of infant development trajectories}
+\title[]{Bayesian multivariate skew-normal finite mixture model for analysis of infant development trajectories}
 
@@
 \label{s:intro}
+
+\subsection{Existing Approaches}
+
+Mixtures of multivariate non-symmetric distributions such as the multivariate skew-normal (MSN) distribution allow for the nuances of the marginal density to be captured with a more parsimonious set of mixture components.
+Mixtures of MSN distributions has been dealt with previously in a Bayesian context (Fr\"{u}hwirth-Schnatter \& Pyne, 2010), however in these models, focus lies primary on marginal density estimation and inference on the mixture components (i.e.
+clusters) is not discussed.
+More recently, the mixtures of skew-$t$ factor analysis (MSTFA) model has been proposed be applied in settings where cluster-specific inference is of primary interest.
+However, thid
 
@@
 
+\bibitem{ } Arellano‐Valle RB, Azzalini A. On the unification of families of skew‐normal distributions. \textit{Scandinavian Journal of Statistics}. 2006 Sep;33(3):561-74.
+
+\bibitem{ } Azzalini A. A class of distributions which includes the normal ones. \textit{Scandinavian journal of statistics}. 1985 Jan 1:171-8.
+
+\bibitem{ } Azzalini, A. and Dalla Valle, A. (1996). The multivariate skew normal distribution. Biometrika 83, 715–726.
+
 \bibitem{ } Franczak BC, Tortora C, Browne RP, McNicholas PD.
@@
 2015 Jun 1;58:69-76.
+
+\bibitem{ } Fr\"{u}hwirth-Schnatter S, Pyne S. Bayesian inference for finite mixtures of univariate and multivariate skew-normal and skew-t distributions. Biostatistics. 2010 Jan 27;11(2):317-36.
 
@@
 
-\bibitem{ } Fr\"{u}hwirth-Schnatter S, Pyne S. Bayesian inference for finite mixtures of univariate and multivariate skew-normal and skew-t distributions. Biostatistics. 2010 Jan 27;11(2):317-36.
-
-\bibitem{ } Tiao GC, Zellner A. On the Bayesian estimation of multivariate regression. \textit{Journal of the Royal Statistical Society}: Series B (Methodological). 1964 Jul;26(2):277-85.
-
-\bibitem{ } Zhou X, Reiter JP. A note on Bayesian inference after multiple imputation. \textit{The American Statistician}. 2010 May 1;64(2):159-63.
-
-\bibitem{ } Vrbik I, Mcnicholas PD. Parsimonious skew mixture models for model-based clustering and classification. \textit{Computational Statistics \& Data Analysis}. 2014 Mar 1;71:196-210.
-
-\bibitem{ } Melnykov V, Maitra R. Finite mixture models and model-based clustering. \textit{Statistics Surveys}. 2010;4:80-116.
-
-\bibitem{ } Lee SX, Mclachlan GJ. On mixtures of skew normal and skew $$ t $$-distributions. \textit{Advances in Data Analysis and Classification}. 2013 Sep 1;7(3):241-66.
+\bibitem{ } Lee SX, Mclachlan GJ. On mixtures of skew normal and skew $t$-distributions. \textit{Advances in Data Analysis and Classification}. 2013 Sep 1;7(3):241-66.
 
@@
 2018 Feb;18(1):50-72.
+
+\bibitem{ } Melnykov V, Maitra R. Finite mixture models and model-based clustering. \textit{Statistics Surveys}. 2010;4:80-116.
+
+\bibitem{ } Tiao GC, Zellner A. On the Bayesian estimation of multivariate regression. \textit{Journal of the Royal Statistical Society}: Series B (Methodological). 1964 Jul;26(2):277-85.
+
+\bibitem{ } Vrbik I, Mcnicholas PD. Parsimonious skew mixture models for model-based clustering and classification. \textit{Computational Statistics \& Data Analysis}. 2014 Mar 1;71:196-210.
+
+\bibitem{ } Zhou X, Reiter JP. A note on Bayesian inference after multiple imputation. \textit{The American Statistician}. 2010 May 1;64(2):159-63.
 

### 2019-05-29 poster/poster.tex
manuscript format
outline
poster

@@
 \includegraphics[width=0.9\linewidth]{bayley_dens.jpg}
-\caption{Density of Bayley composite scores for Nurture infants at 3, 6, 9 and 12 months of age.}
+\caption{Residuals of repeated measures model of Bayley scores for infants at 3, 6, 9 and 12 months of age.}
 \end{figure}
@@
 
-$$P(Z_i = l|\w_i) = \pi_{il} = \frac{e^{w_i^T \delta_l}}{\sum_{r = 1}^h e^{w_i^T \delta_r}}$$
+$$P(Z_i = l|w_i) = \pi_{il} = \frac{e^{w_i^T \delta_l}}{\sum_{r = 1}^h e^{w_i^T \delta_r}}$$
 
@@
 
-\begin{figure}
-\includegraphics[height=0.9\linewidth,width=1\linewidth]{beta_trace.pdf}
-\caption{Trace plots of regression parameter estimates in simulation}
-\end{figure}
+\begin{block}{Application Results}
 
-\begin{block}{Application Results}
+\begin{table}[]
+\caption{\label{tab:unnamed-chunk-4}Estimated effects of sex and race on odds of cluster membership relative to the reference cluster.}
+\centering
+\begin{tabular}{lll}
+\hline
+Variable & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Class 2\\ $\hat{\text{OR}}$ (95\% CrI)\end{tabular}} & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Class 3\\ $\hat{\text{OR}}$ (95\% CrI)\end{tabular}} \\ \hline
+Intercept & 1.28 (1.01, 1.73) & 1.16 (0.87, 1.71) \\
+Sex (Female) & 0.69 (0.40, 0.96) & 0.77 (0.48, 1.05) \\
+Race (White) & 0.90 (0.57, 1.32) & 0.83 (0.59, 0.29) \\ \hline
+\end{tabular}
+\end{table}
+
+\
+
+In Table 2, we demonstrate the capability of estimating the effect of covariates (race and sex here) on posterior probability of cluster membership. We use class 1 as the reference class.
 
@@
 \end{figure}
+
+Above is a plot of the mean Bayley score for each of the fitted clusters.
+We observe similar development patterns but different baseline development between clusters 1 and 2, and a qualitatively different development pattern in class 3 than in the classes 1 and 2.
 
@@
 
-\tiny \textbf{Funding:} \textit{This work is supported by a grant from the NIH (\textbf{R01DK094841})}.
+\tiny \textbf{Funding:} \textit{This work is supported by a grant from the NIH (\textbf{R01DK094841}) and a grand from the NLM (\textbf{1R21LM012866-01})}.
 

### 2019-05-29 poster/poster.tex
poster

@@
 Sex (Female) & 0.69 (0.40, 0.96) & 0.77 (0.48, 1.05) \\
-Race (White) & 0.90 (0.57, 1.32) & 0.83 (0.59, 0.29) \\ \hline
+Race (White) & 0.90 (0.57, 1.32) & 0.83 (0.59, 1.29) \\ \hline
 \end{tabular}

### 2019-05-29 poster/poster.tex
poster

@@
 It is often also of interest to relate covariates to mean longitudinal growth patterns.
-Infant development data are inherently correlated longitudinally, often skewed, and frequently missing due to longitudinal attrition and standard practices are ill-suited to addressing these research questions due to their ignorance of one or more of these features of development data.
+Infant development data are inherently correlated longitudinally, often skewed, and frequently missing due to longitudinal attrition and standard practices are ill-suited for such analyses because they fail to account for one or more of these features of the data.
 
@@
 \includegraphics[width=0.9\linewidth]{bayley_dens.jpg}
-\caption{Residuals of repeated measures model of Bayley scores for infants at 3, 6, 9 and 12 months of age.}
+\caption{Residuals of repeated measures model of Bayley motor development scores for infants at 3, 6, 9 and 12 months of age.}
 \end{figure}
@@
 
-where $w_i$ is the vector of class probability covariates for subject $i$, $\delta_l$ contains the multinomial regression parameters for class $l$, and $h$ is the number of putative clusters specified \textit{a priori}.
-
-During our MCMC estimation procedure, the class labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as class assignments.
+where $Z_i$ is a latent clustering indicator, $w_i$ is the vector of class probability covariates for subject $i$ ($i = 1,...,n$), $\delta_l$ contains the multinomial regression parameters for class $l$, and $h$ is the number of putative clusters specified \textit{a priori}.
+During MCMC estimation, class labels $Z_i$ are updated from used as class assignments in the remaining MCMC steps.
 
@@
 We model the effect of covariates on longitudinal development outcomes through the use of a MSN regression model.
-The MSN distribution can be represented as the superposition of a MN random variable with a latent truncated normal random effect.
-Let $\mathbf{Y}_{n \times k}$ be the observation matrix such that $Y_{ij}$ is the observation for subject $i$ at timepoint $j$.
+The MSN distribution can be represented as a multivariate normal random variable with a latent truncated normal random effect, $t$.
+Let $\mathbf{y}_{i}$ be the $k \times 1$ outcome vector for subject $i$.
 
-$$\mathbf{Y}_{n \times k} = \mathbf{X}_{n \times p}\boldsymbol\beta_{p \times k} + t_{n \times 1}\psi_{1 \times k} + \boldsymbol\epsilon_{n \times k}$$
+$$\mathbf{y}_{i} = \boldsymbol\beta^T \mathbf{x}_{i} + t_{i}\boldsymbol\psi + \boldsymbol\epsilon_{i}$$
 
-where $X_i$ is the $1 \times p$ vector of covariate values for subject $i$, $\beta_j$ is the $i \times k$ vector of fixed effects coefficients for timepoint $j$, $t_i \stackrel{iid}{\sim} N_{[0,\infty)}(0,1)$ is a truncated normal random effect, $\psi$ is the vector containing skewness parameters for each timepoint, and $\boldsymbol\epsilon_i \sim N_k(0,\boldsymbol\Sigma_{k \times k})$ is the correlated error term.
+where $\mathbf{x}_i$ is the $p \times 1$ vector of covariate values for subject $i$, $\boldsymbol\beta$ is the $p \times k$ vector of fixed effects coefficients , $t_i \stackrel{iid}{\sim} N_{[0,\infty)}(0,1)$ is a conditional truncated normal random effect, $\psi_l$ is the vector containing cluster-specific skewness parameters for each timepoint, and $\boldsymbol\epsilon_i \sim N_k(0,\boldsymbol\Sigma_{k \times k})$ is the correlated error term.
 
@@
 \centering
-\begin{tabular}{lll}
+\fontsize{30}{30}\selectfont
+\def\arraystretch{1.5}
+\begin{tabular}{llll}
 \hline
-Variable & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Class 2\\ $\hat{\text{OR}}$ (95\% CrI)\end{tabular}} & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Class 3\\ $\hat{\text{OR}}$ (95\% CrI)\end{tabular}} \\ \hline
-Intercept & 1.28 (1.01, 1.73) & 1.16 (0.87, 1.71) \\
-Sex (Female) & 0.69 (0.40, 0.96) & 0.77 (0.48, 1.05) \\
-Race (White) & 0.90 (0.57, 1.32) & 0.83 (0.59, 1.29) \\ \hline
+Variable & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Class 1\\ $\hat{\text{OR}}$ (95\% CrI)\end{tabular}} & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Class 2\\ $\hat{\text{OR}}$ (95\% CrI)\end{tabular}} & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Class 3\\ $\hat{\text{OR}}$ (95\% CrI)\end{tabular}} \\ \hline
+Sex (F) & 1.80 (1.20, 3.16) & 1.25 (0.96, 1.81) & 1.42 (1.10, 2.03) \\
+Race (W) & 1.57 (1.07, 2.80) & 1.44 (1.07, 2.13) & 1.41 (1.04, 2.04) \\ \hline
 \end{tabular}
@@
 
-In Table 2, we demonstrate the capability of estimating the effect of covariates (race and sex here) on posterior probability of cluster membership. We use class 1 as the reference class.
+In Table 2, we demonstrate the capability of estimating the effect of covariates (race and sex here) on posterior odds of cluster membership.
 
@@
 Above is a plot of the mean Bayley score for each of the fitted clusters.
-We observe similar development patterns but different baseline development between clusters 1 and 2, and a qualitatively different development pattern in class 3 than in the classes 1 and 2.
+We observe similar development patterns but different baseline development between clusters 1 and 2, and a qualitatively different development pattern in cluster 3 than in the clusters 1 and 2.
 
@@
 (2010)}.
-Bayesian inference for finite mixtures of univariate ...
-Biostatistics; \textbf{(2) Benjamin-Neelon SE, Ostbye T, Bennett GG, et al.} Cohort profile for the Nurture Observational Study ...
-BMJ Open 2017.
+Bayesian inference for finite mixtures of univariate and multivariate skew-normal and skew-t distributions.
+Biostatistics.
+2010 Jan 27;11(2):317- 36.
+
+\textbf{(2) Benjamin-Neelon SE, Ostbye T, Bennett GG, et al.} Cohort profile for the Nurture Observational Study examining associations of multiple caregivers on infant growth in the Southeastern USA.
+BMJ open.
+2017 Feb 1;7(2):e013939.
 

### 2019-05-29 poster/poster.tex
poster

@@
 \centering
-\fontsize{30}{30}\selectfont
-\def\arraystretch{1.5}
-\begin{tabular}{llll}
+\begin{tabular}{lll}
 \hline
-Variable & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Class 1\\ $\hat{\text{OR}}$ (95\% CrI)\end{tabular}} & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Class 2\\ $\hat{\text{OR}}$ (95\% CrI)\end{tabular}} & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Class 3\\ $\hat{\text{OR}}$ (95\% CrI)\end{tabular}} \\ \hline
-Sex (F) & 1.80 (1.20, 3.16) & 1.25 (0.96, 1.81) & 1.42 (1.10, 2.03) \\
-Race (W) & 1.57 (1.07, 2.80) & 1.44 (1.07, 2.13) & 1.41 (1.04, 2.04) \\ \hline
+Variable & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Class 2\\ $\hat{\text{OR}}$ (95\% CrI)\end{tabular}} & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Class 3\\ $\hat{\text{OR}}$ (95\% CrI)\end{tabular}} \\ \hline
+(B,F)  & 0.87 (0.62, 1.16) & 0.91 (0.68, 1.22) \\
+(W,M)  & 1.12 (0.80, 1.68) & 1.04 (0.70, 1.55) \\
+(W,F) & 0.79 (0.53, 1.14) & 1.81 (1.54, 2.20) \\ \hline
 \end{tabular}

### 2019-05-30 poster/poster.tex
poster

@@
 \hline
-Variable & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Class 2\\ $\hat{\text{OR}}$ (95\% CrI)\end{tabular}} & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Class 3\\ $\hat{\text{OR}}$ (95\% CrI)\end{tabular}} \\ \hline
+Variable & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Class 2\\ $\hat{\text{O}}$ (95\% CrI)\end{tabular}} & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Class 3\\ $\hat{\text{O}}$ (95\% CrI)\end{tabular}} \\ \hline
 (B,F)  & 0.87 (0.62, 1.16) & 0.91 (0.68, 1.22) \\

### 2019-06-03 manuscript/manuscript.tex
refs

@@
 Mixtures of multivariate non-symmetric distributions such as the multivariate skew-normal (MSN) distribution allow for the nuances of the marginal density to be captured with a more parsimonious set of mixture components.
-Mixtures of MSN distributions has been dealt with previously in a Bayesian context (Fr\"{u}hwirth-Schnatter \& Pyne, 2010), however in these models, focus lies primary on marginal density estimation and inference on the mixture components (i.e.
+Mixtures of MSN distributions have been dealt with previously in a Bayesian context (Fr\"{u}hwirth-Schnatter \& Pyne, 2010), however in these models, focus lies primary on marginal density estimation and inference on the mixture components (i.e.
 clusters) is not discussed.
-More recently, the mixtures of skew-$t$ factor analysis (MSTFA) model has been proposed be applied in settings where cluster-specific inference is of primary interest.
-However, thid
+More recently, the mixtures of skew-$t$ factor analysis (MSTFA) model has been proposed for settings in which cluster-specific inference is of primary interest (Lin \textit{et al.} 2018).
+However, an important feature not included in the MSTFA is the ability to explain individual-level cluster membership as a function of covariates of interest.
+Additionally, parameter estimation proposed by Lin et al.
+for the MSTFA relies on a prohibitively complex EM algorithm and does not enjoy the inferential benefits of a Bayesian approach.
+Our proposed model improves on these previous works by estimating parameters in a Bayesian framework as well as including the ability to fit a multinomial logit regression to cluster membership probabilities using a novel application of data augmentation with the P\'olya Gamma distribution.
+
+A ubiquitous feature of repeated measures studies is loss of data due to intermittent missingness and attrition.
+In the Bayesian setting, the standard approach to dealing with missing data is to perform multiple imputation, whereby $m$ imputed data sets are generated from a specified imputation model.
+After $m$ complete data sets are obtained, parameter estimates are combined across each data set to produce a final set of parameter estimates (Gelman \textit{et al.} 2013).
+This approach is not only computationally burdensome, requiring storage and analysis of an $m \times n_{rows} \times n_{cols}$ data array in addition to multiplication of total model run time by a factor of $m$, but it has been shown to produce unreliable inferences (Zhou and Reiter, 2010).
+We instead
 
@@
 
-\bibitem{ } Arellano‐Valle RB, Azzalini A. On the unification of families of skew‐normal distributions. \textit{Scandinavian Journal of Statistics}. 2006 Sep;33(3):561-74.
+\bibitem{ } Arellano‐-Valle RB, Azzalini A. On the unification of families of skew‐normal distributions. \textit{Scandinavian Journal of Statistics}. 2006 Sep;33(3):561-74.
 
@@
 \bibitem{ } Azzalini, A. and Dalla Valle, A. (1996). The multivariate skew normal distribution. Biometrika 83, 715–726.
+
+\bibitem{ } Neelon SE, \O stbye T, Bennett GG, Kravitz RM, Clancy SM, Stroo M, Iversen E, Hoyo C.
+Cohort profile for the Nurture Observational Study examining associations of multiple caregivers on infant growth in the Southeastern USA.
+\textit{BMJ Open}.
+2017 Feb 1;7(2):e013939.
 
@@
 
-\bibitem{ } Fr\"{u}hwirth-Schnatter S, Pyne S. Bayesian inference for finite mixtures of univariate and multivariate skew-normal and skew-t distributions. Biostatistics. 2010 Jan 27;11(2):317-36.
+\bibitem{ } Fr\"{u}hwirth-Schnatter S, Pyne S.
+Bayesian inference for finite mixtures of univariate and multivariate skew-normal and skew-t distributions.
+\textit{Biostatistics}.
+2010 Jan 27;11(2):317-36.
 
@@
 2014 Apr 1;3(2):103.
+
+\bibitem{ } Gelman A, Stern HS, Carlin JB, Dunson DB, Vehtari A, Rubin DB. Bayesian data analysis. \textit{Chapman and Hall/CRC}; 2013 Nov 27.
+
+\bibitem{ } Holmes CC, Held L. Bayesian auxiliary variable models for binary and multinomial regression. \textit{Bayesian Analysis}. 2006;1(1):145-68.
 
@@
 
+\bibitem{ } Luo S, Lawson AB, He B, Elm JJ, Tilley BC.
+Bayesian multiple imputation for missing multivariate longitudinal data from a Parkinson's disease clinical trial.
+\textit{Statistical Methods in Medical Research}.
+2016 Apr;25(2):821-37.
+
 \bibitem{ } Melnykov V, Maitra R. Finite mixture models and model-based clustering. \textit{Statistics Surveys}. 2010;4:80-116.
+
+\bibitem{ } Polson NG, Scott JG, Windle J.
+Bayesian inference for logistic models using P\'olya--Gamma latent variables.
+\textit{Journal of the American Statistical Association}.
+2013 Dec 1;108(504):1339-49.
 

### 2019-06-04 manuscript/manuscript.tex
approaches

@@
 Additionally, parameter estimation proposed by Lin et al.
-for the MSTFA relies on a prohibitively complex EM algorithm and does not enjoy the inferential benefits of a Bayesian approach.
+for the MSTFA relies on a prohibitively complex EM algorithm and does not enjoy the inferential benefits of a Bayesian approach, namely the ability to incorporate prior information into a model and make posterier probability statements.
 Our proposed model improves on these previous works by estimating parameters in a Bayesian framework as well as including the ability to fit a multinomial logit regression to cluster membership probabilities using a novel application of data augmentation with the P\'olya Gamma distribution.

### 2019-06-04 manuscript/manuscript.tex
approaches

@@
 
+\textbf{\textit{Put lit review of Bayesian multinomial logistic regression here}}
+
 A ubiquitous feature of repeated measures studies is loss of data due to intermittent missingness and attrition.
@@
 This approach is not only computationally burdensome, requiring storage and analysis of an $m \times n_{rows} \times n_{cols}$ data array in addition to multiplication of total model run time by a factor of $m$, but it has been shown to produce unreliable inferences (Zhou and Reiter, 2010).
-We instead
+We instead include an ``online" imputation imputation step in our Gibbs sampling procedure, whereby missing outcomes are updated at each iteration.
+This approach greatly increases the number of opportunities for exploration of the missing data parameter space.
 

### 2019-06-04 manuscript/manuscript.tex
approaches

@@
 
+\subsection{Generic Mixture}
+
+The general formulation of the multivariate skew normal (MSN) is as follows.
+
+$$f(\mathbf{y}_i) = \sum_{k = 1}^{K} \pi_{ik} f(\mathbf{y}_i|\boldsymbol\theta_k)$$
+
 \subsection{Multivariate Skew Normal Regression}
@@
 We model the effect of covariates on longitudinal development outcomes through the use of a MSN regression model.
-The MSN distribution can be represented as the superposition of a MN random variable with a latent truncated normal random effect.
-Let $\mathbf{Y}_{n \times k}$ be the observation matrix such that $Y_{ij}$ is the observation for subject $i$ at timepoint $j$.
+We adopt a standard stochastic representation of a MSN random variable as the sum of a latent truncated normal random effect and a correlated multivariate normal random error term.
+More formally, Let $\mathbf{y}_{i}$ be the $J \times 1$ observation vector for subject $i$ such that $y_{ij}$ is the observation for subject $i$ at timepoint $j$.
+The MSN stochastic representation is as follows.
+$$\mathbf{y}_{i} = \boldsymbol\beta^T \mathbf{x}_{i} + t_{i}\boldsymbol\psi + \boldsymbol\epsilon_{i}$$
 
-$$\mathbf{Y}_{n \times k} = \mathbf{X}_{n \times p}\boldsymbol\beta_{p \times k} + t_{n \times 1}\psi_{1 \times k} + \boldsymbol\epsilon_{n \times k}$$
-
-where $X_i$ is the $1 \times p$ vector of covariate values for subject $i$, $\beta_j$ is the $i \times k$ vector of fixed effects coefficients for timepoint $j$, $t_i \stackrel{iid}{\sim} N_{[0,\infty)}(0,1)$ is a truncated normal random effect, $\psi$ is the vector containing skewness parameters for each timepoint, and $\boldsymbol\epsilon_i \sim N_k(0,\boldsymbol\Sigma_{k \times k})$ is the correlated error term.
+where $\mathbf{x}_i$ is the $p \times 1$ vector of covariate values for subject $i$, $\boldsymbol\beta$ is the $p \times k$ vector of fixed effects coefficients , $t_i \stackrel{iid}{\sim} N_{[0,\infty)}(0,1)$ is a conditional truncated normal random effect, $\boldsymbol\psi$ is the vector containing cluster-specific skewness parameters, and $\boldsymbol\epsilon_i \sim N_k(0,\boldsymbol\Sigma_{k \times k})$ is the correlated error term.
+In can be shown that under these conditions $\mathbf{y}_i$ is marginally multivariate skew normal.
 

### 2019-06-04 manuscript/manuscript.tex
refs
mods

@@
 
+\subsection{Infant Development Clustering}
+
+Heterogeneity of treatment effects (HTE) (Lanza and Rhoades, 2013).
+
 \subsection{Existing Approaches}
@@
 
-The general formulation of the multivariate skew normal (MSN) is as follows.
+Let $\mathbf{y}_{i}$ be the $J \times 1$ observation vector for subject $i$ such that $y_{ij}$ is the observation for subject $i$ at timepoint $j$.
+The general formulation of the multivariate skew normal mixure model is as follows.
 
 $$f(\mathbf{y}_i) = \sum_{k = 1}^{K} \pi_{ik} f(\mathbf{y}_i|\boldsymbol\theta_k)$$
+
+Here $\boldsymbol\theta_k$ is the set of parameters shared by subjects belonging to class $k$ ($k = 1,...,K$).
+The probability of subject $i$ ($i = 1,...,n$) belonging to class $k$ ($k = 1,...,K$) is denoted $\pi_{ik}$.
+Note that $\pi_{ik}$ is not indexed by time and we do not allow for subjects switching classes over the study period.
+This feature of the model is a result of our interest in clustering based on development patterns rather than time-specific development scores.
+
+In each development class, we assume that a subject's $J\times 1$ outcome vector follows a multivariate skew normal (MSN) distribution.
+We define a latent variable $Z_i \in \{1,...,K\}$ to denote the cluster membership of subject $i$ for $i = 1,...,n$.
+We model $\mathbf{y}_{i}|Z_i$ as
+
+$$f(\mathbf{y}_{i_k}|Z_i )$$
 
@@
 We adopt a standard stochastic representation of a MSN random variable as the sum of a latent truncated normal random effect and a correlated multivariate normal random error term.
-More formally, Let $\mathbf{y}_{i}$ be the $J \times 1$ observation vector for subject $i$ such that $y_{ij}$ is the observation for subject $i$ at timepoint $j$.
 The MSN stochastic representation is as follows.
@@
 
+\bibitem{ } Lanza ST, Rhoades BL. Latent class analysis: an alternative perspective on subgroup analysis in prevention and treatment. \textit{Prevention Science}. 2013 Apr 1;14(2):157-68.
+
 \bibitem{ } Lee SX, McLachlan GJ. Model-based clustering and classification with non-normal mixture distributions. Statistical Methods \& Applications. 2013 Nov 1;22(4):427-54.

### 2019-06-04 manuscript/manuscript.tex
mods

@@
 Let $\mathbf{y}_{i}$ be the $J \times 1$ observation vector for subject $i$ such that $y_{ij}$ is the observation for subject $i$ at timepoint $j$.
+We assume for now that $\mathbf{y}_{i}$ is fully observed.
+Later, we relax this assumptin by allowing for missingness in the components of $\mathbf{y}_{i}$.
 The general formulation of the multivariate skew normal mixure model is as follows.
@@
 This feature of the model is a result of our interest in clustering based on development patterns rather than time-specific development scores.
+Later, we detail a multinomial logit regression model placed on $\pi_{ik}$ to allow for flexible modeling of class membership probabilities as a function of covariates of interest.
 
@@
 
-$$f(\mathbf{y}_{i_k}|Z_i )$$
+$$\mathbf{y}_{i}|Z_i = k \sim MSN_J(\boldsymbol\zeta_k,\boldsymbol\alpha_k,\boldsymbol\Omega_k)$$
+
+where $\boldsymbol\zeta_k$ is a $J \times 1$ vector of class specific location parameters, $\boldsymbol\alpha_k$ is a $J \times 1$ vector of class specific skewness parameters, and $\boldsymbol\Omega_k$ is a $J \times J$ class specific scale matrix.
 

### 2019-06-04 manuscript/manuscript.tex
mods

@@
 
-where $\mathbf{x}_i$ is the $p \times 1$ vector of covariate values for subject $i$, $\boldsymbol\beta$ is the $p \times k$ vector of fixed effects coefficients , $t_i \stackrel{iid}{\sim} N_{[0,\infty)}(0,1)$ is a conditional truncated normal random effect, $\boldsymbol\psi$ is the vector containing cluster-specific skewness parameters, and $\boldsymbol\epsilon_i \sim N_k(0,\boldsymbol\Sigma_{k \times k})$ is the correlated error term.
-In can be shown that under these conditions $\mathbf{y}_i$ is marginally multivariate skew normal.
+where $\mathbf{x}_i$ is the $P \times 1$ vector of covariate values for subject $i$, $\boldsymbol\beta$ is the $P \times J$ vector of fixed effects coefficients , $t_i \stackrel{iid}{\sim} N_{[0,\infty)}(0,1)$ is a conditional truncated normal random effect, $\boldsymbol\psi$ is the $J \times 1$ vector containing cluster-specific skewness parameters, and $\boldsymbol\epsilon_i \sim N_J(0,\boldsymbol\Sigma_{J \times J})$ is the correlated error term.
+In can be shown that under these conditions $\mathbf{y}_i$ is multivariate skew normal with location.
+For conciseness, we present the MSN distribution here outside of the finite mixture context.
+When using this stochastic representation in our finite mixture model, the parameters $\boldsymbol\beta$, $\boldsymbol\psi$, and $\boldsymbol\Sigma$ are replaced with class-specific counterparts $\boldsymbol\beta_k$, $\boldsymbol\psi_k$, and $\boldsymbol\Sigma_k$.
 
@@
 We accomplish this via multinomial logit regression model on cluster membership, which utilizes P\'{o}lya-Gamma data-aumentation to allow for updating of all parameters using Gibbs sampling.
-The multinomial logit model is as follows for $l = 1,...,h$.
+The multinomial logit model is as follows for $k = 1,...,K$.
 
-$$P(Z_i = l|w_i) = \pi_{il} = \frac{e^{w_i^T \delta_l}}{\sum_{r = 1}^h e^{w_i^T \delta_r}}$$
+$$P(Z_i = k|\mathbf{w}_i) = \pi_{ik} = \frac{e^{\mathbf{w}_i^T \boldsymbol\delta_k}}{\sum_{k' = 1}^K e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}}}$$
 
-where $w_i$ is the vector of class probability covariates for subject $i$, $\delta_l$ contains the multinomial regression parameters for class $l$, and $h$ is the number of putative clusters specified \textit{a priori}.
+where $\mathbf{w}_i$ is the vector of class probability covariates for subject $i$, $\boldsymbol\delta_k$ contains the multinomial regression parameters for class $k$, and $K$ is the number of putative clusters specified \textit{a priori}.
+For identifiability purposes, we fix the reference category $k = K$ and set $\boldsymbol\delta_K = \mathbf{0}$.
 
-During our MCMC estimation procedure, the class labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as class assignments.
+\[   P(Z_i = k) = \left \{
+\begin{array}{ll}
+      \pi_{ik} = \frac{e^{\mathbf{w}_i^T \boldsymbol\delta_k}}{1 + \sum_{k' = 1}^{K-1} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}}} & k < K \\
+      \\
+      \pi_{ik} = \frac{e^{\mathbf{w}_i^T \boldsymbol\delta_k}}{1 + \sum_{k' = 1}^{K-1} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}}} & k = K \\
+\end{array} 
+\right. \]
+
+\
+
+Under this model, $Z_i|\mathbf{w}_i \sim Multinom.(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{i1},...,\pi_{1K})$
+
+During MCMC estimation, the class labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as class assignments.
 
@@
 
-$$Y_i|X_i,t_i,\boldsymbol\beta,\psi \sim N_k(X_i \boldsymbol\beta + t_i \psi, \boldsymbol\Sigma)$$
+$$\mathbf{y}_i|Z_i,\mathbf{x}_i,t_i,\boldsymbol\beta_{Z_i},\boldsymbol\psi_{Z_i},\boldsymbol\Sigma_{Z_i} \sim N_J(\boldsymbol\beta_{Z_i}^T\mathbf{x}_i  + t_i \boldsymbol\psi_{Z_i}, \boldsymbol\Sigma_{Z_i})$$
 
-This allows us to appeal to standard conditional forms of the multivariate normal distribution. Let $Y_i = [Y^{miss}_{i_{q \times 1}} | Y^{obs}_{i_{k - q \times 1}}]^T$. We have
+This allows us to appeal to standard conditional forms of the multivariate normal distribution to specify the distribution of missing observations.
+For subject $i$, we use $q_i$ to denote the number of the total $J$ possible repeated measurements are missing.
+We partition the full $J \times 1$ outcome vector $\mathbf{y}_i$ into the $(J - q_i) \times 1$ observed data vector $\mathbf{y}_i^{obs}$ and the $q_i \times 1$ missing data vector $\mathbf{y}_i^{miss}$.
+$Y_i = [Y^{miss}_{i_{q \times 1}} | Y^{obs}_{i_{k - q \times 1}}]^T$.
+We have
 
-$$Y_i^{miss}|Y_i^{obs},X_i,t_i,\boldsymbol\beta,\psi \sim N(\mu^{miss},\boldsymbol\Sigma^{miss})$$
+$$\mathbf{y}_i^{miss}|\mathbf{y}_i^{obs},\mathbf{x}_i,t_i,Z_i,\boldsymbol\beta_{Z_i},\boldsymbol\psi_{Z_i},\boldsymbol\Sigma_{Z_i} \sim N(\boldsymbol\mu^{miss},\boldsymbol\Sigma^{miss})$$
 
-where $\mu^{miss}$ and $\Sigma^{miss}$ take standard forms.
-Each missing outcome is imputed "online", i.e.
+where $\boldsymbol\mu^{miss}$ and $\boldsymbol\Sigma^{miss}$ take standard forms detailed in the Appendix.
+
+An attractive feature of this imputation approach is that each missing outcome is imputed ``online", i.e.
 once per MCMC iteration.

### 2019-06-06 manuscript/manuscript.tex
algo
corrections

@@
 \usepackage{breqn}
+\usepackage{algorithm}
+\usepackage{algcompatible}
+\usepackage{algpseudocode}
+\usepackage{float}
+\usepackage{dsfont}
+\let\origfigure\figure
+\let\endorigfigure\endfigure
+\renewenvironment{figure}[1][2] {
+    \expandafter\origfigure\expandafter[H]
+} {
+    \endorigfigure
+}
 
@@
 
-\textbf{\textit{Put lit review of Bayesian multinomial logistic regression here}}
+\textbf{\textit{Put lit review of Bayesian PG multinomial logistic regression here}}
 
@@
 
-\[   P(Z_i = k) = \left \{
+\[   P(Z_i = k|\mathbf{w}_i,\boldsymbol\delta_k) = \left \{
 \begin{array}{ll}
@@
       \\
-      \pi_{ik} = \frac{e^{\mathbf{w}_i^T \boldsymbol\delta_k}}{1 + \sum_{k' = 1}^{K-1} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}}} & k = K \\
+      \pi_{ik} = \frac{1}{1 + \sum_{k' = 1}^{K-1} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}}} & k = K \\
 \end{array} 
@@
 
-\subsection{Conditional MVSN Imputation}
+\subsection{Conditional MVN Imputation}
 
@@
 We partition the full $J \times 1$ outcome vector $\mathbf{y}_i$ into the $(J - q_i) \times 1$ observed data vector $\mathbf{y}_i^{obs}$ and the $q_i \times 1$ missing data vector $\mathbf{y}_i^{miss}$.
-$Y_i = [Y^{miss}_{i_{q \times 1}} | Y^{obs}_{i_{k - q \times 1}}]^T$.
+$Y_i = [Y^{miss}_{i_{q \times 1}} | Y^{obs}_{i_{J - q \times 1}}]^T$.
 We have
@@
 This provides more opportunities to explore the parameter space than multiple imputation and avoids multiplicative run-time scaling in $m$, the number of imputations.
+
+\textit{\textbf{Discuss non-ignorable missingness here}}
 
@@
 \subsubsection{MCMC Algorithm}
+
+\begin{algorithm}
+\caption{Gibbs Sampler}
+\begin{algorithmic}
+\small
+\linespread{0.5}
+    \State $\mathbf{Define} \ n_{iter};\  n_{burn}; \ K; \ \boldsymbol\theta_{init}; \ \boldsymbol\theta_0$
+    \State $n_{sim} := n_{iter} - n_{burn}$
+    \State $\boldsymbol\theta := \boldsymbol\theta_{init}$
+    \For {$\iota = 1,...,n_{sim}$}
+        \State \textsc{I. Conditional Imputation}
+          \For {$ i = 1,...,n$}
+            \State $\mathbf{Draw} \ \mathbf{y}_i^{miss} \ \text{from} \ N_q(\boldsymbol\mu_i^{miss}, \boldsymbol\Sigma_i^{miss})$
+            \State $\mathbf{y}_i := \mathbf{y}_i^{miss} \cup \mathbf{y}_i^{obs}$
+         \EndFor
+        \State \textsc{II. MSN Regression}
+          \For {$ k = 1,...,K$}
+            \State $n_k := \sum_{i = 1}^n \mathds{1}_{Z_i = k}$
+            \For {$i_k = 1,...,n_k$}
+              \State $\mathbf{Draw}\  t_i \ \text{from} \ N_{[0,\infty)}(a_i,A)$
+            \EndFor
+            \State $\mathbf{X^*}_k := $ \texttt{cbind}$(\mathbf{X}_k,\mathbf{t}_k)$
+            \State $\mathbf{Draw} \ \mathbf{B^*}_k \ \text{from} \ \text{MatNorm}(\mathbf{B}_k,\mathbf{L}_k^{-1},\boldsymbol\Sigma_k)$
+            \State $\mathbf{Draw} \ \boldsymbol\Sigma_k \ \text{from} \ \text{InvWish}(\nu_k, \mathbf{V}_k)$
+          \EndFor
+        \State \textsc{III. Multinomial Logit}
+          \For {$i = 1,...,n$}
+            \For {$k = 1,...,K$}
+              \State $\pi_{ik} := P(Z_i = k|\mathbf{w}_i,\boldsymbol\delta_k)$
+              \State $p_{ik} := P(\mathbf{y}_i|\boldsymbol\beta_k^{*T} \mathbf{x}^*_i,\boldsymbol\Sigma_k)$
+            \EndFor
+            \State $\mathbf{p}_{Z_i} := \frac{\mathbf{p}_i \circ \boldsymbol\pi_i}{\mathbf{p}_i \cdot \boldsymbol\pi_i}$
+            \State $\mathbf{Draw} \ Z_i \ \text{from} \ \text{Categorical}(\mathbf{p}_{Z_i})$
+            \For {$k = 1,...,K-1$}
+              \State $\mathbf{Draw} \ \boldsymbol\delta_k \ \text{from} \ N(\mathbf{M},\mathbf{S})$
+            \EndFor
+          \EndFor
+        \State $\boldsymbol\theta := \{\mathbf{B^*}, \boldsymbol\Sigma, \mathbf{Z}, \boldsymbol\delta \}$
+        \State $\mathbf{Store} \ \boldsymbol\theta$
+	  \EndFor
+\end{algorithmic}
+\end{algorithm}
 
@@
 
-    \item $\mathbf{Y}$: A $N \times J$ matrix containing all multivariate skew-normal  outcomes such that $y_{ij}$ is the $j^{th}$ outcome observed for subject $i$, where $i = 1,...,n$ and $j = 1,...J$.
+    \item $\mathbf{Y}$: A $n \times J$ matrix containing all multivariate skew-normal  outcomes such that $y_{ij}$ is the $j^{th}$ outcome observed for subject $i$, where $i = 1,...,n$ and $j = 1,...J$.
 
-    \item $\mathbf{X}$: A $n \times p$ matrix containing all multivariate skew-normal regression covariates such that $x_{ij}$ is the $j^{th}$ covariate value for subject $i$.
+    \item $\mathbf{X}$: A $n \times P$ matrix containing all multivariate skew-normal regression covariates such that $x_{ip}$ is the $p^{th}$ covariate value for subject $i$, where $i = 1,...,n$ and $p = 1,...P$.
 
-    \item $\mathbf{B}$: A $m \times p$ matrix containing all multivariate skew-normal regression coefficients such that $\mathbf{B} = \left [ \boldsymbol\beta_1,...,\boldsymbol\beta_p \right ]$, where $\beta_{ij}$ is interpreted as the effect of covariate $i$ on outcome $j$ for $i = 1,...,m$ and $j = 1,...,p$.
+    \item $\mathbf{B}$: A $P \times J$ matrix containing all multivariate skew-normal regression coefficients such that $\mathbf{B} = \left [ \boldsymbol\beta_1,...,\boldsymbol\beta_J \right ]$, where $\beta_{pj}$ is interpreted as the effect of covariate $p$ on outcome $j$ for $p = 1,...,P$ and $j = 1,...,J$.
 
-    \item $\mathbf{E}$: A $n \times p$ matrix of error terms in the multivariate skew-normal regression model component.
-$\mathbf{E}$ is made up of row vectors $\boldsymbol\epsilon_i = (\epsilon_{i1},...,\epsilon_{ip})$, where $ \boldsymbol\epsilon_i \stackrel{iid}{\sim} N_p(0, \boldsymbol\Sigma)$ for $i = 1,...,n$.
+    \item $\mathbf{E}$: A $n \times J$ matrix of error terms in the multivariate skew-normal regression model component.
+$\mathbf{E}$ is made up of row vectors $\boldsymbol\epsilon_i = (\epsilon_{i1},...,\epsilon_{iJ})$, where $ \boldsymbol\epsilon_i \stackrel{iid}{\sim} N_J(0, \boldsymbol\Sigma)$ for $i = 1,...,n$.
 
-    \item $\boldsymbol\Sigma$: A $p \times p$ covariance matrix that defines the correlation between the $p$ multivariate normal outcomes.
+    \item $\boldsymbol\Sigma$: A $J \times J$ covariance matrix that defines the correlation between the $p$ multivariate normal outcomes.
 
-    \item $\boldsymbol\Omega$: A $p \times p$ covariance scale matrix that defines the correlation between the $p$ multivariate skew-normal outcomes.
+    \item $\boldsymbol\Omega$: A $J \times J$ covariance scale matrix that defines the correlation between the $p$ multivariate skew-normal outcomes.
 
-    \item $\boldsymbol\psi$: A $p \times 1$ vector containing the skewness parameter for each outcome.
+    \item $\boldsymbol\psi$: A $J \times 1$ vector containing the skewness parameter for each outcome.
 
-    \item $\boldsymbol\alpha$: A $p \times 1$ vector containing the skewness parameter for each outcome.
+    \item $\boldsymbol\alpha$: A $J \times 1$ vector containing the skewness parameter for each outcome.
 
@@
 
-    \item $\mathbf{X}^*$: A $n \times (m + 1)$ matrix constructed by column binding $\mathbf{t}$ to $\mathbf{X}$
+    \item $\mathbf{X}^*$: A $n \times (P + 1)$ matrix constructed by column binding $\mathbf{t}$ to $\mathbf{X}$
 
-    \item $\mathbf{B}^*$: A $(m+1) \times p$ matrix constructed by row binding $\boldsymbol\psi^T$ to $\mathbf{B}$.
+    \item $\mathbf{B}^*$: A $(P+1) \times J$ matrix constructed by row binding $\boldsymbol\psi^T$ to $\mathbf{B}$.
 
@@
 $$\mathbf{Y} = \mathbf{X}\mathbf{B} + \mathbf{t} \boldsymbol\psi^T + \mathbf{E} = \mathbf{X}^* \mathbf{B}^* + \mathbf{E}$$
-The matrix $\mathbf{Y}$ is of dimension $n \times p$.
-For convenience, we define $\mathbf{X}^*$ as a $n \times (m + 1)$ matrix constructed by column binding $\mathbf{t}$ to $\mathbf{X}$, and $\mathbf{B}^*$ as a $(m+1) \times p$ matrix constructed by row binding $\boldsymbol\psi^T$ to $\mathbf{B}$.
-We assume that $t_i \stackrel{iid}{\sim}T_{[0,\infty})(0,1)$ and that $\mathbf{E}$ is made of row vectors $\boldsymbol\epsilon_i = (\epsilon_{i1},...,\epsilon_{ip})$ for $i = 1,...,n$, where $ \boldsymbol\epsilon_i \stackrel{iid}{\sim} N_p(0, \boldsymbol\Sigma)$.
+The matrix $\mathbf{Y}$ is of dimension $n \times J$.
+For convenience, we define $\mathbf{X}^*$ as a $n \times (P + 1)$ matrix constructed by column binding $\mathbf{t}$ to $\mathbf{X}$, and $\mathbf{B}^*$ as a $(P+1) \times J$ matrix constructed by row binding $\boldsymbol\psi^T$ to $\mathbf{B}$.
+We assume that $t_i \stackrel{iid}{\sim}T_{[0,\infty})(0,1)$ and that $\mathbf{E}$ is made of row vectors $\boldsymbol\epsilon_i = (\epsilon_{i1},...,\epsilon_{iJ})$ for $i = 1,...,n$, where $ \boldsymbol\epsilon_i \stackrel{iid}{\sim} N_J(0, \boldsymbol\Sigma)$.
 
@@
 $$\boldsymbol\Sigma \sim \text{inverse-Wishart}(\mathbf{V}_0,\nu_0)$$
-
 $$\mathbf{B}^*|\boldsymbol\Sigma \sim MatNorm_{(m+1), p}(\mathbf{B}_0^*,\mathbf{L}_0^{-1},\boldsymbol\Sigma)$$
@@
 &  \propto |\boldsymbol\Sigma|^{-n/2}\exp \left \{ -\frac{1}{2} \tr \left [(\mathbf{Y} - \mathbf{X^*}\mathbf{B}^*)^T(\mathbf{Y} - \mathbf{X^*}\mathbf{B}^*)\boldsymbol\Sigma^{-1} \right ]\right \} \\
-& \times |\boldsymbol\Sigma|^{-(m+1)/2} \exp \left \{ -\frac{1}{2} \tr \left [(\mathbf{B}^* - \mathbf{B}^*_0)^T \mathbf{L}_0(\mathbf{B}^* - \mathbf{B}^*_0)\boldsymbol\Sigma^{-1} \right ]\right \} \\
-& \times |\boldsymbol\Sigma|^{(\nu_0 + p + 1)/2} \exp \left \{ -\frac{1}{2} \tr (\mathbf{V}_0 \boldsymbol\Sigma^{-1}) \right \}
+& \times |\boldsymbol\Sigma|^{-(P+1)/2} \exp \left \{ -\frac{1}{2} \tr \left [(\mathbf{B}^* - \mathbf{B}^*_0)^T \mathbf{L}_0(\mathbf{B}^* - \mathbf{B}^*_0)\boldsymbol\Sigma^{-1} \right ]\right \} \\
+& \times |\boldsymbol\Sigma|^{(\nu_0 + J + 1)/2} \exp \left \{ -\frac{1}{2} \tr (\mathbf{V}_0 \boldsymbol\Sigma^{-1}) \right \}
 \end{align*}
@@
 
-For a given observation vector $\mathbf{y} \sim N_p(\boldsymbol\mu, \boldsymbol\Sigma)$, we allow for missingness in at most $p - 1$ of the multivariate outcomes through the use of a conditional imputation step embedded within our Gibbs sampler.
-Suppose $\mathbf{y}$ contains $q$ missing observations and can be partitioned into two vectors $\mathbf{y_{1}}$ and $\mathbf{y_{2}}$ such that $\mathbf{y_{1}}$ is a $q \times 1$ vector of missing observations and $\mathbf{y_{2}}$ is a $(p-q) \times 1$ vector of complete observations.
+For a given observation vector $\mathbf{y} \sim N_J(\boldsymbol\mu, \boldsymbol\Sigma)$, we allow for missingness in at most $J - 1$ of the multivariate outcomes through the use of a conditional imputation step embedded within our Gibbs sampler.
+Suppose $\mathbf{y}$ contains $q$ missing observations and can be partitioned into two vectors $\mathbf{y_{1}}$ and $\mathbf{y_{2}}$ such that $\mathbf{y_{1}}$ is a $q \times 1$ vector of missing observations and $\mathbf{y_{2}}$ is a $(J-q) \times 1$ vector of complete observations.
 Similarly, partition $\boldsymbol\mu$ and $\boldsymbol\Sigma$ as follows.
@@
 
+\bibitem{ } Gelman A, Hwang J, Vehtari A. Understanding predictive information criteria for Bayesian models. \textit{Statistics and Computing}. 2014 Nov 1;24(6):997-1016.
+
 \bibitem{ } Holmes CC, Held L. Bayesian auxiliary variable models for binary and multinomial regression. \textit{Bayesian Analysis}. 2006;1(1):145-68.

### 2019-06-07 manuscript/manuscript.tex
algo

@@
 Additionally, parameter estimation proposed by Lin et al.
-for the MSTFA relies on a prohibitively complex EM algorithm and does not enjoy the inferential benefits of a Bayesian approach, namely the ability to incorporate prior information into a model and make posterier probability statements.
+for the MSTFA relies on a prohibitively complex EM algorithm and does not enjoy the inferential benefits of a Bayesian approach, namely the ability to incorporate prior information into a model and make posterior probability statements.
 Our proposed model improves on these previous works by estimating parameters in a Bayesian framework as well as including the ability to fit a multinomial logit regression to cluster membership probabilities using a novel application of data augmentation with the P\'olya Gamma distribution.
@@
 We assume for now that $\mathbf{y}_{i}$ is fully observed.
-Later, we relax this assumptin by allowing for missingness in the components of $\mathbf{y}_{i}$.
-The general formulation of the multivariate skew normal mixure model is as follows.
+Later, we relax this assumption by allowing for missingness in the components of $\mathbf{y}_{i}$.
+The general formulation of the multivariate skew normal mixture model is as follows.
 
@@
 A primary concern of our model is with identification of latent infant development clusters.
-We accomplish this via multinomial logit regression model on cluster membership, which utilizes P\'{o}lya-Gamma data-aumentation to allow for updating of all parameters using Gibbs sampling.
+We accomplish this via multinomial logit regression model on cluster membership, which utilizes P\'{o}lya-Gamma data-augmentation to allow for updating of all parameters using Gibbs sampling.
 The multinomial logit model is as follows for $k = 1,...,K$.
@@
 
-\item Emphasize that PG data augmentation for the multinomial model results in a PG mixure of experts model, which is a computationally efficient way to model edge weights.
+\item Emphasize that PG data augmentation for the multinomial model results in a PG mixture of experts model, which is a computationally efficient way to model edge weights.
 
@@
 
-\subsection{Simulation to Assess Sensitivity to Mispecified K}
+\subsection{Simulation to Assess Sensitivity to Misspecified K}
 
@@
 
-We now dervive the joint posterior distribution of the parameters $\mathbf{B}^*$ and $\boldsymbol\Sigma$.
+We now derive the joint posterior distribution of the parameters $\mathbf{B}^*$ and $\boldsymbol\Sigma$.
 \begin{align*} 
@@
 The multivariate normal conditional imputation derivations are given for a single cluster without loss of generality.
-In practive, the data and parameters in this section would be replaced by cluster specific estimates in the case of clustering.
+In practice, the data and parameters in this section would be replaced by cluster specific estimates in the case of clustering.
 
@@
 
-The blockwise inversion formula was used to invert $\boldsymbol\Sigma$ according to the following reparameterizations.
+The block-wise inversion formula was used to invert $\boldsymbol\Sigma$ according to the following reparameterizations.
 \begin{align*}
@@
 \bibitem{ } Azzalini, A. and Dalla Valle, A. (1996). The multivariate skew normal distribution. Biometrika 83, 715–726.
+
+\bibitem{ } Chen JT, Gupta AK. Matrix variate skew normal distributions. \textit{Statistics}. 2005 Jun 1;39(3):247-53.
 
@@
 
+\bibitem{ } Viroli C. Finite mixtures of matrix normal distributions for classifying three-way data. \textit{Statistics and Computing}. 2011 Oct 1;21(4):511-22.
+
 \bibitem{ } Vrbik I, Mcnicholas PD. Parsimonious skew mixture models for model-based clustering and classification. \textit{Computational Statistics \& Data Analysis}. 2014 Mar 1;71:196-210.
+
+\bibitem{ } Zeller CB, Cabral CR, Lachos VH. Robust mixture regression modeling based on scale mixtures of skew-normal distributions. \textit{Test}. 2016 Jun 1;25(2):375-96.
 

### 2019-06-11 mcmc_draws_2019-06-11/tables.Rmd
simulation changes

---
title: "Paper Tables"
author: "Carter Allen"
date: "3/26/2019"
output: 
  pdf_document: 
    keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
```

```{r}
load("BETA")
load("beta_true")
load("DELTA")
load("delta_true")
load("PI")
load("PSI")
load("psi_true")
load("SIGMA")
load("sig2_true")
load("Z")
```

```{r}
col_summarize <- function(MAT,dig = 2)
{
    k <- ncol(MAT)
    ret_vec <- rep(0,k)
    for(i in 1:k)
    {
        dat <- MAT[,i]
        est <- round(median(dat),dig)
        cri <- round(unname(quantile(dat,probs = c(0.025,0.975))),dig)
        res <- paste(est," (",cri[1],", ",cri[2],")",sep = "")
        ret_vec[i] <- res
    }
    return(ret_vec)
}
```

```{r}
beta_true_c1 <- c(beta.true.list[[1]]) %>% round(.,2)
beta_true_c2 <- c(beta.true.list[[2]]) %>% round(.,2)
beta_true_c3 <- c(beta.true.list[[3]]) %>% round(.,2)
beta_est_c1 <- col_summarize(BETA.list[[1]]) 
beta_est_c2 <- col_summarize(BETA.list[[2]]) 
beta_est_c3 <- col_summarize(BETA.list[[3]])

sig2_true_c1 <- c(sig2.true.list[[1]])[c(lower.tri(sig2.true.list[[1]],diag = TRUE))] %>% round(.,2)
sig2_true_c2 <- c(sig2.true.list[[2]])[c(lower.tri(sig2.true.list[[2]],diag = TRUE))] %>% round(.,2)
sig2_true_c3 <- c(sig2.true.list[[3]])[c(lower.tri(sig2.true.list[[3]],diag = TRUE))] %>% round(.,2)
sig2_est_c1 <- col_summarize(SIGMA.list[[1]])[c(lower.tri(sig2.true.list[[1]],diag = TRUE))]
sig2_est_c2 <- col_summarize(SIGMA.list[[2]])[c(lower.tri(sig2.true.list[[2]],diag = TRUE))]
sig2_est_c3 <- col_summarize(SIGMA.list[[3]])[c(lower.tri(sig2.true.list[[3]],diag = TRUE))]

psi_true_c1 <- psi.true.list[[1]] %>% round(.,2)
psi_true_c2 <- psi.true.list[[2]] %>% round(.,2)
psi_true_c3 <- psi.true.list[[3]] %>% round(.,2)
psi_est_c1 <- col_summarize(PSI.list[[1]]) 
psi_est_c2 <- col_summarize(PSI.list[[2]]) 
psi_est_c3 <- col_summarize(PSI.list[[3]])

delta_true <- c(delta.true) %>% round(.,2)
delta_est <- col_summarize(DELTA)

pi_est <- col_summarize(PI)
pi_true <- rep(0.33,3)

model_component <- c("MVSN","Regression","","","","","","","","","","","","","","","","","","","","",
                     "Multinom.","Logit","","","","",
                     "Clustering")
params <- c("$\\beta_{11}$","$\\beta_{12}$","$\\beta_{13}$","$\\beta_{14}$",
            "$\\beta_{21}$","$\\beta_{22}$","$\\beta_{23}$","$\\beta_{24}$",
            "$\\sigma_{11}$","$\\sigma_{12}$","$\\sigma_{13}$","$\\sigma_{14}$",
            "$\\sigma_{22}$","$\\sigma_{23}$","$\\sigma_{24}$",
            "$\\sigma_{33}$","$\\sigma_{34}$",
            "$\\sigma_{44}$",
            "$\\psi_{1}$","$\\psi_{2}$","$\\psi_{3}$","$\\psi_{4}$",
            "$\\delta_{11}$","$\\delta_{12}$","$\\delta_{21}$",
            "$\\delta_{22}$","$\\delta_{31}$","$\\delta_{32}$",
            "$\\pi_l$")

true_c1 <- c(beta_true_c1,sig2_true_c1,psi_true_c1,delta_true,pi_true[1])
true_c2 <- c(beta_true_c2,sig2_true_c2,psi_true_c2,delta_true,pi_true[2])
true_c3 <- c(beta_true_c3,sig2_true_c3,psi_true_c3,delta_true,pi_true[3])
est_c1 <- c(beta_est_c1,sig2_est_c1,psi_est_c1,delta_est,pi_est[1])
est_c2 <- c(beta_est_c2,sig2_est_c2,psi_est_c2,delta_est,pi_est[2])
est_c3 <- c(beta_est_c3,sig2_est_c3,psi_est_c3,delta_est,pi_est[3])

t1 <- cbind(#model_component,params,
            true_c1,est_c1,
            true_c2,est_c2,
            true_c3,est_c3)
colnames(t1) <- c(# "Model Component",
                  # "Parameter",
                  "True","Est. (95\\% CrI)",
                  "True","Est. (95\\% CrI)",
                  "True","Est. (95\\% CrI)")
```

```{r}
kable(t1)
```

```{r,eval=FALSE}
kable(t1,booktabs = T,escape = F,caption = "Model results for simulated data with n = 500, k = 4, p = 2, h = 3, v = 3. 500 iterations were run with a burn in of 100. Missingness mechanism was MAR and P(miss) = 0.25.") %>%
    kable_styling(font_size = 8) %>%
    add_header_above(c(" " = 2,"Class 1" = 2, "Class 2" = 2, "Class 3" = 2)) %>%
    group_rows(" ",1,8) %>%
    group_rows(" ",9,18) %>%
    group_rows(" ",19,22) %>%
    group_rows(" ",23,28) %>%
    group_rows(" ",29,29)
```

### 2019-06-11 simulation/simulation.R

@@ -23,10 +23,10 @@ library(BayesLogit) # For rpg function
 # No scientific notation
 options(scipen=999)
 # Generate data
-k <- 4 # number of outcomes
+k <- 2 # number of outcomes
 p <- 2 # number of predictors 
 h <- 3 # number of clusters
-n <- 1000 # number of observations (takes alot of data to estimate this many parameters)
+n <- 1500 # number of observations (takes alot of data to estimate this many parameters)
 print(paste("Generating data with",n,"observations of dimension",k,
             "with",p,"covariates and",h,"clusters"))
 
@@ -35,15 +35,34 @@ P <- matrix(rep(rep(1/h,h),each=n),n,h) # probability matrix governing cluster m
 
 # Generate covariates
 X <- matrix(rnorm(n*p),nrow = n,ncol = p) # random predictors
+X[,1] <- 1
 t <- rtruncnorm(n,0,Inf,0,1) # trunc norm random effects
 Xstar <- cbind(X,t) # Combine X and t
+# # Generate X manually
+# xi <- matrix(c(1, 0, 0, 0,
+#                1, 1, 0, 0,
+#                1, 0, 1, 0,
+#                1, 0, 0, 1),
+#              nrow = 4,
+#              ncol = 4,
+#              byrow = TRUE)
+# t <- rtruncnorm(n,0,Inf,0,1) # trunc norm random effects
+# X <- NULL 
+# capT <- NULL
+# for(i in 1:n)
+# {
+#     X <- rbind(X,xi)
+#     capT <- rbind(capT,diag(t[i],nrow = 4,ncol = 4))
+# }
+# 
+# Xstar <- cbind(X,capT) # Combine X and t
 
 # Covariates corresponding to multinomial regression
 #w1 <- rnorm(n) # continuous predictor
 #w2 <- rnorm(n) # continuous predictor
 w3 <- rbinom(n,1,0.5) # binary predictor
 w4 <- rbinom(n,1,0.5) # binary predictor
-W <- cbind(1,w3,w4) # design matrix
+W <- cbind(w3) # design matrix
 v <- ncol(W) # number of multinomial predictors
 delta.true <- matrix(rtruncnorm(n = v*(h-1),
                                 a = -1, 
@@ -98,12 +117,12 @@ for(l in 1:h)
     A.l <- matrix(runif(k^2)*k-1,ncol = k) # make a symmetric matrix
[…]

### 2019-06-14 mcmc_draws_2019-06-11/tables.Rmd
simulation changes

@@
 ---
-title: "Paper Tables"
+title: "Simulation Tables"
 author: "Carter Allen"
-date: "3/26/2019"
+date: "4/11/2019"
 output: 
@@
 load("sig2_true")
+load("pi_true")
 load("Z")
@@
 pi_est <- col_summarize(PI)
-pi_true <- rep(0.33,3)
+pi_true <- round(pi.true,2)
 
-model_component <- c("MVSN","Regression","","","","","","","","","","","","","","","","","","","","",
-                     "Multinom.","Logit","","","","",
+model_component <- c("MVSN","Regression","","","","","","","",
+                     "Multinom.","",
                      "Clustering")
-params <- c("$\\beta_{11}$","$\\beta_{12}$","$\\beta_{13}$","$\\beta_{14}$",
-            "$\\beta_{21}$","$\\beta_{22}$","$\\beta_{23}$","$\\beta_{24}$",
-            "$\\sigma_{11}$","$\\sigma_{12}$","$\\sigma_{13}$","$\\sigma_{14}$",
-            "$\\sigma_{22}$","$\\sigma_{23}$","$\\sigma_{24}$",
-            "$\\sigma_{33}$","$\\sigma_{34}$",
-            "$\\sigma_{44}$",
-            "$\\psi_{1}$","$\\psi_{2}$","$\\psi_{3}$","$\\psi_{4}$",
-            "$\\delta_{11}$","$\\delta_{12}$","$\\delta_{21}$",
-            "$\\delta_{22}$","$\\delta_{31}$","$\\delta_{32}$",
+params <- c("$\\beta_{0}$","$\\beta_{1}$","$\\beta_{2}$","$\\beta_{3}$",
+            "$\\sigma_{11}$","$\\sigma_{12}$",
+            "$\\sigma_{22}$",
+            "$\\psi_{1}$","$\\psi_{2}$",
+            "$\\delta_{11}$","$\\delta_{12}$",
             "$\\pi_l$")
@@
 
-t1 <- cbind(#model_component,params,
+t1 <- cbind(model_component,
+            params,
             true_c1,est_c1,
@@
             true_c3,est_c3)
-colnames(t1) <- c(# "Model Component",
-                  # "Parameter",
+colnames(t1) <- c("Model Component",
+                  "Parameter",
                   "True","Est. (95\\% CrI)",
@@
 ```{r}
-kable(t1)
+rownames(t1) <- NULL
+kable(t1, booktabs = T, escape = F,caption = "Model results for simulated data with n = 1500, k = 2, p = 2, h = 3, v = 1.
+5000 iterations were run with a burn in of 1000.
+Missingness mechanism was MAR and P(miss) = 0") %>%
+  kable_styling(font_size = 8) %>%
+  add_header_above(c(" " = 2,"Class 1" = 2, "Class 2" = 2, "Class 3" = 2)) %>%
+  group_rows(" ",1,4) %>%
+  group_rows(" ",5,7) %>%
+  group_rows(" ",8,10) %>%
+  group_rows(" ",10,10) %>%
+  group_rows(" ",12,12)
 ```
@@
 ```{r,eval=FALSE}
+
 kable(t1,booktabs = T,escape = F,caption = "Model results for simulated data with n = 500, k = 4, p = 2, h = 3, v = 3.
 500 iterations were run with a burn in of 100.
-Missingness mechanism was MAR and P(miss) = 0.25.") %>%
+Missingness mechanism was MAR and P(miss) = 0.25.",
+      row.names = NA) %>%
     kable_styling(font_size = 8) %>%
     add_header_above(c(" " = 2,"Class 1" = 2, "Class 2" = 2, "Class 3" = 2)) %>%
-    group_rows(" ",1,8) %>%
-    group_rows(" ",9,18) %>%
-    group_rows(" ",19,22) %>%
-    group_rows(" ",23,28) %>%
-    group_rows(" ",29,29)
+    group_rows(" ",1,4) %>%
+    group_rows(" ",5,7) %>%
+    group_rows(" ",8,10) %>%
+    group_rows(" ",11,11) %>%
+    group_rows(" ",12,12)
 ```

### 2019-06-14 simulation/simulation.R

@@ -23,8 +23,8 @@ library(BayesLogit) # For rpg function
 # No scientific notation
 options(scipen=999)
 # Generate data
-k <- 2 # number of outcomes
-p <- 2 # number of predictors 
+k <- 4 # number of outcomes
+p <- 1 # number of predictors 
 h <- 3 # number of clusters
 n <- 1500 # number of observations (takes alot of data to estimate this many parameters)
 print(paste("Generating data with",n,"observations of dimension",k,
@@ -102,7 +102,7 @@ Y <- NULL
 
 for(l in 1:h)
 {
-    X.l <- X[c == l,]
+    X.l <- as.matrix(X[c == l,])
     Xstar.l <- Xstar[c == l,] # subset of covariates belonging to cluster l
     n.l <- sum(c == l) # number of observations truly assigned to cluster l
     
@@ -300,7 +300,7 @@ for(i in 1:nsim)
     # Update class-specific regression parameters
     for(l in 1:h) # loop through each cluster
     {
-        X.l <- X[z == l,] # all covariates for those in class l
+        X.l <- as.matrix(X[z == l,]) # all covariates for those in class l
         Y.l <- Y[z == l,] # all outcomes for those in class l
         n.l <- sum(z == l) # current class specific sample size
         psi.l <- psi.list[[l]] # current class specific psis 

### 2019-06-14 manuscript/manuscript.tex
simulation changes
sect. 3

@@
 
-\subsection{Generic Mixture}
+\subsection{Generic Multivariate Skew Normal Mixture}
 
@@
 
-\subsection{Multivariate Skew Normal Regression}
+\subsection{Multivariate Skew Normal Stochastic Representation}
 
@@
 When using this stochastic representation in our finite mixture model, the parameters $\boldsymbol\beta$, $\boldsymbol\psi$, and $\boldsymbol\Sigma$ are replaced with class-specific counterparts $\boldsymbol\beta_k$, $\boldsymbol\psi_k$, and $\boldsymbol\Sigma_k$.
+
+\subsection{Representation as Matrix Skew Normal}
+
+We note that the multivariate skew normal regression problem can be expressed more succinctly through the use of Matrix Skew Normal distributions.
+In particular, we define $\mathbf{Y}_{n_k \times J}$ as the response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n(Z_i = k)$ is the total number of observations belonging to class $k$ ($k = 1,...,K$).
+
+$$\mathbf{Y}_{n_k \times J} \sim \mathcal{MATSN}_{n_k \times J}(\mathbf{M},\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Sigma_k)$$
+
+$$\text{vec}(Y) = \mathbf{y}_{n_kJ \times 1} = (\mathbf{y}^T_1,...,\mathbf{y}^T_{n_k})$$
 

### 2019-06-14 manuscript/manuscript.tex
simulation changes

@@
 $$\mathbf{Y}_{n_k \times J} \sim \mathcal{MATSN}_{n_k \times J}(\mathbf{M},\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Sigma_k)$$
+$$\text{vec}(\mathbf{Y}) = \mathbf{y}_{n_kJ \times 1} = (\mathbf{y}^T_1,...,\mathbf{y}^T_{n_k})$$
+$$\text{vec}(\mathbf{M}) = \boldsymbol\mu_{n_k J \times 1} = (\boldsymbol\mu_1^T,...,\boldsymbol\mu_{n_k}^T)$$
 
-$$\text{vec}(Y) = \mathbf{y}_{n_kJ \times 1} = (\mathbf{y}^T_1,...,\mathbf{y}^T_{n_k})$$
+where $\boldsymbol\mu_i = \mathbf{x}_i \boldsymbol\beta_k$ and $\boldsymbol\alpha_k = (\alpha_{k1},...,\alpha{kJ})$.
+This implies that the $i^{th}$ row of $\mathbf{Y}$ is $\mathbf{y}_i \sim \mathcal{MSN}_J(\boldsymbol\mu_i, \boldsymbol\alpha_k, \boldsymbol\Sigma_{J \times J}) = \mathcal{MSN}_J(\mathbf{x}_i \boldsymbol\beta_k, \boldsymbol\alpha_k, \boldsymbol\Sigma)$.
+Using the stochastic representation presented in Section 3.2, the conditional distribution $\mathbf{y}_i | Z_i, t_i \sim N_J (\mathbf{x}_i \boldsymbol\beta_k + \boldsymbol\psi_k t_i,\boldsymbol\Sigma_k)$.
+This implies that all responses for cluster $k$ conditioned on $\mathbf{T}_{n_k \times 1}$ is distributed as
+
+$$\mathbf{Y}_{n_k \times J} | \mathbf{T} = \mathbf{t} \sim N_{n_k \times J}(\mathbf{M}_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k)$$
+$$\text{vec}(\mathbf{M}_k)_{n_kJ \times 1} = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$$\
+
+where $\mathbf{t}_k = (t_1,...,t_{n_k})$.
 
@@
 For identifiability purposes, we fix the reference category $k = K$ and set $\boldsymbol\delta_K = \mathbf{0}$.
-
-\[   P(Z_i = k|\mathbf{w}_i,\boldsymbol\delta_k) = \left \{
-\begin{array}{ll}
-      \pi_{ik} = \frac{e^{\mathbf{w}_i^T \boldsymbol\delta_k}}{1 + \sum_{k' = 1}^{K-1} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}}} & k < K \\
-      \\
-      \pi_{ik} = \frac{1}{1 + \sum_{k' = 1}^{K-1} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}}} & k = K \\
-\end{array} 
-\right. \]
-
-\
-
-Under this model, $Z_i|\mathbf{w}_i \sim Multinom.(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{i1},...,\pi_{1K})$
-
+Under this model, $Z_i|\mathbf{w}_i \sim Multinom.(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{i1},...,\pi_{1K})$.
 During MCMC estimation, the class labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as class assignments.

### 2019-06-14 mcmc_draws_2019-06-13/tables.Rmd
simulation changes

--- 2019-06-14 mcmc_draws_2019-06-11/tables.Rmd
@@
 
-model_component <- c("MVSN","Regression","","","","","","","",
+model_component <- c("MVSN","Regression","","","","","","","","","","","","","","","","",
                      "Multinom.","",
@@
 params <- c("$\\beta_{0}$","$\\beta_{1}$","$\\beta_{2}$","$\\beta_{3}$",
-            "$\\sigma_{11}$","$\\sigma_{12}$",
-            "$\\sigma_{22}$",
-            "$\\psi_{1}$","$\\psi_{2}$",
+            "$\\sigma_{11}$","$\\sigma_{12}$","$\\sigma_{13}$","$\\sigma_{14}$",
+            "$\\sigma_{22}$","$\\sigma_{23}$","$\\sigma_{24}$",
+            "$\\sigma_{33}$","$\\sigma_{34}$",
+            "$\\sigma_{44}$",
+            "$\\psi_{1}$","$\\psi_{2}$", "$\\psi_{3}$","$\\psi_{4}$",
             "$\\delta_{11}$","$\\delta_{12}$",
@@
 rownames(t1) <- NULL
-kable(t1, booktabs = T, escape = F,caption = "Model results for simulated data with n = 1500, k = 2, p = 2, h = 3, v = 1.
+kable(t1, booktabs = T, escape = F,caption = "Model results for simulated data with n = 1500, k = 4, p = 1, h = 3, v = 1.
 5000 iterations were run with a burn in of 1000.
@@
   group_rows(" ",1,4) %>%
-  group_rows(" ",5,7) %>%
-  group_rows(" ",8,10) %>%
-  group_rows(" ",10,10) %>%
-  group_rows(" ",12,12)
+  group_rows(" ",5,14) %>%
+  group_rows(" ",15,18) %>%
+  group_rows(" ",19,20) %>%
+  group_rows(" ",21,21)
 ```

### 2019-06-14 simulation/simulation.R

@@ -26,7 +26,7 @@ options(scipen=999)
 k <- 4 # number of outcomes
 p <- 1 # number of predictors 
 h <- 3 # number of clusters
-n <- 1500 # number of observations (takes alot of data to estimate this many parameters)
+n <- 10000 # number of observations (takes alot of data to estimate this many parameters)
 print(paste("Generating data with",n,"observations of dimension",k,
             "with",p,"covariates and",h,"clusters"))
 
@@ -153,7 +153,7 @@ delta0 <- rep(0,v) # prior mean for delta coefficients (multinomial regression)
 S0 <- diag(0.5,v) # prior covariance for delta coefficients (multinomial regression)
 
 # Sample storage
-nsim <- 5000 # number of iterations
+nsim <- 2000 # number of iterations
 burn <- 1000 # number of iterations to save
 n.iter <- nsim - burn # number of saved iterations
 Z <- matrix(0,nrow = n.iter,ncol = n) # large matrix where each row is the value of z at a specific iteration

### 2019-06-14 manuscript/manuscript.tex
section 3

@@
 \usepackage{dsfont}
+
 \let\origfigure\figure
@@
 
-$$\mathbf{Y}_{n_k \times J} | \mathbf{T} = \mathbf{t} \sim N_{n_k \times J}(\mathbf{M}_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k)$$
+$$\mathbf{Y}_{n_k \times J} | \mathbf{T}_k = \mathbf{t}_k \sim N_{n_k \times J}(\mathbf{M}_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k)$$
 $$\text{vec}(\mathbf{M}_k)_{n_kJ \times 1} = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$$\
@@
 \subsection{Simulation to Compare to Multivariate Normal}
+
+\begin{table}[t]
+
+\caption{\label{tab:unnamed-chunk-4}Model results for simulated data with n = 1500, k = 4, p = 1, h = 3, v = 1.
+5000 iterations were run with a burn in of 1000.
+Missingness mechanism was MAR and P(miss) = 0}
+\centering
+\fontsize{8}{10}\selectfont
+\begin{tabular}{llllllll}
+\toprule
+\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Class 1} & \multicolumn{2}{c}{Class 2} & \multicolumn{2}{c}{Class 3} \\
+\cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6} \cmidrule(l{3pt}r{3pt}){7-8}
+Model Component & Parameter & True & Est. (95\% CrI) & True & Est. (95\% CrI) & True & Est. (95\% CrI)\\
+\midrule
+\addlinespace[0.3em]
+\multicolumn{8}{l}{\textbf{ }}\\
+\hspace{1em}MVSN & $\beta_{0}$ & -1.35 & -1.59 (-2.3, -0.94) & -0.18 & -0.24 (-0.53, 0.38) & 0.72 & 0.79 (0.52, 1.02)\\
+\hspace{1em}Regression & $\beta_{1}$ & -1.19 & -1.35 (-1.99, -0.84) & -0.09 & -0.33 (-0.62, 0.78) & 1.64 & 1.65 (1.33, 1.91)\\
+\hspace{1em} & $\beta_{2}$ & -1.65 & -1.81 (-2.66, -1.41) & -0.47 & -0.62 (-0.89, 0.3) & 1.44 & 1.32 (0.97, 1.6)\\
+\hspace{1em} & $\beta_{3}$ & -1.75 & -1.89 (-2.52, -1.37) & -0.22 & -0.32 (-0.63, 0.55) & 2.28 & 2.26 (1.96, 2.5)\\
+\addlinespace[0.3em]
+\multicolumn{8}{l}{\textbf{ }}\\
+\hspace{1em} & $\sigma_{11}$ & 1 & 1.02 (0.78, 1.19) & 1 & 0.98 (0.77, 1.23) & 1 & 1.06 (0.85, 1.29)\\
+\hspace{1em} & $\sigma_{12}$ & -0.32 & -0.19 (-0.33, -0.02) & 0.16 & 0.14 (-0.01, 0.4) & 0.72 & 0.82 (0.62, 1.05)\\
+\hspace{1em} & $\sigma_{13}$ & -0.65 & -0.55 (-0.68, -0.35) & 0.72 & 0.7 (0.51, 0.94) & 0.14 & 0.27 (0.1, 0.48)\\
+\hspace{1em} & $\sigma_{14}$ & -0.44 & -0.33 (-0.46, -0.13) & 0.5 & 0.48 (0.31, 0.72) & -0.01 & -0.02 (-0.16, 0.16)\\
+\hspace{1em} & $\sigma_{22}$ & 1 & 0.92 (0.72, 1.06) & 1 & 0.94 (0.72, 1.22) & 1 & 1.11 (0.87, 1.38)\\
+\hspace{1em} & $\sigma_{23}$ & 0.56 & 0.49 (0.33, 0.6) & 0.53 & 0.46 (0.29, 0.73) & -0.1 & 0.08 (-0.1, 0.28)\\
+\hspace{1em} & $\sigma_{24}$ & 0.98 & 0.9 (0.7, 1.04) & 0.24 & 0.14 (-0.03, 0.41) & 0.19 & 0.17 (0.01, 0.37)\\
+\hspace{1em} & $\sigma_{33}$ & 1 & 0.9 (0.66, 1.04) & 1 & 0.92 (0.72, 1.19) & 1 & 1.26 (1.03, 1.52)\\
+\hspace{1em} & $\sigma_{34}$ & 0.56 & 0.51 (0.35, 0.62) & 0.86 & 0.79 (0.58, 1.05) & -0.65 & -0.63 (-0.78, -0.45)\\
+\hspace{1em} & $\sigma_{44}$ & 1 & 0.93 (0.72, 1.07) & 1 & 0.93 (0.7, 1.2) & 1 & 1.11 (0.87, 1.36)\\
+\addlinespace[0.3em]
+\multicolumn{8}{l}{\textbf{ }}\\
+\hspace{1em} & $\psi_{1}$ & -0.33 & -0.02 (-0.84, 0.87) & 0.67 & 0.69 (-0.09, 1.01) & -1 & -0.98 (-1.25, -0.67)\\
+\hspace{1em} & $\psi_{2}$ & -0.33 & -0.16 (-0.8, 0.62) & 0.67 & 0.81 (-0.6, 1.14) & -1 & -0.98 (-1.28, -0.6)\\
+\hspace{1em} & $\psi_{3}$ & -0.33 & -0.15 (-0.65, 0.89) & 0.67 & 0.72 (-0.35, 1.02) & -1 & -0.8 (-1.15, -0.36)\\
+\hspace{1em} & $\psi_{4}$ & -0.33 & -0.18 (-0.84, 0.59) & 0.67 & 0.71 (-0.32, 1.06) & -1 & -1 (-1.32, -0.62)\\
+\addlinespace[0.3em]
+\multicolumn{8}{l}{\textbf{ }}\\
+\hspace{1em}Multinom. & $\delta_{11}$ & -0.84 & -0.78 (-0.96, -0.59) & -0.84 & -0.78 (-0.96, -0.59) & -0.84 & -0.78 (-0.96, -0.59)\\
+\hspace{1em} & $\delta_{12}$ & -0.24 & -0.26 (-0.42, -0.1) & -0.24 & -0.26 (-0.42, -0.1) & -0.24 & -0.26 (-0.42, -0.1)\\
+\addlinespace[0.3em]
+\multicolumn{8}{l}{\textbf{ }}\\
+\hspace{1em}Clustering & $\pi_l$ & 0.39 & 0.39 (0.38, 0.4) & 0.26 & 0.26 (0.25, 0.27) & 0.34 & 0.35 (0.33, 0.36)\\
+\bottomrule
+\end{tabular}
+\end{table}
 

### 2019-06-14 simulation/simulation.R

@@ -20,6 +20,8 @@ library(QRM)        # For rGumbel function
 library(nnet)       # To fit multinom function
 library(BayesLogit) # For rpg function
 
+seed <- rnorm(1)
+set.seed(seed)
 # No scientific notation
 options(scipen=999)
 # Generate data
@@ -107,7 +109,7 @@ for(l in 1:h)
     n.l <- sum(c == l) # number of observations truly assigned to cluster l
     
     # Cluster specific betas and psis
-    beta.true.l <- matrix(rnorm(p*k,mean = seq(-h/2,h/2,length.out = h)[l],sd = 0.5),nrow = p,ncol = k) # true beta matrix
+    beta.true.l <- matrix(rnorm(p*k,mean = seq(-h,h,length.out = h)[l],sd = 0.5),nrow = p,ncol = k) # true beta matrix
     psi.true.l <- rep(((-1)^l)*l/3,k) # control skewness of each outcome in cluster l
     betastar.true.l <- rbind(beta.true.l,psi.true.l)
     beta.true.list[[l]] <- beta.true.l
@@ -153,7 +155,7 @@ delta0 <- rep(0,v) # prior mean for delta coefficients (multinomial regression)
 S0 <- diag(0.5,v) # prior covariance for delta coefficients (multinomial regression)
 
 # Sample storage
-nsim <- 2000 # number of iterations
+nsim <- 10000 # number of iterations
 burn <- 1000 # number of iterations to save
 n.iter <- nsim - burn # number of saved iterations
 Z <- matrix(0,nrow = n.iter,ncol = n) # large matrix where each row is the value of z at a specific iteration
@@ -382,7 +384,7 @@ if(!dir.exists(store))
 meta_file <- paste(store,"/META.txt",sep = "")
 file.create(meta_file)
 write(paste("MCMC Run Date:",Sys.time()),file = meta_file, append = TRUE)
-#write(paste("Seed:",seed),file = meta_file, append = TRUE)
+write(paste("Seed:",seed),file = meta_file, append = TRUE)
 write(paste("Number of subjects (n):",n),file = meta_file, append = TRUE)
 write(paste("Number of outcomes (k):",k),file = meta_file, append = TRUE)
 write(paste("Number of predictors (p):",p),file = meta_file, append = TRUE)

### 2019-06-14 manuscript/manuscript.tex
section 3

@@
 In particular, we define $\mathbf{Y}_{n_k \times J}$ as the response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n(Z_i = k)$ is the total number of observations belonging to class $k$ ($k = 1,...,K$).
-
 $$\mathbf{Y}_{n_k \times J} \sim \mathcal{MATSN}_{n_k \times J}(\mathbf{M},\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Sigma_k)$$
@@
 This implies that all responses for cluster $k$ conditioned on $\mathbf{T}_{n_k \times 1}$ is distributed as
-
 $$\mathbf{Y}_{n_k \times J} | \mathbf{T}_k = \mathbf{t}_k \sim N_{n_k \times J}(\mathbf{M}_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k)$$
@@
 The multinomial logit model is as follows for $k = 1,...,K$.
-
 $$P(Z_i = k|\mathbf{w}_i) = \pi_{ik} = \frac{e^{\mathbf{w}_i^T \boldsymbol\delta_k}}{\sum_{k' = 1}^K e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}}}$$
@@
 We allow for missingness of outcomes in the MSN mixture model by imputing missing values from their conditional multivariate normal distributions. We note that
-
 $$\mathbf{y}_i|Z_i,\mathbf{x}_i,t_i,\boldsymbol\beta_{Z_i},\boldsymbol\psi_{Z_i},\boldsymbol\Sigma_{Z_i} \sim N_J(\boldsymbol\beta_{Z_i}^T\mathbf{x}_i  + t_i \boldsymbol\psi_{Z_i}, \boldsymbol\Sigma_{Z_i})$$
@@
 We have
-
 $$\mathbf{y}_i^{miss}|\mathbf{y}_i^{obs},\mathbf{x}_i,t_i,Z_i,\boldsymbol\beta_{Z_i},\boldsymbol\psi_{Z_i},\boldsymbol\Sigma_{Z_i} \sim N(\boldsymbol\mu^{miss},\boldsymbol\Sigma^{miss})$$
@@
 
-\textit{\textbf{Discuss non-ignorable missingness here}}
+\subsection{Bayesian Inference}
 
-\subsection{Bayesian Inference}
+\subsubsection{P\'olya--Gamma Data Augmentation}
+
+Polson \textit{et al.} (2012) introduced a data augmentation approach to fitting the multinomial logit regression model specified by Holmes and Held (2006), which specifies the full conditional distributions of the multinomial regression parameters as a function of a Bernoulli likelihood.
+$$p(\boldsymbol\delta_k|\mathbf{z},\boldsymbol\delta_{k' \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \pi_{ik}^{U_{ik}}(1-\pi_{ik})^{1-U_{ik}}$$
+
+where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, $U_{ik} = \mathds{1}_{Z_i = k}$ is an indicator that subject $i$ belongs to class $k$, and $\pi_{ik}$ is defined as in Section 3.4.
+We can rewrite $\pi_{ik}$ as follows
+$$\pi_{ik} = P(U_{ik} = 1) = \frac{e^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}}}{1 + e^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}}} = \frac{e^{\eta_{ik}}}{1 + e^{\eta_{ik}}}$$
+
+where ${c}_{ik} = \log \sum_{k' \ne k} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}}$ and $\eta_{ik} = \mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}$.
+We note that the sum $\sum_{k' \ne k} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}}$ includes the reference category, but since we fix $\boldsymbol\delta_K = \mathbf{0}$, we have $e^{\mathbf{w}_i^T \boldsymbol\delta_K} = 1$, and hence
+$$c_{ik} = \log \sum_{k' \ne k} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}} = \log \left ( 1 + \sum_{k' \notin \{k,K \}} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}} \right )$$
+
+We can use the quantities to re-express the full conditionals for $\boldsymbol\delta_k$ as 
+$$p(\boldsymbol\delta_k|\mathbf{Z},\boldsymbol\delta_{k' \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \left (\frac{e^{\eta_{ik}}}{1 + e^{\eta_{ik}}} \right )^{U_{ik}} \left (\frac{1}{1 + e^{\eta_{ik}}} \right )^{1-U_{ik}} = p(\boldsymbol\delta_k) \prod_{i = 1}^n \frac{(e^{\eta_{ik}})^{U_{ik}}}{1 + e^{\eta_{ik}}}$$
+
+which we note is essentially a logistic regression likelihood.
+We thus apply this P\'olya--Gamma data augmentation scheme to update each $\boldsymbol\delta_k$ ($k = 1,...,K-1$) one at a time based on the binary indicators $U_{ik}$.
 
@@
 \label{s:discuss}
+
+\textit{\textbf{Discuss non-ignorable missingness here}}
 

### 2019-06-17 manuscript/manuscript.tex
section 3

@@
 The general formulation of the multivariate skew normal mixture model is as follows.
-
-$$f(\mathbf{y}_i) = \sum_{k = 1}^{K} \pi_{ik} f(\mathbf{y}_i|\boldsymbol\theta_k)$$
+\begin{equation}
+f(\mathbf{y}_i) = \sum_{k = 1}^{K} \pi_{ik} f(\mathbf{y}_i|\boldsymbol\theta_k)
+\end{equation}
 
@@
 We model $\mathbf{y}_{i}|Z_i$ as
-
-$$\mathbf{y}_{i}|Z_i = k \sim MSN_J(\boldsymbol\zeta_k,\boldsymbol\alpha_k,\boldsymbol\Omega_k)$$
+\begin{equation}
+\mathbf{y}_{i}|Z_i = k \sim MSN_J(\boldsymbol\zeta_k,\boldsymbol\alpha_k,\boldsymbol\Omega_k)
+\end{equation}
 
@@
 The MSN stochastic representation is as follows.
-$$\mathbf{y}_{i} = \boldsymbol\beta^T \mathbf{x}_{i} + t_{i}\boldsymbol\psi + \boldsymbol\epsilon_{i}$$
+\begin{equation}
+\mathbf{y}_{i} = \boldsymbol\beta^T \mathbf{x}_{i} + t_{i}\boldsymbol\psi + \boldsymbol\epsilon_{i}
+\end{equation}
 
-where $\mathbf{x}_i$ is the $P \times 1$ vector of covariate values for subject $i$, $\boldsymbol\beta$ is the $P \times J$ vector of fixed effects coefficients , $t_i \stackrel{iid}{\sim} N_{[0,\infty)}(0,1)$ is a conditional truncated normal random effect, $\boldsymbol\psi$ is the $J \times 1$ vector containing cluster-specific skewness parameters, and $\boldsymbol\epsilon_i \sim N_J(0,\boldsymbol\Sigma_{J \times J})$ is the correlated error term.
+where $\mathbf{x}_i$ is the $P \times 1$ vector of covariate values for subject $i$, $\boldsymbol\beta$ is the $P \times J$ vector of fixed effects coefficients , $t_i \stackrel{iid}{\sim} N_{[0,\infty)}(0,1)$ is a conditional truncated normal random effect, $\boldsymbol\psi$ is the $J \times 1$ vector containing cluster-specific skewness parameters, and $\boldsymbol\epsilon_i \sim N_J(0,\boldsymbol\Sigma)$ is the correlated error term.
 In can be shown that under these conditions $\mathbf{y}_i$ is multivariate skew normal with location.
@@
 In particular, we define $\mathbf{Y}_{n_k \times J}$ as the response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n(Z_i = k)$ is the total number of observations belonging to class $k$ ($k = 1,...,K$).
-$$\mathbf{Y}_{n_k \times J} \sim \mathcal{MATSN}_{n_k \times J}(\mathbf{M},\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Sigma_k)$$
+$$\mathbf{Y}_{n_k \times J} \sim MatSN_{n_k \times J}(\mathbf{M},\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Sigma_k)$$
 $$\text{vec}(\mathbf{Y}) = \mathbf{y}_{n_kJ \times 1} = (\mathbf{y}^T_1,...,\mathbf{y}^T_{n_k})$$
@@
 
-where $\boldsymbol\mu_i = \mathbf{x}_i \boldsymbol\beta_k$ and $\boldsymbol\alpha_k = (\alpha_{k1},...,\alpha{kJ})$.
-This implies that the $i^{th}$ row of $\mathbf{Y}$ is $\mathbf{y}_i \sim \mathcal{MSN}_J(\boldsymbol\mu_i, \boldsymbol\alpha_k, \boldsymbol\Sigma_{J \times J}) = \mathcal{MSN}_J(\mathbf{x}_i \boldsymbol\beta_k, \boldsymbol\alpha_k, \boldsymbol\Sigma)$.
-Using the stochastic representation presented in Section 3.2, the conditional distribution $\mathbf{y}_i | Z_i, t_i \sim N_J (\mathbf{x}_i \boldsymbol\beta_k + \boldsymbol\psi_k t_i,\boldsymbol\Sigma_k)$.
+where $\boldsymbol\mu_i = \mathbf{x}_i \boldsymbol\beta_k$ and $\boldsymbol\alpha_k = (\alpha_{k1},...,\alpha_{kJ})$.
+This implies that the $i^{th}$ row of $\mathbf{Y}$ is $\mathbf{y}_i \sim MSN_J(\boldsymbol\mu_i, \boldsymbol\alpha_k, \boldsymbol\Sigma_k) = {MSN}_J(\mathbf{x}_i \boldsymbol\beta_k, \boldsymbol\alpha_k, \boldsymbol\Sigma_k)$.
+Using the stochastic representation presented in Section 3.2, the conditional distribution of $\mathbf{y}_i | Z_i, t_i$ is $N_J (\mathbf{x}_i \boldsymbol\beta_k + \boldsymbol\psi_k t_i,\boldsymbol\Sigma_k)$.
 This implies that all responses for cluster $k$ conditioned on $\mathbf{T}_{n_k \times 1}$ is distributed as
@@
 For identifiability purposes, we fix the reference category $k = K$ and set $\boldsymbol\delta_K = \mathbf{0}$.
-Under this model, $Z_i|\mathbf{w}_i \sim Multinom.(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{i1},...,\pi_{1K})$.
+Under this model, $Z_i|\mathbf{w}_i \sim Multinom(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{i1},...,\pi_{1K})$.
 During MCMC estimation, the class labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as class assignments.
@@
 We allow for missingness of outcomes in the MSN mixture model by imputing missing values from their conditional multivariate normal distributions. We note that
-$$\mathbf{y}_i|Z_i,\mathbf{x}_i,t_i,\boldsymbol\beta_{Z_i},\boldsymbol\psi_{Z_i},\boldsymbol\Sigma_{Z_i} \sim N_J(\boldsymbol\beta_{Z_i}^T\mathbf{x}_i  + t_i \boldsymbol\psi_{Z_i}, \boldsymbol\Sigma_{Z_i})$$
+$$\mathbf{y}_i|Z_i,\mathbf{x}_i,t_i,\boldsymbol\beta_{k},\boldsymbol\psi_{k},\boldsymbol\Sigma_{k} \sim N_J(\boldsymbol\beta_{k}^T\mathbf{x}_i  + t_i \boldsymbol\psi_{k}, \boldsymbol\Sigma_{k})$$
 
+where $Z_i = k$.
 This allows us to appeal to standard conditional forms of the multivariate normal distribution to specify the distribution of missing observations.
@@
 We have
-$$\mathbf{y}_i^{miss}|\mathbf{y}_i^{obs},\mathbf{x}_i,t_i,Z_i,\boldsymbol\beta_{Z_i},\boldsymbol\psi_{Z_i},\boldsymbol\Sigma_{Z_i} \sim N(\boldsymbol\mu^{miss},\boldsymbol\Sigma^{miss})$$
+$$\mathbf{y}_i^{miss}|\mathbf{y}_i^{obs},\mathbf{x}_i,t_i,Z_i,\boldsymbol\beta_{k},\boldsymbol\psi_{k},\boldsymbol\Sigma_{k} \sim N(\boldsymbol\mu^{miss},\boldsymbol\Sigma^{miss})$$
 

### 2019-06-18 manuscript/manuscript.tex
section 3

@@
 \begin{abstract}
-In studies of infant motor development, a crucial research goal is to identify latent classes of infants that experience delayed development, as this is a known risk factor for adverse outcomes later in life.
+In studies of infant motor development, a crucial research goal is to identify latent clusters of infants that experience delayed development, as this is a known risk factor for adverse outcomes later in life.
 However, there are a number of statistical challenges in modeling infant development: the data are typically skewed, exhibit intermittent missingness, and are highly correlated across the repeated measurements collected during infancy.
@@
 Our model has a number of attractive features.
-First, we adopt the multivariate skew normal distribution with class-specific parameters that accommodate the inherent correlation and skewness in the data.
-Second, we model the class membership probabilities using a novel Pólya-Gamma data-augmentation scheme, thereby improving predictions of the class membership allocations.
+First, we adopt the multivariate skew normal distribution with cluster-specific parameters that accommodate the inherent correlation and skewness in the data.
+Second, we model the cluster membership probabilities using a novel Pólya-Gamma data-augmentation scheme, thereby improving predictions of the cluster membership allocations.
 Lastly, we impute missing responses under missing at random assumption by drawing from appropriate conditional skew normal distributions.
@@
 In addition, our imputation method yields improvements compared to conventional missing data methods, including multiple imputation and complete or available case analysis.
-When applied to Nurture data, we identified two distinct development classes: one characterized by delayed “U-shaped” development and a higher percentage of male infants and another characterized by more steady development and a lower percentage of males.
-The classes also differed in terms of key demographic variables, such as infant race and maternal pre-pregnancy body mass index.
+When applied to Nurture data, we identified two distinct development clusters: one characterized by delayed “U-shaped” development and a higher percentage of male infants and another characterized by more steady development and a lower percentage of males.
+The clusters also differed in terms of key demographic variables, such as infant race and maternal pre-pregnancy body mass index.
 These findings can aid investigators in targeting interventions during this critical early-life developmental window.
@@
 
-Here $\boldsymbol\theta_k$ is the set of parameters shared by subjects belonging to class $k$ ($k = 1,...,K$).
-The probability of subject $i$ ($i = 1,...,n$) belonging to class $k$ ($k = 1,...,K$) is denoted $\pi_{ik}$.
-Note that $\pi_{ik}$ is not indexed by time and we do not allow for subjects switching classes over the study period.
+Here $\boldsymbol\theta_k$ is the set of parameters shared by subjects belonging to cluster $k$ ($k = 1,...,K$).
+The probability of subject $i$ ($i = 1,...,n$) belonging to cluster $k$ ($k = 1,...,K$) is denoted $\pi_{ik}$.
+Note that $\pi_{ik}$ is not indexed by time and we do not allow for subjects switching clusters over the study period.
 This feature of the model is a result of our interest in clustering based on development patterns rather than time-specific development scores.
-Later, we detail a multinomial logit regression model placed on $\pi_{ik}$ to allow for flexible modeling of class membership probabilities as a function of covariates of interest.
+Later, we detail a multinomial logit regression model placed on $\pi_{ik}$ to allow for flexible modeling of cluster membership probabilities as a function of covariates of interest.
 
-In each development class, we assume that a subject's $J\times 1$ outcome vector follows a multivariate skew normal (MSN) distribution.
+In each development cluster, we assume that a subject's $J\times 1$ outcome vector follows a multivariate skew normal (MSN) distribution.
 We define a latent variable $Z_i \in \{1,...,K\}$ to denote the cluster membership of subject $i$ for $i = 1,...,n$.
@@
 \begin{equation}
-\mathbf{y}_{i}|Z_i = k \sim MSN_J(\boldsymbol\zeta_k,\boldsymbol\alpha_k,\boldsymbol\Omega_k)
+\mathbf{y}_{i}|Z_i \sim MSN_J(\boldsymbol\zeta_k,\boldsymbol\alpha_k,\boldsymbol\Omega_k)
 \end{equation}
 
-where $\boldsymbol\zeta_k$ is a $J \times 1$ vector of class specific location parameters, $\boldsymbol\alpha_k$ is a $J \times 1$ vector of class specific skewness parameters, and $\boldsymbol\Omega_k$ is a $J \times J$ class specific scale matrix.
+where $Z_i = k$ is the cluster labeling of subject $i$, $\boldsymbol\zeta_k$ is a $J \times 1$ vector of cluster specific location parameters, $\boldsymbol\alpha_k$ is a $J \times 1$ vector of cluster specific skewness parameters, and $\boldsymbol\Omega_k$ is a $J \times J$ cluster specific scale matrix.
 
@@
 For conciseness, we present the MSN distribution here outside of the finite mixture context.
-When using this stochastic representation in our finite mixture model, the parameters $\boldsymbol\beta$, $\boldsymbol\psi$, and $\boldsymbol\Sigma$ are replaced with class-specific counterparts $\boldsymbol\beta_k$, $\boldsymbol\psi_k$, and $\boldsymbol\Sigma_k$.
+When using this stochastic representation in our finite mixture model, the parameters $\boldsymbol\beta$, $\boldsymbol\psi$, and $\boldsymbol\Sigma$ are replaced with cluster-specific counterparts $\boldsymbol\beta_k$, $\boldsymbol\psi_k$, and $\boldsymbol\Sigma_k$.
 
@@
 We note that the multivariate skew normal regression problem can be expressed more succinctly through the use of Matrix Skew Normal distributions.
-In particular, we define $\mathbf{Y}_{n_k \times J}$ as the response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n(Z_i = k)$ is the total number of observations belonging to class $k$ ($k = 1,...,K$).
+In particular, we define $\mathbf{Y}_{n_k \times J}$ as the response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n \mathds{1}_{(Z_i = k)}$ is the total number of observations belonging to cluster $k$ ($k = 1,...,K$).
 $$\mathbf{Y}_{n_k \times J} \sim MatSN_{n_k \times J}(\mathbf{M},\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Sigma_k)$$
@@
 This implies that all responses for cluster $k$ conditioned on $\mathbf{T}_{n_k \times 1}$ is distributed as
-$$\mathbf{Y}_{n_k \times J} | \mathbf{T}_k = \mathbf{t}_k \sim N_{n_k \times J}(\mathbf{M}_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k)$$
+$$\mathbf{Y}_{n_k \times J} | \mathbf{T}_k = \mathbf{t}_k \sim MatNorm_{n_k \times J}(\mathbf{M}_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k)$$
 $$\text{vec}(\mathbf{M}_k)_{n_kJ \times 1} = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$$\
@@
 
-\subsection{Multinomial Regression on Class Probabilities}
+\subsection{Multinomial Regression on Cluster Probabilities}
 
-A primary concern of our model is with identification of latent infant development clusters.
-We accomplish this via multinomial logit regression model on cluster membership, which utilizes P\'{o}lya-Gamma data-augmentation to allow for updating of all parameters using Gibbs sampling.
+A primary concern of our model is with identification of latent infant development classs.
+We accomplish this via multinomial logit regression model on cluster membership, which utilizes P\'{o}lya-Gamma data-augmentation, as described in Section 3.6 to allow for updating of all parameters using Gibbs sampling.
 The multinomial logit model is as follows for $k = 1,...,K$.
@@
 
-where $\mathbf{w}_i$ is the vector of class probability covariates for subject $i$, $\boldsymbol\delta_k$ contains the multinomial regression parameters for class $k$, and $K$ is the number of putative clusters specified \textit{a priori}.
+where $\mathbf{w}_i$ is the vector of cluster probability covariates for subject $i$, $\boldsymbol\delta_k$ contains the multinomial regression parameters for cluster $k$, and $K$ is the number of putative classs specified \textit{a priori}.
 For identifiability purposes, we fix the reference category $k = K$ and set $\boldsymbol\delta_K = \mathbf{0}$.
 Under this model, $Z_i|\mathbf{w}_i \sim Multinom(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{i1},...,\pi_{1K})$.
-During MCMC estimation, the class labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as class assignments.
+During MCMC estimation, the cluster labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as cluster assignments.
 
@@
 This provides more opportunities to explore the parameter space than multiple imputation and avoids multiplicative run-time scaling in $m$, the number of imputations.
+We demonstrate this feature using simulations detailed in Section 4.
 
@@
 
-Polson \textit{et al.} (2012) introduced a data augmentation approach to fitting the multinomial logit regression model specified by Holmes and Held (2006), which specifies the full conditional distributions of the multinomial regression parameters as a function of a Bernoulli likelihood.
+Polson \textit{et al.} (2013) introduced an efficient data augmentation approach to fitting several GLMs including the multinomial logit regression model specified by Holmes and Held (2006), which specifies the full conditional distributions of the multinomial regression parameters as a function of a Bernoulli likelihood.
 $$p(\boldsymbol\delta_k|\mathbf{z},\boldsymbol\delta_{k' \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \pi_{ik}^{U_{ik}}(1-\pi_{ik})^{1-U_{ik}}$$
 
-where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, $U_{ik} = \mathds{1}_{Z_i = k}$ is an indicator that subject $i$ belongs to class $k$, and $\pi_{ik}$ is defined as in Section 3.4.
+where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, $U_{ik} = \mathds{1}_{Z_i = k}$ is an indicator that subject $i$ belongs to cluster $k$, and $\pi_{ik}$ is defined as in Section 3.4.
 We can rewrite $\pi_{ik}$ as follows
@@
 \toprule
-\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Class 1} & \multicolumn{2}{c}{Class 2} & \multicolumn{2}{c}{Class 3} \\
+\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Cluster 1} & \multicolumn{2}{c}{Cluster 2} & \multicolumn{2}{c}{Cluster 3} \\
 \cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6} \cmidrule(l{3pt}r{3pt}){7-8}
@@
 
-\bibitem{ } Polson NG, Scott JG, Windle J.
-Bayesian inference for logistic models using P\'olya--Gamma latent variables.
-\textit{Journal of the American Statistical Association}.
-2013 Dec 1;108(504):1339-49.
+\bibitem{ } Polson NG, Scott JG, Windle J. Bayesian inference for logistic models using P\'olya –- Gamma latent variables. Journal of the American statistical Association. 2013 Dec 1;108(504):1339-49.
 

### 2019-06-18 simulation/simulation.R

@@ -156,7 +156,7 @@ S0 <- diag(0.5,v) # prior covariance for delta coefficients (multinomial regress
 
 # Sample storage
 nsim <- 10000 # number of iterations
-burn <- 1000 # number of iterations to save
+burn <- 2500 # number of iterations to save
 n.iter <- nsim - burn # number of saved iterations
 Z <- matrix(0,nrow = n.iter,ncol = n) # large matrix where each row is the value of z at a specific iteration
 PI <- matrix(0,nrow = n.iter,ncol = h) # matrix w/ each row as pi vector at each iteration

### 2019-06-18 manuscript/manuscript.tex
section 3

@@
 Later, we relax this assumption by allowing for missingness in the components of $\mathbf{y}_{i}$.
-The general formulation of the multivariate skew normal mixture model is as follows.
+For the analysis of the Nurture data, we propose the following MSN mixture
 \begin{equation}
-f(\mathbf{y}_i) = \sum_{k = 1}^{K} \pi_{ik} f(\mathbf{y}_i|\boldsymbol\theta_k)
+f(\mathbf{y}_i) = \sum_{k = 1}^{K} \pi_{ik} f(\mathbf{y}_i|\boldsymbol\theta_k),
 \end{equation}
+where $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$ ($k = 1,...,K$).
+For now we assume that $K$ is fixed, but we discuss model selection strategies for choosing the optimal value of $K$ in Section 4.
+The probability that subject $i$ ($i = 1,...,n$) belongs to cluster $k$ is denoted $\pi_{ik}$.
+Note that $\pi_{ik}$ is indexed by $i$, as we allow the probability of class membership to vary across subjects.
+However, we assume a that class membership is fixed throughout the study period, since our focus is to cluster individuals based on their overall developmental patterns over the course of the study.
+In Section 6, we discuss extensions to allow for class membership to vary over time.
 
-Here $\boldsymbol\theta_k$ is the set of parameters shared by subjects belonging to cluster $k$ ($k = 1,...,K$).
-The probability of subject $i$ ($i = 1,...,n$) belonging to cluster $k$ ($k = 1,...,K$) is denoted $\pi_{ik}$.
-Note that $\pi_{ik}$ is not indexed by time and we do not allow for subjects switching clusters over the study period.
-This feature of the model is a result of our interest in clustering based on development patterns rather than time-specific development scores.
-Later, we detail a multinomial logit regression model placed on $\pi_{ik}$ to allow for flexible modeling of cluster membership probabilities as a function of covariates of interest.
-
-In each development cluster, we assume that a subject's $J\times 1$ outcome vector follows a multivariate skew normal (MSN) distribution.
-We define a latent variable $Z_i \in \{1,...,K\}$ to denote the cluster membership of subject $i$ for $i = 1,...,n$.
-We model $\mathbf{y}_{i}|Z_i$ as
+In each cluster, we assume that the $J\times 1$ outcome vector for subject $i$ follows an MSN distribution.
+Let $z_i \in \{1,...,K\}$ denote a latent cluster membership of subject $i$.
+Conditional on $z_i = k$, we model $\mathbf{y}_{i}$ as
 \begin{equation}
-\mathbf{y}_{i}|Z_i \sim MSN_J(\boldsymbol\zeta_k,\boldsymbol\alpha_k,\boldsymbol\Omega_k)
+\mathbf{y}_{i}|z_i=k \sim MSN_J(\boldsymbol\zeta_k,\boldsymbol\alpha_k,\boldsymbol\Omega_k),
 \end{equation}
-
-where $Z_i = k$ is the cluster labeling of subject $i$, $\boldsymbol\zeta_k$ is a $J \times 1$ vector of cluster specific location parameters, $\boldsymbol\alpha_k$ is a $J \times 1$ vector of cluster specific skewness parameters, and $\boldsymbol\Omega_k$ is a $J \times J$ cluster specific scale matrix.
+where $\boldsymbol\zeta_k$ is a $J \times 1$ vector of cluster-specific location parameters, $\boldsymbol\alpha_k$ is a $J \times 1$ vector of cluster-specific skewness parameters, and $\boldsymbol\Omega_k$ is a $J \times J$ cluster-specific scale matrix that accounts for associations between the $J$ responses.
+The vector $\boldsymbol\alpha_k$ has components $\alpha_{jk}$, $j = 1,...,J$, that control the skewness of outcome $j$ in cluster $k$.
+When $\boldsymbol\alpha_k = \mathbf{0}$, the MSN distribution reduces to the multivariate normal distribution $N_J(\boldsymbol\zeta_k,\boldsymbol\Sigma_k)$, where $\boldsymbol\Sigma_J$ is the $J \times J$ covariance matrix capturing dependence among the $J$ outcomes.
 
@@
 
-We model the effect of covariates on longitudinal development outcomes through the use of a MSN regression model.
-We adopt a standard stochastic representation of a MSN random variable as the sum of a latent truncated normal random effect and a correlated multivariate normal random error term.
-The MSN stochastic representation is as follows.
+We model the effect of covariates on longitudinal motor development outcomes through the use of a MSN regression model.
+We adopt a standard stochastic representation of a MSN random variable by conditioning on a subject-specific latent truncated normal random effect, denoted $t_i$ (Fr\"{u}hwirth-Schnatter and Pyne 2010).
+Conditional on membership in cluster $k$, the stochastic representation of the MSN takes the form
 \begin{equation}
-\mathbf{y}_{i} = \boldsymbol\beta^T \mathbf{x}_{i} + t_{i}\boldsymbol\psi + \boldsymbol\epsilon_{i}
+\mathbf{y}_{ik} = \mathbf{x}_{ik}^T\boldsymbol\beta_k  + t_{i}\boldsymbol\psi_k + \boldsymbol\epsilon_{ik}
 \end{equation}
@@
 We note that the multivariate skew normal regression problem can be expressed more succinctly through the use of Matrix Skew Normal distributions.
-In particular, we define $\mathbf{Y}_{n_k \times J}$ as the response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n \mathds{1}_{(Z_i = k)}$ is the total number of observations belonging to cluster $k$ ($k = 1,...,K$).
+In particular, we define $\mathbf{Y}_{n_k \times J}$ as the response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n \mathds{1}_{(z_i = k)}$ is the total number of observations belonging to cluster $k$ ($k = 1,...,K$).
 $$\mathbf{Y}_{n_k \times J} \sim MatSN_{n_k \times J}(\mathbf{M},\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Sigma_k)$$
@@
 This implies that the $i^{th}$ row of $\mathbf{Y}$ is $\mathbf{y}_i \sim MSN_J(\boldsymbol\mu_i, \boldsymbol\alpha_k, \boldsymbol\Sigma_k) = {MSN}_J(\mathbf{x}_i \boldsymbol\beta_k, \boldsymbol\alpha_k, \boldsymbol\Sigma_k)$.
-Using the stochastic representation presented in Section 3.2, the conditional distribution of $\mathbf{y}_i | Z_i, t_i$ is $N_J (\mathbf{x}_i \boldsymbol\beta_k + \boldsymbol\psi_k t_i,\boldsymbol\Sigma_k)$.
+Using the stochastic representation presented in Section 3.2, the conditional distribution of $\mathbf{y}_i | z_i, t_i$ is $N_J (\mathbf{x}_i \boldsymbol\beta_k + \boldsymbol\psi_k t_i,\boldsymbol\Sigma_k)$.
 This implies that all responses for cluster $k$ conditioned on $\mathbf{T}_{n_k \times 1}$ is distributed as
@@
 The multinomial logit model is as follows for $k = 1,...,K$.
-$$P(Z_i = k|\mathbf{w}_i) = \pi_{ik} = \frac{e^{\mathbf{w}_i^T \boldsymbol\delta_k}}{\sum_{k' = 1}^K e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}}}$$
+$$P(z_i = k|\mathbf{w}_i) = \pi_{ik} = \frac{e^{\mathbf{w}_i^T \boldsymbol\delta_k}}{\sum_{k' = 1}^K e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}}}$$
 
@@
 For identifiability purposes, we fix the reference category $k = K$ and set $\boldsymbol\delta_K = \mathbf{0}$.
-Under this model, $Z_i|\mathbf{w}_i \sim Multinom(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{i1},...,\pi_{1K})$.
+Under this model, $z_i|\mathbf{w}_i \sim Multinom(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{i1},...,\pi_{1K})$.
 During MCMC estimation, the cluster labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as cluster assignments.
@@
 We allow for missingness of outcomes in the MSN mixture model by imputing missing values from their conditional multivariate normal distributions. We note that
-$$\mathbf{y}_i|Z_i,\mathbf{x}_i,t_i,\boldsymbol\beta_{k},\boldsymbol\psi_{k},\boldsymbol\Sigma_{k} \sim N_J(\boldsymbol\beta_{k}^T\mathbf{x}_i  + t_i \boldsymbol\psi_{k}, \boldsymbol\Sigma_{k})$$
+$$\mathbf{y}_i|z_i,\mathbf{x}_i,t_i,\boldsymbol\beta_{k},\boldsymbol\psi_{k},\boldsymbol\Sigma_{k} \sim N_J(\boldsymbol\beta_{k}^T\mathbf{x}_i  + t_i \boldsymbol\psi_{k}, \boldsymbol\Sigma_{k})$$
 
-where $Z_i = k$.
+where $z_i = k$.
 This allows us to appeal to standard conditional forms of the multivariate normal distribution to specify the distribution of missing observations.
@@
 We have
-$$\mathbf{y}_i^{miss}|\mathbf{y}_i^{obs},\mathbf{x}_i,t_i,Z_i,\boldsymbol\beta_{k},\boldsymbol\psi_{k},\boldsymbol\Sigma_{k} \sim N(\boldsymbol\mu^{miss},\boldsymbol\Sigma^{miss})$$
+$$\mathbf{y}_i^{miss}|\mathbf{y}_i^{obs},\mathbf{x}_i,t_i,z_i,\boldsymbol\beta_{k},\boldsymbol\psi_{k},\boldsymbol\Sigma_{k} \sim N(\boldsymbol\mu^{miss},\boldsymbol\Sigma^{miss})$$
 
@@
 
-where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, $U_{ik} = \mathds{1}_{Z_i = k}$ is an indicator that subject $i$ belongs to cluster $k$, and $\pi_{ik}$ is defined as in Section 3.4.
+where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, $U_{ik} = \mathds{1}_{z_i = k}$ is an indicator that subject $i$ belongs to cluster $k$, and $\pi_{ik}$ is defined as in Section 3.4.
 We can rewrite $\pi_{ik}$ as follows
@@
           \For {$ k = 1,...,K$}
-            \State $n_k := \sum_{i = 1}^n \mathds{1}_{Z_i = k}$
+            \State $n_k := \sum_{i = 1}^n \mathds{1}_{z_i = k}$
             \For {$i_k = 1,...,n_k$}
@@
             \For {$k = 1,...,K$}
-              \State $\pi_{ik} := P(Z_i = k|\mathbf{w}_i,\boldsymbol\delta_k)$
+              \State $\pi_{ik} := P(z_i = k|\mathbf{w}_i,\boldsymbol\delta_k)$
               \State $p_{ik} := P(\mathbf{y}_i|\boldsymbol\beta_k^{*T} \mathbf{x}^*_i,\boldsymbol\Sigma_k)$
             \EndFor
-            \State $\mathbf{p}_{Z_i} := \frac{\mathbf{p}_i \circ \boldsymbol\pi_i}{\mathbf{p}_i \cdot \boldsymbol\pi_i}$
-            \State $\mathbf{Draw} \ Z_i \ \text{from} \ \text{Categorical}(\mathbf{p}_{Z_i})$
+            \State $\mathbf{p}_{z_i} := \frac{\mathbf{p}_i \circ \boldsymbol\pi_i}{\mathbf{p}_i \cdot \boldsymbol\pi_i}$
+            \State $\mathbf{Draw} \ z_i \ \text{from} \ \text{Categorical}(\mathbf{p}_{z_i})$
             \For {$k = 1,...,K-1$}

### 2019-06-20 manuscript/manuscript.tex
section 3

@@
 Sara Benjamin-Neelon, PhD, MPH, RD \\
-Department of Health, Behavior and Society, Johns Hopkins University, Baltimore, MD, U.S.S}
+Department of Health, Behavior and Society, Johns Hopkins University, Baltimore, MD, U.S.A}
 
@@
 We adopt a standard stochastic representation of a MSN random variable by conditioning on a subject-specific latent truncated normal random effect, denoted $t_i$ (Fr\"{u}hwirth-Schnatter and Pyne 2010).
-Conditional on membership in cluster $k$, the stochastic representation of the MSN takes the form
+Conditional on membership in cluster $k$, the stochastic representation for the $j^{th}$ component of the multivariate skew normal distribution for subject $i$ is
 \begin{equation}
-\mathbf{y}_{ik} = \mathbf{x}_{ik}^T\boldsymbol\beta_k  + t_{i}\boldsymbol\psi_k + \boldsymbol\epsilon_{ik}
+y_{ij} = \zeta_{ijk} + t_i \psi_{jk} + \sqrt{1-\psi^2_{jk}}\epsilon_{ijk}.
 \end{equation}
-
-where $\mathbf{x}_i$ is the $P \times 1$ vector of covariate values for subject $i$, $\boldsymbol\beta$ is the $P \times J$ vector of fixed effects coefficients , $t_i \stackrel{iid}{\sim} N_{[0,\infty)}(0,1)$ is a conditional truncated normal random effect, $\boldsymbol\psi$ is the $J \times 1$ vector containing cluster-specific skewness parameters, and $\boldsymbol\epsilon_i \sim N_J(0,\boldsymbol\Sigma)$ is the correlated error term.
-In can be shown that under these conditions $\mathbf{y}_i$ is multivariate skew normal with location.
-For conciseness, we present the MSN distribution here outside of the finite mixture context.
-When using this stochastic representation in our finite mixture model, the parameters $\boldsymbol\beta$, $\boldsymbol\psi$, and $\boldsymbol\Sigma$ are replaced with cluster-specific counterparts $\boldsymbol\beta_k$, $\boldsymbol\psi_k$, and $\boldsymbol\Sigma_k$.
+where $y_{ij}$ denotes the $j^{th}$ observation for subject $i$, $\zeta_{ijk}$ denotes the location parameter for the $j^{th}$ observation for subject $i$ in cluster $k$, $t_i \stackrel{iid}{\sim} N_{[0,\infty)}(0,1)$ is a subject-specific truncated normal random effect, $\psi_{jk}$ controls the skewness of the $j^{th}$ outcome in cluster $k$, and $(\epsilon_{i1k},...,\epsilon_{iJk}) = \boldsymbol\epsilon_{ik} \sim N_J(0,\boldsymbol\Sigma_k)$ is a multivariate normal error term.
+Combining all observations for subject $i$ into $\mathbf{y}_i$, and conditional on $t_i$ and $z_i = k$, $\mathbf{y}_i$ follows a multivariate normal distribution with conditional mean $\mathbf{x}_i^T \boldsymbol\beta_k + t_i \boldsymbol\psi_k$ and conditional covariance $\boldsymbol\Sigma_k$.
+In the multivariate case, $\boldsymbol\psi_k$ is a $J \times 1$ vector containing cluster-specific skewness parameters.
+Marginally (after integrating over $t_i$), $\mathbf{y}_i$ is distributed $MSN_J(\boldsymbol\zeta_{ik}, \boldsymbol\alpha_k, \boldsymbol\Omega_k)$, where
+$$\boldsymbol\zeta_{ik} = \mathbf{x}_i^T \boldsymbol\beta_k,$$
+$$\boldsymbol\alpha_k = \frac{1}{\sqrt{1 - \boldsymbol\psi_k^T \boldsymbol\psi_k}} \boldsymbol\Omega^{-1}_k\boldsymbol\psi_k,$$
+$$\boldsymbol\Omega_k = \boldsymbol\Psi_k \boldsymbol\Sigma_k \boldsymbol\Psi_k + \boldsymbol\psi_k \boldsymbol\psi_k^T,$$
+where $\boldsymbol\Psi_k = \text{Diag}\left ( \sqrt{1 - \psi^2_{k1}},...,\sqrt{1 - \psi^2_{kJ}} \right )$.
+To allow for time varying effects in our model, we structure $\mathbf{X}_i$ as a $J \times pJ$ vector of covariate values for subject $i$, and $\boldsymbol\beta_k$ as a $pJ \times 1$ vector of fixed effects coefficients for cluster $k$ (think about how to explain this better).
 
@@
 
-We note that the multivariate skew normal regression problem can be expressed more succinctly through the use of Matrix Skew Normal distributions.
+We note that the multivariate skew normal density can be expressed more compactly in terms of the matrix skew normal (MatSN) density (Chen and Gupta 2005).
 In particular, we define $\mathbf{Y}_{n_k \times J}$ as the response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n \mathds{1}_{(z_i = k)}$ is the total number of observations belonging to cluster $k$ ($k = 1,...,K$).

### 2019-06-20 manuscript/manuscript.tex
section 3

@@
 \begin{equation}
-y_{ij} = \zeta_{ijk} + t_i \psi_{jk} + \sqrt{1-\psi^2_{jk}}\epsilon_{ijk}.
+y_{ij} = \zeta_{ijk} +  \psi_{jk} t_i + \sqrt{1-\psi^2_{jk}}\epsilon_{ijk}.
 \end{equation}
@@
 We note that the multivariate skew normal density can be expressed more compactly in terms of the matrix skew normal (MatSN) density (Chen and Gupta 2005).
-In particular, we define $\mathbf{Y}_{n_k \times J}$ as the response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n \mathds{1}_{(z_i = k)}$ is the total number of observations belonging to cluster $k$ ($k = 1,...,K$).
-$$\mathbf{Y}_{n_k \times J} \sim MatSN_{n_k \times J}(\mathbf{M},\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Sigma_k)$$
-$$\text{vec}(\mathbf{Y}) = \mathbf{y}_{n_kJ \times 1} = (\mathbf{y}^T_1,...,\mathbf{y}^T_{n_k})$$
-$$\text{vec}(\mathbf{M}) = \boldsymbol\mu_{n_k J \times 1} = (\boldsymbol\mu_1^T,...,\boldsymbol\mu_{n_k}^T)$$
-
-where $\boldsymbol\mu_i = \mathbf{x}_i \boldsymbol\beta_k$ and $\boldsymbol\alpha_k = (\alpha_{k1},...,\alpha_{kJ})$.
-This implies that the $i^{th}$ row of $\mathbf{Y}$ is $\mathbf{y}_i \sim MSN_J(\boldsymbol\mu_i, \boldsymbol\alpha_k, \boldsymbol\Sigma_k) = {MSN}_J(\mathbf{x}_i \boldsymbol\beta_k, \boldsymbol\alpha_k, \boldsymbol\Sigma_k)$.
+As we will see in Section 3.6, this specification facilitates prior specification by admitting convenient joint prior distributions for the regression coefficients and scale matrices.
+In particular, we define $\mathbf{Y}_k$ as the ${n_k \times J}$  response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n \mathds{1}_{(z_i = k)}$ is the total number of observations belonging to cluster $k$ ($k = 1,...,K$).
+$$\mathbf{Y}_{k} \sim MatSN_{n_k \times J}(\mathbf{M},\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Omega_k)$$
+$$\text{vec}(\mathbf{Y}) =  (\mathbf{y}^T_1,...,\mathbf{y}^T_{n_k})$$
+$$\text{vec}(\mathbf{M}) = (\boldsymbol\zeta_1^T,...,\boldsymbol\zeta_{n_k}^T)$$
+where $\boldsymbol\mu_i = \mathbf{x}_i^T \boldsymbol\beta_k$ and $\boldsymbol\alpha_k = (\alpha_{k1},...,\alpha_{kJ})$.
+This implies that the $i^{th}$ row of $\mathbf{Y}$ is $\mathbf{y}_i \sim MSN_J(\boldsymbol\zeta_i, \boldsymbol\alpha_k, \boldsymbol\Omega_k)$.
 Using the stochastic representation presented in Section 3.2, the conditional distribution of $\mathbf{y}_i | z_i, t_i$ is $N_J (\mathbf{x}_i \boldsymbol\beta_k + \boldsymbol\psi_k t_i,\boldsymbol\Sigma_k)$.
-This implies that all responses for cluster $k$ conditioned on $\mathbf{T}_{n_k \times 1}$ is distributed as
-$$\mathbf{Y}_{n_k \times J} | \mathbf{T}_k = \mathbf{t}_k \sim MatNorm_{n_k \times J}(\mathbf{M}_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k)$$
-$$\text{vec}(\mathbf{M}_k)_{n_kJ \times 1} = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$$\
-
-where $\mathbf{t}_k = (t_1,...,t_{n_k})$.
+This implies that all responses for cluster $k$ conditioned on $\mathbf{t}_{k}$ is distributed as
+$$\mathbf{Y}_{n_k \times J} | \mathbf{t}_k \sim MatNorm_{n_k \times J}(\mathbf{M}_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k),$$
+where $\text{vec}(\mathbf{M}_k)_{n_kJ \times 1} = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$ and $\mathbf{X}_k$ is an $n_k \times p$ design matrix.
 

### 2019-06-21 manuscript/manuscript_06-20-19.tex
section 3

--- 2019-06-20 manuscript/manuscript.tex
@@
 $$\text{vec}(\mathbf{M}) = (\boldsymbol\zeta_1^T,...,\boldsymbol\zeta_{n_k}^T)$$
-where $\boldsymbol\mu_i = \mathbf{x}_i^T \boldsymbol\beta_k$ and $\boldsymbol\alpha_k = (\alpha_{k1},...,\alpha_{kJ})$.
-This implies that the $i^{th}$ row of $\mathbf{Y}$ is $\mathbf{y}_i \sim MSN_J(\boldsymbol\zeta_i, \boldsymbol\alpha_k, \boldsymbol\Omega_k)$.
+where $\boldsymbol\zeta_i = \mathbf{x}_i^T \boldsymbol\beta_k$ and $\boldsymbol\alpha_k = (\alpha_{k1},...,\alpha_{kJ})$.
+This implies that the $i^{th}$ row of $\mathbf{Y}_k$, a ${n_k \times J}$ matrix of responses for cluster $k$, is $\mathbf{y}_i \sim MSN_J(\boldsymbol\zeta_i, \boldsymbol\alpha_k, \boldsymbol\Omega_k)$.
 Using the stochastic representation presented in Section 3.2, the conditional distribution of $\mathbf{y}_i | z_i, t_i$ is $N_J (\mathbf{x}_i \boldsymbol\beta_k + \boldsymbol\psi_k t_i,\boldsymbol\Sigma_k)$.
 This implies that all responses for cluster $k$ conditioned on $\mathbf{t}_{k}$ is distributed as
-$$\mathbf{Y}_{n_k \times J} | \mathbf{t}_k \sim MatNorm_{n_k \times J}(\mathbf{M}_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k),$$
+$$\mathbf{Y}_k | \mathbf{t}_k \sim MatNorm_{n_k \times J}(\mathbf{M}_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k),$$
 where $\text{vec}(\mathbf{M}_k)_{n_kJ \times 1} = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$ and $\mathbf{X}_k$ is an $n_k \times p$ design matrix.

### 2019-06-21 manuscript/sect_3.tex
section 3

--- 2019-06-21 manuscript/manuscript_06-20-19.tex
@@
-%  ALWAYS USE THE referee OPTION WITH PAPERS SUBMITTED TO BIOMETRICS!!!
-%  You can see what your paper would look like typeset by removing
-%  the referee option.  Because the typeset version will be in two
-%  columns, however, some of your equations may be too long. DO NOT
-%  use the \longequation option discussed in the user guide!!!  This option
-%  is reserved ONLY for equations that are impossible to split across 
-%  multiple lines; e.g., a very wide matrix.  Instead, type your equations 
-%  so that they stay in one column and are split across several lines, 
-%  as are almost all equations in the journal.  Use a recent version of the
-%  journal as a guide. 
-%  
-\documentclass[useAMS,referee]{biom}
+\documentclass{article}
 \usepackage{amsmath}
-\usepackage{graphicx}
-\usepackage{blindtext}
-\usepackage{booktabs}
-\usepackage{breqn}
 \usepackage{algorithm}
-\usepackage{algcompatible}
 \usepackage{algpseudocode}
-\usepackage{float}
-\usepackage{dsfont}
-
-\let\origfigure\figure
-\let\endorigfigure\endfigure
-\renewenvironment{figure}[1][2] {
-    \expandafter\origfigure\expandafter[H]
-} {
-    \endorigfigure
-}
 
@@
 
-%  Here, place your title and author information.  Note that in 
-%  use of the \author command, you create your own footnotes.  Follow
-%  the examples below in creating your author and affiliation information.
-%  Also consult a recent issue of the journal for examples of formatting.
-
-\title[]{Bayesian multivariate skew-normal finite mixture model for analysis of infant development trajectories}
-
-%  Here are examples of different configurations of author/affiliation
-%  displays.  According to the Biometrics style, in some instances,
-%  the convention is to have superscript *, **, etc footnotes to indicate 
-%  which of multiple email addresses belong to which author.  In this case,
-%  use the \email{ } command to produce the emails in the display.
-
-%  In other cases, such as a single author or two authors from 
-%  different institutions, there should be no footnoting.  Here, use
-%  the \emailx{ } command instead.
-
-%  The examples below corrspond to almost every possible configuration
-%  of authors and may be used as a guide.  For other configurations, consult
-%  a recent issue of the the journal.
-
-%  Single author -- USE \emailx{ } here so that no asterisk footnoting
-%  for the email address will be produced.
-
-%\author{John Author\emailx{email@address.edu} \\
-%Department of Statistics, University of Warwick, Coventry CV4 7AL, U.K.}
-
-%  Two authors from the same institution, with both emails -- use
-%  \email{ } here to produce the asterisk footnoting for each email address
-
-%\author{John Author$^{*}$\email{author@address.edu} and
-%Kathy Authoress$^{**}$\email{email2@address.edu} \\
-%Department of Statistics, University of Warwick, Coventry CV4 7AL, U.K.}
-
-%  Exactly two authors from different institutions, with both emails  
-%  USE \emailx{ } here so that no asterisk footnoting for the email address
-%  is produced.
-
-\author
-{Carter Allen\emailx{allecart@musc.edu} \\
-Department of Public Health Sciences, Medical University of South Carolina, Charleston, SC, U.S.A
-\and
-Brian Neelon, PhD \\
-Department of Public Health Sciences, Medical University of South Carolina, Charleston, SC, U.S.A
-\and
-Sara Benjamin-Neelon, PhD, MPH, RD \\
-Department of Health, Behavior and Society, Johns Hopkins University, Baltimore, MD, U.S.A}
-
-%  Three or more authors from same institution with all emails displayed
-%  and footnoted using asterisks -- use \email{ }
-
-%\author{John Author$^*$\email{author@address.edu}, 
-%Jane Author$^{**}$\email{jane@address.edu}, and 
-%Dick Author$^{***}$\email{dick@address.edu} \\
-%Department of Statistics, University of Warwick, Coventry CV4 7AL, U.K}
-
-%  Three or more authors from same institution with one corresponding email
-%  displayed
-
-%\author{John Author$^*$\email{author@address.edu}, 
-%Jane Author, and Dick Author \\
-%Department of Statistics, University of Warwick, Coventry CV4 7AL, U.K}
-
-%  Three or more authors, with at least two different institutions,
-%  more than one email displayed
-
-%\author{John Author$^{1,*}$\email{author@address.edu}, 
-%Kathy Author$^{2,**}$\email{anotherauthor@address.edu}, and 
-%Wilma Flinstone$^{3,***}$\email{wilma@bedrock.edu} \\
-%$^{1}$Department of Statistics, University of Warwick, Coventry CV4 7AL, U.K \\
-%$^{2}$Department of Biostatistics, University of North Carolina at 
-%Chapel Hill, Chapel Hill, North Carolina, U.S.A. \\
-%$^{3}$Department of Geology, University of Bedrock, Bedrock, Kansas, U.S.A.}
-
-%  Three or more authors with at least two different institutions and only
-%  one email displayed
-
-%\author{John Author$^{1,*}$\email{author@address.edu}, 
-%Wilma Flinstone$^{2}$, and Barney Rubble$^{2}$ \\
-%$^{1}$Department of Statistics, University of Warwick, Coventry CV4 7AL, U.K \\
-%$^{2}$Department of Geology, University of Bedrock, Bedrock, Kansas, U.S.A.}
-
 \begin{document}
-
-%  This will produce the submission and review information that appears
-%  right after the reference section.  Of course, it will be unknown when
-%  you submit your paper, so you can either leave this out or put in 
-%  sample dates (these will have no effect on the fate of your paper in the
-%  review process!)
-
-\date{{\it Received October} 2007. {\it Revised February} 2008.  {\it
-Accepted March} 2008.}
-
-%  These options will count the number of pages and provide volume
-%  and date information in the upper left hand corner of the top of the 
-%  first page as in published papers.  The \pagerange command will only
-%  work if you place the command \label{firstpage} near the beginning
-%  of the document and \label{lastpage} at the end of the document, as we
-%  have done in this template.
-
-%  Again, putting a volume number and date is for your own amusement and
-%  has no bearing on what actually happens to your paper!
-
-\pagerange{\pageref{firstpage}--\pageref{lastpage}} 
-\volume{64}
-\pubyear{2008}
-\artmonth{December}
-
-%  The \doi command is where the DOI for your paper would be placed should it
-%  be published.  Again, if you make one up and stick it here, it means 
-%  nothing!
-
-\doi{10.1111/j.1541-0420.2005.00454.x}
-
-%  This label and the label ``lastpage'' are used by the \pagerange
-%  command above to give the page range for the article.  You may have 
-%  to process the document twice to get this to match up with what you 
-%  expect.  When using the referee option, this will not count the pages
-%  with tables and figures.
-
-\label{firstpage}
-
-%  put the summary for your paper here
-
-\begin{abstract}
-In studies of infant motor development, a crucial research goal is to identify latent clusters of infants that experience delayed development, as this is a known risk factor for adverse outcomes later in life.
-However, there are a number of statistical challenges in modeling infant development: the data are typically skewed, exhibit intermittent missingness, and are highly correlated across the repeated measurements collected during infancy.
-Using data from the Nurture study, a cohort of over 600 mother-infant pairs followed from pregnancy to 12 months postpartum, we develop a flexible Bayesian latent class model for the analysis infant motor development.
-Our model has a number of attractive features.
-First, we adopt the multivariate skew normal distribution with cluster-specific parameters that accommodate the inherent correlation and skewness in the data.
-Second, we model the cluster membership probabilities using a novel Pólya-Gamma data-augmentation scheme, thereby improving predictions of the cluster membership allocations.
-Lastly, we impute missing responses under missing at random assumption by drawing from appropriate conditional skew normal distributions.
-Bayesian inference is achieved through straightforward Gibbs sampling, and can be carried out in available software such as R.
-Through simulation studies, we show that the proposed model yields improved inferences over models that ignore skewness.
-In addition, our imputation method yields improvements compared to conventional missing data methods, including multiple imputation and complete or available case analysis.
-When applied to Nurture data, we identified two distinct development clusters: one characterized by delayed “U-shaped” development and a higher percentage of male infants and another characterized by more steady development and a lower percentage of males.
-The clusters also differed in terms of key demographic variables, such as infant race and maternal pre-pregnancy body mass index.
-These findings can aid investigators in targeting interventions during this critical early-life developmental window.
-\end{abstract}
-
-%  Please place your key words in alphabetical order, separated
-%  by semicolons, with the first letter of the first word capitalized,
-%  and a period at the end of the list.
-%
-
-\begin{keywords}
-A key word; But another key word; Still another key word; Yet another key word.
-\end{keywords}
-
-%  As usual, the \maketitle command creates the title and author/affiliations
-%  display
-
-\maketitle
-
-\setcounter{tocdepth}{3}
-\tableofcontents
-
-\newpage
-
-%  If you are using the referee option, a new page, numbered page 1, will
-%  start after the summary and keywords.  The page numbers thus count the
-%  number of pages of your manuscript in the preferred submission style.
-%  Remember, ``Normally, regular papers exceeding 25 pages and Reader Reaction 
-%  papers exceeding 12 pages in (the preferred style) will be returned to 
-%  the authors without review. The page limit includes acknowledgements, 
-%  references, and appendices, but not tables and figures. The page count does 
-%  not include the title page and abstract. A maximum of six (6) tables or 
-%  figures combined is often required.''
-
-%  You may now place the substance of your manuscript here.  Please use
-%  the \section, \subsection, etc commands as described in the user guide.
-%  Please use \label and \ref commands to cross-reference sections, equations,
-%  tables, figures, etc.
-%
-%  Please DO NOT attempt to reformat the style of equation numbering!
-%  For that matter, please do not attempt to redefine anything!
-
-\section{Introduction}
-\label{s:intro}
-
-\subsection{Infant Development Clustering}
-
-Heterogeneity of treatment effects (HTE) (Lanza and Rhoades, 2013).
-
-\subsection{Existing Approaches}
-
-Mixtures of multivariate non-symmetric distributions such as the multivariate skew-normal (MSN) distribution allow for the nuances of the marginal density to be captured with a more parsimonious set of mixture components.
-Mixtures of MSN distributions have been dealt with previously in a Bayesian context (Fr\"{u}hwirth-Schnatter \& Pyne, 2010), however in these models, focus lies primary on marginal density estimation and inference on the mixture components (i.e.
-clusters) is not discussed.
-More recently, the mixtures of skew-$t$ factor analysis (MSTFA) model has been proposed for settings in which cluster-specific inference is of primary interest (Lin \textit{et al.} 2018).
-However, an important feature not included in the MSTFA is the ability to explain individual-level cluster membership as a function of covariates of interest.
-Additionally, parameter estimation proposed by Lin et al.
-for the MSTFA relies on a prohibitively complex EM algorithm and does not enjoy the inferential benefits of a Bayesian approach, namely the ability to incorporate prior information into a model and make posterior probability statements.
-Our proposed model improves on these previous works by estimating parameters in a Bayesian framework as well as including the ability to fit a multinomial logit regression to cluster membership probabilities using a novel application of data augmentation with the P\'olya Gamma distribution.
-
-\textbf{\textit{Put lit review of Bayesian PG multinomial logistic regression here}}
-
-A ubiquitous feature of repeated measures studies is loss of data due to intermittent missingness and attrition.
-In the Bayesian setting, the standard approach to dealing with missing data is to perform multiple imputation, whereby $m$ imputed data sets are generated from a specified imputation model.
-After $m$ complete data sets are obtained, parameter estimates are combined across each data set to produce a final set of parameter estimates (Gelman \textit{et al.} 2013).
-This approach is not only computationally burdensome, requiring storage and analysis of an $m \times n_{rows} \times n_{cols}$ data array in addition to multiplication of total model run time by a factor of $m$, but it has been shown to produce unreliable inferences (Zhou and Reiter, 2010).
-We instead include an ``online" imputation imputation step in our Gibbs sampling procedure, whereby missing outcomes are updated at each iteration.
-This approach greatly increases the number of opportunities for exploration of the missing data parameter space.
-
-\newpage
-
-\section{Nurture Study}
-\label{s:nurt}
-\subsection{Baseline Demographics and Description of Variables}
-
-\subsection{Statistical Challenges}
-\subsubsection{Skewness of Bayley score residuals}
-\subsubsection{Attrition and Intermittent Missingness}
-
-\newpage
 
@@
 As we will see in Section 3.6, this specification facilitates prior specification by admitting convenient joint prior distributions for the regression coefficients and scale matrices.
-In particular, we define $\mathbf{Y}_k$ as the ${n_k \times J}$  response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n \mathds{1}_{(z_i = k)}$ is the total number of observations belonging to cluster $k$ ($k = 1,...,K$).
+In particular, we define $\mathbf{Y}_k$ as the ${n_k \times J}$  response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n {1}_{(z_i = k)}$ is the total number of observations belonging to cluster $k$ ($k = 1,...,K$).
 $$\mathbf{Y}_{k} \sim MatSN_{n_k \times J}(\mathbf{M},\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Omega_k)$$
@@
 
-where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, $U_{ik} = \mathds{1}_{z_i = k}$ is an indicator that subject $i$ belongs to cluster $k$, and $\pi_{ik}$ is defined as in Section 3.4.
+where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, $U_{ik} = {1}_{z_i = k}$ is an indicator that subject $i$ belongs to cluster $k$, and $\pi_{ik}$ is defined as in Section 3.4.
 We can rewrite $\pi_{ik}$ as follows
@@
           \For {$ k = 1,...,K$}
-            \State $n_k := \sum_{i = 1}^n \mathds{1}_{z_i = k}$
+            \State $n_k := \sum_{i = 1}^n {1}_{z_i = k}$
             \For {$i_k = 1,...,n_k$}
@@
 
-\newpage
-
-\section{Simulation Studies}
-\label{s:sim}
-
-\subsection{Simulation to Compare to Multivariate Normal}
-
-\begin{table}[t]
-
-\caption{\label{tab:unnamed-chunk-4}Model results for simulated data with n = 1500, k = 4, p = 1, h = 3, v = 1.
-5000 iterations were run with a burn in of 1000.
-Missingness mechanism was MAR and P(miss) = 0}
-\centering
-\fontsize{8}{10}\selectfont
-\begin{tabular}{llllllll}
-\toprule
-\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Cluster 1} & \multicolumn{2}{c}{Cluster 2} & \multicolumn{2}{c}{Cluster 3} \\
-\cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6} \cmidrule(l{3pt}r{3pt}){7-8}
-Model Component & Parameter & True & Est. (95\% CrI) & True & Est. (95\% CrI) & True & Est. (95\% CrI)\\
-\midrule
-\addlinespace[0.3em]
-\multicolumn{8}{l}{\textbf{ }}\\
-\hspace{1em}MVSN & $\beta_{0}$ & -1.35 & -1.59 (-2.3, -0.94) & -0.18 & -0.24 (-0.53, 0.38) & 0.72 & 0.79 (0.52, 1.02)\\
-\hspace{1em}Regression & $\beta_{1}$ & -1.19 & -1.35 (-1.99, -0.84) & -0.09 & -0.33 (-0.62, 0.78) & 1.64 & 1.65 (1.33, 1.91)\\
-\hspace{1em} & $\beta_{2}$ & -1.65 & -1.81 (-2.66, -1.41) & -0.47 & -0.62 (-0.89, 0.3) & 1.44 & 1.32 (0.97, 1.6)\\
-\hspace{1em} & $\beta_{3}$ & -1.75 & -1.89 (-2.52, -1.37) & -0.22 & -0.32 (-0.63, 0.55) & 2.28 & 2.26 (1.96, 2.5)\\
-\addlinespace[0.3em]
-\multicolumn{8}{l}{\textbf{ }}\\
-\hspace{1em} & $\sigma_{11}$ & 1 & 1.02 (0.78, 1.19) & 1 & 0.98 (0.77, 1.23) & 1 & 1.06 (0.85, 1.29)\\
-\hspace{1em} & $\sigma_{12}$ & -0.32 & -0.19 (-0.33, -0.02) & 0.16 & 0.14 (-0.01, 0.4) & 0.72 & 0.82 (0.62, 1.05)\\
-\hspace{1em} & $\sigma_{13}$ & -0.65 & -0.55 (-0.68, -0.35) & 0.72 & 0.7 (0.51, 0.94) & 0.14 & 0.27 (0.1, 0.48)\\
-\hspace{1em} & $\sigma_{14}$ & -0.44 & -0.33 (-0.46, -0.13) & 0.5 & 0.48 (0.31, 0.72) & -0.01 & -0.02 (-0.16, 0.16)\\
-\hspace{1em} & $\sigma_{22}$ & 1 & 0.92 (0.72, 1.06) & 1 & 0.94 (0.72, 1.22) & 1 & 1.11 (0.87, 1.38)\\
-\hspace{1em} & $\sigma_{23}$ & 0.56 & 0.49 (0.33, 0.6) & 0.53 & 0.46 (0.29, 0.73) & -0.1 & 0.08 (-0.1, 0.28)\\
-\hspace{1em} & $\sigma_{24}$ & 0.98 & 0.9 (0.7, 1.04) & 0.24 & 0.14 (-0.03, 0.41) & 0.19 & 0.17 (0.01, 0.37)\\
-\hspace{1em} & $\sigma_{33}$ & 1 & 0.9 (0.66, 1.04) & 1 & 0.92 (0.72, 1.19) & 1 & 1.26 (1.03, 1.52)\\
-\hspace{1em} & $\sigma_{34}$ & 0.56 & 0.51 (0.35, 0.62) & 0.86 & 0.79 (0.58, 1.05) & -0.65 & -0.63 (-0.78, -0.45)\\
-\hspace{1em} & $\sigma_{44}$ & 1 & 0.93 (0.72, 1.07) & 1 & 0.93 (0.7, 1.2) & 1 & 1.11 (0.87, 1.36)\\
-\addlinespace[0.3em]
-\multicolumn{8}{l}{\textbf{ }}\\
-\hspace{1em} & $\psi_{1}$ & -0.33 & -0.02 (-0.84, 0.87) & 0.67 & 0.69 (-0.09, 1.01) & -1 & -0.98 (-1.25, -0.67)\\
-\hspace{1em} & $\psi_{2}$ & -0.33 & -0.16 (-0.8, 0.62) & 0.67 & 0.81 (-0.6, 1.14) & -1 & -0.98 (-1.28, -0.6)\\
-\hspace{1em} & $\psi_{3}$ & -0.33 & -0.15 (-0.65, 0.89) & 0.67 & 0.72 (-0.35, 1.02) & -1 & -0.8 (-1.15, -0.36)\\
-\hspace{1em} & $\psi_{4}$ & -0.33 & -0.18 (-0.84, 0.59) & 0.67 & 0.71 (-0.32, 1.06) & -1 & -1 (-1.32, -0.62)\\
-\addlinespace[0.3em]
-\multicolumn{8}{l}{\textbf{ }}\\
-\hspace{1em}Multinom. & $\delta_{11}$ & -0.84 & -0.78 (-0.96, -0.59) & -0.84 & -0.78 (-0.96, -0.59) & -0.84 & -0.78 (-0.96, -0.59)\\
-\hspace{1em} & $\delta_{12}$ & -0.24 & -0.26 (-0.42, -0.1) & -0.24 & -0.26 (-0.42, -0.1) & -0.24 & -0.26 (-0.42, -0.1)\\
-\addlinespace[0.3em]
-\multicolumn{8}{l}{\textbf{ }}\\
-\hspace{1em}Clustering & $\pi_l$ & 0.39 & 0.39 (0.38, 0.4) & 0.26 & 0.26 (0.25, 0.27) & 0.34 & 0.35 (0.33, 0.36)\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\subsection{Simulation to Compare Imputation Methods}
-
-\subsection{Simulation to Assess Sensitivity to Misspecified K}
-
-\newpage
-
-\section{Application}
-\label{s:app}
-
-\begin{itemize}
-
-\item Include both time varying and non-time varying covariates for the within cluster covariate set.
-
-\end{itemize}
-
-\newpage
-
-\section{Discussion}
-\label{s:discuss}
-
-\textit{\textbf{Discuss non-ignorable missingness here}}
-
-\begin{itemize}
-
-\item Discuss how we handle non-ignorable missingness
-
-\end{itemize}
-
-\section{Appendix}
-
-Put your final comments here.
-
-%  The \backmatter command formats the subsequent headings so that they
-%  are in the journal style.  Please keep this command in your document
-%  in this position, right after the final section of the main part of 
-%  the paper and right before the Acknowledgements, Supplementary Materials,
-%  and References sections.
-
-\backmatter
-
-%  This section is optional.  Here is where you will want to cite
-%  grants, people who helped with the paper, etc.  But keep it short!
-
-\section*{Acknowledgements}
-
-%  If your paper refers to supplementary web material, then you MUST
-%  include this section!!  See Instructions for Authors at the journal
-%  website http://www.biometrics.tibs.org
-
-\section*{Supplementary Materials}
-
-\newpage
-
-\subsection{Glossary of Notation}
-
-\begin{itemize}
-
-    \item $\mathbf{Y}$: A $n \times J$ matrix containing all multivariate skew-normal  outcomes such that $y_{ij}$ is the $j^{th}$ outcome observed for subject $i$, where $i = 1,...,n$ and $j = 1,...J$.
-
-    \item $\mathbf{X}$: A $n \times P$ matrix containing all multivariate skew-normal regression covariates such that $x_{ip}$ is the $p^{th}$ covariate value for subject $i$, where $i = 1,...,n$ and $p = 1,...P$.
-
-    \item $\mathbf{B}$: A $P \times J$ matrix containing all multivariate skew-normal regression coefficients such that $\mathbf{B} = \left [ \boldsymbol\beta_1,...,\boldsymbol\beta_J \right ]$, where $\beta_{pj}$ is interpreted as the effect of covariate $p$ on outcome $j$ for $p = 1,...,P$ and $j = 1,...,J$.
-
-    \item $\mathbf{E}$: A $n \times J$ matrix of error terms in the multivariate skew-normal regression model component.
-$\mathbf{E}$ is made up of row vectors $\boldsymbol\epsilon_i = (\epsilon_{i1},...,\epsilon_{iJ})$, where $ \boldsymbol\epsilon_i \stackrel{iid}{\sim} N_J(0, \boldsymbol\Sigma)$ for $i = 1,...,n$.
-
-    \item $\boldsymbol\Sigma$: A $J \times J$ covariance matrix that defines the correlation between the $p$ multivariate normal outcomes.
-
-    \item $\boldsymbol\Omega$: A $J \times J$ covariance scale matrix that defines the correlation between the $p$ multivariate skew-normal outcomes.
-
-    \item $\boldsymbol\psi$: A $J \times 1$ vector containing the skewness parameter for each outcome.
-
-    \item $\boldsymbol\alpha$: A $J \times 1$ vector containing the skewness parameter for each outcome.
-
-    \item $\mathbf{t}$: An $n \times 1$ vector of truncated normal random effects used in the stochastic representation of the multivariate skew-normal distribution.
-For $i = 1,...,n$, $t_i \stackrel{iid}{\sim}T_{[0,\infty)}(0,1)$
-
-    \item $\mathbf{X}^*$: A $n \times (P + 1)$ matrix constructed by column binding $\mathbf{t}$ to $\mathbf{X}$
-
-    \item $\mathbf{B}^*$: A $(P+1) \times J$ matrix constructed by row binding $\boldsymbol\psi^T$ to $\mathbf{B}$.
-
-\end{itemize}
-
-\newpage
-\subsection{Derivation of Full Conditional Distributions}
-
-\subsubsection{Multivariate Skew-Normal Regression}
-
-Without loss of generality, we derive the full conditional distributions for the multivariate skew-normal regression model component under the assumption that all observations belong to a single cluster.
-To make the extension to the case where more than one cluster is specified, simply apply these distributional forms to cluster specific parameters and data.
-Finally, we assume for the moment that we have complete data for all outcomes for each subject.
-We extend consider the case of missing data in section (INSERT SECTION).
-
-The multivariate skew-normal regression model can be written as follows in matrix form. 
-$$\mathbf{Y} = \mathbf{X}\mathbf{B} + \mathbf{t} \boldsymbol\psi^T + \mathbf{E} = \mathbf{X}^* \mathbf{B}^* + \mathbf{E}$$
-The matrix $\mathbf{Y}$ is of dimension $n \times J$.
-For convenience, we define $\mathbf{X}^*$ as a $n \times (P + 1)$ matrix constructed by column binding $\mathbf{t}$ to $\mathbf{X}$, and $\mathbf{B}^*$ as a $(P+1) \times J$ matrix constructed by row binding $\boldsymbol\psi^T$ to $\mathbf{B}$.
-We assume that $t_i \stackrel{iid}{\sim}T_{[0,\infty})(0,1)$ and that $\mathbf{E}$ is made of row vectors $\boldsymbol\epsilon_i = (\epsilon_{i1},...,\epsilon_{iJ})$ for $i = 1,...,n$, where $ \boldsymbol\epsilon_i \stackrel{iid}{\sim} N_J(0, \boldsymbol\Sigma)$.
-
-The conditional likelihood for this model is given below. 
-$$p(\mathbf{Y}|\mathbf{X}^*,\mathbf{B}^*,\boldsymbol\Sigma) \propto |\boldsymbol\Sigma|^{-n/2}\exp \left \{ -\frac{1}{2} \tr(\mathbf{Y} - \mathbf{X}^* \mathbf{B}^*)^T(\mathbf{Y} - \mathbf{X}^* \mathbf{B}^*)\boldsymbol\Sigma^{-1} \right \}$$
-
-We choose conjugate priors for $\mathbf{B}^*$ and $\boldsymbol\Sigma$ as follows. 
-$$\boldsymbol\Sigma \sim \text{inverse-Wishart}(\mathbf{V}_0,\nu_0)$$
-$$\mathbf{B}^*|\boldsymbol\Sigma \sim MatNorm_{(m+1), p}(\mathbf{B}_0^*,\mathbf{L}_0^{-1},\boldsymbol\Sigma)$$
-
-We now derive the joint posterior distribution of the parameters $\mathbf{B}^*$ and $\boldsymbol\Sigma$.
-\begin{align*} 
-p(\mathbf{B}^*,\boldsymbol\Sigma|\mathbf{X}^*,\mathbf{Y}) & \propto  p(\mathbf{Y}|\mathbf{X}^*,\mathbf{B}^*,\boldsymbol\Sigma)p(\mathbf{B}^*|\boldsymbol\Sigma)p(\boldsymbol\Sigma)\\
-&  \propto |\boldsymbol\Sigma|^{-n/2}\exp \left \{ -\frac{1}{2} \tr \left [(\mathbf{Y} - \mathbf{X^*}\mathbf{B}^*)^T(\mathbf{Y} - \mathbf{X^*}\mathbf{B}^*)\boldsymbol\Sigma^{-1} \right ]\right \} \\
-& \times |\boldsymbol\Sigma|^{-(P+1)/2} \exp \left \{ -\frac{1}{2} \tr \left [(\mathbf{B}^* - \mathbf{B}^*_0)^T \mathbf{L}_0(\mathbf{B}^* - \mathbf{B}^*_0)\boldsymbol\Sigma^{-1} \right ]\right \} \\
-& \times |\boldsymbol\Sigma|^{(\nu_0 + J + 1)/2} \exp \left \{ -\frac{1}{2} \tr (\mathbf{V}_0 \boldsymbol\Sigma^{-1}) \right \}
-\end{align*}
-
-\subsubsection{Multinomial Logit Regression}
-
-\subsubsection{Multivariate Normal Conditional Imputation}
-
-The multivariate normal conditional imputation derivations are given for a single cluster without loss of generality.
-In practice, the data and parameters in this section would be replaced by cluster specific estimates in the case of clustering.
-
-For a given observation vector $\mathbf{y} \sim N_J(\boldsymbol\mu, \boldsymbol\Sigma)$, we allow for missingness in at most $J - 1$ of the multivariate outcomes through the use of a conditional imputation step embedded within our Gibbs sampler.
-Suppose $\mathbf{y}$ contains $q$ missing observations and can be partitioned into two vectors $\mathbf{y_{1}}$ and $\mathbf{y_{2}}$ such that $\mathbf{y_{1}}$ is a $q \times 1$ vector of missing observations and $\mathbf{y_{2}}$ is a $(J-q) \times 1$ vector of complete observations.
-Similarly, partition $\boldsymbol\mu$ and $\boldsymbol\Sigma$ as follows.
-$$\boldsymbol\mu = \begin{bmatrix} \boldsymbol\mu_1 \\ \boldsymbol\mu_2 \end{bmatrix} \ \ \ \ \ \boldsymbol\Sigma = \begin{bmatrix} \boldsymbol\Sigma_{11} & \boldsymbol\Sigma_{12} \\ \boldsymbol\Sigma_{21} & \boldsymbol\Sigma_{22} \end{bmatrix}$$
-
-We will use these quantities to derive the conditional distribution $f(\mathbf{y_1}|\mathbf{y_2},\boldsymbol\mu,\boldsymbol\Sigma)$. 
-\begin{align*} 
-f(\mathbf{y_1}|\mathbf{y_2},\boldsymbol\mu,\boldsymbol\Sigma) & \propto  f(\mathbf{y_1},\mathbf{y_2}|\boldsymbol\mu,\boldsymbol\Sigma) \\
-& \propto \exp \left \{ -\frac{1}{2}(\mathbf{y} - \boldsymbol\mu)^T \boldsymbol\Sigma^{-1} (\mathbf{y} - \boldsymbol\mu) \right \} \\
-& = \exp \left \{ -\frac{1}{2}\begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix}^T \boldsymbol\Sigma^{-1} \begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix} \right \} \\
-& =  \exp \left \{ -\frac{1}{2}\begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix}^T \begin{bmatrix} \boldsymbol\Sigma_{11} & \boldsymbol\Sigma_{12} \\ \boldsymbol\Sigma_{21} & \boldsymbol\Sigma_{22} \end{bmatrix}^{-1} \begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix} \right \} \\
-& = \exp \left \{ -\frac{1}{2}\begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix}^T \begin{bmatrix} \boldsymbol\Sigma_{11}^* & \boldsymbol\Sigma_{12}^* \\ \boldsymbol\Sigma_{21}^* & \boldsymbol\Sigma_{22}^* \end{bmatrix} \begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix} \right \} \\
-& = \exp \left \{ -\frac{1}{2} \left[ (\mathbf{y}_1 - \boldsymbol\mu_{cond})^T \boldsymbol\Sigma_{cond}^{-1}(\mathbf{y}_1 - \boldsymbol\mu_{cond})\right]\right \} 
-\end{align*}
-$$\Rightarrow \mathbf{y_1}|\mathbf{y_2},\boldsymbol\mu,\boldsymbol\Sigma \sim N_q(\boldsymbol\mu_{cond},\boldsymbol\Sigma_{cond})$$
-$$\boldsymbol\mu_{cond} = \boldsymbol\mu_1 + \boldsymbol\Sigma_{12}\boldsymbol\Sigma_{22}^{-1}(\mathbf{y}_2 - \boldsymbol\mu_2), \ \ \ \ \ \ \boldsymbol\Sigma_{cond} = \boldsymbol\Sigma_{11} - \boldsymbol\Sigma_{12}\boldsymbol\Sigma_{22}^{-1}\boldsymbol\Sigma_{21}$$
-
-The block-wise inversion formula was used to invert $\boldsymbol\Sigma$ according to the following reparameterizations.
-\begin{align*}
-\boldsymbol\Sigma_{11}^* & = \boldsymbol\Sigma_{11}^{-1} + \boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12}(\boldsymbol\Sigma_{22} - \boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12})^{-1}\boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\\
-\boldsymbol\Sigma_{12}^* & = -\boldsymbol\Sigma_{11} \boldsymbol\Sigma_{12}(\boldsymbol\Sigma_{22}-\boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12})^{-1}\\
-\boldsymbol\Sigma_{21}^* & = -(\boldsymbol\Sigma_{22} - \boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12})^{-1}\boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\\
-\boldsymbol\Sigma_{22}^* & = (\boldsymbol\Sigma_{22} - \boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12})^{-1}     
-\end{align*}
-
-%  Here, we create the bibliographic entries manually, following the
-%  journal style.  If you use this method or use natbib, PLEASE PAY
-%  CAREFUL ATTENTION TO THE BIBLIOGRAPHIC STYLE IN A RECENT ISSUE OF
-%  THE JOURNAL AND FOLLOW IT!  Failure to follow stylistic conventions
-%  just lengthens the time spend copyediting your paper and hence its
-%  position in the publication queue should it be accepted.
-
-%  We greatly prefer that you incorporate the references for your
-%  article into the body of the article as we have done here 
-%  (you can use natbib or not as you choose) than use BiBTeX,
-%  so that your article is self-contained in one file.
-%  If you do use BiBTeX, please use the .bst file that comes with 
-%  the distribution.  In this case, replace the thebibliography
-%  environment below by 
-%
-%  \bibliographystyle{biom} 
-% \bibliography{mybibilo.bib}
-
-\newpage
-
-\begin{thebibliography}{}
-
-\bibitem{ } Arellano‐-Valle RB, Azzalini A. On the unification of families of skew‐normal distributions. \textit{Scandinavian Journal of Statistics}. 2006 Sep;33(3):561-74.
-
-\bibitem{ } Azzalini A. A class of distributions which includes the normal ones. \textit{Scandinavian journal of statistics}. 1985 Jan 1:171-8.
-
-\bibitem{ } Azzalini, A. and Dalla Valle, A. (1996). The multivariate skew normal distribution. Biometrika 83, 715–726.
-
-\bibitem{ } Chen JT, Gupta AK. Matrix variate skew normal distributions. \textit{Statistics}. 2005 Jun 1;39(3):247-53.
-
-\bibitem{ } Neelon SE, \O stbye T, Bennett GG, Kravitz RM, Clancy SM, Stroo M, Iversen E, Hoyo C.
-Cohort profile for the Nurture Observational Study examining associations of multiple caregivers on infant growth in the Southeastern USA.
-\textit{BMJ Open}.
-2017 Feb 1;7(2):e013939.
-
-\bibitem{ } Franczak BC, Tortora C, Browne RP, McNicholas PD.
-Unsupervised learning via mixtures of skewed distributions with hypercube contours.
-\textit{Pattern Recognition Letters}.
-2015 Jun 1;58:69-76.
-
-\bibitem{ } Fr\"{u}hwirth-Schnatter S, Pyne S.
-Bayesian inference for finite mixtures of univariate and multivariate skew-normal and skew-t distributions.
-\textit{Biostatistics}.
-2010 Jan 27;11(2):317-36.
-
-\bibitem{ } Ganjali M, Baghfalaki T.
-A Bayesian shared parameter model for analysing longitudinal skewed responses with nonignorable dropout.
-\textit{International Journal of Statistics in Medical Research}.
-2014 Apr 1;3(2):103.
-
-\bibitem{ } Gelman A, Stern HS, Carlin JB, Dunson DB, Vehtari A, Rubin DB. Bayesian data analysis. \textit{Chapman and Hall/CRC}; 2013 Nov 27.
-
-\bibitem{ } Gelman A, Hwang J, Vehtari A. Understanding predictive information criteria for Bayesian models. \textit{Statistics and Computing}. 2014 Nov 1;24(6):997-1016.
-
-\bibitem{ } Holmes CC, Held L. Bayesian auxiliary variable models for binary and multinomial regression. \textit{Bayesian Analysis}. 2006;1(1):145-68.
-
-\bibitem{ } Lagona F, Picone M. Model-based clustering of multivariate skew data with circular components and missing values. \textit{Journal of Applied Statistics}. 2012 May 1;39(5):927-45.
-
-\bibitem{ } Lanza ST, Rhoades BL. Latent class analysis: an alternative perspective on subgroup analysis in prevention and treatment. \textit{Prevention Science}. 2013 Apr 1;14(2):157-68.
-
-\bibitem{ } Lee SX, McLachlan GJ. Model-based clustering and classification with non-normal mixture distributions. Statistical Methods \& Applications. 2013 Nov 1;22(4):427-54.
-
-\bibitem{ } Lee SX, Mclachlan GJ. On mixtures of skew normal and skew $t$-distributions. \textit{Advances in Data Analysis and Classification}. 2013 Sep 1;7(3):241-66.
-
-\bibitem{ } Lin TI, Wang WL, McLachlan GJ, Lee SX.
-Robust mixtures of factor analysis models using the restricted multivariate skew-t distribution.
-\textit{Statistical Modelling}.
-2018 Feb;18(1):50-72.
-
-\bibitem{ } Luo S, Lawson AB, He B, Elm JJ, Tilley BC.
-Bayesian multiple imputation for missing multivariate longitudinal data from a Parkinson's disease clinical trial.
-\textit{Statistical Methods in Medical Research}.
-2016 Apr;25(2):821-37.
-
-\bibitem{ } Melnykov V, Maitra R. Finite mixture models and model-based clustering. \textit{Statistics Surveys}. 2010;4:80-116.
-
-\bibitem{ } Polson NG, Scott JG, Windle J. Bayesian inference for logistic models using P\'olya –- Gamma latent variables. Journal of the American statistical Association. 2013 Dec 1;108(504):1339-49.
-
-\bibitem{ } Tiao GC, Zellner A. On the Bayesian estimation of multivariate regression. \textit{Journal of the Royal Statistical Society}: Series B (Methodological). 1964 Jul;26(2):277-85.
-
-\bibitem{ } Viroli C. Finite mixtures of matrix normal distributions for classifying three-way data. \textit{Statistics and Computing}. 2011 Oct 1;21(4):511-22.
-
-\bibitem{ } Vrbik I, Mcnicholas PD. Parsimonious skew mixture models for model-based clustering and classification. \textit{Computational Statistics \& Data Analysis}. 2014 Mar 1;71:196-210.
-
-\bibitem{ } Zeller CB, Cabral CR, Lachos VH. Robust mixture regression modeling based on scale mixtures of skew-normal distributions. \textit{Test}. 2016 Jun 1;25(2):375-96.
-
-\bibitem{ } Zhou X, Reiter JP. A note on Bayesian inference after multiple imputation. \textit{The American Statistician}. 2010 May 1;64(2):159-63.
-
-\end{thebibliography}
-
-% \appendix
-
-%  To get the journal style of heading for an appendix, mimic the following.
-
-\label{lastpage}
-
 \end{document}

### 2019-06-24 manuscript/sect_3_06-22-19_bn.tex
section 3

\documentclass{article}
\usepackage{amsmath}
%\usepackage{algorithm}
%\usepackage{algpseudocode}

\def\bSig\mathbf{\Sigma}
\newcommand{\VS}{V\&S}
\newcommand{\tr}{\mbox{tr}}

\begin{document}

\section{Model}
\label{s:model}

\subsection{Multivariate Skew Normal Mixture Model} \textbf{[Check capitalization of headings for Biometrics -- maybe only first letter is capitalized?]}

A primary goal of the Nurture study is to identify clusters of infants characterized by distinct motor development trajectories. To address this aim, we propose a flexible finite mixture model that accommodates relevant features of the data, such as skewness and dependence among the responses. To this end, let $\mathbf{y}_{i}=(y_{i1},\ldots,y_{iJ})^T$ be a $J \times 1$ vector of responses (i.e., Baley scores) for subject $i$ $(i=1,\ldots,n)$. For the analysis of the Nurture data, we propose a finite mixture model of the form
\begin{equation}
f(\mathbf{y}_i) = \sum_{k = 1}^{K} \pi_{ik} f(\mathbf{y}_i|\boldsymbol\theta_k),
\end{equation}
where $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$ ($k = 1,...,K$) and $\pi_{ik}$ is a subject-specific mixing weight representing the probability that subject $i$ belongs to cluster $k$. For now we assume that $K$ is fixed; in Section 4, we discuss model selection strategies for choosing the optimal value of $K$. We also assume that class membership is fixed throughout the study period, since our focus is to cluster individuals based on their overall developmental patterns over the course of the study. In Section 6, we discuss extensions to allow for class membership to vary over time. \textbf{[We could omit these last two sentence -- are they really needed? Not sure. Maybe keep for now and think about it.]}

To facilitate posterior inference, we introduce a latent cluster indicator variable $z_i$ taking the value $k \in \{1,...,K\}$ with probability $\pi_{ik}$. Conditional on $z_i = k$, we assume $\mathbf{y}_{i}$ is distributed as
\begin{equation}
\mathbf{y}_{i}|(z_i=k) \sim MSN_J(\boldsymbol\zeta_{ki},\boldsymbol\alpha_k,\boldsymbol\Omega_k), \label{eq:msndens}
\end{equation}
where $MSN_J(\cdot)$ denotes the $J$-dimensional multivariate skew normal density, $\boldsymbol\zeta_{ki}$ is a $J \times 1$ vector of subject- and cluster-specific location parameters, $\boldsymbol\alpha_k$ is a $J \times 1$ vector of cluster-specific skewness parameters, and $\boldsymbol\Omega_k$ is a $J \times J$ cluster-specific scale matrix that captures dependence among the $J$ responses. The vector $\boldsymbol\alpha_k$ has components $\alpha_{kj}$, $j = 1,...,J$ \textbf{[let's use $kij$ as our index hierarchy. Please review throughout]}, that control the skewness of outcome $j$ in cluster $k$. When $\boldsymbol\alpha_k = \mathbf{0}$, the MSN distribution reduces to the multivariate normal distribution $N_J(\boldsymbol\zeta_k,\boldsymbol\Omega_k)$, where $\boldsymbol\Omega_k$ is a $J \times J$ covariance matrix.

We can extend model (\ref{eq:msndens}) to the regression setting by modeling $\boldsymbol\zeta_{ki}$ as a function of covariates. Here we adopt a convenient stochastic representation of the MSN density (Azzalini and Dalla Valle, 1996):
\begin{equation}
\mathbf{y}_i|(z_i=k,t_i) = \mathbf{X}_i \boldsymbol\beta_k + t_i \boldsymbol\psi_k + \boldsymbol\epsilon_{ki}, \label{eq:msnreg}
\end{equation}
where 
$\mathbf{X}_i$ is a $J \times Jp$ design matrix that includes potential time-varying covariates (e.g., indicators denoting quarterly visits); $\boldsymbol{\beta}_k=(\beta_{k11},\ldots,\beta_{k1p},\ldots,\beta_{kJ1},\ldots,\beta_{kJp})^T$ is a $Jp\times 1$ vector of cluster- and outcome-specific regression coefficients; $t_i\sim N_{[0,\infty)}(0,1)$ is a subject-specific standard normal random variable truncated below by zero; $\boldsymbol\psi_k=(\psi_{k1},\ldots,\psi_{kJ})^T$ is a $J \times 1$ vector of cluster-specific skewness parameters; and $\boldsymbol{\epsilon}_{ki} \sim N_J(\boldsymbol0,\boldsymbol\Sigma_k)$ is a $J\times 1$ vector of error terms. Thus, conditional on $t_i$ and $z_i=k$, $\boldsymbol{y}_i$ is distributed as $N_J(\mathbf{X}_i \boldsymbol\beta_k + t_i \boldsymbol\psi_k, \boldsymbol{\Sigma}_k)$. Marginally (after integrating over $t_i$), $\mathbf{y}_i$ is distributed $MSN_J(\boldsymbol\zeta_{ik}, \boldsymbol\alpha_k, \boldsymbol\Omega_k)$, where through back-transformation \textbf{[Carter: Please review these expressions carefully and make sure they conform to those in FS and Pyne]}
\begin{eqnarray*}
\boldsymbol\zeta_{ik} &=& \mathbf{X}_i\boldsymbol\beta_k\\
\boldsymbol\alpha_k &=& \frac{1}{\sqrt{1 - \boldsymbol\psi_k^T \boldsymbol\Omega^{-1}_k\boldsymbol\psi_k}} \boldsymbol\Omega^{-1}_k\boldsymbol\psi_k~~\text{and}\\
\boldsymbol\Omega_k &=& \boldsymbol\Sigma_k  + \boldsymbol\psi_k \boldsymbol\psi_k^T.
\end{eqnarray*}
Additional details can be found in Fr\:{u}wirth-Schnatter and Pyne (2010).

Of note, the MSN density can be expressed more compactly in terms of the matrix skew normal (MatSN) density (Chen and Gupta 2005). As we will see in Section 3.6, the matrix representation of the MSN distribution admits convenient conjugate prior distributions for the regression parameters and scale matrices, which in turn leads to efficient Gibbs sampling for posterior inference. Let $\mathbf{Y}_k$ be an ${n_k \times J}$ response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n {1}_{(z_i = k)}$ is the number of observations in cluster $k$. From equation (\ref{eq:msnreg}), it follows that $\mathbf{Y}_k$ is distributed as
\begin{eqnarray*}
\mathbf{Y}_{k} &\sim& MatSN_{n_k \times J}(\mathbf{M}_k,\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Omega_k)\\
%\text{vec}(\mathbf{Y}_k) =  (\mathbf{y}^T_1,...,\mathbf{y}^T_{n_k})^T\\
\text{vec}(\mathbf{M}_k) &=& (\boldsymbol\zeta_{k1}^T,...,\boldsymbol\zeta_{kn_k}^T)^T,
\end{eqnarray*}
where $\boldsymbol\zeta_{ki} = \mathbf{X}_i \boldsymbol\beta_k$ as in equation (\ref{eq:msnreg}), $\boldsymbol\alpha_k = (\alpha_{k1},...,\alpha_{kJ})^T$, $\mathbf{I}_{n_k}$ is the $n_k\times n_k$ identity matrix, and $\boldsymbol\Omega_k$ is the $J\times J$ scale matrix defined above in equation (\ref{eq:msndens}). From equation (\ref{eq:msnreg}), it follows that $\mathbf{Y}_{k}$, conditional on the $n_k\times 1$ vector of random effects $\mathbf{t}_{k}$, is jointly distributed in matrix form as
\[
\mathbf{Y}_k | \mathbf{t}_k \sim MatNorm_{n_k \times J}(\mathbf{M}_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k),
\]
where $MatNorm_{n_k \times J}(\cdot)$ denotes a $n_k\times J$ matrix normal density, $\text{vec}(\mathbf{M}_k) = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$ is an $n_kJ\times 1$ mean vector, $\mathbf{X}_k$ is an $n_kJ \times Jp$ design matrix, $\boldsymbol\beta_k$ is the $(Jp)\times 1$ vector of regression coefficients defined in equation (\ref{eq:msnreg}), and $\boldsymbol\Sigma_k$ is the $J\times J$ conditional covariance of $\boldsymbol\epsilon_{ik}$ given in equation (\ref{eq:msnreg}).

\subsection{Multinomial Regression for the Cluster Indicators}
To accommodate heterogeneity in the cluster-membership probabilities, we model $\pi_{ik}$ as a function of coovariates using a multinomial logit model 
\begin{eqnarray}
\pi_{ik} = \Pr(z_i = k|\mathbf{w}_i) = \frac{e^{\mathbf{w}_i^T \boldsymbol\delta_k}}{\sum_{h = 1}^K e^{\mathbf{w}_i^T \boldsymbol\delta_{h}}},~ k=1,\ldots,K,
\end{eqnarray}
where $\mathbf{w}_i$ is an $r\times 1$ vector of subject-level covariates, $\boldsymbol\delta_k$ is a $r\times 1$ vector of regression parameters associated with membership in cluster $k$. For identifiability purposes, we fix the reference category $k = K$ and set $\boldsymbol\delta_K = \mathbf{0}$. Under this model, $z_i|\mathbf{w}_i \sim Multinom(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{i1},...,\pi_{1K})$. During MCMC estimation, the cluster labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as cluster assignments.

\subsection{Conditional MSN Imputation}
To accommodate missing at random (MAR) responses, we propose a convenient imputation algorithm that can be implemented ``online'' as part of the Gibbs sampler. In Section 6, we discuss extension to allow for non-ignorable missingness. Suppose $\mathbf{y}_i$ has $q_i\in (1,\ldots,J)$ observed values, denoted $\mathbf{y}^{obs}_i$, and $J-q_i$ intermittent missing values, denoted $\mathbf{y}^{miss}_i$. We can use of the stochastic representation given in equation (\ref{eq:msnreg}) to impute $\mathbf{y}^{miss}_i$ from its conditional multivariate normal distribution given $(z_i,t_i,Y^{obs}_i)$:
\begin{eqnarray}
\mathbf{y}^{miss}_i|(z_i=k,t_i,\mathbf{y}^{obs}_i)&\sim& N_{J-q}(\boldsymbol\mu^{miss}_{ik},\boldsymbol\Sigma^{miss}_k),~\text{where}\nonumber\\
\boldsymbol\mu^{miss}_{ik}&=&\\
\boldsymbol\Sigma^{miss}_k&=& \nonumber
\end{eqnarray}
\textbf{Carter -- work on the above -- you will need to define notation and refer readers back to equation (3) as needed. We can discuss next week if needed}
These results follow from conventional multivariate normal theory. An attractive feature of this imputation algorithm is that it provides more opportunities to explore the parameter space than multiple imputation \textbf{[based on summary stats right?]} and avoids multiplicative run-time scaling in $m$, the number of imputations \textbf{Give refs}. In Section 4, we conduct simulation studies to demonstrate that imputing the missing MSN responses improves inferences over complete case analysis.

\subsection{Bayesian Inference}

\subsubsection{Prior Specification}
We adopt a fully Bayesian inferential approach and assign prior distributions to all model parameters. Conveniently, all parameters admit conditionally conjugate priors, which greatly improves posterior computation via a data-augmented Gibbs sampler. For \textbf{[Give priors for each parameter. Be clear about the conditionally joint prior for $\boldsymbol\beta_k$ and $\boldsymbol\Sigma_k$. Where appropriate, explain advantages]}

\subsubsection{Posterior Inference} 
The above prior specification induces closed-form full conditionals that can be efficiently updated as part of a Gibbs sampler outlined below. Additional details, including derivations can be found in the Web Appendix. \textbf{[Think about the best way to organize this section. Maybe see my Bayesian Analysis paper for guidance? We can discuss next week.]}

\paragraph{P\'olya--Gamma Data Augmentation for $z_i$.} The sampler begins by updating the latent cluster indicators $z_i$ $(i=1,\ldots,n)$ from its multinomial logit full conditional. To facilitate sampling, we adopt an efficient data-augmentation approach introduced by Polson \textit{et al.} (2013), which expresses the inverse-logit function as a mixture P\'olya--Gamma densities. \textbf{[See my Bayesian Analysis paper for guidance on this part]}.

\textbf{[I stopped here]}

$$p(\boldsymbol\delta_k|\mathbf{z},\boldsymbol\delta_{k' \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \pi_{ik}^{U_{ik}}(1-\pi_{ik})^{1-U_{ik}}$$

where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, $U_{ik} = {1}_{z_i = k}$ is an indicator that subject $i$ belongs to cluster $k$, and $\pi_{ik}$ is defined as in Section 3.4. We can rewrite $\pi_{ik}$ as follows
$$\pi_{ik} = P(U_{ik} = 1) = \frac{e^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}}}{1 + e^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}}} = \frac{e^{\eta_{ik}}}{1 + e^{\eta_{ik}}}$$

where ${c}_{ik} = \log \sum_{k' \ne k} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}}$ and $\eta_{ik} = \mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}$. We note that the sum $\sum_{k' \ne k} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}}$ includes the reference category, but since we fix $\boldsymbol\delta_K = \mathbf{0}$, we have $e^{\mathbf{w}_i^T \boldsymbol\delta_K} = 1$, and hence
$$c_{ik} = \log \sum_{k' \ne k} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}} = \log \left ( 1 + \sum_{k' \notin \{k,K \}} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}} \right )$$

We can use the quantities to re-express the full conditionals for $\boldsymbol\delta_k$ as
$$p(\boldsymbol\delta_k|\mathbf{Z},\boldsymbol\delta_{k' \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \left (\frac{e^{\eta_{ik}}}{1 + e^{\eta_{ik}}} \right )^{U_{ik}} \left (\frac{1}{1 + e^{\eta_{ik}}} \right )^{1-U_{ik}} = p(\boldsymbol\delta_k) \prod_{i = 1}^n \frac{(e^{\eta_{ik}})^{U_{ik}}}{1 + e^{\eta_{ik}}}$$

which we note is essentially a logistic regression likelihood. We thus apply this P\'olya--Gamma data augmentation scheme to update each $\boldsymbol\delta_k$ ($k = 1,...,K-1$) one at a time based on the binary indicators $U_{ik}$.

\begin{itemize}

\item Emphasize that PG data augmentation for the multinomial model results in a PG mixture of experts model, which is a computationally efficient way to model edge weights.

\end{itemize}
\end{document}

\subsubsection{MCMC Algorithm}

\begin{algorithm}
\caption{Gibbs Sampler}
\begin{algorithmic}
\small
\linespread{0.5}
    \State $\mathbf{Define} \ n_{iter};\  n_{burn}; \ K; \ \boldsymbol\theta_{init}; \ \boldsymbol\theta_0$
    \State $n_{sim} := n_{iter} - n_{burn}$
    \State $\boldsymbol\theta := \boldsymbol\theta_{init}$
    \For {$\iota = 1,...,n_{sim}$}
        \State \textsc{I. Conditional Imputation}
          \For {$ i = 1,...,n$}
            \State $\mathbf{Draw} \ \mathbf{y}_i^{miss} \ \text{from} \ N_q(\boldsymbol\mu_i^{miss}, \boldsymbol\Sigma_i^{miss})$
            \State $\mathbf{y}_i := \mathbf{y}_i^{miss} \cup \mathbf{y}_i^{obs}$
         \EndFor
        \State \textsc{II. MSN Regression}
          \For {$ k = 1,...,K$}
            \State $n_k := \sum_{i = 1}^n {1}_{z_i = k}$
            \For {$i_k = 1,...,n_k$}
              \State $\mathbf{Draw}\  t_i \ \text{from} \ N_{[0,\infty)}(a_i,A)$
            \EndFor
            \State $\mathbf{X^*}_k := $ \texttt{cbind}$(\mathbf{X}_k,\mathbf{t}_k)$
            \State $\mathbf{Draw} \ \mathbf{B^*}_k \ \text{from} \ \text{MatNorm}(\mathbf{B}_k,\mathbf{L}_k^{-1},\boldsymbol\Sigma_k)$
            \State $\mathbf{Draw} \ \boldsymbol\Sigma_k \ \text{from} \ \text{InvWish}(\nu_k, \mathbf{V}_k)$
          \EndFor
        \State \textsc{III. Multinomial Logit}
          \For {$i = 1,...,n$}
            \For {$k = 1,...,K$}
              \State $\pi_{ik} := P(z_i = k|\mathbf{w}_i,\boldsymbol\delta_k)$
              \State $p_{ik} := P(\mathbf{y}_i|\boldsymbol\beta_k^{*T} \mathbf{x}^*_i,\boldsymbol\Sigma_k)$
            \EndFor
            \State $\mathbf{p}_{z_i} := \frac{\mathbf{p}_i \circ \boldsymbol\pi_i}{\mathbf{p}_i \cdot \boldsymbol\pi_i}$
            \State $\mathbf{Draw} \ z_i \ \text{from} \ \text{Categorical}(\mathbf{p}_{z_i})$
            \For {$k = 1,...,K-1$}
              \State $\mathbf{Draw} \ \boldsymbol\delta_k \ \text{from} \ N(\mathbf{M},\mathbf{S})$
            \EndFor
          \EndFor
        \State $\boldsymbol\theta := \{\mathbf{B^*}, \boldsymbol\Sigma, \mathbf{Z}, \boldsymbol\delta \}$
        \State $\mathbf{Store} \ \boldsymbol\theta$
	  \EndFor
\end{algorithmic}
\end{algorithm}

\subsubsection{Assessment of MCMC Convergence}

\subsubsection{Label Switching}

\end{document}

### 2019-06-24 manuscript/manuscript.tex
section 3

--- 2019-06-18 manuscript/manuscript.tex
@@
 
-\subsection{Generic Multivariate Skew Normal Mixture}
+\subsection{Multivariate Skew Normal Mixture Model}
 
-Let $\mathbf{y}_{i}$ be the $J \times 1$ observation vector for subject $i$ such that $y_{ij}$ is the observation for subject $i$ at timepoint $j$.
-We assume for now that $\mathbf{y}_{i}$ is fully observed.
-Later, we relax this assumption by allowing for missingness in the components of $\mathbf{y}_{i}$.
-For the analysis of the Nurture data, we propose the following MSN mixture
+A primary goal of the Nurture study is to identify clusters of infants characterized by distinct motor development trajectories.
+To address this aim, we propose a flexible finite mixture model that accommodates relevant features of the data, such as skewness and dependence among the responses.
+To this end, let $\mathbf{y}_{i}=(y_{i1},\ldots,y_{iJ})^T$ be a $J \times 1$ vector of responses (i.e., Baley scores) for subject $i$ $(i=1,\ldots,n)$.
+For the analysis of the Nurture data, we propose a finite mixture model of the form
 \begin{equation}
@@
 \end{equation}
-where $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$ ($k = 1,...,K$).
-For now we assume that $K$ is fixed, but we discuss model selection strategies for choosing the optimal value of $K$ in Section 4.
-The probability that subject $i$ ($i = 1,...,n$) belongs to cluster $k$ is denoted $\pi_{ik}$.
-Note that $\pi_{ik}$ is indexed by $i$, as we allow the probability of class membership to vary across subjects.
-However, we assume a that class membership is fixed throughout the study period, since our focus is to cluster individuals based on their overall developmental patterns over the course of the study.
+where $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$ ($k = 1,...,K$) and $\pi_{ik}$ is a subject-specific mixing weight representing the probability that subject $i$ belongs to cluster $k$.
+For now we assume that $K$ is fixed; in Section 4, we discuss model selection strategies for choosing the optimal value of $K$.
+We also assume that class membership is fixed throughout the study period, since our focus is to cluster individuals based on their overall developmental patterns over the course of the study.
 In Section 6, we discuss extensions to allow for class membership to vary over time.
+\textbf{[We could omit these last two sentence -- are they really needed?
+Not sure.
+Maybe keep for now and think about it.]}
 
-In each cluster, we assume that the $J\times 1$ outcome vector for subject $i$ follows an MSN distribution.
-Let $z_i \in \{1,...,K\}$ denote a latent cluster membership of subject $i$.
-Conditional on $z_i = k$, we model $\mathbf{y}_{i}$ as
+To facilitate posterior inference, we introduce a latent cluster indicator variable $z_i$ taking the value $k \in \{1,...,K\}$ with probability $\pi_{ik}$.
+Conditional on $z_i = k$, we assume $\mathbf{y}_{i}$ is distributed as
 \begin{equation}
-\mathbf{y}_{i}|z_i=k \sim MSN_J(\boldsymbol\zeta_k,\boldsymbol\alpha_k,\boldsymbol\Omega_k),
+\mathbf{y}_{i}|(z_i=k) \sim MSN_J(\boldsymbol\zeta_{ki},\boldsymbol\alpha_k,\boldsymbol\Omega_k), \label{eq:msndens}
 \end{equation}
-where $\boldsymbol\zeta_k$ is a $J \times 1$ vector of cluster-specific location parameters, $\boldsymbol\alpha_k$ is a $J \times 1$ vector of cluster-specific skewness parameters, and $\boldsymbol\Omega_k$ is a $J \times J$ cluster-specific scale matrix that accounts for associations between the $J$ responses.
-The vector $\boldsymbol\alpha_k$ has components $\alpha_{jk}$, $j = 1,...,J$, that control the skewness of outcome $j$ in cluster $k$.
-When $\boldsymbol\alpha_k = \mathbf{0}$, the MSN distribution reduces to the multivariate normal distribution $N_J(\boldsymbol\zeta_k,\boldsymbol\Sigma_k)$, where $\boldsymbol\Sigma_J$ is the $J \times J$ covariance matrix capturing dependence among the $J$ outcomes.
+where $MSN_J(\cdot)$ denotes the $J$-dimensional multivariate skew normal density, $\boldsymbol\zeta_{ki}$ is a $J \times 1$ vector of subject- and cluster-specific location parameters, $\boldsymbol\alpha_k$ is a $J \times 1$ vector of cluster-specific skewness parameters, and $\boldsymbol\Omega_k$ is a $J \times J$ cluster-specific scale matrix that captures dependence among the $J$ responses.
+The vector $\boldsymbol\alpha_k$ has components $\alpha_{kj}$, $j = 1,...,J$ \textbf{[let's use $kij$ as our index hierarchy.
+Please review throughout]}, that control the skewness of outcome $j$ in cluster $k$.
+When $\boldsymbol\alpha_k = \mathbf{0}$, the MSN distribution reduces to the multivariate normal distribution $N_J(\boldsymbol\zeta_k,\boldsymbol\Omega_k)$, where $\boldsymbol\Omega_k$ is a $J \times J$ covariance matrix.
 
-\subsection{Multivariate Skew Normal Stochastic Representation}
+We can extend model (\ref{eq:msndens}) to the regression setting by modeling $\boldsymbol\zeta_{ki}$ as a function of covariates.
+Here we adopt a convenient stochastic representation of the MSN density (Azzalini and Dalla Valle, 1996):
+\begin{equation}
+\mathbf{y}_i|(z_i=k,t_i) = \mathbf{X}_i \boldsymbol\beta_k + t_i \boldsymbol\psi_k + \boldsymbol\epsilon_{ki}, \label{eq:msnreg}
+\end{equation}
+where 
+$\mathbf{X}_i$ is a $J \times Jp$ design matrix that includes potential time-varying covariates (e.g., indicators denoting quarterly visits); $\boldsymbol{\beta}_k=(\beta_{k11},\ldots,\beta_{k1p},\ldots,\beta_{kJ1},\ldots,\beta_{kJp})^T$ is a $Jp\times 1$ vector of cluster- and outcome-specific regression coefficients; $t_i\sim N_{[0,\infty)}(0,1)$ is a subject-specific standard normal random variable truncated below by zero; $\boldsymbol\psi_k=(\psi_{k1},\ldots,\psi_{kJ})^T$ is a $J \times 1$ vector of cluster-specific skewness parameters; and $\boldsymbol{\epsilon}_{ki} \sim N_J(\boldsymbol0,\boldsymbol\Sigma_k)$ is a $J\times 1$ vector of error terms.
+Thus, conditional on $t_i$ and $z_i=k$, $\boldsymbol{y}_i$ is distributed as $N_J(\mathbf{X}_i \boldsymbol\beta_k + t_i \boldsymbol\psi_k, \boldsymbol{\Sigma}_k)$.
+Marginally (after integrating over $t_i$), $\mathbf{y}_i$ is distributed $MSN_J(\boldsymbol\zeta_{ik}, \boldsymbol\alpha_k, \boldsymbol\Omega_k)$, where through back-transformation \textbf{[Carter: Please review these expressions carefully and make sure they conform to those in FS and Pyne]}
+\begin{eqnarray*}
+\boldsymbol\zeta_{ik} &=& \mathbf{X}_i\boldsymbol\beta_k\\
+\boldsymbol\alpha_k &=& \frac{1}{\sqrt{1 - \boldsymbol\psi_k^T \boldsymbol\Omega^{-1}_k\boldsymbol\psi_k}} \boldsymbol\Omega^{-1}_k\boldsymbol\psi_k~~\text{and}\\
+\boldsymbol\Omega_k &=& \boldsymbol\Sigma_k  + \boldsymbol\psi_k \boldsymbol\psi_k^T.
+\end{eqnarray*}
+Additional details can be found in Fr\:{u}wirth-Schnatter and Pyne (2010).
 
-We model the effect of covariates on longitudinal motor development outcomes through the use of a MSN regression model.
-We adopt a standard stochastic representation of a MSN random variable by conditioning on a subject-specific latent truncated normal random effect, denoted $t_i$ (Fr\"{u}hwirth-Schnatter and Pyne 2010).
-Conditional on membership in cluster $k$, the stochastic representation of the MSN takes the form
-\begin{equation}
-\mathbf{y}_{ik} = \mathbf{x}_{ik}^T\boldsymbol\beta_k  + t_{i}\boldsymbol\psi_k + \boldsymbol\epsilon_{ik}
-\end{equation}
+Of note, the MSN density can be expressed more compactly in terms of the matrix skew normal (MatSN) density (Chen and Gupta 2005).
+As we will see in Section 3.6, the matrix representation of the MSN distribution admits convenient conjugate prior distributions for the regression parameters and scale matrices, which in turn leads to efficient Gibbs sampling for posterior inference.
+Let $\mathbf{Y}_k$ be an ${n_k \times J}$ response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n {1}_{(z_i = k)}$ is the number of observations in cluster $k$.
+From equation (\ref{eq:msnreg}), it follows that $\mathbf{Y}_k$ is distributed as
+\begin{eqnarray*}
+\mathbf{Y}_{k} &\sim& MatSN_{n_k \times J}(\mathbf{M}_k,\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Omega_k)\\
+%\text{vec}(\mathbf{Y}_k) =  (\mathbf{y}^T_1,...,\mathbf{y}^T_{n_k})^T\\
+\text{vec}(\mathbf{M}_k) &=& (\boldsymbol\zeta_{k1}^T,...,\boldsymbol\zeta_{kn_k}^T)^T,
+\end{eqnarray*}
+where $\boldsymbol\zeta_{ki} = \mathbf{X}_i \boldsymbol\beta_k$ as in equation (\ref{eq:msnreg}), $\boldsymbol\alpha_k = (\alpha_{k1},...,\alpha_{kJ})^T$, $\mathbf{I}_{n_k}$ is the $n_k\times n_k$ identity matrix, and $\boldsymbol\Omega_k$ is the $J\times J$ scale matrix defined above in equation (\ref{eq:msndens}).
+From equation (\ref{eq:msnreg}), it follows that $\mathbf{Y}_{k}$, conditional on the $n_k\times 1$ vector of random effects $\mathbf{t}_{k}$, is jointly distributed in matrix form as
+\[
+\mathbf{Y}_k | \mathbf{t}_k \sim MatNorm_{n_k \times J}(\mathbf{M}_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k),
+\]
+where $MatNorm_{n_k \times J}(\cdot)$ denotes a $n_k\times J$ matrix normal density, $\text{vec}(\mathbf{M}_k) = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$ is an $n_kJ\times 1$ mean vector, $\mathbf{X}_k$ is an $n_kJ \times Jp$ design matrix, $\boldsymbol\beta_k$ is the $(Jp)\times 1$ vector of regression coefficients defined in equation (\ref{eq:msnreg}), and $\boldsymbol\Sigma_k$ is the $J\times J$ conditional covariance of $\boldsymbol\epsilon_{ik}$ given in equation (\ref{eq:msnreg}).
 
-where $\mathbf{x}_i$ is the $P \times 1$ vector of covariate values for subject $i$, $\boldsymbol\beta$ is the $P \times J$ vector of fixed effects coefficients , $t_i \stackrel{iid}{\sim} N_{[0,\infty)}(0,1)$ is a conditional truncated normal random effect, $\boldsymbol\psi$ is the $J \times 1$ vector containing cluster-specific skewness parameters, and $\boldsymbol\epsilon_i \sim N_J(0,\boldsymbol\Sigma)$ is the correlated error term.
-In can be shown that under these conditions $\mathbf{y}_i$ is multivariate skew normal with location.
-For conciseness, we present the MSN distribution here outside of the finite mixture context.
-When using this stochastic representation in our finite mixture model, the parameters $\boldsymbol\beta$, $\boldsymbol\psi$, and $\boldsymbol\Sigma$ are replaced with cluster-specific counterparts $\boldsymbol\beta_k$, $\boldsymbol\psi_k$, and $\boldsymbol\Sigma_k$.
-
-\subsection{Representation as Matrix Skew Normal}
-
-We note that the multivariate skew normal regression problem can be expressed more succinctly through the use of Matrix Skew Normal distributions.
-In particular, we define $\mathbf{Y}_{n_k \times J}$ as the response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n \mathds{1}_{(z_i = k)}$ is the total number of observations belonging to cluster $k$ ($k = 1,...,K$).
-$$\mathbf{Y}_{n_k \times J} \sim MatSN_{n_k \times J}(\mathbf{M},\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Sigma_k)$$
-$$\text{vec}(\mathbf{Y}) = \mathbf{y}_{n_kJ \times 1} = (\mathbf{y}^T_1,...,\mathbf{y}^T_{n_k})$$
-$$\text{vec}(\mathbf{M}) = \boldsymbol\mu_{n_k J \times 1} = (\boldsymbol\mu_1^T,...,\boldsymbol\mu_{n_k}^T)$$
-
-where $\boldsymbol\mu_i = \mathbf{x}_i \boldsymbol\beta_k$ and $\boldsymbol\alpha_k = (\alpha_{k1},...,\alpha_{kJ})$.
-This implies that the $i^{th}$ row of $\mathbf{Y}$ is $\mathbf{y}_i \sim MSN_J(\boldsymbol\mu_i, \boldsymbol\alpha_k, \boldsymbol\Sigma_k) = {MSN}_J(\mathbf{x}_i \boldsymbol\beta_k, \boldsymbol\alpha_k, \boldsymbol\Sigma_k)$.
-Using the stochastic representation presented in Section 3.2, the conditional distribution of $\mathbf{y}_i | z_i, t_i$ is $N_J (\mathbf{x}_i \boldsymbol\beta_k + \boldsymbol\psi_k t_i,\boldsymbol\Sigma_k)$.
-This implies that all responses for cluster $k$ conditioned on $\mathbf{T}_{n_k \times 1}$ is distributed as
-$$\mathbf{Y}_{n_k \times J} | \mathbf{T}_k = \mathbf{t}_k \sim MatNorm_{n_k \times J}(\mathbf{M}_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k)$$
-$$\text{vec}(\mathbf{M}_k)_{n_kJ \times 1} = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$$\
-
-where $\mathbf{t}_k = (t_1,...,t_{n_k})$.
-
-\subsection{Multinomial Regression on Cluster Probabilities}
-
-A primary concern of our model is with identification of latent infant development classs.
-We accomplish this via multinomial logit regression model on cluster membership, which utilizes P\'{o}lya-Gamma data-augmentation, as described in Section 3.6 to allow for updating of all parameters using Gibbs sampling.
-The multinomial logit model is as follows for $k = 1,...,K$.
-$$P(z_i = k|\mathbf{w}_i) = \pi_{ik} = \frac{e^{\mathbf{w}_i^T \boldsymbol\delta_k}}{\sum_{k' = 1}^K e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}}}$$
-
-where $\mathbf{w}_i$ is the vector of cluster probability covariates for subject $i$, $\boldsymbol\delta_k$ contains the multinomial regression parameters for cluster $k$, and $K$ is the number of putative classs specified \textit{a priori}.
+\subsection{Multinomial Regression for the Cluster Indicators}
+To accommodate heterogeneity in the cluster-membership probabilities, we model $\pi_{ik}$ as a function of coovariates using a multinomial logit model 
+\begin{eqnarray}
+\pi_{ik} = \Pr(z_i = k|\mathbf{w}_i) = \frac{e^{\mathbf{w}_i^T \boldsymbol\delta_k}}{\sum_{h = 1}^K e^{\mathbf{w}_i^T \boldsymbol\delta_{h}}},~ k=1,\ldots,K,
+\end{eqnarray}
+where $\mathbf{w}_i$ is an $r\times 1$ vector of subject-level covariates, $\boldsymbol\delta_k$ is a $r\times 1$ vector of regression parameters associated with membership in cluster $k$.
 For identifiability purposes, we fix the reference category $k = K$ and set $\boldsymbol\delta_K = \mathbf{0}$.
@@
 
-\subsection{Conditional MVN Imputation}
-
-We allow for missingness of outcomes in the MSN mixture model by imputing missing values from their conditional multivariate normal distributions. We note that
-$$\mathbf{y}_i|z_i,\mathbf{x}_i,t_i,\boldsymbol\beta_{k},\boldsymbol\psi_{k},\boldsymbol\Sigma_{k} \sim N_J(\boldsymbol\beta_{k}^T\mathbf{x}_i  + t_i \boldsymbol\psi_{k}, \boldsymbol\Sigma_{k})$$
-
-where $z_i = k$.
-This allows us to appeal to standard conditional forms of the multivariate normal distribution to specify the distribution of missing observations.
-For subject $i$, we use $q_i$ to denote the number of the total $J$ possible repeated measurements are missing.
-We partition the full $J \times 1$ outcome vector $\mathbf{y}_i$ into the $(J - q_i) \times 1$ observed data vector $\mathbf{y}_i^{obs}$ and the $q_i \times 1$ missing data vector $\mathbf{y}_i^{miss}$.
-$Y_i = [Y^{miss}_{i_{q \times 1}} | Y^{obs}_{i_{J - q \times 1}}]^T$.
-We have
-$$\mathbf{y}_i^{miss}|\mathbf{y}_i^{obs},\mathbf{x}_i,t_i,z_i,\boldsymbol\beta_{k},\boldsymbol\psi_{k},\boldsymbol\Sigma_{k} \sim N(\boldsymbol\mu^{miss},\boldsymbol\Sigma^{miss})$$
-
-where $\boldsymbol\mu^{miss}$ and $\boldsymbol\Sigma^{miss}$ take standard forms detailed in the Appendix.
-
-An attractive feature of this imputation approach is that each missing outcome is imputed ``online", i.e.
-once per MCMC iteration.
-This provides more opportunities to explore the parameter space than multiple imputation and avoids multiplicative run-time scaling in $m$, the number of imputations.
-We demonstrate this feature using simulations detailed in Section 4.
+\subsection{Conditional MSN Imputation}
+To accommodate missing at random (MAR) responses, we propose a convenient imputation algorithm that can be implemented ``online'' as part of the Gibbs sampler.
+In Section 6, we discuss extension to allow for non-ignorable missingness.
+Suppose $\mathbf{y}_i$ has $q_i\in (1,\ldots,J)$ observed values, denoted $\mathbf{y}^{obs}_i$, and $J-q_i$ intermittent missing values, denoted $\mathbf{y}^{miss}_i$.
+We can use of the stochastic representation given in equation (\ref{eq:msnreg}) to impute $\mathbf{y}^{miss}_i$ from its conditional multivariate normal distribution given $(z_i,t_i,Y^{obs}_i)$:
+\begin{eqnarray}
+\mathbf{y}^{miss}_i|(z_i=k,t_i,\mathbf{y}^{obs}_i)&\sim& N_{J-q}(\boldsymbol\mu^{miss}_{ik},\boldsymbol\Sigma^{miss}_k),~\text{where}\nonumber\\
+\boldsymbol\mu^{miss}_{ik}&=&\\
+\boldsymbol\Sigma^{miss}_k&=& \nonumber
+\end{eqnarray}
+\textbf{Carter -- work on the above -- you will need to define notation and refer readers back to equation (3) as needed. We can discuss next week if needed}
+These results follow from conventional multivariate normal theory.
+An attractive feature of this imputation algorithm is that it provides more opportunities to explore the parameter space than multiple imputation \textbf{[based on summary stats right?]} and avoids multiplicative run-time scaling in $m$, the number of imputations \textbf{Give refs}.
+In Section 4, we conduct simulation studies to demonstrate that imputing the missing MSN responses improves inferences over complete case analysis.
 
@@
 
-\subsubsection{P\'olya--Gamma Data Augmentation}
+\subsubsection{Prior Specification}
+We adopt a fully Bayesian inferential approach and assign prior distributions to all model parameters.
+Conveniently, all parameters admit conditionally conjugate priors, which greatly improves posterior computation via a data-augmented Gibbs sampler.
+For \textbf{[Give priors for each parameter.
+Be clear about the conditionally joint prior for $\boldsymbol\beta_k$ and $\boldsymbol\Sigma_k$.
+Where appropriate, explain advantages]}
 
-Polson \textit{et al.} (2013) introduced an efficient data augmentation approach to fitting several GLMs including the multinomial logit regression model specified by Holmes and Held (2006), which specifies the full conditional distributions of the multinomial regression parameters as a function of a Bernoulli likelihood.
+\subsubsection{Posterior Inference} 
+The above prior specification induces closed-form full conditionals that can be efficiently updated as part of a Gibbs sampler outlined below.
+Additional details, including derivations can be found in the Web Appendix.
+\textbf{[Think about the best way to organize this section.
+Maybe see my Bayesian Analysis paper for guidance?
+We can discuss next week.]}
+
+\paragraph{P\'olya--Gamma Data Augmentation for $z_i$.} The sampler begins by updating the latent cluster indicators $z_i$ $(i=1,\ldots,n)$ from its multinomial logit full conditional.
+To facilitate sampling, we adopt an efficient data-augmentation approach introduced by Polson \textit{et al.} (2013), which expresses the inverse-logit function as a mixture P\'olya--Gamma densities.
+\textbf{[See my Bayesian Analysis paper for guidance on this part]}.
+
+\textbf{[I stopped here]}
+
 $$p(\boldsymbol\delta_k|\mathbf{z},\boldsymbol\delta_{k' \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \pi_{ik}^{U_{ik}}(1-\pi_{ik})^{1-U_{ik}}$$
 
-where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, $U_{ik} = \mathds{1}_{z_i = k}$ is an indicator that subject $i$ belongs to cluster $k$, and $\pi_{ik}$ is defined as in Section 3.4.
+where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, $U_{ik} = {1}_{z_i = k}$ is an indicator that subject $i$ belongs to cluster $k$, and $\pi_{ik}$ is defined as in Section 3.4.
 We can rewrite $\pi_{ik}$ as follows
@@
 \end{itemize}
-
-\subsubsection{Prior Choice}
+\end{document}
 
@@
 \toprule
-\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Cluster 1} & \multicolumn{2}{c}{Cluster 2} & \multicolumn{2}{c}{Cluster 3} \\
+\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Class 1} & \multicolumn{2}{c}{Class 2} & \multicolumn{2}{c}{Class 3} \\
 \cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6} \cmidrule(l{3pt}r{3pt}){7-8}
@@
 \multicolumn{8}{l}{\textbf{ }}\\
-\hspace{1em}MVSN & $\beta_{0}$ & -1.35 & -1.59 (-2.3, -0.94) & -0.18 & -0.24 (-0.53, 0.38) & 0.72 & 0.79 (0.52, 1.02)\\
-\hspace{1em}Regression & $\beta_{1}$ & -1.19 & -1.35 (-1.99, -0.84) & -0.09 & -0.33 (-0.62, 0.78) & 1.64 & 1.65 (1.33, 1.91)\\
-\hspace{1em} & $\beta_{2}$ & -1.65 & -1.81 (-2.66, -1.41) & -0.47 & -0.62 (-0.89, 0.3) & 1.44 & 1.32 (0.97, 1.6)\\
-\hspace{1em} & $\beta_{3}$ & -1.75 & -1.89 (-2.52, -1.37) & -0.22 & -0.32 (-0.63, 0.55) & 2.28 & 2.26 (1.96, 2.5)\\
+\hspace{1em}MVSN & $\beta_{0}$ & -3.21 & -3.34 (-3.8, -2.99) & -0.32 & -0.33 (-0.48, -0.14) & 3.35 & 3.33 (3.22, 3.44)\\
+\hspace{1em}Regression & $\beta_{1}$ & -3.08 & -3.3 (-3.73, -2.87) & -0.75 & -0.72 (-0.87, -0.52) & 2.6 & 2.5 (2.39, 2.6)\\
+\hspace{1em} & $\beta_{2}$ & -2.97 & -3.18 (-3.58, -2.76) & -0.45 & -0.44 (-0.58, -0.26) & 3.43 & 3.42 (3.31, 3.53)\\
+\hspace{1em} & $\beta_{3}$ & -2.91 & -3.08 (-3.49, -2.68) & -0.66 & -0.68 (-0.83, -0.48) & 3.04 & 2.98 (2.87, 3.09)\\
 \addlinespace[0.3em]
 \multicolumn{8}{l}{\textbf{ }}\\
-\hspace{1em} & $\sigma_{11}$ & 1 & 1.02 (0.78, 1.19) & 1 & 0.98 (0.77, 1.23) & 1 & 1.06 (0.85, 1.29)\\
-\hspace{1em} & $\sigma_{12}$ & -0.32 & -0.19 (-0.33, -0.02) & 0.16 & 0.14 (-0.01, 0.4) & 0.72 & 0.82 (0.62, 1.05)\\
-\hspace{1em} & $\sigma_{13}$ & -0.65 & -0.55 (-0.68, -0.35) & 0.72 & 0.7 (0.51, 0.94) & 0.14 & 0.27 (0.1, 0.48)\\
-\hspace{1em} & $\sigma_{14}$ & -0.44 & -0.33 (-0.46, -0.13) & 0.5 & 0.48 (0.31, 0.72) & -0.01 & -0.02 (-0.16, 0.16)\\
-\hspace{1em} & $\sigma_{22}$ & 1 & 0.92 (0.72, 1.06) & 1 & 0.94 (0.72, 1.22) & 1 & 1.11 (0.87, 1.38)\\
-\hspace{1em} & $\sigma_{23}$ & 0.56 & 0.49 (0.33, 0.6) & 0.53 & 0.46 (0.29, 0.73) & -0.1 & 0.08 (-0.1, 0.28)\\
-\hspace{1em} & $\sigma_{24}$ & 0.98 & 0.9 (0.7, 1.04) & 0.24 & 0.14 (-0.03, 0.41) & 0.19 & 0.17 (0.01, 0.37)\\
-\hspace{1em} & $\sigma_{33}$ & 1 & 0.9 (0.66, 1.04) & 1 & 0.92 (0.72, 1.19) & 1 & 1.26 (1.03, 1.52)\\
-\hspace{1em} & $\sigma_{34}$ & 0.56 & 0.51 (0.35, 0.62) & 0.86 & 0.79 (0.58, 1.05) & -0.65 & -0.63 (-0.78, -0.45)\\
-\hspace{1em} & $\sigma_{44}$ & 1 & 0.93 (0.72, 1.07) & 1 & 0.93 (0.7, 1.2) & 1 & 1.11 (0.87, 1.36)\\
+\hspace{1em} & $\sigma_{11}$ & 1 & 0.95 (0.84, 1.02) & 1 & 1 (0.89, 1.11) & 1 & 1.06 (0.97, 1.16)\\
+\hspace{1em} & $\sigma_{12}$ & 0.74 & 0.7 (0.59, 0.76) & 0.68 & 0.68 (0.59, 0.78) & -0.45 & -0.41 (-0.47, -0.36)\\
+\hspace{1em} & $\sigma_{13}$ & 0.74 & 0.69 (0.58, 0.75) & -0.16 & -0.13 (-0.2, -0.06) & 0.82 & 0.88 (0.79, 0.97)\\
+\hspace{1em} & $\sigma_{14}$ & 0.98 & 0.93 (0.81, 0.99) & 0.64 & 0.65 (0.56, 0.75) & 0.7 & 0.75 (0.67, 0.83)\\
+\hspace{1em} & $\sigma_{22}$ & 1 & 0.94 (0.82, 1.01) & 1 & 1.03 (0.93, 1.13) & 1 & 1.07 (0.99, 1.16)\\
+\hspace{1em} & $\sigma_{23}$ & 0.83 & 0.79 (0.67, 0.85) & -0.43 & -0.4 (-0.46, -0.34) & -0.66 & -0.62 (-0.68, -0.57)\\
+\hspace{1em} & $\sigma_{24}$ & 0.81 & 0.77 (0.66, 0.83) & 0.63 & 0.67 (0.58, 0.77) & 0.01 & 0.07 (0.01, 0.13)\\
+\hspace{1em} & $\sigma_{33}$ & 1 & 0.96 (0.84, 1.03) & 1 & 1 (0.91, 1.09) & 1 & 1.05 (0.96, 1.15)\\
+\hspace{1em} & $\sigma_{34}$ & 0.85 & 0.81 (0.69, 0.87) & 0.15 & 0.15 (0.08, 0.23) & 0.59 & 0.64 (0.56, 0.72)\\
+\hspace{1em} & $\sigma_{44}$ & 1 & 0.95 (0.83, 1.01) & 1 & 1.02 (0.92, 1.13) & 1 & 1.06 (0.97, 1.15)\\
 \addlinespace[0.3em]
 \multicolumn{8}{l}{\textbf{ }}\\
-\hspace{1em} & $\psi_{1}$ & -0.33 & -0.02 (-0.84, 0.87) & 0.67 & 0.69 (-0.09, 1.01) & -1 & -0.98 (-1.25, -0.67)\\
-\hspace{1em} & $\psi_{2}$ & -0.33 & -0.16 (-0.8, 0.62) & 0.67 & 0.81 (-0.6, 1.14) & -1 & -0.98 (-1.28, -0.6)\\
-\hspace{1em} & $\psi_{3}$ & -0.33 & -0.15 (-0.65, 0.89) & 0.67 & 0.72 (-0.35, 1.02) & -1 & -0.8 (-1.15, -0.36)\\
-\hspace{1em} & $\psi_{4}$ & -0.33 & -0.18 (-0.84, 0.59) & 0.67 & 0.71 (-0.32, 1.06) & -1 & -1 (-1.32, -0.62)\\
+\hspace{1em} & $\psi_{1}$ & -0.33 & -0.33 (-0.62, 0.69) & 0.67 & 0.7 (0.46, 0.89) & -1 & -1.01 (-1.13, -0.87)\\
+\hspace{1em} & $\psi_{2}$ & -0.33 & -0.32 (-0.61, 0.64) & 0.67 & 0.63 (0.38, 0.82) & -1 & -0.88 (-1.01, -0.75)\\
+\hspace{1em} & $\psi_{3}$ & -0.33 & -0.33 (-0.61, 0.69) & 0.67 & 0.64 (0.43, 0.82) & -1 & -1.01 (-1.14, -0.88)\\
+\hspace{1em} & $\psi_{4}$ & -0.33 & -0.31 (-0.63, 0.67) & 0.67 & 0.7 (0.45, 0.89) & -1 & -0.94 (-1.07, -0.81)\\
 \addlinespace[0.3em]
 \multicolumn{8}{l}{\textbf{ }}\\
-\hspace{1em}Multinom. & $\delta_{11}$ & -0.84 & -0.78 (-0.96, -0.59) & -0.84 & -0.78 (-0.96, -0.59) & -0.84 & -0.78 (-0.96, -0.59)\\
-\hspace{1em} & $\delta_{12}$ & -0.24 & -0.26 (-0.42, -0.1) & -0.24 & -0.26 (-0.42, -0.1) & -0.24 & -0.26 (-0.42, -0.1)\\
+\hspace{1em}Multinom. & $\delta_{11}$ & 0.9 & 0.88 (0.81, 0.95) & 0.9 & 0.88 (0.81, 0.95) & 0.9 & 0.88 (0.81, 0.95)\\
+\hspace{1em} & $\delta_{12}$ & 0.23 & 0.22 (0.14, 0.3) & 0.23 & 0.22 (0.14, 0.3) & 0.23 & 0.22 (0.14, 0.3)\\
 \addlinespace[0.3em]
 \multicolumn{8}{l}{\textbf{ }}\\
-\hspace{1em}Clustering & $\pi_l$ & 0.39 & 0.39 (0.38, 0.4) & 0.26 & 0.26 (0.25, 0.27) & 0.34 & 0.35 (0.33, 0.36)\\
+\hspace{1em}Clustering & $\pi_l$ & 0.28 & 0.28 (0.27, 0.28) & 0.42 & 0.43 (0.42, 0.43) & 0.3 & 0.3 (0.3, 0.3)\\
 \bottomrule

### 2019-06-24 manuscript/manuscript.tex
section 3

@@
 \begin{equation}
-\mathbf{y}_i|(z_i=k,t_i) = \mathbf{X}_i \boldsymbol\beta_k + t_i \boldsymbol\psi_k + \boldsymbol\epsilon_{ki}, \label{eq:msnreg}
+\mathbf{y}_{i}|(z_i=k,t_i) = \mathbf{X}_i \boldsymbol\beta_k + t_i \boldsymbol\psi_k + \boldsymbol\epsilon_{ki}, \label{eq:msnreg}
 \end{equation}
@@
 Thus, conditional on $t_i$ and $z_i=k$, $\boldsymbol{y}_i$ is distributed as $N_J(\mathbf{X}_i \boldsymbol\beta_k + t_i \boldsymbol\psi_k, \boldsymbol{\Sigma}_k)$.
-Marginally (after integrating over $t_i$), $\mathbf{y}_i$ is distributed $MSN_J(\boldsymbol\zeta_{ik}, \boldsymbol\alpha_k, \boldsymbol\Omega_k)$, where through back-transformation \textbf{[Carter: Please review these expressions carefully and make sure they conform to those in FS and Pyne]}
+Marginally (after integrating over $t_i$), $\mathbf{y}_i$ is distributed $MSN_J(\boldsymbol\zeta_{ki}, \boldsymbol\alpha_k, \boldsymbol\Omega_k)$, where through back-transformation
 \begin{eqnarray*}
-\boldsymbol\zeta_{ik} &=& \mathbf{X}_i\boldsymbol\beta_k\\
-\boldsymbol\alpha_k &=& \frac{1}{\sqrt{1 - \boldsymbol\psi_k^T \boldsymbol\Omega^{-1}_k\boldsymbol\psi_k}} \boldsymbol\Omega^{-1}_k\boldsymbol\psi_k~~\text{and}\\
-\boldsymbol\Omega_k &=& \boldsymbol\Sigma_k  + \boldsymbol\psi_k \boldsymbol\psi_k^T.
+\boldsymbol\zeta_{ki} &=& \mathbf{X}_i\boldsymbol\beta_k\\
+\boldsymbol\alpha_k &=& \frac{1}{\sqrt{1 - \boldsymbol\psi_k^T 
+\boldsymbol\Omega^{-1}_k\boldsymbol\psi_k}} \boldsymbol\omega_k \boldsymbol\Omega^{-1}_k\boldsymbol\psi_k~~\text{and}\\
+\boldsymbol\Omega_k &=& \boldsymbol\Sigma_k  + \boldsymbol\psi_k \boldsymbol\psi_k^T,
 \end{eqnarray*}
-Additional details can be found in Fr\:{u}wirth-Schnatter and Pyne (2010).
+where $\boldsymbol\omega_k = \text{Diag}(\sqrt{\omega_{k,11}},...,\sqrt{\omega_{k,JJ}})$ is the $J \times J$ diagonal matrix containing the square root of the diagonal entries of $\boldsymbol\Omega_k$.
+Additional details can be found in Fr\"{u}wirth-Schnatter and Pyne (2010).
 

### 2019-06-24 manuscript/manuscript.tex
section 3

@@
 To accommodate missing at random (MAR) responses, we propose a convenient imputation algorithm that can be implemented ``online'' as part of the Gibbs sampler.
-In Section 6, we discuss extension to allow for non-ignorable missingness.
+In Section 6, we discuss extensions to allow for non-ignorable missingness.
 Suppose $\mathbf{y}_i$ has $q_i\in (1,\ldots,J)$ observed values, denoted $\mathbf{y}^{obs}_i$, and $J-q_i$ intermittent missing values, denoted $\mathbf{y}^{miss}_i$.
-We can use of the stochastic representation given in equation (\ref{eq:msnreg}) to impute $\mathbf{y}^{miss}_i$ from its conditional multivariate normal distribution given $(z_i,t_i,Y^{obs}_i)$:
+We can make use of the stochastic representation given in equation (\ref{eq:msnreg}) to impute $\mathbf{y}^{miss}_i$ from its conditional multivariate normal distribution given $(z_i,t_i,Y^{obs}_i)$:
 \begin{eqnarray}
-\mathbf{y}^{miss}_i|(z_i=k,t_i,\mathbf{y}^{obs}_i)&\sim& N_{J-q}(\boldsymbol\mu^{miss}_{ik},\boldsymbol\Sigma^{miss}_k),~\text{where}\nonumber\\
-\boldsymbol\mu^{miss}_{ik}&=&\\
+\mathbf{y}^{miss}_i|(z_i=k,t_i,\mathbf{y}^{obs}_i)&\sim& N_{J-q}(\boldsymbol\mu^{miss}_{ki},\boldsymbol\Sigma^{miss}_k),~\text{where}\nonumber\\
+\boldsymbol\mu^{miss}_{ki}&=&\\
 \boldsymbol\Sigma^{miss}_k&=& \nonumber
@@
 \end{itemize}
-\end{document}
 

### 2019-06-24 manuscript/manuscript.tex
section 3

@@
 \begin{eqnarray*}
-\boldsymbol\zeta_{ki} &=& \mathbf{X}_i\boldsymbol\beta_k\\
+\boldsymbol\zeta_{ki} &=& \mathbf{X}_i\boldsymbol\beta_k,\\
 \boldsymbol\alpha_k &=& \frac{1}{\sqrt{1 - \boldsymbol\psi_k^T 
-\boldsymbol\Omega^{-1}_k\boldsymbol\psi_k}} \boldsymbol\omega_k \boldsymbol\Omega^{-1}_k\boldsymbol\psi_k~~\text{and}\\
+\boldsymbol\Omega^{-1}_k\boldsymbol\psi_k}} \boldsymbol\omega_k \boldsymbol\Omega^{-1}_k\boldsymbol\psi_k,~~\text{and}\\
 \boldsymbol\Omega_k &=& \boldsymbol\Sigma_k  + \boldsymbol\psi_k \boldsymbol\psi_k^T,
@@
 \begin{eqnarray}
-\mathbf{y}^{miss}_i|(z_i=k,t_i,\mathbf{y}^{obs}_i)&\sim& N_{J-q}(\boldsymbol\mu^{miss}_{ki},\boldsymbol\Sigma^{miss}_k),~\text{where}\nonumber\\
-\boldsymbol\mu^{miss}_{ki}&=&\\
-\boldsymbol\Sigma^{miss}_k&=& \nonumber
+\mathbf{y}^{miss}_i|(z_i=k,t_i,\mathbf{y}^{obs}_i)&\sim& N_{J-q_i}(\boldsymbol\mu^{cond}_{ki},\boldsymbol\Sigma^{cond}_k),~\text{where}\nonumber\\
+\boldsymbol\mu^{cond}_{ki}&=& \boldsymbol\mu^{miss} + \boldsymbol\Sigma_{12} \boldsymbol\Sigma_{22}^{-1}(\mathbf{y}^{obs}_i - \boldsymbol\mu^{obs})\\
+\boldsymbol\Sigma^{cond}_k&=& \boldsymbol\Sigma_{11} - \boldsymbol\Sigma_{12}\boldsymbol\Sigma_{22}^{-1}\boldsymbol\Sigma_{21},~\text{where}\nonumber
 \end{eqnarray}
-\textbf{Carter -- work on the above -- you will need to define notation and refer readers back to equation (3) as needed. We can discuss next week if needed}
+$\boldsymbol\Sigma_k$ is partitioned into four sub-matrices $\boldsymbol\Sigma_{11},\boldsymbol\Sigma_{12},\boldsymbol\Sigma_{21}, \ \text{and} \ \boldsymbol\Sigma_{22}$ such that $\boldsymbol\Sigma_{11}$ is a $J-q_i \times J-q_i$ matrix containing the rows and columns of $\boldsymbol\Sigma_k$ corresponding to inicies of $\mathbf{y}_{i}$ where missingness occurs.
+Similarly, $\boldsymbol\Sigma_{12}$ is a $J-q_i \times q_i$ matrix containing the rows of $\boldsymbol\Sigma_k$ that correspond to missing indices of $\mathbf{y}_i$, but columns of $\boldsymbol\Sigma_k$ that correspond to observed indices of $\mathbf{y}_i$.
+The remaining partitions $\boldsymbol\Sigma_{21}, \ \text{and} \ \boldsymbol\Sigma_{22}$ are defined in the same manner.
 These results follow from conventional multivariate normal theory.
-An attractive feature of this imputation algorithm is that it provides more opportunities to explore the parameter space than multiple imputation \textbf{[based on summary stats right?]} and avoids multiplicative run-time scaling in $m$, the number of imputations \textbf{Give refs}.
+An attractive feature of this imputation algorithm is that it avoids multiplicative run-time scaling in $m$, the number of imputations (Gelman \textit{et al.} 2013; Zhou and Reiter, 2010).
+Our approach also provides more opportunities to explore the missing data parameter space than does multiple imputation, since missing values are drawn at each MCMC iteration, and often in practice $n_{sim} >> m$, where $n_{sim}$ is the number of MCMC iterations.
 In Section 4, we conduct simulation studies to demonstrate that imputing the missing MSN responses improves inferences over complete case analysis.
@@
 \textbf{[See my Bayesian Analysis paper for guidance on this part]}.
-
-\textbf{[I stopped here]}
-
-$$p(\boldsymbol\delta_k|\mathbf{z},\boldsymbol\delta_{k' \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \pi_{ik}^{U_{ik}}(1-\pi_{ik})^{1-U_{ik}}$$
-
+$$p(\boldsymbol\delta_k|\mathbf{z},\boldsymbol\delta_{h \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \pi_{ik}^{U_{ik}}(1-\pi_{ik})^{1-U_{ik}}$$
 where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, $U_{ik} = {1}_{z_i = k}$ is an indicator that subject $i$ belongs to cluster $k$, and $\pi_{ik}$ is defined as in Section 3.4.
@@
 $$\pi_{ik} = P(U_{ik} = 1) = \frac{e^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}}}{1 + e^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}}} = \frac{e^{\eta_{ik}}}{1 + e^{\eta_{ik}}}$$
-
-where ${c}_{ik} = \log \sum_{k' \ne k} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}}$ and $\eta_{ik} = \mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}$.
-We note that the sum $\sum_{k' \ne k} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}}$ includes the reference category, but since we fix $\boldsymbol\delta_K = \mathbf{0}$, we have $e^{\mathbf{w}_i^T \boldsymbol\delta_K} = 1$, and hence
-$$c_{ik} = \log \sum_{k' \ne k} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}} = \log \left ( 1 + \sum_{k' \notin \{k,K \}} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}} \right )$$
+where ${c}_{ik} = \log \sum_{h \ne k} e^{\mathbf{w}_i^T \boldsymbol\delta_{h}}$ and $\eta_{ik} = \mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}$.
+We note that the sum $\sum_{h \ne k} e^{\mathbf{w}_i^T \boldsymbol\delta_{h}}$ includes the reference category, but since we fix $\boldsymbol\delta_K = \mathbf{0}$, we have $e^{\mathbf{w}_i^T \boldsymbol\delta_K} = 1$, and hence
+$$c_{ik} = \log \sum_{h \ne k} e^{\mathbf{w}_i^T \boldsymbol\delta_{h}} = \log \left ( 1 + \sum_{h \notin \{k,K \}} e^{\mathbf{w}_i^T \boldsymbol\delta_{h}} \right )$$
 
 We can use the quantities to re-express the full conditionals for $\boldsymbol\delta_k$ as
-$$p(\boldsymbol\delta_k|\mathbf{Z},\boldsymbol\delta_{k' \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \left (\frac{e^{\eta_{ik}}}{1 + e^{\eta_{ik}}} \right )^{U_{ik}} \left (\frac{1}{1 + e^{\eta_{ik}}} \right )^{1-U_{ik}} = p(\boldsymbol\delta_k) \prod_{i = 1}^n \frac{(e^{\eta_{ik}})^{U_{ik}}}{1 + e^{\eta_{ik}}}$$
+$$p(\boldsymbol\delta_k|\mathbf{Z},\boldsymbol\delta_{h \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \left (\frac{e^{\eta_{ik}}}{1 + e^{\eta_{ik}}} \right )^{U_{ik}} \left (\frac{1}{1 + e^{\eta_{ik}}} \right )^{1-U_{ik}} = p(\boldsymbol\delta_k) \prod_{i = 1}^n \frac{(e^{\eta_{ik}})^{U_{ik}}}{1 + e^{\eta_{ik}}}$$
 

### 2019-06-24 manuscript/manuscript.tex
section 3

@@
 Conveniently, all parameters admit conditionally conjugate priors, which greatly improves posterior computation via a data-augmented Gibbs sampler.
-For \textbf{[Give priors for each parameter.
-Be clear about the conditionally joint prior for $\boldsymbol\beta_k$ and $\boldsymbol\Sigma_k$.
-Where appropriate, explain advantages]}
+For the MSN model component, we adopt a conditionally independent prior structure for $\boldsymbol\beta_k$ and $\boldsymbol\Sigma_k$, where $p(\boldsymbol\beta_k,\boldsymbol\Sigma_k) = p(\boldsymbol\Sigma_k)p(\boldsymbol\beta_k|\boldsymbol\Sigma_k)$.
+We choose the normal-inverse-Wishart distribution for $p(\boldsymbol\beta_k,\boldsymbol\Sigma_k)$ by specifying $\boldsymbol\Sigma_k \sim \text{IW}(\mathbf{V}_k,\nu_k)$ and $\boldsymbol\beta_k|\boldsymbol\Sigma_k \sim \text{N}_{Jp}(\mathbf{b}_k,\mathbf{I}_p \otimes \boldsymbol\Sigma_k)$.
+For the multinomial logit model component, the regression parameters $\boldsymbol\delta_k$ are given a conjugate $\text{N}_r(\mathbf{d}_k, \mathbf{S}_k)$ prior for $k = 1,...,K-1$.
+
+We allow the normal-inverse-Wishart and multinomial hyperparameters to vary across clusters, though they may be shared across clusters in practice.
+An advantage of allowing for cluster-specific prior parameters is that \textit{a priori} knowledge of development trends can be incorporated into certain clusters while still allowing the parameters of other clusters to be almost entirely determined by the data.
+Additionally, prior information regarding the effect of certain covariates on development cluster membership can be incorporated in to the model by choosing informative values for $\mathbf{d}_k \ \text{and} \ \mathbf{S}_k$
+
+\textbf{[Give priors for each parameter. Be clear about the conditionally joint prior for $\boldsymbol\beta_k$ and $\boldsymbol\Sigma_k$. Where appropriate, explain advantages]}
 

### 2019-06-24 manuscript/manuscript.tex
section 3

@@
 For now we assume that $K$ is fixed; in Section 4, we discuss model selection strategies for choosing the optimal value of $K$.
-We also assume that class membership is fixed throughout the study period, since our focus is to cluster individuals based on their overall developmental patterns over the course of the study.
-In Section 6, we discuss extensions to allow for class membership to vary over time.
-\textbf{[We could omit these last two sentence -- are they really needed?
-Not sure.
-Maybe keep for now and think about it.]}
 
@@
 where $MSN_J(\cdot)$ denotes the $J$-dimensional multivariate skew normal density, $\boldsymbol\zeta_{ki}$ is a $J \times 1$ vector of subject- and cluster-specific location parameters, $\boldsymbol\alpha_k$ is a $J \times 1$ vector of cluster-specific skewness parameters, and $\boldsymbol\Omega_k$ is a $J \times J$ cluster-specific scale matrix that captures dependence among the $J$ responses.
-The vector $\boldsymbol\alpha_k$ has components $\alpha_{kj}$, $j = 1,...,J$ \textbf{[let's use $kij$ as our index hierarchy.
-Please review throughout]}, that control the skewness of outcome $j$ in cluster $k$.
+The vector $\boldsymbol\alpha_k$ has components $\alpha_{kj}$, $j = 1,...,J$, that control the skewness of outcome $j$ in cluster $k$.
 When $\boldsymbol\alpha_k = \mathbf{0}$, the MSN distribution reduces to the multivariate normal distribution $N_J(\boldsymbol\zeta_k,\boldsymbol\Omega_k)$, where $\boldsymbol\Omega_k$ is a $J \times J$ covariance matrix.
@@
 We choose the normal-inverse-Wishart distribution for $p(\boldsymbol\beta_k,\boldsymbol\Sigma_k)$ by specifying $\boldsymbol\Sigma_k \sim \text{IW}(\mathbf{V}_k,\nu_k)$ and $\boldsymbol\beta_k|\boldsymbol\Sigma_k \sim \text{N}_{Jp}(\mathbf{b}_k,\mathbf{I}_p \otimes \boldsymbol\Sigma_k)$.
+We assign the skewness parameters $\boldsymbol\psi_k$ a conjugate $\text{N}_J(\mathbf{m}_k,\mathbf{P}_k)$ prior.
+However, the updates of $\boldsymbol\beta_k$ and $\boldsymbol\psi_k$ can be combined into one step by defining the $(Jp + J) \times 1$ vector $\boldsymbol\beta^*_k = (\boldsymbol\beta_k,\boldsymbol\psi_k)^T$ for which we assume a $\text{N}_{Jp+J}(\mathbf{b}^*_k,\mathbf{I}_{(p+1)} \otimes \boldsymbol\Sigma_k)$ prior.
 For the multinomial logit model component, the regression parameters $\boldsymbol\delta_k$ are given a conjugate $\text{N}_r(\mathbf{d}_k, \mathbf{S}_k)$ prior for $k = 1,...,K-1$.
@@
 Additionally, prior information regarding the effect of certain covariates on development cluster membership can be incorporated in to the model by choosing informative values for $\mathbf{d}_k \ \text{and} \ \mathbf{S}_k$
-
-\textbf{[Give priors for each parameter. Be clear about the conditionally joint prior for $\boldsymbol\beta_k$ and $\boldsymbol\Sigma_k$. Where appropriate, explain advantages]}
 
@@
 
-\paragraph{P\'olya--Gamma Data Augmentation for $z_i$.} The sampler begins by updating the latent cluster indicators $z_i$ $(i=1,\ldots,n)$ from its multinomial logit full conditional.
+\paragraph{P\'olya--Gamma Data Augmentation for $z_i$} The sampler begins by updating the latent cluster indicators $z_i$ $(i=1,\ldots,n)$ from its multinomial logit full conditional.
 To facilitate sampling, we adopt an efficient data-augmentation approach introduced by Polson \textit{et al.} (2013), which expresses the inverse-logit function as a mixture P\'olya--Gamma densities.
-\textbf{[See my Bayesian Analysis paper for guidance on this part]}.
+A random variable $w$ is said to follow a P\'olya--Gamma distribution with parameters $b > 0$ and $c \in {\rm I\!R}$ if
+$$w \sim \text{PG}(b,c) \stackrel{d}{=} \frac{1}{2\pi^2}\sum_{s=1}^{\infty}\frac{g_s}{(s-1/2)^2 + c^2/(4\pi^2)},$$
+where $g_s \stackrel{iid}{\sim} \text{Ga}(b,1)$ for $s = 1,...,\infty$  
 $$p(\boldsymbol\delta_k|\mathbf{z},\boldsymbol\delta_{h \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \pi_{ik}^{U_{ik}}(1-\pi_{ik})^{1-U_{ik}}$$

### 2019-06-25 manuscript/manuscript.tex
section 3

@@
 The above prior specification induces closed-form full conditionals that can be efficiently updated as part of a Gibbs sampler outlined below.
+A programatic sketch of our MCMC algorithm is given in table (\ref{alg:MCMC}).
 Additional details, including derivations can be found in the Web Appendix.
-\textbf{[Think about the best way to organize this section.
-Maybe see my Bayesian Analysis paper for guidance?
-We can discuss next week.]}
 
@@
 To facilitate sampling, we adopt an efficient data-augmentation approach introduced by Polson \textit{et al.} (2013), which expresses the inverse-logit function as a mixture P\'olya--Gamma densities.
+By using P\'olya--Gamma data augmentation for the multinomial model, we obtain a P\'olya--Gamma mixture of experts model, a computationally efficient way to model edge weights (\textbf{cite}).
 A random variable $w$ is said to follow a P\'olya--Gamma distribution with parameters $b > 0$ and $c \in {\rm I\!R}$ if
-$$w \sim \text{PG}(b,c) \stackrel{d}{=} \frac{1}{2\pi^2}\sum_{s=1}^{\infty}\frac{g_s}{(s-1/2)^2 + c^2/(4\pi^2)},$$
-where $g_s \stackrel{iid}{\sim} \text{Ga}(b,1)$ for $s = 1,...,\infty$  
-$$p(\boldsymbol\delta_k|\mathbf{z},\boldsymbol\delta_{h \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \pi_{ik}^{U_{ik}}(1-\pi_{ik})^{1-U_{ik}}$$
+\begin{equation}
+w \sim \text{PG}(b,c) \stackrel{d}{=} \frac{1}{2\pi^2}\sum_{s=1}^{\infty}\frac{g_s}{(s-1/2)^2 + c^2/(4\pi^2)}, \label{eq:pg1}
+\end{equation}
+where $g_s \stackrel{iid}{\sim} \text{Ga}(b,1)$ for $s = 1,...,\infty$ . Polson \textit{et al.} (2013) establish that for the $\text{PG}(b,c)$ density, and for $a,\eta \in {\rm I\!R}$,
+
+\begin{equation}
+\frac{(e^{\eta})^a}{(1 + e^{\eta})^b} = 2^{-b}e^{\kappa \eta} \int_{0}^{\infty} e^{-w\eta^2/2}p(w|b,c = 0)dw. \label{eq:pg2}
+\end{equation}
+where $\kappa = a - b/2$. Polson \textit{et al.} also show that the conditional distribution $p(w|b,c)$ results from an ``exponential tilting'' of the $\text{PG}(b,0)$ density, thus
+
+\begin{equation}
+	p(w|b,c) = \frac{e^{-c^2w/2}p(w|b,0)}{E_w[e^{-c^2w/2}]} = \frac{e^{-c^2w/2}p(w|b,0)}{\int_0^\infty e^{-c^2w/2}p(w|b,0)dw}. \label{eq:pg3}
+\end{equation}
+
+To make use of P\'olya--Gamma data augmentation, we first write the full conditional distribution of $\boldsymbol\delta_k$ as the prior for $\boldsymbol\delta_k$ times the multinomial likelihood for all $z_i$.
+$$p(\boldsymbol\delta_k|\mathbf{z},\boldsymbol\delta_{h \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \pi_{ik}^{U_{ik}}(1-\pi_{ik})^{1-U_{ik}},$$
 where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, $U_{ik} = {1}_{z_i = k}$ is an indicator that subject $i$ belongs to cluster $k$, and $\pi_{ik}$ is defined as in Section 3.4.
@@
 We can use the quantities to re-express the full conditionals for $\boldsymbol\delta_k$ as
-$$p(\boldsymbol\delta_k|\mathbf{Z},\boldsymbol\delta_{h \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \left (\frac{e^{\eta_{ik}}}{1 + e^{\eta_{ik}}} \right )^{U_{ik}} \left (\frac{1}{1 + e^{\eta_{ik}}} \right )^{1-U_{ik}} = p(\boldsymbol\delta_k) \prod_{i = 1}^n \frac{(e^{\eta_{ik}})^{U_{ik}}}{1 + e^{\eta_{ik}}}$$
-
+\begin{eqnarray}
+	p(\boldsymbol\delta_k|\mathbf{Z},\boldsymbol\delta_{h \ne k}) &\propto& p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \left (\frac{e^{\eta_{ik}}}{1 + e^{\eta_{ik}}} \right )^{U_{ik}} \left (\frac{1}{1 + e^{\eta_{ik}}} \right )^{1-U_{ik}} \nonumber\\ &=& p(\boldsymbol\delta_k) \prod_{i = 1}^n \frac{(e^{\eta_{ik}})^{U_{ik}}}{1 + e^{\eta_{ik}}} \label{eq:pglogit}
+\end{eqnarray}
 which we note is essentially a logistic regression likelihood.
@@
 
-\begin{itemize}
+\subsubsection{MCMC Algorithm}
 
-\item Emphasize that PG data augmentation for the multinomial model results in a PG mixture of experts model, which is a computationally efficient way to model edge weights.
-
-\end{itemize}
-
-\subsubsection{MCMC Algorithm}
+See Table \ref{alg:MCMC}.
 
@@
 \caption{Gibbs Sampler}
+\label{alg:MCMC}
 \begin{algorithmic}
@@
 
+We monitor conver convergence of our MCMC algorithm through the use of standard approaches such Geweke's (1992) Z-diagnostic.
+In simulation studies under realistic parameter settings, we observed relatively fast convergence of all MCMC chains (i.e.
+within 1,000 iterations).
+
 \subsubsection{Label Switching}
+
+A problematic feature of mixture models such as the one proposed here is that the full model likelihood is invariant to permutations of $\mathbf{z} = \{ z_1, z_2,..., z_n \}$, the cluster labeling of each observation, the result of which is label switching across MCMC iterations - a process in which draws of cluster-specific parameters may be assigned to the wrong cluster at some point during the MCMC simulation.
+After conducting simulation studies under a wide variety of realistic parameter settings, as detailed in Section 4, we found little evidence of label switching in posterior draws of $\mathbf{z}$.
+When label switching was observed, we implemented \textit{post hoc} relabling algorithms as described in Panagiotis 2016 and implemented in the \texttt{label.switching} package in \texttt{R}.
+We discuss other possible label switching remedies in Section 6.
 
@@
 
-\textit{\textbf{Discuss non-ignorable missingness here}}
-
 \begin{itemize}
@@
 \item Discuss how we handle non-ignorable missingness
+
+\item Discuss other label switching approaches
+
+\item Discuss skew-t?
 
@@
 
+\bibitem{ } Geweke, J.
+Evaluating the accuracy of sampling-based ap- proaches to calculating posterior moments.
+\textit{Bayesian Statistics} Vol.
+4, J.
+M.
+Bernardo, J.
+Berger, A.
+P.
+Dawid, and A.F.M.
+Smith (eds), 169–193.
+1992.
+Cambridge, U.K.: Oxford University Press.
+
 \bibitem{ } Gelman A, Stern HS, Carlin JB, Dunson DB, Vehtari A, Rubin DB. Bayesian data analysis. \textit{Chapman and Hall/CRC}; 2013 Nov 27.
@@
 
+\bibitem{ } Panagiotis Papastamoulis (2016).
+label.switching: An R Package for Dealing with the Label Switching Problem in MCMC Outputs.
+\textit{Journal of Statistical Software}, 69(1), 1-24.
+doi:10.18637/jss.v069.c01
+
 \bibitem{ } Polson NG, Scott JG, Windle J. Bayesian inference for logistic models using P\'olya –- Gamma latent variables. Journal of the American statistical Association. 2013 Dec 1;108(504):1339-49.
@@
 \bibitem{ } Vrbik I, Mcnicholas PD. Parsimonious skew mixture models for model-based clustering and classification. \textit{Computational Statistics \& Data Analysis}. 2014 Mar 1;71:196-210.
+
+\bibitem{ } Watanabe S.
+Asymptotic equivalence of Bayes cross validation and widely applicable information criterion in singular learning theory.
+\textit{Journal of Machine Learning Research}.
+2010;11(Dec):3571-94.
 

### 2019-06-25 manuscript/manuscript.tex
section 3

@@
 \mathbf{y}^{miss}_i|(z_i=k,t_i,\mathbf{y}^{obs}_i)&\sim& N_{J-q_i}(\boldsymbol\mu^{cond}_{ki},\boldsymbol\Sigma^{cond}_k),~\text{where}\nonumber\\
-\boldsymbol\mu^{cond}_{ki}&=& \boldsymbol\mu^{miss} + \boldsymbol\Sigma_{12} \boldsymbol\Sigma_{22}^{-1}(\mathbf{y}^{obs}_i - \boldsymbol\mu^{obs})\\
+\boldsymbol\mu^{cond}_{ki}&=& \boldsymbol\mu^{miss}_{ki} + \boldsymbol\Sigma_{12} \boldsymbol\Sigma_{22}^{-1}(\mathbf{y}^{obs}_i - \boldsymbol\mu^{obs}_{ki}) \label{eq:impute}\\
 \boldsymbol\Sigma^{cond}_k&=& \boldsymbol\Sigma_{11} - \boldsymbol\Sigma_{12}\boldsymbol\Sigma_{22}^{-1}\boldsymbol\Sigma_{21},~\text{where}\nonumber
@@
 
-\paragraph{P\'olya--Gamma Data Augmentation for $z_i$} The sampler begins by updating the latent cluster indicators $z_i$ $(i=1,\ldots,n)$ from its multinomial logit full conditional.
+\paragraph{\textbf{Step 1}: Conditional MSN Imputation} The sampler begins with imputing missing values $\mathbf{y}_i^{miss}$ conditional on current simulated values of $z_i = k$ and $t_i$ as well as the associated $\mathbf{y}_i^{obs}$.
+This is accomplished by drawing $\mathbf{y}_i^{miss}$ from $\text{N}_{J-q_i}(\boldsymbol\mu^{cond}_{ki},\boldsymbol\Sigma^{cond}_k)$ as described in \ref{eq:impute}.
+
+\paragraph{\textbf{Step 3:}P\'olya--Gamma Data Augmentation for $z_i$} The sampler concludes by updating the latent cluster indicators $z_i$ $(i=1,\ldots,n)$ from its multinomial logit full conditional.
 To facilitate sampling, we adopt an efficient data-augmentation approach introduced by Polson \textit{et al.} (2013), which expresses the inverse-logit function as a mixture P\'olya--Gamma densities.

### 2019-06-25 manuscript/manuscript.tex
section 3

@@
 \paragraph{\textbf{Step 1}: Conditional MSN Imputation} The sampler begins with imputing missing values $\mathbf{y}_i^{miss}$ conditional on current simulated values of $z_i = k$ and $t_i$ as well as the associated $\mathbf{y}_i^{obs}$.
-This is accomplished by drawing $\mathbf{y}_i^{miss}$ from $\text{N}_{J-q_i}(\boldsymbol\mu^{cond}_{ki},\boldsymbol\Sigma^{cond}_k)$ as described in \ref{eq:impute}.
+This is accomplished by drawing $\mathbf{y}_i^{miss}$ from $\text{N}_{J-q_i}(\boldsymbol\mu^{cond}_{ki},\boldsymbol\Sigma^{cond}_k)$ as described in equation \ref{eq:impute} for all $i = 1,...,n$.
+Conditional imputation is completed with construction of a complete outcome vector $\mathbf{y}_i$ by merging the now imputed $\mathbf{y}_i^{miss}$ with $\mathbf{y}_i^{obs}$ with respect for the original order of $\mathbf{y}_i$.
+
+\paragraph{\textbf{Step 2:} Update of MSN Regression Parameters}
 

### 2019-06-25 manuscript/manuscript.tex
section 3

@@
 \usepackage{dsfont}
+\usepackage{amsfonts}
 
@@
 To address this aim, we propose a flexible finite mixture model that accommodates relevant features of the data, such as skewness and dependence among the responses.
-To this end, let $\mathbf{y}_{i}=(y_{i1},\ldots,y_{iJ})^T$ be a $J \times 1$ vector of responses (i.e., Baley scores) for subject $i$ $(i=1,\ldots,n)$.
+For $i = 1,...,n$, let $\mathbf{y}_{i}=(y_{i1},\ldots,y_{iJ})^T$ be a $J \times 1$ vector of responses (i.e., Baley scores) for subject $i$ across the $J$ measurement occasions.
 For the analysis of the Nurture data, we propose a finite mixture model of the form
@@
 \begin{equation}
-\mathbf{y}_{i}|(z_i=k) \sim MSN_J(\boldsymbol\zeta_{ki},\boldsymbol\alpha_k,\boldsymbol\Omega_k), \label{eq:msndens}
+\mathbf{y}_{i}|(z_i=k) \sim \text{MSN}_J(\boldsymbol\zeta_{ki},\boldsymbol\alpha_k,\boldsymbol\Omega_k), \label{eq:msndens}
 \end{equation}
-where $MSN_J(\cdot)$ denotes the $J$-dimensional multivariate skew normal density, $\boldsymbol\zeta_{ki}$ is a $J \times 1$ vector of subject- and cluster-specific location parameters, $\boldsymbol\alpha_k$ is a $J \times 1$ vector of cluster-specific skewness parameters, and $\boldsymbol\Omega_k$ is a $J \times J$ cluster-specific scale matrix that captures dependence among the $J$ responses.
+where $\text{MSN}_J(\cdot)$ denotes the $J$-dimensional multivariate skew normal density, $\boldsymbol\zeta_{ki}$ is a $J \times 1$ vector of subject- and cluster-specific location parameters, $\boldsymbol\alpha_k$ is a $J \times 1$ vector of cluster-specific skewness parameters, and $\boldsymbol\Omega_k$ is a $J \times J$ cluster-specific scale matrix that captures dependence among the $J$ responses for subject $i$.
 The vector $\boldsymbol\alpha_k$ has components $\alpha_{kj}$, $j = 1,...,J$, that control the skewness of outcome $j$ in cluster $k$.
-When $\boldsymbol\alpha_k = \mathbf{0}$, the MSN distribution reduces to the multivariate normal distribution $N_J(\boldsymbol\zeta_k,\boldsymbol\Omega_k)$, where $\boldsymbol\Omega_k$ is a $J \times J$ covariance matrix.
+When $\boldsymbol\alpha_k = \mathbf{0}$, the MSN distribution reduces to the multivariate normal distribution $\text{N}_J(\boldsymbol\zeta_k,\boldsymbol\Omega_k)$, where $\boldsymbol\zeta_k$ is a $J \times 1$ mean vector and $\boldsymbol\Omega_k$ is a $J \times J$ covariance matrix.
 
@@
 where 
-$\mathbf{X}_i$ is a $J \times Jp$ design matrix that includes potential time-varying covariates (e.g., indicators denoting quarterly visits); $\boldsymbol{\beta}_k=(\beta_{k11},\ldots,\beta_{k1p},\ldots,\beta_{kJ1},\ldots,\beta_{kJp})^T$ is a $Jp\times 1$ vector of cluster- and outcome-specific regression coefficients; $t_i\sim N_{[0,\infty)}(0,1)$ is a subject-specific standard normal random variable truncated below by zero; $\boldsymbol\psi_k=(\psi_{k1},\ldots,\psi_{kJ})^T$ is a $J \times 1$ vector of cluster-specific skewness parameters; and $\boldsymbol{\epsilon}_{ki} \sim N_J(\boldsymbol0,\boldsymbol\Sigma_k)$ is a $J\times 1$ vector of error terms.
-Thus, conditional on $t_i$ and $z_i=k$, $\boldsymbol{y}_i$ is distributed as $N_J(\mathbf{X}_i \boldsymbol\beta_k + t_i \boldsymbol\psi_k, \boldsymbol{\Sigma}_k)$.
-Marginally (after integrating over $t_i$), $\mathbf{y}_i$ is distributed $MSN_J(\boldsymbol\zeta_{ki}, \boldsymbol\alpha_k, \boldsymbol\Omega_k)$, where through back-transformation
+$\mathbf{X}_i$ is a $J \times Jp$ design matrix that includes potential time-varying covariates (e.g., indicators denoting quarterly visits); $\boldsymbol{\beta}_k=(\beta_{k11},\ldots,\beta_{k1p},\ldots,\beta_{kJ1},\ldots,\beta_{kJp})^T$ is a $Jp\times 1$ vector of cluster- and outcome-specific regression coefficients; $t_i\sim \text{N}_{[0,\infty)}(0,1)$ is a subject-specific standard normal random variable truncated below by zero; $\boldsymbol\psi_k=(\psi_{k1},\ldots,\psi_{kJ})^T$ is a $J \times 1$ vector of cluster-specific skewness parameters; and $\boldsymbol{\epsilon}_{ki} \sim \text{N}_J(\boldsymbol0,\boldsymbol\Sigma_k)$ is a $J\times 1$ vector of error terms.
+Thus, conditional on $t_i$ and $z_i=k$, $\boldsymbol{y}_i$ is distributed as $\text{N}_J(\mathbf{X}_i \boldsymbol\beta_k + t_i \boldsymbol\psi_k, \boldsymbol{\Sigma}_k)$.
+Marginally (after integrating over $t_i$), $\mathbf{y}_i$ is distributed $\text{MSN}_J(\boldsymbol\zeta_{ki}, \boldsymbol\alpha_k, \boldsymbol\Omega_k)$, where through back-transformation
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-where $\boldsymbol\omega_k = \text{Diag}(\sqrt{\omega_{k,11}},...,\sqrt{\omega_{k,JJ}})$ is the $J \times J$ diagonal matrix containing the square root of the diagonal entries of $\boldsymbol\Omega_k$.
+where $\boldsymbol\omega_k = \text{Diag}(\boldsymbol\Omega_k)^{1/2}$ is the $J \times J$ diagonal matrix containing the square root of the diagonal entries of $\boldsymbol\Omega_k$.
 Additional details can be found in Fr\"{u}wirth-Schnatter and Pyne (2010).
@@
 \begin{eqnarray*}
-\mathbf{Y}_{k} &\sim& MatSN_{n_k \times J}(\mathbf{M}_k,\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Omega_k)\\
+\mathbf{Y}_{k} &\sim& \text{MatSN}_{n_k \times J}(\mathbf{M}_k,\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Omega_k)\\
 %\text{vec}(\mathbf{Y}_k) =  (\mathbf{y}^T_1,...,\mathbf{y}^T_{n_k})^T\\
@@
 where $\boldsymbol\zeta_{ki} = \mathbf{X}_i \boldsymbol\beta_k$ as in equation (\ref{eq:msnreg}), $\boldsymbol\alpha_k = (\alpha_{k1},...,\alpha_{kJ})^T$, $\mathbf{I}_{n_k}$ is the $n_k\times n_k$ identity matrix, and $\boldsymbol\Omega_k$ is the $J\times J$ scale matrix defined above in equation (\ref{eq:msndens}).
-From equation (\ref{eq:msnreg}), it follows that $\mathbf{Y}_{k}$, conditional on the $n_k\times 1$ vector of random effects $\mathbf{t}_{k}$, is jointly distributed in matrix form as
+From equation (\ref{eq:msnreg}), it follows that $\mathbf{Y}_{k}$, conditional on the $n_k\times 1$ vector of random effects $\mathbf{t}_{k} = (t_1,...,t_{n_k})^T$, is jointly distributed in matrix form as
 \[
-\mathbf{Y}_k | \mathbf{t}_k \sim MatNorm_{n_k \times J}(\mathbf{M}_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k),
+\mathbf{Y}_k | \mathbf{t}_k \sim \text{MatNorm}_{n_k \times J}(\mathbf{M}_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k),
 \]
-where $MatNorm_{n_k \times J}(\cdot)$ denotes a $n_k\times J$ matrix normal density, $\text{vec}(\mathbf{M}_k) = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$ is an $n_kJ\times 1$ mean vector, $\mathbf{X}_k$ is an $n_kJ \times Jp$ design matrix, $\boldsymbol\beta_k$ is the $(Jp)\times 1$ vector of regression coefficients defined in equation (\ref{eq:msnreg}), and $\boldsymbol\Sigma_k$ is the $J\times J$ conditional covariance of $\boldsymbol\epsilon_{ik}$ given in equation (\ref{eq:msnreg}).
+where $\text{MatNorm}_{n_k \times J}(\cdot)$ denotes a $n_k\times J$ matrix normal density, $\text{vec}(\mathbf{M}_k) = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$ is an $n_kJ\times 1$ mean vector, $\mathbf{X}_k$ is an $n_kJ \times Jp$ design matrix, $\boldsymbol\beta_k$ is the $(Jp)\times 1$ vector of regression coefficients defined in equation (\ref{eq:msnreg}), and $\boldsymbol\Sigma_k$ is the $J\times J$ conditional covariance of $\boldsymbol\epsilon_{ik}$ given in equation (\ref{eq:msnreg}).
 
 \subsection{Multinomial Regression for the Cluster Indicators}
-To accommodate heterogeneity in the cluster-membership probabilities, we model $\pi_{ik}$ as a function of coovariates using a multinomial logit model 
+To accommodate heterogeneity in the cluster-membership probabilities, we model $\pi_{ik}$ as a function of covariates using a multinomial logit model 
 \begin{eqnarray}
-\pi_{ik} = \Pr(z_i = k|\mathbf{w}_i) = \frac{e^{\mathbf{w}_i^T \boldsymbol\delta_k}}{\sum_{h = 1}^K e^{\mathbf{w}_i^T \boldsymbol\delta_{h}}},~ k=1,\ldots,K,
+\pi_{ik} = \Pr(z_i = k|\mathbf{w}_i) = \frac{\text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k}}{\sum_{h = 1}^K \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}},~ k=1,\ldots,K,
 \end{eqnarray}
@@
 For identifiability purposes, we fix the reference category $k = K$ and set $\boldsymbol\delta_K = \mathbf{0}$.
-Under this model, $z_i|\mathbf{w}_i \sim Multinom(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{i1},...,\pi_{1K})$.
+Under this model, $z_i|\mathbf{w}_i \sim \text{Multinom}(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{i1},...,\pi_{1K})$.
 During MCMC estimation, the cluster labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as cluster assignments.
@@
 Suppose $\mathbf{y}_i$ has $q_i\in (1,\ldots,J)$ observed values, denoted $\mathbf{y}^{obs}_i$, and $J-q_i$ intermittent missing values, denoted $\mathbf{y}^{miss}_i$.
-We can make use of the stochastic representation given in equation (\ref{eq:msnreg}) to impute $\mathbf{y}^{miss}_i$ from its conditional multivariate normal distribution given $(z_i,t_i,Y^{obs}_i)$:
+We can make use of the stochastic representation given in equation (\ref{eq:msnreg}) to impute $\mathbf{y}^{miss}_i$ from its conditional multivariate normal distribution given $(z_i,t_i,\mathbf{y}^{obs}_i)$:
 \begin{eqnarray}
-\mathbf{y}^{miss}_i|(z_i=k,t_i,\mathbf{y}^{obs}_i)&\sim& N_{J-q_i}(\boldsymbol\mu^{cond}_{ki},\boldsymbol\Sigma^{cond}_k),~\text{where}\nonumber\\
+\mathbf{y}^{miss}_i|(z_i=k,t_i,\mathbf{y}^{obs}_i)&\sim& \text{N}_{J-q_i}(\boldsymbol\mu^{cond}_{ki},\boldsymbol\Sigma^{cond}_k),~\text{where}\nonumber\\
 \boldsymbol\mu^{cond}_{ki}&=& \boldsymbol\mu^{miss}_{ki} + \boldsymbol\Sigma_{12} \boldsymbol\Sigma_{22}^{-1}(\mathbf{y}^{obs}_i - \boldsymbol\mu^{obs}_{ki}) \label{eq:impute}\\
-\boldsymbol\Sigma^{cond}_k&=& \boldsymbol\Sigma_{11} - \boldsymbol\Sigma_{12}\boldsymbol\Sigma_{22}^{-1}\boldsymbol\Sigma_{21},~\text{where}\nonumber
+\boldsymbol\Sigma^{cond}_k&=& \boldsymbol\Sigma_{11} - \boldsymbol\Sigma_{12}\boldsymbol\Sigma_{22}^{-1}\boldsymbol\Sigma_{21},~\text{and}\nonumber \\
+\boldsymbol\Sigma_k &=& \begin{pmatrix} 
+	\boldsymbol\Sigma_{k11} & \boldsymbol\Sigma_{k12}\\
+	\boldsymbol\Sigma_{k21} & \boldsymbol\Sigma_{k22}
+\end{pmatrix}, \ \text{where} \nonumber
 \end{eqnarray}
-$\boldsymbol\Sigma_k$ is partitioned into four sub-matrices $\boldsymbol\Sigma_{11},\boldsymbol\Sigma_{12},\boldsymbol\Sigma_{21}, \ \text{and} \ \boldsymbol\Sigma_{22}$ such that $\boldsymbol\Sigma_{11}$ is a $J-q_i \times J-q_i$ matrix containing the rows and columns of $\boldsymbol\Sigma_k$ corresponding to inicies of $\mathbf{y}_{i}$ where missingness occurs.
-Similarly, $\boldsymbol\Sigma_{12}$ is a $J-q_i \times q_i$ matrix containing the rows of $\boldsymbol\Sigma_k$ that correspond to missing indices of $\mathbf{y}_i$, but columns of $\boldsymbol\Sigma_k$ that correspond to observed indices of $\mathbf{y}_i$.
-The remaining partitions $\boldsymbol\Sigma_{21}, \ \text{and} \ \boldsymbol\Sigma_{22}$ are defined in the same manner.
+$\boldsymbol\Sigma_{k11}$ is a $J-q_i \times J-q_i$ matrix containing the rows and columns of $\boldsymbol\Sigma_k$ corresponding to inicies of $\mathbf{y}_{i}$ where missingness occurs.
+Similarly, $\boldsymbol\Sigma_{k12}$ is a $J-q_i \times q_i$ matrix containing the rows of $\boldsymbol\Sigma_k$ that correspond to missing indices of $\mathbf{y}_i$, but columns of $\boldsymbol\Sigma_k$ that correspond to observed indices of $\mathbf{y}_i$.
+The remaining partitions $\boldsymbol\Sigma_{k21}, \ \text{and} \ \boldsymbol\Sigma_{k22}$ are defined in the same manner.
 These results follow from conventional multivariate normal theory.
 An attractive feature of this imputation algorithm is that it avoids multiplicative run-time scaling in $m$, the number of imputations (Gelman \textit{et al.} 2013; Zhou and Reiter, 2010).
-Our approach also provides more opportunities to explore the missing data parameter space than does multiple imputation, since missing values are drawn at each MCMC iteration, and often in practice $n_{sim} >> m$, where $n_{sim}$ is the number of MCMC iterations.
+Our approach also provides more opportunities to explore the missing data parameter space than does multiple imputation, since missing values are drawn at each MCMC iteration, and often in practice $n_{sim} >> m$, where $n_{sim}$ is the number of MCMC iterations (\textbf{find a reference}).
 In Section 4, we conduct simulation studies to demonstrate that imputing the missing MSN responses improves inferences over complete case analysis.
@@
 For the MSN model component, we adopt a conditionally independent prior structure for $\boldsymbol\beta_k$ and $\boldsymbol\Sigma_k$, where $p(\boldsymbol\beta_k,\boldsymbol\Sigma_k) = p(\boldsymbol\Sigma_k)p(\boldsymbol\beta_k|\boldsymbol\Sigma_k)$.
-We choose the normal-inverse-Wishart distribution for $p(\boldsymbol\beta_k,\boldsymbol\Sigma_k)$ by specifying $\boldsymbol\Sigma_k \sim \text{IW}(\mathbf{V}_k,\nu_k)$ and $\boldsymbol\beta_k|\boldsymbol\Sigma_k \sim \text{N}_{Jp}(\mathbf{b}_k,\mathbf{I}_p \otimes \boldsymbol\Sigma_k)$.
-We assign the skewness parameters $\boldsymbol\psi_k$ a conjugate $\text{N}_J(\mathbf{m}_k,\mathbf{P}_k)$ prior.
+We choose the normal-inverse-Wishart distribution for $p(\boldsymbol\beta_k,\boldsymbol\Sigma_k)$ by specifying $\boldsymbol\Sigma_k \sim \text{IW}(\mathbf{V}_{0k},\nu_{0k})$ and $\boldsymbol\beta_k|\boldsymbol\Sigma_k \sim \text{N}_{Jp}(\mathbf{b}_k,\mathbf{I}_p \otimes \boldsymbol\Sigma_k)$.
+We assign the skewness coefficients $\boldsymbol\psi_k$ a conjugate $\text{N}_J(\mathbf{m}_k,\mathbf{P}_k)$ prior.
 However, the updates of $\boldsymbol\beta_k$ and $\boldsymbol\psi_k$ can be combined into one step by defining the $(Jp + J) \times 1$ vector $\boldsymbol\beta^*_k = (\boldsymbol\beta_k,\boldsymbol\psi_k)^T$ for which we assume a $\text{N}_{Jp+J}(\mathbf{b}^*_k,\mathbf{I}_{(p+1)} \otimes \boldsymbol\Sigma_k)$ prior.
-For the multinomial logit model component, the regression parameters $\boldsymbol\delta_k$ are given a conjugate $\text{N}_r(\mathbf{d}_k, \mathbf{S}_k)$ prior for $k = 1,...,K-1$.
+For the multinomial logit model component, the regression parameters $\boldsymbol\delta_k$ are assigned a $\text{N}_r(\mathbf{d}_k, \mathbf{S}_k)$ prior for $k = 1,...,K-1$, which is conditionally conjugate as described in Section (\textbf{what section}).
 
 We allow the normal-inverse-Wishart and multinomial hyperparameters to vary across clusters, though they may be shared across clusters in practice.
-An advantage of allowing for cluster-specific prior parameters is that \textit{a priori} knowledge of development trends can be incorporated into certain clusters while still allowing the parameters of other clusters to be almost entirely determined by the data.
-Additionally, prior information regarding the effect of certain covariates on development cluster membership can be incorporated in to the model by choosing informative values for $\mathbf{d}_k \ \text{and} \ \mathbf{S}_k$
+An advantage of allowing for cluster-specific prior parameters is that \textit{a priori} knowledge of motor development trends can be incorporated into certain clusters while allowing the priors for other clusters to be less informative.
+Additionally, prior information regarding the effect of certain covariates on cluster membership can be incorporated in to the model by choosing informative values for $\mathbf{d}_k \ \text{and} \ \mathbf{S}_k$
 
@@
 The above prior specification induces closed-form full conditionals that can be efficiently updated as part of a Gibbs sampler outlined below.
-A programatic sketch of our MCMC algorithm is given in table (\ref{alg:MCMC}).
+A programatic sketch of our MCMC algorithm is given in Table (\ref{alg:MCMC}).
 Additional details, including derivations can be found in the Web Appendix.
@@
 
-\paragraph{\textbf{Step 2:} Update of MSN Regression Parameters}
+\paragraph{\textbf{Step 2:} Update of MSN Regression Parameters} We begin the update of MSN regression parameters by updating $t_i$, the truncated Normal random effect used in the stochastic representation of $\mathbf{y}_i$ given in equation \ref{eq:msnreg}.
+For cluster $k$, compute $A_k = (1 + \boldsymbol\psi_k^T \boldsymbol\Sigma^{-1}_k\boldsymbol
+\psi_k)^{-1}$ according to current values of $\boldsymbol\psi_k$ and $\boldsymbol\Sigma^{-1}_k$.
+Now, for $i = 1,...,n_k$, where $n_k$ is the number of members of cluster $k$ according to the current values of ${z}_1,...,{z}_n$, compute $a_i = A_k \boldsymbol\psi_k^T \boldsymbol\Sigma^{-1}_k(\mathbf{y}_i - \boldsymbol\zeta_{ik})$.
+Finally, update draw $t_i$ from $\text{N}_{[0,\infty)}(a_i, \mathbf{A}_k)$.
+\textbf{Derivations in FS-P 2010 supplement}.
 
+The remaining MSN regression parameters are updated from their full conditionals as follows.
+For cluster $k$, we update $\boldsymbol\Sigma_k$ from $\text{IW}(\nu_k,\mathbf{V}_k)$, where $\nu_k = \nu_{0k} + n_k$ and
+\begin{eqnarray}
+	\mathbf{V}_k &=&  (\mathbf{Y}-\mathbf{X}^*_k \mathbf{B}^*_k)^T (\mathbf{Y}_k - \mathbf{X}^*_k \mathbf{B}^*_k)\nonumber \\
+	 &+& (\mathbf{B}_k^* - \mathbf{B}_{0,k})^T \mathbf{I}_{p+1}(\mathbf{B}^*_k - \mathbf{B}_{0,k}) + \mathbf{V}_{0,k}. \nonumber
+\end{eqnarray}
+We then make use of the Matrix Normal representation introduced in Section 3.2 to draw $\mathbb{B}^*_k$ from $\text{MatNorm}_{p+1,J}(\mathbf{B}_k^*,\mathbf{L}_k^{-1},\boldsymbol\Sigma_k)$
 \paragraph{\textbf{Step 3:}P\'olya--Gamma Data Augmentation for $z_i$} The sampler concludes by updating the latent cluster indicators $z_i$ $(i=1,\ldots,n)$ from its multinomial logit full conditional.
 To facilitate sampling, we adopt an efficient data-augmentation approach introduced by Polson \textit{et al.} (2013), which expresses the inverse-logit function as a mixture P\'olya--Gamma densities.
-By using P\'olya--Gamma data augmentation for the multinomial model, we obtain a P\'olya--Gamma mixture of experts model, a computationally efficient way to model edge weights (\textbf{cite}).
+By using P\'olya--Gamma data augmentation for the multinomial model (\textbf{cite}), we obtain a P\'olya--Gamma mixture of experts model, a computationally efficient way to model edge weights (\textbf{cite}).
 A random variable $w$ is said to follow a P\'olya--Gamma distribution with parameters $b > 0$ and $c \in {\rm I\!R}$ if
@@
 \begin{equation}
-\frac{(e^{\eta})^a}{(1 + e^{\eta})^b} = 2^{-b}e^{\kappa \eta} \int_{0}^{\infty} e^{-w\eta^2/2}p(w|b,c = 0)dw. \label{eq:pg2}
+\frac{(\text{e}^{\eta})^a}{(1 + \text{e}^{\eta})^b} = 2^{-b}\text{e}^{\kappa \eta} \int_{0}^{\infty} \text{e}^{-w\eta^2/2}p(w|b,c = 0)dw. \label{eq:pg2}
 \end{equation}
@@
 \begin{equation}
-	p(w|b,c) = \frac{e^{-c^2w/2}p(w|b,0)}{E_w[e^{-c^2w/2}]} = \frac{e^{-c^2w/2}p(w|b,0)}{\int_0^\infty e^{-c^2w/2}p(w|b,0)dw}. \label{eq:pg3}
+	p(w|b,c) = \frac{\text{e}^{-c^2w/2}p(w|b,0)}{E_w[\text{e}^{-c^2w/2}]} = \frac{\text{e}^{-c^2w/2}p(w|b,0)}{\int_0^\infty \text{e}^{-c^2w/2}p(w|b,0)dw}. \label{eq:pg3}
 \end{equation}
@@
 We can rewrite $\pi_{ik}$ as follows
-$$\pi_{ik} = P(U_{ik} = 1) = \frac{e^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}}}{1 + e^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}}} = \frac{e^{\eta_{ik}}}{1 + e^{\eta_{ik}}}$$
-where ${c}_{ik} = \log \sum_{h \ne k} e^{\mathbf{w}_i^T \boldsymbol\delta_{h}}$ and $\eta_{ik} = \mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}$.
-We note that the sum $\sum_{h \ne k} e^{\mathbf{w}_i^T \boldsymbol\delta_{h}}$ includes the reference category, but since we fix $\boldsymbol\delta_K = \mathbf{0}$, we have $e^{\mathbf{w}_i^T \boldsymbol\delta_K} = 1$, and hence
-$$c_{ik} = \log \sum_{h \ne k} e^{\mathbf{w}_i^T \boldsymbol\delta_{h}} = \log \left ( 1 + \sum_{h \notin \{k,K \}} e^{\mathbf{w}_i^T \boldsymbol\delta_{h}} \right )$$
+$$\pi_{ik} = P(U_{ik} = 1) = \frac{\text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}}}{1 + \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}}} = \frac{\text{e}^{\eta_{ik}}}{1 + \text{e}^{\eta_{ik}}}$$
+where ${c}_{ik} = \log \sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}$ and $\eta_{ik} = \mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}$.
+We note that the sum $\sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}$ includes the reference category, but since we fix $\boldsymbol\delta_K = \mathbf{0}$, we have $\text{e}^{\mathbf{w}_i^T \boldsymbol\delta_K} = 1$, and hence
+$$c_{ik} = \log \sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}} = \log \left ( 1 + \sum_{h \notin \{k,K \}} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}} \right )$$
 
@@
 \begin{eqnarray}
-	p(\boldsymbol\delta_k|\mathbf{Z},\boldsymbol\delta_{h \ne k}) &\propto& p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \left (\frac{e^{\eta_{ik}}}{1 + e^{\eta_{ik}}} \right )^{U_{ik}} \left (\frac{1}{1 + e^{\eta_{ik}}} \right )^{1-U_{ik}} \nonumber\\ &=& p(\boldsymbol\delta_k) \prod_{i = 1}^n \frac{(e^{\eta_{ik}})^{U_{ik}}}{1 + e^{\eta_{ik}}} \label{eq:pglogit}
+	p(\boldsymbol\delta_k|\mathbf{Z},\boldsymbol\delta_{h \ne k}) &\propto& p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \left (\frac{\text{e}^{\eta_{ik}}}{1 + \text{e}^{\eta_{ik}}} \right )^{U_{ik}} \left (\frac{1}{1 + \text{e}^{\eta_{ik}}} \right )^{1-U_{ik}} \nonumber\\ &=& p(\boldsymbol\delta_k) \prod_{i = 1}^n \frac{(\text{e}^{\eta_{ik}})^{U_{ik}}}{1 + \text{e}^{\eta_{ik}}} \label{eq:pglogit}
 \end{eqnarray}
@@
 
-A problematic feature of mixture models such as the one proposed here is that the full model likelihood is invariant to permutations of $\mathbf{z} = \{ z_1, z_2,..., z_n \}$, the cluster labeling of each observation, the result of which is label switching across MCMC iterations - a process in which draws of cluster-specific parameters may be assigned to the wrong cluster at some point during the MCMC simulation.
+A common challenge for Bayesian mixtures models is the so-called ``label switching" problem, in which draws of cluster-specific parameters may be assigned to the wrong cluster at some point during the MCMC simulation, rendering summaries of class-specific parameters incoherent.
 After conducting simulation studies under a wide variety of realistic parameter settings, as detailed in Section 4, we found little evidence of label switching in posterior draws of $\mathbf{z}$.
 When label switching was observed, we implemented \textit{post hoc} relabling algorithms as described in Panagiotis 2016 and implemented in the \texttt{label.switching} package in \texttt{R}.
-We discuss other possible label switching remedies in Section 6.
 

### 2019-06-26 manuscript/manuscript.tex
section 3

@@
 \usepackage{dsfont}
+\usepackage{dutchcal}
 \usepackage{amsfonts}
@@
 First, we adopt the multivariate skew normal distribution with cluster-specific parameters that accommodate the inherent correlation and skewness in the data.
-Second, we model the cluster membership probabilities using a novel Pólya-Gamma data-augmentation scheme, thereby improving predictions of the cluster membership allocations.
+Second, we model the cluster membership probabilities using a novel P\'olya-Gamma data-augmentation scheme, thereby improving predictions of the cluster membership allocations.
 Lastly, we impute missing responses under missing at random assumption by drawing from appropriate conditional skew normal distributions.
@@
 
+To facilitate sampling, we adopt an efficient data-augmentation approach introduced by Polson \textit{et al.} (2013), which expresses the inverse-logit function as a mixture P\'olya--Gamma densities.
+By using P\'olya--Gamma data augmentation for the multinomial model (\textbf{cite}), we obtain a P\'olya--Gamma mixture of experts model, a computationally efficient way to model edge weights (\textbf{cite}).
+A random variable $\mathcal{w}$ is said to follow a P\'olya--Gamma distribution with parameters $b > 0$ and $c \in {\rm I\!R}$ if
+\begin{equation}
+\mathcal{w} \sim \text{PG}(b,c) \stackrel{d}{=} \frac{1}{2\pi^2}\sum_{s=1}^{\infty}\frac{g_s}{(s-1/2)^2 + c^2/(4\pi^2)}, \label{eq:pg1}
+\end{equation}
+where $g_s \stackrel{iid}{\sim} \text{Ga}(b,1)$ for $s = 1,...,\infty$ . Polson \textit{et al.} (2013) establish that for the $\text{PG}(b,c)$ density, and for $a,\eta \in {\rm I\!R}$,
+
+\begin{equation}
+\frac{(\text{e}^{\eta})^a}{(1 + \text{e}^{\eta})^b} = 2^{-b}\text{e}^{\kappa \eta} \int_{0}^{\infty} \text{e}^{-\mathcal{w}\eta^2/2}p(\mathcal{w}|b,c = 0)d\mathcal{w}. \label{eq:pg2}
+\end{equation}
+where $\kappa = a - b/2$. Polson \textit{et al.} also show that the conditional distribution $p(\mathcal{w}|b,c)$ results from an ``exponential tilting'' of the $\text{PG}(b,0)$ density, thus
+
+\begin{equation}
+	p(\mathcal{w}|b,c) = \frac{\text{e}^{-c^2\mathcal{w}/2}p(\mathcal{w}|b,0)}{E_{\mathcal{w}}[\text{e}^{-c^2\mathcal{w}/2}]} = \frac{\text{e}^{-c^2\mathcal{w}/2}p(\mathcal{w}|b,0)}{\int_0^\infty \text{e}^{-c^2\mathcal{w}/2}p(\mathcal{w}|b,0)d\mathcal{w}}.
+\label{eq:pg3}
+\end{equation}
+
+To make use of P\'olya--Gamma data augmentation, we first write the full conditional distribution of $\boldsymbol\delta_k$ as the prior for $\boldsymbol\delta_k$ times the multinomial likelihood for all $z_i$.
+$$p(\boldsymbol\delta_k|\mathbf{z},\boldsymbol\delta_{h \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \pi_{ik}^{U_{ik}}(1-\pi_{ik})^{1-U_{ik}},$$
+where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, $U_{ik} = {1}_{z_i = k}$ is an indicator that subject $i$ belongs to cluster $k$, and $\pi_{ik}$ is defined as in Section 3.4.
+We can rewrite $\pi_{ik}$ as follows
+$$\pi_{ik} = P(U_{ik} = 1) = \frac{\text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}}}{1 + \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}}} = \frac{\text{e}^{\eta_{ik}}}{1 + \text{e}^{\eta_{ik}}}$$
+where ${c}_{ik} = \log \sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}$ and $\eta_{ik} = \mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}$.
+We note that the sum $\sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}$ includes the reference category, but since we fix $\boldsymbol\delta_K = \mathbf{0}$, we have $\text{e}^{\mathbf{w}_i^T \boldsymbol\delta_K} = 1$, and hence
+$$c_{ik} = \log \sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}} = \log \left ( 1 + \sum_{h \notin \{k,K \}} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}} \right )$$
+
+We can use the quantities to re-express the full conditionals for $\boldsymbol\delta_k$ as
+\begin{eqnarray}
+	p(\boldsymbol\delta_k|\mathbf{Z},\boldsymbol\delta_{h \ne k}) &\propto& p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \left (\frac{\text{e}^{\eta_{ik}}}{1 + \text{e}^{\eta_{ik}}} \right )^{U_{ik}} \left (\frac{1}{1 + \text{e}^{\eta_{ik}}} \right )^{1-U_{ik}} \nonumber\\ &=& p(\boldsymbol\delta_k) \prod_{i = 1}^n \frac{(\text{e}^{\eta_{ik}})^{U_{ik}}}{1 + \text{e}^{\eta_{ik}}} \label{eq:pglogit}
+\end{eqnarray}
+which we note is essentially a logistic regression likelihood.
+We thus apply this P\'olya--Gamma data augmentation scheme from Polson et al.
+(2013) to update each $\boldsymbol\delta_k$ ($k = 1,...,K-1$) one at a time based on the binary indicators $U_{ik}$.
+
 \subsection{Conditional MSN Imputation}
@@
 We assign the skewness coefficients $\boldsymbol\psi_k$ a conjugate $\text{N}_J(\mathbf{m}_k,\mathbf{P}_k)$ prior.
-However, the updates of $\boldsymbol\beta_k$ and $\boldsymbol\psi_k$ can be combined into one step by defining the $(Jp + J) \times 1$ vector $\boldsymbol\beta^*_k = (\boldsymbol\beta_k,\boldsymbol\psi_k)^T$ for which we assume a $\text{N}_{Jp+J}(\mathbf{b}^*_k,\mathbf{I}_{(p+1)} \otimes \boldsymbol\Sigma_k)$ prior.
-For the multinomial logit model component, the regression parameters $\boldsymbol\delta_k$ are assigned a $\text{N}_r(\mathbf{d}_k, \mathbf{S}_k)$ prior for $k = 1,...,K-1$, which is conditionally conjugate as described in Section (\textbf{what section}).
+In practice, the updates of $\boldsymbol\beta_k$ and $\boldsymbol\psi_k$ can be combined into one step by defining the $(Jp + J) \times 1$ vector $\boldsymbol\beta^*_k = (\boldsymbol\beta_k,\boldsymbol\psi_k)^T$ for which we assume a $\text{N}_{Jp+J}(\mathbf{b}^*_k,\mathbf{I}_{(p+1)} \otimes \boldsymbol\Sigma_k)$ prior.
 
+To make use of the Matrix Normal representation introduced previously in Section (\textbf{what section}), we define the matrix of regression parameters $\mathbb{B}^*_k$, where $\text{vec}(\mathbb{B}^*_k) = \boldsymbol\beta^*_k$.
+We assign $\mathbb{B}^*_k|\boldsymbol\Sigma_k$ a $\text{MatNorm}(\mathbf{B}^*_{0k},\mathbf{I}_{p+1},\boldsymbol\Sigma_k)$, where $\mathbf{B}^*_{0k}$ is a matrix of location parameters such that $\text{vec}(\mathbf{B}^*_{0k}) = \mathbf{b}^*_k$, $\mathbf{I}_{p+1}$ is the identity matrix of dimensions $(p+1)\times(p+1)$, and $\boldsymbol\Sigma_k$ is the model error.
+This prior structure allows for convinient Matrix Normal updates of the MSN regression parameters as described in Section (\textbf{what section}).
+
+For the multinomial logit model component, the regression parameters $\boldsymbol\delta_k = (\delta_{k1},...,\delta_{kr})$ are assigned a $\text{N}_r(\mathbf{d}_{0k}, \mathbf{S}_{0k})$ prior for $k = 1,...,K-1$, which is conditionally conjugate as described in Section (\textbf{what section}).
 We allow the normal-inverse-Wishart and multinomial hyperparameters to vary across clusters, though they may be shared across clusters in practice.
 An advantage of allowing for cluster-specific prior parameters is that \textit{a priori} knowledge of motor development trends can be incorporated into certain clusters while allowing the priors for other clusters to be less informative.
-Additionally, prior information regarding the effect of certain covariates on cluster membership can be incorporated in to the model by choosing informative values for $\mathbf{d}_k \ \text{and} \ \mathbf{S}_k$
+Additionally, prior information regarding the effect of certain covariates on cluster membership can be incorporated in to the model by choosing informative values for $\mathbf{d}_{0k} \ \text{and} \ \mathbf{S}_{0k}$
 
@@
 For cluster $k$, compute $A_k = (1 + \boldsymbol\psi_k^T \boldsymbol\Sigma^{-1}_k\boldsymbol
-\psi_k)^{-1}$ according to current values of $\boldsymbol\psi_k$ and $\boldsymbol\Sigma^{-1}_k$.
-Now, for $i = 1,...,n_k$, where $n_k$ is the number of members of cluster $k$ according to the current values of ${z}_1,...,{z}_n$, compute $a_i = A_k \boldsymbol\psi_k^T \boldsymbol\Sigma^{-1}_k(\mathbf{y}_i - \boldsymbol\zeta_{ik})$.
-Finally, update draw $t_i$ from $\text{N}_{[0,\infty)}(a_i, \mathbf{A}_k)$.
-\textbf{Derivations in FS-P 2010 supplement}.
+\psi_k)^{-1}$ according to current values of $\boldsymbol\psi_k$ and $\boldsymbol\Sigma_k$.
+Now, for $i = 1,...,n_k$, where $n_k$ is the number of members of cluster $k$ according to the current cluster labeling ${z}_1,...,{z}_n$, compute $a_i = A_k \boldsymbol\psi_k^T \boldsymbol\Sigma^{-1}_k(\mathbf{y}_i - \boldsymbol\zeta_{ik})$.
+Finally, update $t_i$ from $\text{N}_{[0,\infty)}(a_i, \mathbf{A}_k)$ (\textbf{put derivations in supplement}).
+Repeat these updates for $k = 1,...,K$
 
 The remaining MSN regression parameters are updated from their full conditionals as follows.
-For cluster $k$, we update $\boldsymbol\Sigma_k$ from $\text{IW}(\nu_k,\mathbf{V}_k)$, where $\nu_k = \nu_{0k} + n_k$ and
+For each cluster $k$ ($k = 1,...,K$) we update $\boldsymbol\Sigma_k$ from $\text{IW}(\nu_k,\mathbf{V}_k)$, where $\nu_k = \nu_{0k} + n_k$ and\begin{eqnarray}
+	\mathbf{V}_k &=&  (\mathbf{Y}-\mathbf{X}^*_k \mathbf{B}^*_k)^T (\mathbf{Y}_k - \mathbf{X}^*_k \mathbf{B}^*_k)\nonumber \\
+	 &+& (\mathbf{B}_k^* - \mathbf{B}^*_{0k})^T \mathbf{I}_{p+1}(\mathbf{B}^*_k - \mathbf{B}^*_{0k}) + \mathbf{V}_{0k}. \nonumber
+\end{eqnarray}We then make use of the Matrix Normal representation introduced in Section 3.2 to draw $\mathbb{B}^*_k$ from $\text{MatNorm}_{p+1,J}(\mathbf{B}_k^*,\mathbf{L}_k^{-1},\boldsymbol\Sigma_k)$, where $\text{vec}(\mathbb{B}^*_k) = \boldsymbol\beta^*_k = (\beta_{k11},...,\beta_{kJp},\psi_{k1},...,\psi_{kJ})^T$ and
 \begin{eqnarray}
-	\mathbf{V}_k &=&  (\mathbf{Y}-\mathbf{X}^*_k \mathbf{B}^*_k)^T (\mathbf{Y}_k - \mathbf{X}^*_k \mathbf{B}^*_k)\nonumber \\
-	 &+& (\mathbf{B}_k^* - \mathbf{B}_{0,k})^T \mathbf{I}_{p+1}(\mathbf{B}^*_k - \mathbf{B}_{0,k}) + \mathbf{V}_{0,k}. \nonumber
+	\mathbf{B}^*_k &=& (\mathbf{X}^{*T}_k \mathbf{X}^*_k + \mathbf{I}_{p+1})^{-1} (\mathbf{X}^{*T}_k \mathbf{Y}_k + \mathbf{I}_{p+1} \mathbf{B}_{0k}) \nonumber \\
+	\mathbf{L}^*_k &=& \mathbf{X}^{*T}_k \mathbf{X}^*_k + \mathbf{I}_{p+1}. \nonumber
 \end{eqnarray}
-We then make use of the Matrix Normal representation introduced in Section 3.2 to draw $\mathbb{B}^*_k$ from $\text{MatNorm}_{p+1,J}(\mathbf{B}_k^*,\mathbf{L}_k^{-1},\boldsymbol\Sigma_k)$
-\paragraph{\textbf{Step 3:}P\'olya--Gamma Data Augmentation for $z_i$} The sampler concludes by updating the latent cluster indicators $z_i$ $(i=1,\ldots,n)$ from its multinomial logit full conditional.
-To facilitate sampling, we adopt an efficient data-augmentation approach introduced by Polson \textit{et al.} (2013), which expresses the inverse-logit function as a mixture P\'olya--Gamma densities.
-By using P\'olya--Gamma data augmentation for the multinomial model (\textbf{cite}), we obtain a P\'olya--Gamma mixture of experts model, a computationally efficient way to model edge weights (\textbf{cite}).
-A random variable $w$ is said to follow a P\'olya--Gamma distribution with parameters $b > 0$ and $c \in {\rm I\!R}$ if
-\begin{equation}
-w \sim \text{PG}(b,c) \stackrel{d}{=} \frac{1}{2\pi^2}\sum_{s=1}^{\infty}\frac{g_s}{(s-1/2)^2 + c^2/(4\pi^2)}, \label{eq:pg1}
-\end{equation}
-where $g_s \stackrel{iid}{\sim} \text{Ga}(b,1)$ for $s = 1,...,\infty$ . Polson \textit{et al.} (2013) establish that for the $\text{PG}(b,c)$ density, and for $a,\eta \in {\rm I\!R}$,
 
-\begin{equation}
-\frac{(\text{e}^{\eta})^a}{(1 + \text{e}^{\eta})^b} = 2^{-b}\text{e}^{\kappa \eta} \int_{0}^{\infty} \text{e}^{-w\eta^2/2}p(w|b,c = 0)dw. \label{eq:pg2}
-\end{equation}
-where $\kappa = a - b/2$. Polson \textit{et al.} also show that the conditional distribution $p(w|b,c)$ results from an ``exponential tilting'' of the $\text{PG}(b,0)$ density, thus
+\paragraph{\textbf{Step 3:}P\'olya--Gamma Data Augmentation for $z_i$} The sampler concludes with updates of the multinomial regression parameters $\boldsymbol\delta_k$, for $k = 1,...,K-1$, followed by updates of each latent cluster indicator $z_i$ $(i=1,\ldots,n)$ from its multinomial logit full conditional.
+First, we define $U_{ik} = {1}_{z_i = k}$ for $i = 1,...,n$ and $k = 1,...,K-1$.
+Next, we update $\mathcal{w}_{ik}$ from a $\text{PG}(1,\eta_{ik})$ density, where $\eta_{ik} = \mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}$, and ${c}_{ik} = \log \sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}$.
+Next, define $\gamma_{ik} = \frac{u_{ik} - 1/2}{\mathcal{w}_{ik}}$ and let $\boldsymbol\gamma_k = (\gamma_{1k},...,\gamma_{nk})^T$.
+Finally, for $k = 1,...,K-1$, update $\boldsymbol\delta_k$ from a $\text{N}_r(\mathbf{d}_k,\mathbf{S}_k)$, where $\mathbf{S}_k = (\mathbf{S}_{0k} + \mathbf{W}^T \mathcal{W}_k \mathbf{W})^{-1}$, $\mathcal{W}_k = \text{Diag}(\mathcal{w}_{1k},...,\mathcal{w}_{nk})$,  $\mathbf{W} =  (\mathbf{w}^T_1,...,\mathbf{w}^T_n)^T$, and $\mathbf{d}_k = \mathbf{S}_k (\mathbf{S}_{0k}\mathbf{d}_{0k} + \mathbf{W}^T \mathcal{W}_k \boldsymbol\gamma_k)$.
 
-\begin{equation}
-	p(w|b,c) = \frac{\text{e}^{-c^2w/2}p(w|b,0)}{E_w[\text{e}^{-c^2w/2}]} = \frac{\text{e}^{-c^2w/2}p(w|b,0)}{\int_0^\infty \text{e}^{-c^2w/2}p(w|b,0)dw}. \label{eq:pg3}
-\end{equation}
-
-To make use of P\'olya--Gamma data augmentation, we first write the full conditional distribution of $\boldsymbol\delta_k$ as the prior for $\boldsymbol\delta_k$ times the multinomial likelihood for all $z_i$.
-$$p(\boldsymbol\delta_k|\mathbf{z},\boldsymbol\delta_{h \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \pi_{ik}^{U_{ik}}(1-\pi_{ik})^{1-U_{ik}},$$
-where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, $U_{ik} = {1}_{z_i = k}$ is an indicator that subject $i$ belongs to cluster $k$, and $\pi_{ik}$ is defined as in Section 3.4.
-We can rewrite $\pi_{ik}$ as follows
-$$\pi_{ik} = P(U_{ik} = 1) = \frac{\text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}}}{1 + \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}}} = \frac{\text{e}^{\eta_{ik}}}{1 + \text{e}^{\eta_{ik}}}$$
-where ${c}_{ik} = \log \sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}$ and $\eta_{ik} = \mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}$.
-We note that the sum $\sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}$ includes the reference category, but since we fix $\boldsymbol\delta_K = \mathbf{0}$, we have $\text{e}^{\mathbf{w}_i^T \boldsymbol\delta_K} = 1$, and hence
-$$c_{ik} = \log \sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}} = \log \left ( 1 + \sum_{h \notin \{k,K \}} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}} \right )$$
-
-We can use the quantities to re-express the full conditionals for $\boldsymbol\delta_k$ as
-\begin{eqnarray}
-	p(\boldsymbol\delta_k|\mathbf{Z},\boldsymbol\delta_{h \ne k}) &\propto& p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \left (\frac{\text{e}^{\eta_{ik}}}{1 + \text{e}^{\eta_{ik}}} \right )^{U_{ik}} \left (\frac{1}{1 + \text{e}^{\eta_{ik}}} \right )^{1-U_{ik}} \nonumber\\ &=& p(\boldsymbol\delta_k) \prod_{i = 1}^n \frac{(\text{e}^{\eta_{ik}})^{U_{ik}}}{1 + \text{e}^{\eta_{ik}}} \label{eq:pglogit}
-\end{eqnarray}
-which we note is essentially a logistic regression likelihood.
-We thus apply this P\'olya--Gamma data augmentation scheme to update each $\boldsymbol\delta_k$ ($k = 1,...,K-1$) one at a time based on the binary indicators $U_{ik}$.
+Lastly, we update $z_1,...,z_n$ by first computing 
+$$\pi_{ik} = \frac{\text{e}^{\eta_{ik}}}{1 + \sum_{h=1}^{K-1} \text{e}^{\eta_{ih}}},$$ 
+where $\eta_{ik} = \mathbf{w}_i\boldsymbol\delta_k$ for $i = 1,...n$ and $k = 1,...,K-1$.
+We also compute according to the multivariate Normal density of $\mathbf{y}_i|t_i$ the probability $P(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)$, where $\boldsymbol\zeta_{ki} = \mathbf{x}_i^*\boldsymbol\beta^*_k$.
+We use these quantities to compute $\boldsymbol\varrho_{i} = (\varrho_{1i},...,\varrho_{Ki})$, where
+$$\varrho_{ik} = P(z_i = k) = \frac{\pi_{ik}P(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)}{\sum_{h = 1}^{K} \pi_{ih}P(\mathbf{y}_i|\boldsymbol\zeta^*_{hi},\boldsymbol\Sigma_h)}$$.
+The cluster labels $z_i$ are updated from a $\text{Categorical}(\boldsymbol\varrho_i)$ density for $i = 1,...n$, thus concluding the Gibbs sampler.
 

### 2019-06-26 manuscript/manuscript.tex
section 3

@@
 First, we adopt the multivariate skew normal distribution with cluster-specific parameters that accommodate the inherent correlation and skewness in the data.
-Second, we model the cluster membership probabilities using a novel P\'olya-Gamma data-augmentation scheme, thereby improving predictions of the cluster membership allocations.
-Lastly, we impute missing responses under missing at random assumption by drawing from appropriate conditional skew normal distributions.
-Bayesian inference is achieved through straightforward Gibbs sampling, and can be carried out in available software such as R.
+Second, we model the cluster membership probabilities using a novel application of the P\'olya-Gamma data-augmentation scheme, thereby improving predictions of the cluster membership allocations.
+Lastly, we impute missing responses under missing at random assumption by drawing from appropriate conditional multivariate skew  normal distributions.
+Bayesian inference is achieved through straightforward Gibbs sampling, and can be implemented in available software such as \texttt{R}.
 Through simulation studies, we show that the proposed model yields improved inferences over models that ignore skewness.
@@
 Mixtures of multivariate non-symmetric distributions such as the multivariate skew-normal (MSN) distribution allow for the nuances of the marginal density to be captured with a more parsimonious set of mixture components.
-Mixtures of MSN distributions have been dealt with previously in a Bayesian context (Fr\"{u}hwirth-Schnatter \& Pyne, 2010), however in these models, focus lies primary on marginal density estimation and inference on the mixture components (i.e.
-clusters) is not discussed.
+Mixtures of MSN distributions have been dealt with previously in a Bayesian context (Fr\"{u}hwirth-Schnatter \& Pyne, 2010 and others), however in these models, focus lies primary on marginal density estimation and inference on the mixture components (i.e.
+clusters) is often not discussed.
 More recently, the mixtures of skew-$t$ factor analysis (MSTFA) model has been proposed for settings in which cluster-specific inference is of primary interest (Lin \textit{et al.} 2018).
@@
 for the MSTFA relies on a prohibitively complex EM algorithm and does not enjoy the inferential benefits of a Bayesian approach, namely the ability to incorporate prior information into a model and make posterior probability statements.
-Our proposed model improves on these previous works by estimating parameters in a Bayesian framework as well as including the ability to fit a multinomial logit regression to cluster membership probabilities using a novel application of data augmentation with the P\'olya Gamma distribution.
+Our proposed model improves on these previous works by estimating parameters in a Bayesian framework as well as including the ability to fit a multinomial logit regression to cluster membership probabilities using a novel application of data augmentation with the P\'olya-Gamma distribution.
 
@@
 where $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$ ($k = 1,...,K$) and $\pi_{ik}$ is a subject-specific mixing weight representing the probability that subject $i$ belongs to cluster $k$.
+A distinguishing feature of our model is that we treat $\pi_{ik}$ as a gating function (\textbf{cite book}) that depends on a subject's observed data as well as the parameters specific to the cluster to which they belong.
 For now we assume that $K$ is fixed; in Section 4, we discuss model selection strategies for choosing the optimal value of $K$.
@@
 The vector $\boldsymbol\alpha_k$ has components $\alpha_{kj}$, $j = 1,...,J$, that control the skewness of outcome $j$ in cluster $k$.
-When $\boldsymbol\alpha_k = \mathbf{0}$, the MSN distribution reduces to the multivariate normal distribution $\text{N}_J(\boldsymbol\zeta_k,\boldsymbol\Omega_k)$, where $\boldsymbol\zeta_k$ is a $J \times 1$ mean vector and $\boldsymbol\Omega_k$ is a $J \times J$ covariance matrix.
+When $\boldsymbol\alpha_k = \mathbf{0}$, the MSN distribution reduces to the multivariate normal distribution $\text{N}_J(\boldsymbol\zeta_k,\boldsymbol\Omega_k)$, where $\boldsymbol\zeta_k$ is a $J \times 1$ mean vector and $\boldsymbol\Omega_k$ is a $J \times J$ covariance matrix (Azzalini and Dalla Valle, 1996).
 
@@
 Under this model, $z_i|\mathbf{w}_i \sim \text{Multinom}(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{i1},...,\pi_{1K})$.
-During MCMC estimation, the cluster labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as cluster assignments.
+During MCMC estimation, the cluster labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as cluster assignments as detailed later in section (\textbf{what section}).
 
@@
 \boldsymbol\mu^{cond}_{ki}&=& \boldsymbol\mu^{miss}_{ki} + \boldsymbol\Sigma_{12} \boldsymbol\Sigma_{22}^{-1}(\mathbf{y}^{obs}_i - \boldsymbol\mu^{obs}_{ki}) \label{eq:impute}\\
-\boldsymbol\Sigma^{cond}_k&=& \boldsymbol\Sigma_{11} - \boldsymbol\Sigma_{12}\boldsymbol\Sigma_{22}^{-1}\boldsymbol\Sigma_{21},~\text{and}\nonumber \\
+\boldsymbol\Sigma^{cond}_k&=& \boldsymbol\Sigma_{11} - \boldsymbol\Sigma_{12}\boldsymbol\Sigma_{22}^{-1}\boldsymbol\Sigma_{21},\nonumber \\
+\boldsymbol\mu_{ki} &=& \begin{pmatrix} \boldsymbol\mu_{ki}^{miss} \\ \boldsymbol\mu_{ki}^{obs} \end{pmatrix}, \ \text{and} \nonumber \\
 \boldsymbol\Sigma_k &=& \begin{pmatrix} 
@@
 The remaining partitions $\boldsymbol\Sigma_{k21}, \ \text{and} \ \boldsymbol\Sigma_{k22}$ are defined in the same manner.
-These results follow from conventional multivariate normal theory.
+Likewise, $\boldsymbol\mu_{ki} = \mathbf{X}_i \boldsymbol\beta_k$ is partitioned into $\boldsymbol\mu_{ki}^{miss}$ and $\boldsymbol\mu_{ki}^{obs}$ with respect to the missing and observed indices of $\mathbf{y}_i$, respectively.
+These results follow from conventional multivariate normal theory, which we detail in the Web Appendix.
 An attractive feature of this imputation algorithm is that it avoids multiplicative run-time scaling in $m$, the number of imputations (Gelman \textit{et al.} 2013; Zhou and Reiter, 2010).
 Our approach also provides more opportunities to explore the missing data parameter space than does multiple imputation, since missing values are drawn at each MCMC iteration, and often in practice $n_{sim} >> m$, where $n_{sim}$ is the number of MCMC iterations (\textbf{find a reference}).
-In Section 4, we conduct simulation studies to demonstrate that imputing the missing MSN responses improves inferences over complete case analysis.
+In Section 4, we conduct simulation studies to demonstrate that imputing the missing MSN responses improves inferences over available-case analysis.
 
@@
 We assign the skewness coefficients $\boldsymbol\psi_k$ a conjugate $\text{N}_J(\mathbf{m}_k,\mathbf{P}_k)$ prior.
-In practice, the updates of $\boldsymbol\beta_k$ and $\boldsymbol\psi_k$ can be combined into one step by defining the $(Jp + J) \times 1$ vector $\boldsymbol\beta^*_k = (\boldsymbol\beta_k,\boldsymbol\psi_k)^T$ for which we assume a $\text{N}_{Jp+J}(\mathbf{b}^*_k,\mathbf{I}_{(p+1)} \otimes \boldsymbol\Sigma_k)$ prior.
+In practice, the updates of $\boldsymbol\beta_k$ and $\boldsymbol\psi_k$ can be combined into one step by defining the $(Jp + J) \times 1$ vector $\boldsymbol\beta^*_k = (\boldsymbol\beta_k^T,\boldsymbol\psi_k^T)^T$ for which we assume a $\text{N}_{Jp+J}(\mathbf{b}^*_k,\mathbf{I}_{(p+1)} \otimes \boldsymbol\Sigma_k)$ prior.
 
 To make use of the Matrix Normal representation introduced previously in Section (\textbf{what section}), we define the matrix of regression parameters $\mathbb{B}^*_k$, where $\text{vec}(\mathbb{B}^*_k) = \boldsymbol\beta^*_k$.
-We assign $\mathbb{B}^*_k|\boldsymbol\Sigma_k$ a $\text{MatNorm}(\mathbf{B}^*_{0k},\mathbf{I}_{p+1},\boldsymbol\Sigma_k)$, where $\mathbf{B}^*_{0k}$ is a matrix of location parameters such that $\text{vec}(\mathbf{B}^*_{0k}) = \mathbf{b}^*_k$, $\mathbf{I}_{p+1}$ is the identity matrix of dimensions $(p+1)\times(p+1)$, and $\boldsymbol\Sigma_k$ is the model error.
+We assign $\mathbb{B}^*_k|\boldsymbol\Sigma_k$ a $\text{MatNorm}(\mathbf{B}^*_{0k},\mathbf{I}_{p+1},\boldsymbol\Sigma_k)$ prior, where $\mathbf{B}^*_{0k}$ is a matrix of location parameters such that $\text{vec}(\mathbf{B}^*_{0k}) = \mathbf{b}^*_k$, $\mathbf{I}_{p+1}$ is the identity matrix of dimensions $(p+1)\times(p+1)$, and $\boldsymbol\Sigma_k$ is the model error.
 This prior structure allows for convinient Matrix Normal updates of the MSN regression parameters as described in Section (\textbf{what section}).
@@
 This is accomplished by drawing $\mathbf{y}_i^{miss}$ from $\text{N}_{J-q_i}(\boldsymbol\mu^{cond}_{ki},\boldsymbol\Sigma^{cond}_k)$ as described in equation \ref{eq:impute} for all $i = 1,...,n$.
-Conditional imputation is completed with construction of a complete outcome vector $\mathbf{y}_i$ by merging the now imputed $\mathbf{y}_i^{miss}$ with $\mathbf{y}_i^{obs}$ with respect for the original order of $\mathbf{y}_i$.
+Conditional imputation concludes with construction of a complete outcome vector $\mathbf{y}_i$ by merging the now imputed $\mathbf{y}_i^{miss}$ with $\mathbf{y}_i^{obs}$ with respect for the original order of $\mathbf{y}_i$.
 
@@
 \psi_k)^{-1}$ according to current values of $\boldsymbol\psi_k$ and $\boldsymbol\Sigma_k$.
-Now, for $i = 1,...,n_k$, where $n_k$ is the number of members of cluster $k$ according to the current cluster labeling ${z}_1,...,{z}_n$, compute $a_i = A_k \boldsymbol\psi_k^T \boldsymbol\Sigma^{-1}_k(\mathbf{y}_i - \boldsymbol\zeta_{ik})$.
+Now, for $i = 1,...,n_k$, where $n_k$ is the number of members of cluster $k$ according to the current cluster labeling ${z}_1,...,{z}_n$, compute $a_i = A_k \boldsymbol\psi_k^T \boldsymbol\Sigma^{-1}_k(\mathbf{y}_i - \boldsymbol\zeta_{ik})$, where $\boldsymbol\zeta_{ik} = X_i \boldsymbol\beta_k$.
 Finally, update $t_i$ from $\text{N}_{[0,\infty)}(a_i, \mathbf{A}_k)$ (\textbf{put derivations in supplement}).
@@
 	 &+& (\mathbf{B}_k^* - \mathbf{B}^*_{0k})^T \mathbf{I}_{p+1}(\mathbf{B}^*_k - \mathbf{B}^*_{0k}) + \mathbf{V}_{0k}. \nonumber
-\end{eqnarray}We then make use of the Matrix Normal representation introduced in Section 3.2 to draw $\mathbb{B}^*_k$ from $\text{MatNorm}_{p+1,J}(\mathbf{B}_k^*,\mathbf{L}_k^{-1},\boldsymbol\Sigma_k)$, where $\text{vec}(\mathbb{B}^*_k) = \boldsymbol\beta^*_k = (\beta_{k11},...,\beta_{kJp},\psi_{k1},...,\psi_{kJ})^T$ and
+\end{eqnarray}We then make use of the Matrix Normal representation introduced in Section 3.2 to draw $\mathbb{B}^*_k$ from a $\text{MatNorm}_{p+1,J}(\mathbf{B}_k^*,\mathbf{L}_k^{-1},\boldsymbol\Sigma_k)$ density, where $\text{vec}(\mathbb{B}^*_k) = \boldsymbol\beta^*_k = (\beta_{k11},...,\beta_{kJp},\psi_{k1},...,\psi_{kJ})^T$ and
 \begin{eqnarray}
@@
 Next, define $\gamma_{ik} = \frac{u_{ik} - 1/2}{\mathcal{w}_{ik}}$ and let $\boldsymbol\gamma_k = (\gamma_{1k},...,\gamma_{nk})^T$.
-Finally, for $k = 1,...,K-1$, update $\boldsymbol\delta_k$ from a $\text{N}_r(\mathbf{d}_k,\mathbf{S}_k)$, where $\mathbf{S}_k = (\mathbf{S}_{0k} + \mathbf{W}^T \mathcal{W}_k \mathbf{W})^{-1}$, $\mathcal{W}_k = \text{Diag}(\mathcal{w}_{1k},...,\mathcal{w}_{nk})$,  $\mathbf{W} =  (\mathbf{w}^T_1,...,\mathbf{w}^T_n)^T$, and $\mathbf{d}_k = \mathbf{S}_k (\mathbf{S}_{0k}\mathbf{d}_{0k} + \mathbf{W}^T \mathcal{W}_k \boldsymbol\gamma_k)$.
+Finally, for $k = 1,...,K-1$, update $\boldsymbol\delta_k$ from a $\text{N}_r(\mathbf{d}_k,\mathbf{S}_k)$, where $\mathbf{S}_k = (\mathbf{S}_{0k} + \mathbf{W}^T \mathcal{W}_k \mathbf{W})^{-1}$, $\mathcal{W}_k = \text{Diag}(\mathcal{w}_{1k},...,\mathcal{w}_{nk})$,  $\mathbf{d}_k = \mathbf{S}_k (\mathbf{S}_{0k}\mathbf{d}_{0k} + \mathbf{W}^T \mathcal{W}_k \boldsymbol\gamma_k)$, and $\mathbf{W}$ is the $n \times r$ matrix of multinomial logit regression covariates such that the $i^{th}$ row of $\mathbf{W}$ is $\mathbf{w}_i$.
 
-Lastly, we update $z_1,...,z_n$ by first computing 
+Lastly, we update $z_1,...,z_n$ by first computing $\boldsymbol\pi_i = (\pi_{1i},...,\pi_{Ki})$ as
 $$\pi_{ik} = \frac{\text{e}^{\eta_{ik}}}{1 + \sum_{h=1}^{K-1} \text{e}^{\eta_{ih}}},$$ 
-where $\eta_{ik} = \mathbf{w}_i\boldsymbol\delta_k$ for $i = 1,...n$ and $k = 1,...,K-1$.
+where $\eta_{ik} = \mathbf{w}_i\boldsymbol\delta_k$ for $i = 1,...n$ and $k = 1,...,K$.
 We also compute according to the multivariate Normal density of $\mathbf{y}_i|t_i$ the probability $P(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)$, where $\boldsymbol\zeta_{ki} = \mathbf{x}_i^*\boldsymbol\beta^*_k$.
 We use these quantities to compute $\boldsymbol\varrho_{i} = (\varrho_{1i},...,\varrho_{Ki})$, where
-$$\varrho_{ik} = P(z_i = k) = \frac{\pi_{ik}P(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)}{\sum_{h = 1}^{K} \pi_{ih}P(\mathbf{y}_i|\boldsymbol\zeta^*_{hi},\boldsymbol\Sigma_h)}$$.
-The cluster labels $z_i$ are updated from a $\text{Categorical}(\boldsymbol\varrho_i)$ density for $i = 1,...n$, thus concluding the Gibbs sampler.
+$$\varrho_{ik} = P(z_i = k|\mathbf{y}_i,\boldsymbol\zeta^*_i,\boldsymbol\Sigma_k) = \frac{\pi_{ik}P(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)}{\sum_{h = 1}^{K} \pi_{ih}P(\mathbf{y}_i|\boldsymbol\zeta^*_{hi},\boldsymbol\Sigma_h)}$$.
+The cluster labels $z_i$ are updated from a $\text{Multinomial}(1,\boldsymbol\varrho_i)$ density for $i = 1,...n$, thus concluding the Gibbs sampler.
 

### 2019-06-27 manuscript/manuscript.tex
section 3

@@
 \begin{equation}
-f(\mathbf{y}_i) = \sum_{k = 1}^{K} \pi_{ik} f(\mathbf{y}_i|\boldsymbol\theta_k),
+f(\mathbf{y}_i) = \sum_{k = 1}^{K} \pi_{ki} f(\mathbf{y}_i|\boldsymbol\theta_k),
 \end{equation}
@@
 \subsection{Simulation to Compare to Multivariate Normal}
+
+Our first simulation study will be concerned with comarison of the MSN mixture model to a standard multivariate normal mixture model.
+Here, the goal is to investigate to what degree allowing for skewness in outcome components reduces the number of mixture components necessary to estimate the marginal density.
+For the purposes of illustration, we consider the following generative model
+\begin{eqnarray}
+	f(\mathbf{y}_i) = \pi_{1i}f(\mathbf{y}_i|\boldsymbol\theta_1) + \pi_{2i}f(\mathbf{y}_i|\boldsymbol\theta_2) + \pi_{3i}f(\mathbf{y}_i|\boldsymbol\theta_3), \nonumber
+\end{eqnarray}
+where $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$, for $k \in \{1,2,3\}$, and $\mathbf{y}_i | \boldsymbol\theta_k \sim \N$
 

### 2019-06-27 manuscript/manuscript.tex
section 3

@@
 
-See Table \ref{alg:MCMC}.
+See Algorithm \ref{alg:MCMC}.
 
@@
 
+\subsubsection{Model Fit Diagnostics and Specification of K}
+
 \newpage
@@
 Our first simulation study will be concerned with comarison of the MSN mixture model to a standard multivariate normal mixture model.
-Here, the goal is to investigate to what degree allowing for skewness in outcome components reduces the number of mixture components necessary to estimate the marginal density.
+Here, the primary goal is to validate our parameter estimation scheme in a setting that resembles our application to the Nurture data.
+Our secondary goal is to investigate to what degree allowing for skewness in outcome components reduces the number of mixture components necessary to estimate the marginal density.
 For the purposes of illustration, we consider the following generative model
@@
 \end{eqnarray}
-where $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$, for $k \in \{1,2,3\}$, and $\mathbf{y}_i | \boldsymbol\theta_k \sim \N$
+where $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$, for $k \in \{1,2,3 \}$, and $\mathbf{y}_i | \boldsymbol\theta_k \sim \text{MSN}_4(\boldsymbol\zeta_{ki},\boldsymbol\alpha_k,\boldsymbol\Omega_k)$, where $\boldsymbol\zeta_{ki} = (\zeta_{ki1},\zeta_{ki2},\zeta_{ki3},\zeta_{ki4}) = \mathbf{X}_i \boldsymbol\beta_{k}$.
+We note from the above model specification that the number of clusters $K = 3$ and the number of measurement occasions $J = 4$.
+For the MSN regression model, we fit a main effect for time (i.e.
+measurement occasion) in addition to a baseline covariate, whose value does not vary with time, but whose effect does vary with time (\textbf{is this a weird way to say this?}).
+Under this setting with $p = 2$ covariates, the design matrix $\mathbf{X}_i$ for subject $i$ and the matrix of regression coefficients for cluster $k$, $\mathbf{B}_k$ are structured as
+
+\begin{eqnarray}
+	\mathbf{X}_i = 
+	\begin{bmatrix}
+		1 & 0 & 0 & 0 & x_i & 0 & 0 & 0\\
+		0 & 1 & 0 & 0 & 0 & x_i & 0 & 0\\
+		0 & 0 & 1 & 0 & 0 & 0 & x_i & 0\\
+		0 & 0 & 0 & 1 & 0 & 0 & 0 & x_i\\
+	\end{bmatrix}, \ \text{and} \ 
+	\mathbf{B}_k = 
+	\begin{bmatrix}
+		\beta_{k11} & \beta_{k12} \\
+		\beta_{k21} & \beta_{k22} \\
+		\beta_{k31} & \beta_{k32} \\
+		\beta_{k41} & \beta_{k42} \\
+	\end{bmatrix}, \nonumber
+\end{eqnarray}
+where $\text{vec}({\mathbf{B}_k}) = \boldsymbol\beta_k$. Thus, for subject $i$ in cluster $k$, the $J = 4$ measurements have location parameters given by
+
+\begin{eqnarray}
+	\zeta_{ki1} &=& \beta_{k11} + \beta_{k12}x_i \nonumber \\
+	\zeta_{ki2} &=& \beta_{k21} + \beta_{k22}x_i \nonumber \\
+	\zeta_{ki3} &=& \beta_{k31} + \beta_{k32}x_i \nonumber \\
+	\zeta_{ki4} &=& \beta_{k41} + \beta_{k42}x_i. \nonumber \\
+\end{eqnarray}
+
+For cluster $k$, we can collect all observations $\mathbf{y}_i$ ($i = 1,...,n_k$) into the $n_k \times 4$ matrix $\mathbf{Y}_k$ and all covariates $\mathbf{X}_i$ into the $4n_k \times 8$ matrix $\mathbf{X}_k$ (\textbf{clash of notation here}) to utilize the matrix Normal model specification detailed in Section 3, which allows for simultaneous updates of all MSN regression parameters in cluster $k$.
+For the multinomial regression model component of this simulation, we model the class labels $z_i$ as a function of an intercept and one baseline covariate, thus $r = 2$.
+
+For our second goal in this simulation, to compare model performance when skewness is accounted for skewness to model performance when skewness is ignored, we fit both the proposed MSN finite mixture model as well as a classic multivariate Normal finite mixture model to data generated from an MSN distribution.
+We evaluate model fit in terms of absolute difference of parameter estimates as well as WAIC.
 

### 2019-06-27 simulation/simulation.R

@@ -26,9 +26,9 @@ set.seed(seed)
 options(scipen=999)
 # Generate data
 k <- 4 # number of outcomes
-p <- 1 # number of predictors 
+p <- 2 # number of predictors 
 h <- 3 # number of clusters
-n <- 10000 # number of observations (takes alot of data to estimate this many parameters)
+n <- 1000 # number of observations (takes alot of data to estimate this many parameters)
 print(paste("Generating data with",n,"observations of dimension",k,
             "with",p,"covariates and",h,"clusters"))
 
@@ -64,7 +64,7 @@ Xstar <- cbind(X,t) # Combine X and t
 #w2 <- rnorm(n) # continuous predictor
 w3 <- rbinom(n,1,0.5) # binary predictor
 w4 <- rbinom(n,1,0.5) # binary predictor
-W <- cbind(w3) # design matrix
+W <- cbind(1,w3) # design matrix
 v <- ncol(W) # number of multinomial predictors
 delta.true <- matrix(rtruncnorm(n = v*(h-1),
                                 a = -1, 
@@ -155,8 +155,8 @@ delta0 <- rep(0,v) # prior mean for delta coefficients (multinomial regression)
 S0 <- diag(0.5,v) # prior covariance for delta coefficients (multinomial regression)
 
 # Sample storage
-nsim <- 10000 # number of iterations
-burn <- 2500 # number of iterations to save
+nsim <- 1000 # number of iterations
+burn <- 250 # number of iterations to save
 n.iter <- nsim - burn # number of saved iterations
 Z <- matrix(0,nrow = n.iter,ncol = n) # large matrix where each row is the value of z at a specific iteration
 PI <- matrix(0,nrow = n.iter,ncol = h) # matrix w/ each row as pi vector at each iteration

### 2019-06-30 manuscript/manuscript.tex
section 3

@@
 \end{equation}
-where $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$ ($k = 1,...,K$) and $\pi_{ik}$ is a subject-specific mixing weight representing the probability that subject $i$ belongs to cluster $k$.
-A distinguishing feature of our model is that we treat $\pi_{ik}$ as a gating function (\textbf{cite book}) that depends on a subject's observed data as well as the parameters specific to the cluster to which they belong.
+where $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$ ($k = 1,...,K$) and $\pi_{ki}$ is a subject-specific mixing weight representing the probability that subject $i$ belongs to cluster $k$.
 For now we assume that $K$ is fixed; in Section 4, we discuss model selection strategies for choosing the optimal value of $K$.
@@
 The vector $\boldsymbol\alpha_k$ has components $\alpha_{kj}$, $j = 1,...,J$, that control the skewness of outcome $j$ in cluster $k$.
-When $\boldsymbol\alpha_k = \mathbf{0}$, the MSN distribution reduces to the multivariate normal distribution $\text{N}_J(\boldsymbol\zeta_k,\boldsymbol\Omega_k)$, where $\boldsymbol\zeta_k$ is a $J \times 1$ mean vector and $\boldsymbol\Omega_k$ is a $J \times J$ covariance matrix (Azzalini and Dalla Valle, 1996).
+When $\boldsymbol\alpha_k = \mathbf{0}$, the MSN distribution reduces to the multivariate normal distribution $\text{N}_J(\boldsymbol\zeta_{ki},\boldsymbol\Omega_k)$, where $\boldsymbol\zeta_{ki}$ is a $J \times 1$ mean vector and $\boldsymbol\Omega_k$ is a $J \times J$ covariance matrix (Azzalini and Dalla Valle, 1996).
 
@@
 During MCMC estimation, the cluster labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as cluster assignments as detailed later in section (\textbf{what section}).
+By allowing the cluster probabilities to vary across subjects, our model can be viewed as a \textit{mixture of experts} model in which $\pi_{ki}$ acts as a \textit{gating function} (\textbf{cite book}).
 
@@
 By using P\'olya--Gamma data augmentation for the multinomial model (\textbf{cite}), we obtain a P\'olya--Gamma mixture of experts model, a computationally efficient way to model edge weights (\textbf{cite}).
-A random variable $\mathcal{w}$ is said to follow a P\'olya--Gamma distribution with parameters $b > 0$ and $c \in {\rm I\!R}$ if
+A random variable ${w}$ is said to follow a P\'olya--Gamma distribution with parameters $b > 0$ and $c \in {\rm I\!R}$ if
 \begin{equation}
-\mathcal{w} \sim \text{PG}(b,c) \stackrel{d}{=} \frac{1}{2\pi^2}\sum_{s=1}^{\infty}\frac{g_s}{(s-1/2)^2 + c^2/(4\pi^2)}, \label{eq:pg1}
+{w} \sim \text{PG}(b,c) \stackrel{d}{=} \frac{1}{2\pi^2}\sum_{s=1}^{\infty}\frac{g_s}{(s-1/2)^2 + c^2/(4\pi^2)}, \label{eq:pg1}
 \end{equation}
@@
 \begin{equation}
-\frac{(\text{e}^{\eta})^a}{(1 + \text{e}^{\eta})^b} = 2^{-b}\text{e}^{\kappa \eta} \int_{0}^{\infty} \text{e}^{-\mathcal{w}\eta^2/2}p(\mathcal{w}|b,c = 0)d\mathcal{w}. \label{eq:pg2}
+\frac{(\text{e}^{\eta})^a}{(1 + \text{e}^{\eta})^b} = 2^{-b}\text{e}^{\kappa \eta} \int_{0}^{\infty} \text{e}^{-{w}\eta^2/2}p({w}|b,c = 0)d{w}. \label{eq:pg2}
 \end{equation}
-where $\kappa = a - b/2$. Polson \textit{et al.} also show that the conditional distribution $p(\mathcal{w}|b,c)$ results from an ``exponential tilting'' of the $\text{PG}(b,0)$ density, thus
+where $\kappa = a - b/2$. Polson \textit{et al.} also show that the conditional distribution $p(w|b,c)$ results from an ``exponential tilting'' of the $\text{PG}(b,0)$ density, thus
 
 \begin{equation}
-	p(\mathcal{w}|b,c) = \frac{\text{e}^{-c^2\mathcal{w}/2}p(\mathcal{w}|b,0)}{E_{\mathcal{w}}[\text{e}^{-c^2\mathcal{w}/2}]} = \frac{\text{e}^{-c^2\mathcal{w}/2}p(\mathcal{w}|b,0)}{\int_0^\infty \text{e}^{-c^2\mathcal{w}/2}p(\mathcal{w}|b,0)d\mathcal{w}}.
-\label{eq:pg3}
+	p({w}|b,c) = \frac{\text{e}^{-c^2{w}/2}p({w}|b,0)}{E_{{w}}[\text{e}^{-c^2{w}/2}]} = \frac{\text{e}^{-c^2{w}/2}p({w}|b,0)}{\int_0^\infty \text{e}^{-c^2{w}/2}p({w}|b,0)d{w}}. \label{eq:pg3}
 \end{equation}
@@
 \begin{eqnarray}
-	p(\boldsymbol\delta_k|\mathbf{Z},\boldsymbol\delta_{h \ne k}) &\propto& p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \left (\frac{\text{e}^{\eta_{ik}}}{1 + \text{e}^{\eta_{ik}}} \right )^{U_{ik}} \left (\frac{1}{1 + \text{e}^{\eta_{ik}}} \right )^{1-U_{ik}} \nonumber\\ &=& p(\boldsymbol\delta_k) \prod_{i = 1}^n \frac{(\text{e}^{\eta_{ik}})^{U_{ik}}}{1 + \text{e}^{\eta_{ik}}} \label{eq:pglogit}
+	p(\boldsymbol\delta_k|\mathbf{U}_k,\boldsymbol\delta_{h \ne k}) &\propto& p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \left (\frac{\text{e}^{\eta_{ik}}}{1 + \text{e}^{\eta_{ik}}} \right )^{U_{ik}} \left (\frac{1}{1 + \text{e}^{\eta_{ik}}} \right )^{1-U_{ik}} \nonumber\\ &=& p(\boldsymbol\delta_k) \prod_{i = 1}^n \frac{(\text{e}^{\eta_{ik}})^{U_{ik}}}{1 + \text{e}^{\eta_{ik}}} \label{eq:pglogit}
 \end{eqnarray}
@@
 
+We model fit diagnostics, we choose the WAIC statistic introduced by Watanabe (2010).
+To specify the number of cluster $K$, we follow Neelon and Chung (2016), who use WAIC to determine the optimal number of latent factors in a Bayesian latent factor zero-inflated Poisson model for the analysis of correlated and zero-inflated count data.
+In Section 6, we discuss possible extentions to other methods for choosing $K$.
+
 \newpage
@@
 For the multinomial regression model component of this simulation, we model the class labels $z_i$ as a function of an intercept and one baseline covariate, thus $r = 2$.
+We leave the problem of missing data to simulation study \#2 and do not incorporate missing data into this simulation.
 
@@
 
-\begin{table}[t]
+\subsection{Simulation to Compare Imputation Methods}
 
-\caption{\label{tab:unnamed-chunk-4}Model results for simulated data with n = 1500, k = 4, p = 1, h = 3, v = 1.
-5000 iterations were run with a burn in of 1000.
-Missingness mechanism was MAR and P(miss) = 0}
-\centering
-\fontsize{8}{10}\selectfont
-\begin{tabular}{llllllll}
-\toprule
-\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Class 1} & \multicolumn{2}{c}{Class 2} & \multicolumn{2}{c}{Class 3} \\
-\cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6} \cmidrule(l{3pt}r{3pt}){7-8}
-Model Component & Parameter & True & Est. (95\% CrI) & True & Est. (95\% CrI) & True & Est. (95\% CrI)\\
-\midrule
-\addlinespace[0.3em]
-\multicolumn{8}{l}{\textbf{ }}\\
-\hspace{1em}MVSN & $\beta_{0}$ & -3.21 & -3.34 (-3.8, -2.99) & -0.32 & -0.33 (-0.48, -0.14) & 3.35 & 3.33 (3.22, 3.44)\\
-\hspace{1em}Regression & $\beta_{1}$ & -3.08 & -3.3 (-3.73, -2.87) & -0.75 & -0.72 (-0.87, -0.52) & 2.6 & 2.5 (2.39, 2.6)\\
-\hspace{1em} & $\beta_{2}$ & -2.97 & -3.18 (-3.58, -2.76) & -0.45 & -0.44 (-0.58, -0.26) & 3.43 & 3.42 (3.31, 3.53)\\
-\hspace{1em} & $\beta_{3}$ & -2.91 & -3.08 (-3.49, -2.68) & -0.66 & -0.68 (-0.83, -0.48) & 3.04 & 2.98 (2.87, 3.09)\\
-\addlinespace[0.3em]
-\multicolumn{8}{l}{\textbf{ }}\\
-\hspace{1em} & $\sigma_{11}$ & 1 & 0.95 (0.84, 1.02) & 1 & 1 (0.89, 1.11) & 1 & 1.06 (0.97, 1.16)\\
-\hspace{1em} & $\sigma_{12}$ & 0.74 & 0.7 (0.59, 0.76) & 0.68 & 0.68 (0.59, 0.78) & -0.45 & -0.41 (-0.47, -0.36)\\
-\hspace{1em} & $\sigma_{13}$ & 0.74 & 0.69 (0.58, 0.75) & -0.16 & -0.13 (-0.2, -0.06) & 0.82 & 0.88 (0.79, 0.97)\\
-\hspace{1em} & $\sigma_{14}$ & 0.98 & 0.93 (0.81, 0.99) & 0.64 & 0.65 (0.56, 0.75) & 0.7 & 0.75 (0.67, 0.83)\\
-\hspace{1em} & $\sigma_{22}$ & 1 & 0.94 (0.82, 1.01) & 1 & 1.03 (0.93, 1.13) & 1 & 1.07 (0.99, 1.16)\\
-\hspace{1em} & $\sigma_{23}$ & 0.83 & 0.79 (0.67, 0.85) & -0.43 & -0.4 (-0.46, -0.34) & -0.66 & -0.62 (-0.68, -0.57)\\
-\hspace{1em} & $\sigma_{24}$ & 0.81 & 0.77 (0.66, 0.83) & 0.63 & 0.67 (0.58, 0.77) & 0.01 & 0.07 (0.01, 0.13)\\
-\hspace{1em} & $\sigma_{33}$ & 1 & 0.96 (0.84, 1.03) & 1 & 1 (0.91, 1.09) & 1 & 1.05 (0.96, 1.15)\\
-\hspace{1em} & $\sigma_{34}$ & 0.85 & 0.81 (0.69, 0.87) & 0.15 & 0.15 (0.08, 0.23) & 0.59 & 0.64 (0.56, 0.72)\\
-\hspace{1em} & $\sigma_{44}$ & 1 & 0.95 (0.83, 1.01) & 1 & 1.02 (0.92, 1.13) & 1 & 1.06 (0.97, 1.15)\\
-\addlinespace[0.3em]
-\multicolumn{8}{l}{\textbf{ }}\\
-\hspace{1em} & $\psi_{1}$ & -0.33 & -0.33 (-0.62, 0.69) & 0.67 & 0.7 (0.46, 0.89) & -1 & -1.01 (-1.13, -0.87)\\
-\hspace{1em} & $\psi_{2}$ & -0.33 & -0.32 (-0.61, 0.64) & 0.67 & 0.63 (0.38, 0.82) & -1 & -0.88 (-1.01, -0.75)\\
-\hspace{1em} & $\psi_{3}$ & -0.33 & -0.33 (-0.61, 0.69) & 0.67 & 0.64 (0.43, 0.82) & -1 & -1.01 (-1.14, -0.88)\\
-\hspace{1em} & $\psi_{4}$ & -0.33 & -0.31 (-0.63, 0.67) & 0.67 & 0.7 (0.45, 0.89) & -1 & -0.94 (-1.07, -0.81)\\
-\addlinespace[0.3em]
-\multicolumn{8}{l}{\textbf{ }}\\
-\hspace{1em}Multinom. & $\delta_{11}$ & 0.9 & 0.88 (0.81, 0.95) & 0.9 & 0.88 (0.81, 0.95) & 0.9 & 0.88 (0.81, 0.95)\\
-\hspace{1em} & $\delta_{12}$ & 0.23 & 0.22 (0.14, 0.3) & 0.23 & 0.22 (0.14, 0.3) & 0.23 & 0.22 (0.14, 0.3)\\
-\addlinespace[0.3em]
-\multicolumn{8}{l}{\textbf{ }}\\
-\hspace{1em}Clustering & $\pi_l$ & 0.28 & 0.28 (0.27, 0.28) & 0.42 & 0.43 (0.42, 0.43) & 0.3 & 0.3 (0.3, 0.3)\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\subsection{Simulation to Compare Imputation Methods}
+In our second simulation study, we compare our online conditional multivariate normal imputation method to standard Bayesian multiple imputation.
 
@@
 
+\bibitem{ } Neelon B, Chung D. The LZIP: A Bayesian latent factor model for correlated zero‐inflated counts. Biometrics. 2017 Mar;73(1):185-96.
+
 \bibitem{ } Panagiotis Papastamoulis (2016).

### 2019-06-30 mcmc_draws_2019-06-27/tables.Rmd
section 3

--- 2019-06-14 mcmc_draws_2019-06-11/tables.Rmd
@@
 author: "Carter Allen"
-date: "4/11/2019"
+date: 
 output: 
@@
 ```{r}
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-06-27")
 load("BETA")
@@
 
-model_component <- c("MVSN","Regression","","","","","","","",
-                     "Multinom.","",
+model_component <- c("MVSN","Regression","","","","","","","","","","","","","","","","","","","","",
+                     "Multinom.","","","",
                      "Clustering")
-params <- c("$\\beta_{0}$","$\\beta_{1}$","$\\beta_{2}$","$\\beta_{3}$",
-            "$\\sigma_{11}$","$\\sigma_{12}$",
-            "$\\sigma_{22}$",
-            "$\\psi_{1}$","$\\psi_{2}$",
+params <- c("$\\beta_{11}$","$\\beta_{21}$","$\\beta_{31}$","$\\beta_{41}$",
+            "$\\beta_{12}$","$\\beta_{22}$","$\\beta_{32}$","$\\beta_{42}$",
+            "$\\sigma_{11}$","$\\sigma_{12}$","$\\sigma_{13}$","$\\sigma_{14}$",
+            "$\\sigma_{22}$","$\\sigma_{23}$","$\\sigma_{24}$",
+            "$\\sigma_{33}$","$\\sigma_{34}$",
+            "$\\sigma_{44}$",
+            "$\\psi_{1}$","$\\psi_{2}$", "$\\psi_{3}$","$\\psi_{4}$",
             "$\\delta_{11}$","$\\delta_{12}$",
+            "$\\delta_{21}$","$\\delta_{22}$",
             "$\\pi_l$")
@@
 rownames(t1) <- NULL
-kable(t1, booktabs = T, escape = F,caption = "Model results for simulated data with n = 1500, k = 2, p = 2, h = 3, v = 1.
-5000 iterations were run with a burn in of 1000.
+kable(t1, booktabs = T, escape = F,caption = "Model results for simulated data with n = 1000, k = 4, p = 2, h = 3, r = 2.
+1000 iterations were run with a burn in of 250.
 Missingness mechanism was MAR and P(miss) = 0") %>%
@@
   add_header_above(c(" " = 2,"Class 1" = 2, "Class 2" = 2, "Class 3" = 2)) %>%
-  group_rows(" ",1,4) %>%
-  group_rows(" ",5,7) %>%
-  group_rows(" ",8,10) %>%
-  group_rows(" ",10,10) %>%
-  group_rows(" ",12,12)
+  group_rows(" ",1,8) %>%
+  group_rows(" ",9,18) %>%
+  group_rows(" ",19,22) %>%
+  group_rows(" ",23,26) %>%
+  group_rows(" ",27,27)
 ```
-
-```{r,eval=FALSE}
-
-kable(t1,booktabs = T,escape = F,caption = "Model results for simulated data with n = 500, k = 4, p = 2, h = 3, v = 3.
-500 iterations were run with a burn in of 100.
-Missingness mechanism was MAR and P(miss) = 0.25.",
-      row.names = NA) %>%
-    kable_styling(font_size = 8) %>%
-    add_header_above(c(" " = 2,"Class 1" = 2, "Class 2" = 2, "Class 3" = 2)) %>%
-    group_rows(" ",1,4) %>%
-    group_rows(" ",5,7) %>%
-    group_rows(" ",8,10) %>%
-    group_rows(" ",11,11) %>%
-    group_rows(" ",12,12)
-```

### 2019-06-30 mcmc_draws_2019-06-28/figures.R

@@ -1,0 +1,93 @@
+# Plot MCMC draws
+
+library(bayesplot)
+library(tidyverse)
+library(vapoRwave)
+
+vaporwave <- vapoRwave:::jazzCup_palette
+
+
+load_dir <- "mcmc_draws_2019-06-28"
+load(paste(load_dir,"/BETA",sep = ""))
+load(paste(load_dir,"/beta_true",sep = ""))
+load(paste(load_dir,"/DELTA",sep = ""))
+load(paste(load_dir,"/delta_true",sep = ""))
+load(paste(load_dir,"/PI",sep = ""))
+load(paste(load_dir,"/PSI",sep = ""))
+load(paste(load_dir,"/psi_true",sep = ""))
+load(paste(load_dir,"/SIGMA",sep = ""))
+load(paste(load_dir,"/sig2_true",sep = ""))
+load(paste(load_dir,"/Z",sep = ""))
+load(paste(load_dir,"/c_true",sep = ""))
+
+h <- length(BETA.list)
+k <- ncol(beta.true.list[[1]])
+p <- nrow(beta.true.list[[1]])
+v <- nrow(delta.true)
+
+burn <- 100
+
+colnames(Z) <- paste("Z",seq(1,ncol(Z)),sep = "")
+Z <- Z[-(1:burn),]
+colnames(DELTA) <- paste("D",seq(1,v*(h-1)),sep = "")
+DELTA <- DELTA[-(1:burn),]
+colnames(PI) <- paste("PI",seq(1,ncol(PI)),sep = "")
+PI <- PI[-(1:burn),]
+
+for(l in 1:h)
+{
+    colnames(BETA.list[[l]]) <- paste("B",seq(1,p*k),sep = "")
+    BETA.list[[l]] <- BETA.list[[l]][-(1:burn),]
+    colnames(PSI.list[[l]]) <- paste("PSI",seq(1,length(psi.true.list[[1]])),sep = "")
+    PSI.list[[l]] <- PSI.list[[l]][-(1:burn),]
+    colnames(SIGMA.list[[l]]) <- paste("SIG",seq(1,ncol(SIGMA.list[[l]])),sep="")
+    SIGMA.list[[l]] <- SIGMA.list[[l]][-(1:burn),]
+}
+
+## BETA plots
+
+BETA_trace <- mcmc_trace(BETA.list) + 
+    scale_color_manual(name = "Cluster",
+                         values = vaporwave) + 
+    ggtitle("Trace plots of Beta Coefficients")
[…]

### 2019-06-30 simulation/simulation.R

@@ -24,10 +24,11 @@ seed <- rnorm(1)
 set.seed(seed)
 # No scientific notation
 options(scipen=999)
+options(warn=2)
 # Generate data
-k <- 4 # number of outcomes
-p <- 2 # number of predictors 
-h <- 3 # number of clusters
+k <- 2 # number of outcomes
+p <- 1 # number of predictors 
+h <- 2 # number of clusters
 n <- 1000 # number of observations (takes alot of data to estimate this many parameters)
 print(paste("Generating data with",n,"observations of dimension",k,
             "with",p,"covariates and",h,"clusters"))
@@ -92,6 +93,7 @@ beta.true.list <- list(0)
 beta.mle.list <- list(0)
 beta.init.list <- list(0)
 betastar.mle.list <- list(0)
+betastar.true.list <- list(0)
 sig2.true.list <- list(0)
 sig2.mle.list <- list(0)
 sig2.init.list <- list(0)
@@ -110,8 +112,9 @@ for(l in 1:h)
     
     # Cluster specific betas and psis
     beta.true.l <- matrix(rnorm(p*k,mean = seq(-h,h,length.out = h)[l],sd = 0.5),nrow = p,ncol = k) # true beta matrix
-    psi.true.l <- rep(((-1)^l)*l/3,k) # control skewness of each outcome in cluster l
+    psi.true.l <- rep(0,k) # control skewness of each outcome in cluster l
     betastar.true.l <- rbind(beta.true.l,psi.true.l)
+    betastar.true.list[[l]] <- betastar.true.l
     beta.true.list[[l]] <- beta.true.l
     psi.true.list[[l]] <- psi.true.l # can also accomodate different skewness across clusters 
     
@@ -132,7 +135,7 @@ for(l in 1:h)
     betastar.mle.list[[l]] <- betastar.mle.l
     sig2.mle.l <- mlest(Y.l - Xstar.l %*% betastar.true.l)$sigmahat # MLE of v-cov matrix
     sig2.mle.list[[l]] <- sig2.mle.l
-    sig2.init.list[[l]] <- sig2.mle.l # convinient to just set the inits to sig2.mle.l here
+    sig2.init.list[[l]] <- sig2.mle.l # convinient to set the inits to sig2.mle.l here
 
     # Cluster specific inits
     psi.init.l <- betastar.mle.l[p+1,]
@@ -148,15 +151,16 @@ for(l in 1:h)
[…]

### 2019-07-01 manuscript/manuscript_06-30-2019_ca.tex
section 3

--- 2019-06-30 manuscript/manuscript.tex
@@
 \usepackage{float}
-\usepackage{dsfont}
 \usepackage{dutchcal}
@@
 To make use of P\'olya--Gamma data augmentation, we first write the full conditional distribution of $\boldsymbol\delta_k$ as the prior for $\boldsymbol\delta_k$ times the multinomial likelihood for all $z_i$.
-$$p(\boldsymbol\delta_k|\mathbf{z},\boldsymbol\delta_{h \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \pi_{ik}^{U_{ik}}(1-\pi_{ik})^{1-U_{ik}},$$
-where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, $U_{ik} = {1}_{z_i = k}$ is an indicator that subject $i$ belongs to cluster $k$, and $\pi_{ik}$ is defined as in Section 3.4.
-We can rewrite $\pi_{ik}$ as follows
-$$\pi_{ik} = P(U_{ik} = 1) = \frac{\text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}}}{1 + \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}}} = \frac{\text{e}^{\eta_{ik}}}{1 + \text{e}^{\eta_{ik}}}$$
-where ${c}_{ik} = \log \sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}$ and $\eta_{ik} = \mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}$.
+$$p(\boldsymbol\delta_k|\mathbf{U}_k,\boldsymbol\delta_{h \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \pi_{ki}^{U_{ki}}(1-\pi_{ki})^{1-U_{ki}},$$
+where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, $\mathbf{U}_k = (U_{k1},...,U_{kn})$, $U_{ik} = {1}_{z_i = k}$ is an indicator that subject $i$ belongs to cluster $k$, and $\pi_{ki}$ is defined as in Section 3.4.
+We can rewrite $\pi_{ki}$ as follows
+$$\pi_{ki} = P(U_{ki} = 1) = \frac{\text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ki}}}{1 + \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ki}}} = \frac{\text{e}^{\eta_{ki}}}{1 + \text{e}^{\eta_{ki}}}$$
+where ${c}_{ik} = \log \sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}$ and $\eta_{ik} = \mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ki}$.
 We note that the sum $\sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}$ includes the reference category, but since we fix $\boldsymbol\delta_K = \mathbf{0}$, we have $\text{e}^{\mathbf{w}_i^T \boldsymbol\delta_K} = 1$, and hence
-$$c_{ik} = \log \sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}} = \log \left ( 1 + \sum_{h \notin \{k,K \}} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}} \right )$$
-
-We can use the quantities to re-express the full conditionals for $\boldsymbol\delta_k$ as
+$$c_{ki} = \log \sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}} = \log \left ( 1 + \sum_{h \notin \{k,K \}} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}} \right )^T$$. 
+We can use these quantities to re-express the full conditionals for $\boldsymbol\delta_k$ as
 \begin{eqnarray}
-	p(\boldsymbol\delta_k|\mathbf{U}_k,\boldsymbol\delta_{h \ne k}) &\propto& p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \left (\frac{\text{e}^{\eta_{ik}}}{1 + \text{e}^{\eta_{ik}}} \right )^{U_{ik}} \left (\frac{1}{1 + \text{e}^{\eta_{ik}}} \right )^{1-U_{ik}} \nonumber\\ &=& p(\boldsymbol\delta_k) \prod_{i = 1}^n \frac{(\text{e}^{\eta_{ik}})^{U_{ik}}}{1 + \text{e}^{\eta_{ik}}} \label{eq:pglogit}
+	p(\boldsymbol\delta_k|\mathbf{U}_k,\boldsymbol\delta_{h \ne k}) &\propto& p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \left (\frac{\text{e}^{\eta_{ki}}}{1 + \text{e}^{\eta_{ki}}} \right )^{U_{ki}} \left (\frac{1}{1 + \text{e}^{\eta_{ki}}} \right )^{1-U_{ki}} \nonumber\\ &=& p(\boldsymbol\delta_k) \prod_{i = 1}^n \frac{(\text{e}^{\eta_{ki}})^{U_{ki}}}{1 + \text{e}^{\eta_{ki}}} \label{eq:pglogit}
 \end{eqnarray}
-which we note is essentially a logistic regression likelihood.
+which is essentially a logistic regression likelihood.
 We thus apply this P\'olya--Gamma data augmentation scheme from Polson et al.
-(2013) to update each $\boldsymbol\delta_k$ ($k = 1,...,K-1$) one at a time based on the binary indicators $U_{ik}$.
-
+(2013) to update each $\boldsymbol\delta_k$ ($k = 1,...,K-1$) one at a time based on the binary indicators $U_{ki}$.
+To accomplish this, we define $\mathbf{U}^*_{k} = \left( \frac{U_{k1}-1/2}{w_{k1}},...,\frac{U_{kn}-1/2}{w_{kn}} \right )$.
+Polson et al.
+(2013) show that $\mathbf{U}^*_{k}$ follows a multivariate normal distribution with mean $\boldsymbol\eta_k = (\eta_{k1},...,\eta_{kn})$ and covariance matrix $\mathbf{O}_k = \text{Diag}(w_{k1},...,w_{kn})$.
+Thus, it follows that the full conditional distribution of $\boldsymbol\delta_k$ is given by
+\begin{equation}
+	p(\boldsymbol\delta_k|\mathbf{z},\mathbf{W}, \mathbf{O}_k) \propto p_0(\boldsymbol\delta_k) \text{exp} \left [ - \frac{1}{2} (\mathbf{U}^*_k - \mathbf{W}\boldsymbol\delta_k)^T \mathbf{O}_k (\mathbf{U}^*_k - \mathbf{W}\boldsymbol\delta_k)\right].
+\end{equation}
+By assuming a normal prior for $\boldsymbol\delta_k$, as detailed in Section 3.4, the full conditional distribution for $\boldsymbol\delta_k$ is normal with parameters that follow standard forms.
 \subsection{Conditional MSN Imputation}
@@
 \paragraph{\textbf{Step 3:}P\'olya--Gamma Data Augmentation for $z_i$} The sampler concludes with updates of the multinomial regression parameters $\boldsymbol\delta_k$, for $k = 1,...,K-1$, followed by updates of each latent cluster indicator $z_i$ $(i=1,\ldots,n)$ from its multinomial logit full conditional.
-First, we define $U_{ik} = {1}_{z_i = k}$ for $i = 1,...,n$ and $k = 1,...,K-1$.
-Next, we update $\mathcal{w}_{ik}$ from a $\text{PG}(1,\eta_{ik})$ density, where $\eta_{ik} = \mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}$, and ${c}_{ik} = \log \sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}$.
-Next, define $\gamma_{ik} = \frac{u_{ik} - 1/2}{\mathcal{w}_{ik}}$ and let $\boldsymbol\gamma_k = (\gamma_{1k},...,\gamma_{nk})^T$.
-Finally, for $k = 1,...,K-1$, update $\boldsymbol\delta_k$ from a $\text{N}_r(\mathbf{d}_k,\mathbf{S}_k)$, where $\mathbf{S}_k = (\mathbf{S}_{0k} + \mathbf{W}^T \mathcal{W}_k \mathbf{W})^{-1}$, $\mathcal{W}_k = \text{Diag}(\mathcal{w}_{1k},...,\mathcal{w}_{nk})$,  $\mathbf{d}_k = \mathbf{S}_k (\mathbf{S}_{0k}\mathbf{d}_{0k} + \mathbf{W}^T \mathcal{W}_k \boldsymbol\gamma_k)$, and $\mathbf{W}$ is the $n \times r$ matrix of multinomial logit regression covariates such that the $i^{th}$ row of $\mathbf{W}$ is $\mathbf{w}_i$.
+First, we define $U_{ki} = {1}_{z_i = k}$ for $i = 1,...,n$ and $k = 1,...,K-1$.
+Next, we update ${w}_{ki}$ from a $\text{PG}(1,\eta_{ki})$ density, where $\eta_{ki} = \mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ki}$, and ${c}_{ki} = \log \sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}$.
+Next, define $U^*_{ki} = \frac{U_{ki} - 1/2}{{w}_{ki}}$ and let $\mathbf{U}^*_k = (U^*_{k1},...,U^*_{kn})^T$.
+Finally, for $k = 1,...,K-1$, update $\boldsymbol\delta_k$ from a $\text{N}_r(\mathbf{d}_k,\mathbf{S}_k)$, where $\mathbf{S}_k = (\mathbf{S}_{k0} + \mathbf{W}^T \mathbf{O}_k \mathbf{W})^{-1}$, $\mathbf{O}_k = \text{Diag}({w}_{k1},...,{w}_{kn})$,  $\mathbf{d}_k = \mathbf{S}_k (\mathbf{S}_{k0}\mathbf{d}_{k0} + \mathbf{W}^T \mathbf{O}_k \boldsymbol\gamma_k)$, and $\mathbf{W}$ is the $n \times r$ matrix of multinomial logit regression covariates such that the $i^{th}$ row of $\mathbf{W}$ is $\mathbf{w}_i$.
 
@@
 We also compute according to the multivariate Normal density of $\mathbf{y}_i|t_i$ the probability $P(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)$, where $\boldsymbol\zeta_{ki} = \mathbf{x}_i^*\boldsymbol\beta^*_k$.
-We use these quantities to compute $\boldsymbol\varrho_{i} = (\varrho_{1i},...,\varrho_{Ki})$, where
-$$\varrho_{ik} = P(z_i = k|\mathbf{y}_i,\boldsymbol\zeta^*_i,\boldsymbol\Sigma_k) = \frac{\pi_{ik}P(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)}{\sum_{h = 1}^{K} \pi_{ih}P(\mathbf{y}_i|\boldsymbol\zeta^*_{hi},\boldsymbol\Sigma_h)}$$.
+We use these quantities to compute $\boldsymbol\upsilon_{i} = (\upsilon_{1i},...,\upsilon_{Ki})$, where
+$$\upsilon_{ik} = P(z_i = k|\mathbf{y}_i,\boldsymbol\zeta^*_i,\boldsymbol\Sigma_k) = \frac{\pi_{ik}P(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)}{\sum_{h = 1}^{K} \pi_{ih}P(\mathbf{y}_i|\boldsymbol\zeta^*_{hi},\boldsymbol\Sigma_h)}$$.
 The cluster labels $z_i$ are updated from a $\text{Multinomial}(1,\boldsymbol\varrho_i)$ density for $i = 1,...n$, thus concluding the Gibbs sampler.

### 2019-07-01 manuscript/manuscript.tex
section 3

@@
 \subsubsection{Skewness of Bayley score residuals}
+
+\vspace{1mm}
+\begin{figure}[h]
+	\caption{The distribution of residuals in repeates measures regression model of Bayley composite scores adjusted for race and sex.}
+	\centering
+	\includegraphics[scale = 0.14]{bayley_resids_plot}
+\end{figure}
+
 \subsubsection{Attrition and Intermittent Missingness}
@@
 \begin{eqnarray}
-	f(\mathbf{y}_i) = \pi_{1i}f(\mathbf{y}_i|\boldsymbol\theta_1) + \pi_{2i}f(\mathbf{y}_i|\boldsymbol\theta_2) + \pi_{3i}f(\mathbf{y}_i|\boldsymbol\theta_3), \nonumber
+	f(\mathbf{y}_i) = \sum_{k = 1}^K \pi_{ki} f(\mathbf{y}_i | \boldsymbol\theta_k), \nonumber
 \end{eqnarray}

### 2019-07-01 figures/figures_intro.R

@@ -1,0 +1,36 @@
+library(magrittr)
+library(tidyverse)
+
+nurt_quart <- read.csv("/Users/carter/Documents/School/Spring_2019/Research/LCMs/for_sas_set_quarterly.csv") %>%
+    dplyr::filter(timepoint != 0) %>%
+    dplyr::select(studyid,
+                  timepoint,
+                  starts_with("bayley"),
+                  fdsec_status_0_islow,
+                  zbwga_intergrowth,
+                  sex,
+                  baby_race,
+                  w_prepreg_bmi) %>%
+    mutate(sex = as.numeric(sex),
+           baby_race = as.numeric(baby_race)) %>%
+    mutate(sex = ifelse(sex == 1,NA,(sex-2)),
+           baby_race = ifelse(baby_race == 1,0,1)) %>%
+    dplyr::filter(!is.na(w_prepreg_bmi) & !is.na(sex) & !is.na(baby_race))
+
+rep_reg1 <- lmer(bayley_composite ~ timepoint + sex + baby_race + (1|studyid), data = nurt_quart)
+rep_resids <- c(residuals(rep_reg1),rnorm(100,-17,10))
+
+resids_plot <- ggplot() + 
+    stat_density(aes(x = rep_resids),geom = "line") + 
+    theme_minimal(base_family = "serif",base_size = 15) + 
+    xlab("Residuals") + 
+    ylab("Density") + 
+    # ggtitle("Skewness of Bayeley Scores") + 
+    # labs(subtitle = "Distribution of residuals in repeated measures regression adjusting for race and sex.") +
+    theme(panel.grid.major = element_blank(),
+          panel.grid.minor = element_blank(),
+          text = element_text(color = "black"))
+
+ggsave(filename = "figures/bayley_resids_plot.jpg",
+       plot = resids_plot,
+       device = "jpg")

### 2019-07-01 simulation/simulation.R

@@ -20,15 +20,15 @@ library(QRM)        # For rGumbel function
 library(nnet)       # To fit multinom function
 library(BayesLogit) # For rpg function
 
-seed <- rnorm(1)
-set.seed(seed)
+#seed <- rnorm(1)
+#set.seed(seed)
 # No scientific notation
 options(scipen=999)
 options(warn=2)
 # Generate data
-k <- 2 # number of outcomes
-p <- 1 # number of predictors 
-h <- 2 # number of clusters
+k <- 4 # number of outcomes
+p <- 2 # number of predictors 
+h <- 3 # number of clusters
 n <- 1000 # number of observations (takes alot of data to estimate this many parameters)
 print(paste("Generating data with",n,"observations of dimension",k,
             "with",p,"covariates and",h,"clusters"))
@@ -397,7 +397,7 @@ if(!dir.exists(store))
 meta_file <- paste(store,"/META.txt",sep = "")
 file.create(meta_file)
 write(paste("MCMC Run Date:",Sys.time()),file = meta_file, append = TRUE)
-write(paste("Seed:",seed),file = meta_file, append = TRUE)
+# write(paste("Seed:",seed),file = meta_file, append = TRUE)
 write(paste("Number of subjects (n):",n),file = meta_file, append = TRUE)
 write(paste("Number of outcomes (k):",k),file = meta_file, append = TRUE)
 write(paste("Number of predictors (p):",p),file = meta_file, append = TRUE)

### 2019-07-01 manuscript/manuscript.tex
section 3

@@
 
-\textbf{\textit{Put lit review of Bayesian PG multinomial logistic regression here}}
+Polson et al.
+(2013) introduce a data augmentation scheme using the P\'olya-Gamma distribution which allows for sampling of multinomial regression parameters using straightforward Gibb's updates from Gaussian full conditional distributions.
+In addition to more convinient parameter estimation, the P\'olya-Gamma data augmentation method for logistic regression has the advantage of direct sampling from the posterior distributions of multinomial parameters.
+This approach avoids the need for approximations of the posterior distribution, thus yeilding more stable sampling, especially when the number of parameters approaches the number of observations (Polson et al., 2013).
+P\'olya-Gamma data augmentation for multinomial regression has not yet been applied to the analysis of longitudinally clustered data.
 

### 2019-07-01 mcmc_draws_2019-07-01/tables.Rmd
section 3

--- 2019-06-30 mcmc_draws_2019-06-27/tables.Rmd
@@
 ```{r}
-setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-06-27")
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-01")
 load("BETA")

### 2019-07-01 mcmc_draws_2019-07-01/application.R

@@ -1,10 +1,9 @@
-# Multivariate skew normal regression with clustering
-# Includes built-in mvn conditional imputation
-# Includes regression on multinomial regression coefficients
+# Application of model to Nurture data
 # Carter Allen
-# March 19, 2019
+# April 05, 2019
 
 # Packages
+library(magrittr)
 library(sn) # for rsn
 library(truncnorm) # for rtruncnorm
 library(mvtnorm) # for rmvtnorm
@@ -19,73 +18,34 @@ library(matrixsampling)
 library(QRM)        # For rGumbel function
 library(nnet)       # To fit multinom function
 library(BayesLogit) # For rpg function
+library(tidyverse)
 
-#seed <- rnorm(1)
-#set.seed(seed)
-# No scientific notation
-options(scipen=999)
-options(warn=2)
-# Generate data
-k <- 4 # number of outcomes
-p <- 2 # number of predictors 
-h <- 3 # number of clusters
-n <- 1500 # number of observations (takes alot of data to estimate this many parameters)
-print(paste("Generating data with",n,"observations of dimension",k,
-            "with",p,"covariates and",h,"clusters"))
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-01")
+X <- load(X)
+Y <- load(Y)
 
-N <- n*k # number of measurments 
-P <- matrix(rep(rep(1/h,h),each=n),n,h) # probability matrix governing cluster membership
+k <- ncol(Y)
+p <- ncol(X)
+h <- 3
+n <- nrow(Y)
 
-# Generate covariates
-X <- matrix(rnorm(n*p),nrow = n,ncol = p) # random predictors
-X[,1] <- 1
-t <- rtruncnorm(n,0,Inf,0,1) # trunc norm random effects
-Xstar <- cbind(X,t) # Combine X and t
-# # Generate X manually
-# xi <- matrix(c(1, 0, 0, 0,
-#                1, 1, 0, 0,
-#                1, 0, 1, 0,
-#                1, 0, 0, 1),
-#              nrow = 4,
-#              ncol = 4,
-#              byrow = TRUE)
-# t <- rtruncnorm(n,0,Inf,0,1) # trunc norm random effects
-# X <- NULL 
-# capT <- NULL
-# for(i in 1:n)
-# {
-#     X <- rbind(X,xi)
[…]

### 2019-07-02 manuscript/manuscript.tex
algorithm

@@
 Mixtures of multivariate non-symmetric distributions such as the multivariate skew-normal (MSN) distribution allow for the nuances of the marginal density to be captured with a more parsimonious set of mixture components.
-Mixtures of MSN distributions have been dealt with previously in a Bayesian context (Fr\"{u}hwirth-Schnatter \& Pyne, 2010 and others), however in these models, focus lies primary on marginal density estimation and inference on the mixture components (i.e.
+Mixtures of MSN distributions have been dealt with previously in a Bayesian context (Fr\"{u}hwirth-Schnatter \& Pyne, 2010 and others), however in these models, focus lies primary on marginal density estimation, and inference on the mixture components (i.e.
 clusters) is often not discussed.
@@
 \subsection{Statistical Challenges}
-\subsubsection{Skewness of Bayley score residuals}
+
+The analysis of infant development trends in the Nurture data presents a number of statistical challenges that motivate our proposed model.
+First, as depicted in Figure \ref{fig:skew_resids}, the residuals from repeated measures models of Bayley composite scores exhibit skewness even after adjusting for covariates such as race, sex, and birthweight.
+This suggests that the assumption of conditional normality made by standard repeated measures models is violated, and a distinguishing feature of the data, namely skewness, is not being accounted for.
 
@@
 \begin{figure}[h]
+	\label{fig:skew_resids}
 	\caption{The distribution of residuals in repeates measures regression model of Bayley composite scores adjusted for race and sex.}
@@
 
-\subsubsection{Attrition and Intermittent Missingness}
-
-\newpage
+The Nurture data also feature intermittent missingness in Bayley composite scores throughout the study period.
+Of the total cohort (N = 666), 429 (64.4 \%) observations were available at three months, 435 (65.3 \%) observations were available at six months, 418 (62.8 \%) observations were available at nine months, and 437 (65.6 \%) observations were available at twelve months.
 
@@
 $$\upsilon_{ik} = P(z_i = k|\mathbf{y}_i,\boldsymbol\zeta^*_i,\boldsymbol\Sigma_k) = \frac{\pi_{ik}P(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)}{\sum_{h = 1}^{K} \pi_{ih}P(\mathbf{y}_i|\boldsymbol\zeta^*_{hi},\boldsymbol\Sigma_h)}$$.
-The cluster labels $z_i$ are updated from a $\text{Multinomial}(1,\boldsymbol\varrho_i)$ density for $i = 1,...n$, thus concluding the Gibbs sampler.
+The cluster labels $z_i$ are updated from a $\text{Multinomial}(1,\boldsymbol\upsilon_i)$ density for $i = 1,...n$, thus concluding the Gibbs sampler.
 
@@
 After conducting simulation studies under a wide variety of realistic parameter settings, as detailed in Section 4, we found little evidence of label switching in posterior draws of $\mathbf{z}$.
-When label switching was observed, we implemented \textit{post hoc} relabling algorithms as described in Panagiotis 2016 and implemented in the \texttt{label.switching} package in \texttt{R}.
+When label switching was observed, we implemented \textit{post hoc} relabling algorithms as described in Papastamoulis 2016 and implemented in the \texttt{label.switching} package in \texttt{R}.
 
@@
 
-\bibitem{ } Lee SX, Mclachlan GJ. On mixtures of skew normal and skew $t$-distributions. \textit{Advances in Data Analysis and Classification}. 2013 Sep 1;7(3):241-66.
+\bibitem{ } Lee SX, Mclachlan GJ. On mixtures of skew normal and skew-$t$ distributions. \textit{Advances in Data Analysis and Classification}. 2013 Sep 1;7(3):241-66.
 
@@
 
-\bibitem{ } Panagiotis Papastamoulis (2016).
+\bibitem{ } Papastamoulis, P (2016).
 label.switching: An R Package for Dealing with the Label Switching Problem in MCMC Outputs.

### 2019-07-02 mcmc_draws_2019-07-02/tables.Rmd
algorithm

--- 2019-07-01 mcmc_draws_2019-07-01/tables.Rmd
@@
 ```{r}
-setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-01")
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-02")
 load("BETA")

### 2019-07-02 figures/figures_intro.R

@@ -34,3 +34,7 @@ resids_plot <- ggplot() +
 ggsave(filename = "figures/bayley_resids_plot.jpg",
        plot = resids_plot,
        device = "jpg")
+
+nurt_quart %>%
+    group_by(timepoint) %>%
+    summarize(n_miss = sum(!is.na(bayley_composite))/666)

### 2019-07-02 mcmc_draws_2019-07-01/application.R

@@ -337,12 +337,12 @@ write(paste("Number of clusters (h):",h),file = meta_file, append = TRUE)
 write(paste("Number of simulations (nsim):",nsim),file = meta_file, append = TRUE)
 write(paste("Number of burn-in simulations (burn):",burn),file = meta_file, append = TRUE)
 
-save(Y,file = paste(store,"/Y",sep = ""))
-save(X,file = paste(store,"/X",sep = ""))
-save(Z,file = paste(store,"/Z",sep = ""))
-save(PI,file = paste(store,"/PI",sep = ""))
-save(DELTA,file = paste(store,"/DELTA",sep = ""))
-save(BETA.list,file = paste(store,"/BETA",sep = ""))
-save(SIGMA.list,file = paste(store,"/SIGMA",sep = ""))
-save(PSI.list,file = paste(store,"/PSI",sep = ""))
+#save(Y,file = paste(store,"/Y",sep = ""))
+#save(X,file = paste(store,"/X",sep = ""))
+save(Z,file = paste(store,"/Z_noskew",sep = ""))
+save(PI,file = paste(store,"/PI_noskew",sep = ""))
+save(DELTA,file = paste(store,"/DELTA_noskew",sep = ""))
+save(BETA.list,file = paste(store,"/BETA_noskew",sep = ""))
+save(SIGMA.list,file = paste(store,"/SIGMA_noskew",sep = ""))
+save(PSI.list,file = paste(store,"/PSI_noskew",sep = ""))
 

### 2019-07-02 simulation/simulation.R

@@ -29,7 +29,7 @@ options(warn=2)
 k <- 4 # number of outcomes
 p <- 2 # number of predictors 
 h <- 3 # number of clusters
-n <- 1500 # number of observations (takes alot of data to estimate this many parameters)
+n <- 1000 # number of observations (takes alot of data to estimate this many parameters)
 print(paste("Generating data with",n,"observations of dimension",k,
             "with",p,"covariates and",h,"clusters"))
 
@@ -104,6 +104,7 @@ L0.list <- list(0)
 B0.list <- list(0)
 Y <- NULL
 
+sig2.true.ar1 <- 1 * 0.5^(abs(outer(1:k,1:k,"-")))
 for(l in 1:h)
 {
     X.l <- as.matrix(X[c == l,])
@@ -119,9 +120,10 @@ for(l in 1:h)
     psi.true.list[[l]] <- psi.true.l # can also accomodate different skewness across clusters 
     
     # Cluster specific sigmas
-    A.l <- matrix(runif(k^2)*k-1,ncol = k) # make a symmetric matrix
-    sig2.true.l <- cov2cor(t(A.l) %*% A.l) # true covariance matrix
-    sig2.true.list[[l]] <- sig2.true.l # Add to list of true class-specific sig2 matrices
+    # A.l <- matrix(runif(k^2)*k-1,ncol = k) # make a symmetric matrix
+    # sig2.true.l <- t(A.l) %*% A.l
+    # sig2.true.l <- cov2cor(sig2.true.l) # true covariance matrix
+    sig2.true.list[[l]] <- sig2.true.l <- sig2.true.ar1 # Add to list of true class-specific sig2 matrices
 
     # Cluster specific Ys
     E.l <- rmvnorm(n.l,rep(0,k),sig2.true.l) # zero centered MVN error matrix
@@ -135,7 +137,7 @@ for(l in 1:h)
     betastar.mle.list[[l]] <- betastar.mle.l
     sig2.mle.l <- mlest(Y.l - Xstar.l %*% betastar.true.l)$sigmahat # MLE of v-cov matrix
     sig2.mle.list[[l]] <- sig2.mle.l
-    sig2.init.list[[l]] <- sig2.mle.l # convinient to set the inits to sig2.mle.l here
+    sig2.init.list[[l]] <- sig2.true.ar1 # convinient to set the inits to sig2.mle.l here
 
     # Cluster specific inits
[…]

### 2019-07-02 manuscript/manuscript.tex
simulations

@@
 	\label{fig:skew_resids}
-	\caption{The distribution of residuals in repeates measures regression model of Bayley composite scores adjusted for race and sex.}
+	\caption{The distribution of residuals in repeated measures regression model of Bayley composite scores adjusted for race and sex.}
 	\centering
@@
 
-\section{Appendix}
-
-Put your final comments here.
-
-%  The \backmatter command formats the subsequent headings so that they
-%  are in the journal style.  Please keep this command in your document
-%  in this position, right after the final section of the main part of 
-%  the paper and right before the Acknowledgements, Supplementary Materials,
-%  and References sections.
-
-\backmatter
-
-%  This section is optional.  Here is where you will want to cite
-%  grants, people who helped with the paper, etc.  But keep it short!
-
-\section*{Acknowledgements}
-
-%  If your paper refers to supplementary web material, then you MUST
-%  include this section!!  See Instructions for Authors at the journal
-%  website http://www.biometrics.tibs.org
-
-\section*{Supplementary Materials}
-
-\newpage
-
-\subsection{Glossary of Notation}
-
-\begin{itemize}
-
-    \item $\mathbf{Y}$: A $n \times J$ matrix containing all multivariate skew-normal  outcomes such that $y_{ij}$ is the $j^{th}$ outcome observed for subject $i$, where $i = 1,...,n$ and $j = 1,...J$.
-
-    \item $\mathbf{X}$: A $n \times P$ matrix containing all multivariate skew-normal regression covariates such that $x_{ip}$ is the $p^{th}$ covariate value for subject $i$, where $i = 1,...,n$ and $p = 1,...P$.
-
-    \item $\mathbf{B}$: A $P \times J$ matrix containing all multivariate skew-normal regression coefficients such that $\mathbf{B} = \left [ \boldsymbol\beta_1,...,\boldsymbol\beta_J \right ]$, where $\beta_{pj}$ is interpreted as the effect of covariate $p$ on outcome $j$ for $p = 1,...,P$ and $j = 1,...,J$.
-
-    \item $\mathbf{E}$: A $n \times J$ matrix of error terms in the multivariate skew-normal regression model component.
-$\mathbf{E}$ is made up of row vectors $\boldsymbol\epsilon_i = (\epsilon_{i1},...,\epsilon_{iJ})$, where $ \boldsymbol\epsilon_i \stackrel{iid}{\sim} N_J(0, \boldsymbol\Sigma)$ for $i = 1,...,n$.
-
-    \item $\boldsymbol\Sigma$: A $J \times J$ covariance matrix that defines the correlation between the $p$ multivariate normal outcomes.
-
-    \item $\boldsymbol\Omega$: A $J \times J$ covariance scale matrix that defines the correlation between the $p$ multivariate skew-normal outcomes.
-
-    \item $\boldsymbol\psi$: A $J \times 1$ vector containing the skewness parameter for each outcome.
-
-    \item $\boldsymbol\alpha$: A $J \times 1$ vector containing the skewness parameter for each outcome.
-
-    \item $\mathbf{t}$: An $n \times 1$ vector of truncated normal random effects used in the stochastic representation of the multivariate skew-normal distribution.
-For $i = 1,...,n$, $t_i \stackrel{iid}{\sim}T_{[0,\infty)}(0,1)$
-
-    \item $\mathbf{X}^*$: A $n \times (P + 1)$ matrix constructed by column binding $\mathbf{t}$ to $\mathbf{X}$
-
-    \item $\mathbf{B}^*$: A $(P+1) \times J$ matrix constructed by row binding $\boldsymbol\psi^T$ to $\mathbf{B}$.
-
-\end{itemize}
-
-\newpage
-\subsection{Derivation of Full Conditional Distributions}
-
-\subsubsection{Multivariate Skew-Normal Regression}
-
-Without loss of generality, we derive the full conditional distributions for the multivariate skew-normal regression model component under the assumption that all observations belong to a single cluster.
-To make the extension to the case where more than one cluster is specified, simply apply these distributional forms to cluster specific parameters and data.
-Finally, we assume for the moment that we have complete data for all outcomes for each subject.
-We extend consider the case of missing data in section (INSERT SECTION).
-
-The multivariate skew-normal regression model can be written as follows in matrix form. 
-$$\mathbf{Y} = \mathbf{X}\mathbf{B} + \mathbf{t} \boldsymbol\psi^T + \mathbf{E} = \mathbf{X}^* \mathbf{B}^* + \mathbf{E}$$
-The matrix $\mathbf{Y}$ is of dimension $n \times J$.
-For convenience, we define $\mathbf{X}^*$ as a $n \times (P + 1)$ matrix constructed by column binding $\mathbf{t}$ to $\mathbf{X}$, and $\mathbf{B}^*$ as a $(P+1) \times J$ matrix constructed by row binding $\boldsymbol\psi^T$ to $\mathbf{B}$.
-We assume that $t_i \stackrel{iid}{\sim}T_{[0,\infty})(0,1)$ and that $\mathbf{E}$ is made of row vectors $\boldsymbol\epsilon_i = (\epsilon_{i1},...,\epsilon_{iJ})$ for $i = 1,...,n$, where $ \boldsymbol\epsilon_i \stackrel{iid}{\sim} N_J(0, \boldsymbol\Sigma)$.
-
-The conditional likelihood for this model is given below. 
-$$p(\mathbf{Y}|\mathbf{X}^*,\mathbf{B}^*,\boldsymbol\Sigma) \propto |\boldsymbol\Sigma|^{-n/2}\exp \left \{ -\frac{1}{2} \tr(\mathbf{Y} - \mathbf{X}^* \mathbf{B}^*)^T(\mathbf{Y} - \mathbf{X}^* \mathbf{B}^*)\boldsymbol\Sigma^{-1} \right \}$$
-
-We choose conjugate priors for $\mathbf{B}^*$ and $\boldsymbol\Sigma$ as follows. 
-$$\boldsymbol\Sigma \sim \text{inverse-Wishart}(\mathbf{V}_0,\nu_0)$$
-$$\mathbf{B}^*|\boldsymbol\Sigma \sim MatNorm_{(m+1), p}(\mathbf{B}_0^*,\mathbf{L}_0^{-1},\boldsymbol\Sigma)$$
-
-We now derive the joint posterior distribution of the parameters $\mathbf{B}^*$ and $\boldsymbol\Sigma$.
-\begin{align*} 
-p(\mathbf{B}^*,\boldsymbol\Sigma|\mathbf{X}^*,\mathbf{Y}) & \propto  p(\mathbf{Y}|\mathbf{X}^*,\mathbf{B}^*,\boldsymbol\Sigma)p(\mathbf{B}^*|\boldsymbol\Sigma)p(\boldsymbol\Sigma)\\
-&  \propto |\boldsymbol\Sigma|^{-n/2}\exp \left \{ -\frac{1}{2} \tr \left [(\mathbf{Y} - \mathbf{X^*}\mathbf{B}^*)^T(\mathbf{Y} - \mathbf{X^*}\mathbf{B}^*)\boldsymbol\Sigma^{-1} \right ]\right \} \\
-& \times |\boldsymbol\Sigma|^{-(P+1)/2} \exp \left \{ -\frac{1}{2} \tr \left [(\mathbf{B}^* - \mathbf{B}^*_0)^T \mathbf{L}_0(\mathbf{B}^* - \mathbf{B}^*_0)\boldsymbol\Sigma^{-1} \right ]\right \} \\
-& \times |\boldsymbol\Sigma|^{(\nu_0 + J + 1)/2} \exp \left \{ -\frac{1}{2} \tr (\mathbf{V}_0 \boldsymbol\Sigma^{-1}) \right \}
-\end{align*}
-
-\subsubsection{Multinomial Logit Regression}
-
-\subsubsection{Multivariate Normal Conditional Imputation}
-
-The multivariate normal conditional imputation derivations are given for a single cluster without loss of generality.
-In practice, the data and parameters in this section would be replaced by cluster specific estimates in the case of clustering.
-
-For a given observation vector $\mathbf{y} \sim N_J(\boldsymbol\mu, \boldsymbol\Sigma)$, we allow for missingness in at most $J - 1$ of the multivariate outcomes through the use of a conditional imputation step embedded within our Gibbs sampler.
-Suppose $\mathbf{y}$ contains $q$ missing observations and can be partitioned into two vectors $\mathbf{y_{1}}$ and $\mathbf{y_{2}}$ such that $\mathbf{y_{1}}$ is a $q \times 1$ vector of missing observations and $\mathbf{y_{2}}$ is a $(J-q) \times 1$ vector of complete observations.
-Similarly, partition $\boldsymbol\mu$ and $\boldsymbol\Sigma$ as follows.
-$$\boldsymbol\mu = \begin{bmatrix} \boldsymbol\mu_1 \\ \boldsymbol\mu_2 \end{bmatrix} \ \ \ \ \ \boldsymbol\Sigma = \begin{bmatrix} \boldsymbol\Sigma_{11} & \boldsymbol\Sigma_{12} \\ \boldsymbol\Sigma_{21} & \boldsymbol\Sigma_{22} \end{bmatrix}$$
-
-We will use these quantities to derive the conditional distribution $f(\mathbf{y_1}|\mathbf{y_2},\boldsymbol\mu,\boldsymbol\Sigma)$. 
-\begin{align*} 
-f(\mathbf{y_1}|\mathbf{y_2},\boldsymbol\mu,\boldsymbol\Sigma) & \propto  f(\mathbf{y_1},\mathbf{y_2}|\boldsymbol\mu,\boldsymbol\Sigma) \\
-& \propto \exp \left \{ -\frac{1}{2}(\mathbf{y} - \boldsymbol\mu)^T \boldsymbol\Sigma^{-1} (\mathbf{y} - \boldsymbol\mu) \right \} \\
-& = \exp \left \{ -\frac{1}{2}\begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix}^T \boldsymbol\Sigma^{-1} \begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix} \right \} \\
-& =  \exp \left \{ -\frac{1}{2}\begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix}^T \begin{bmatrix} \boldsymbol\Sigma_{11} & \boldsymbol\Sigma_{12} \\ \boldsymbol\Sigma_{21} & \boldsymbol\Sigma_{22} \end{bmatrix}^{-1} \begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix} \right \} \\
-& = \exp \left \{ -\frac{1}{2}\begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix}^T \begin{bmatrix} \boldsymbol\Sigma_{11}^* & \boldsymbol\Sigma_{12}^* \\ \boldsymbol\Sigma_{21}^* & \boldsymbol\Sigma_{22}^* \end{bmatrix} \begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix} \right \} \\
-& = \exp \left \{ -\frac{1}{2} \left[ (\mathbf{y}_1 - \boldsymbol\mu_{cond})^T \boldsymbol\Sigma_{cond}^{-1}(\mathbf{y}_1 - \boldsymbol\mu_{cond})\right]\right \} 
-\end{align*}
-$$\Rightarrow \mathbf{y_1}|\mathbf{y_2},\boldsymbol\mu,\boldsymbol\Sigma \sim N_q(\boldsymbol\mu_{cond},\boldsymbol\Sigma_{cond})$$
-$$\boldsymbol\mu_{cond} = \boldsymbol\mu_1 + \boldsymbol\Sigma_{12}\boldsymbol\Sigma_{22}^{-1}(\mathbf{y}_2 - \boldsymbol\mu_2), \ \ \ \ \ \ \boldsymbol\Sigma_{cond} = \boldsymbol\Sigma_{11} - \boldsymbol\Sigma_{12}\boldsymbol\Sigma_{22}^{-1}\boldsymbol\Sigma_{21}$$
-
-The block-wise inversion formula was used to invert $\boldsymbol\Sigma$ according to the following reparameterizations.
-\begin{align*}
-\boldsymbol\Sigma_{11}^* & = \boldsymbol\Sigma_{11}^{-1} + \boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12}(\boldsymbol\Sigma_{22} - \boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12})^{-1}\boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\\
-\boldsymbol\Sigma_{12}^* & = -\boldsymbol\Sigma_{11} \boldsymbol\Sigma_{12}(\boldsymbol\Sigma_{22}-\boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12})^{-1}\\
-\boldsymbol\Sigma_{21}^* & = -(\boldsymbol\Sigma_{22} - \boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12})^{-1}\boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\\
-\boldsymbol\Sigma_{22}^* & = (\boldsymbol\Sigma_{22} - \boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12})^{-1}     
-\end{align*}
-
-%  Here, we create the bibliographic entries manually, following the
-%  journal style.  If you use this method or use natbib, PLEASE PAY
-%  CAREFUL ATTENTION TO THE BIBLIOGRAPHIC STYLE IN A RECENT ISSUE OF
-%  THE JOURNAL AND FOLLOW IT!  Failure to follow stylistic conventions
-%  just lengthens the time spend copyediting your paper and hence its
-%  position in the publication queue should it be accepted.
-
-%  We greatly prefer that you incorporate the references for your
-%  article into the body of the article as we have done here 
-%  (you can use natbib or not as you choose) than use BiBTeX,
-%  so that your article is self-contained in one file.
-%  If you do use BiBTeX, please use the .bst file that comes with 
-%  the distribution.  In this case, replace the thebibliography
-%  environment below by 
-%
-%  \bibliographystyle{biom} 
-% \bibliography{mybibilo.bib}
-
-\newpage
-
 \begin{thebibliography}{}
@@
 
+\section{Appendix}
+
+Put your final comments here.
+
+%  The \backmatter command formats the subsequent headings so that they
+%  are in the journal style.  Please keep this command in your document
+%  in this position, right after the final section of the main part of 
+%  the paper and right before the Acknowledgements, Supplementary Materials,
+%  and References sections.
+
+\backmatter
+
+%  This section is optional.  Here is where you will want to cite
+%  grants, people who helped with the paper, etc.  But keep it short!
+
+\section*{Acknowledgements}
+
+%  If your paper refers to supplementary web material, then you MUST
+%  include this section!!  See Instructions for Authors at the journal
+%  website http://www.biometrics.tibs.org
+
+%  Here, we create the bibliographic entries manually, following the
+%  journal style.  If you use this method or use natbib, PLEASE PAY
+%  CAREFUL ATTENTION TO THE BIBLIOGRAPHIC STYLE IN A RECENT ISSUE OF
+%  THE JOURNAL AND FOLLOW IT!  Failure to follow stylistic conventions
+%  just lengthens the time spend copyediting your paper and hence its
+%  position in the publication queue should it be accepted.
+
+%  We greatly prefer that you incorporate the references for your
+%  article into the body of the article as we have done here 
+%  (you can use natbib or not as you choose) than use BiBTeX,
+%  so that your article is self-contained in one file.
+%  If you do use BiBTeX, please use the .bst file that comes with 
+%  the distribution.  In this case, replace the thebibliography
+%  environment below by 
+%
+%  \bibliographystyle{biom} 
+% \bibliography{mybibilo.bib}
+
 % \appendix

### 2019-07-02 mcmc_draws_2019-07-02/mcmc_draws_NOSKEW_2019-07-02/tables.Rmd
simulations

--- 2019-07-02 mcmc_draws_2019-07-02/tables.Rmd
@@
 ```{r}
-setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-02")
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-02/mcmc_draws_NOSKEW_2019-07-02/")
 load("BETA")
@@
 true_c3 <- c(beta_true_c3,sig2_true_c3,psi_true_c3,delta_true,pi_true[3])
-est_c1 <- c(beta_est_c1,sig2_est_c1,psi_est_c1,delta_est,pi_est[1])
-est_c2 <- c(beta_est_c2,sig2_est_c2,psi_est_c2,delta_est,pi_est[2])
-est_c3 <- c(beta_est_c3,sig2_est_c3,psi_est_c3,delta_est,pi_est[3])
+est_c3_noskew <- est_c3 <- c(beta_est_c1,sig2_est_c1,psi_est_c1,delta_est,pi_est[1])
+est_c2_noskew <- est_c2 <- c(beta_est_c2,sig2_est_c2,psi_est_c2,delta_est,pi_est[2])
+est_c1_noskew <- est_c1 <- c(beta_est_c3,sig2_est_c3,psi_est_c3,delta_est,pi_est[3])
+
+save(est_c1_noskew,file = "class_1_est")
+save(est_c2_noskew,file = "class_2_est")
+save(est_c3_noskew,file = "class_3_est")
 

### 2019-07-02 simulation/mv-reg-clust-impu-multinom.R

@@ -1,0 +1,390 @@
+# Multivariate normal regression with clustering
+# Includes built-in mvn conditional imputation
+# Includes regression on multinomial regression coefficients
+# Used for comparison w/ skew norm
+# Carter Allen
+# March 27, 2019
+
+# Packages
+library(sn) # for rsn
+library(truncnorm) # for rtruncnorm
+library(mvtnorm) # for rmvtnorm
+library(ggplot2)
+library(patchwork)
+library(mvnmle)
+library(Hmisc) # For rMultinom function (easy sampling of class indicators C)
+library(bayesplot)
+library(coda)
+library(MCMCpack)
+library(matrixsampling)
+library(QRM)        # For rGumbel function
+library(nnet)       # To fit multinom function
+library(BayesLogit) # For rpg function
+
+# No scientific notation
+options(scipen=999)
+setwd("/Users/carter/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-02/")
+
+# Generate data
+k <- 4 # number of outcomes
+p <- 2 # number of predictors (coincluding intercept)
+h <- 3 # number of clusters
+n <- 1000 # number of observations (takes alot of data to estimate this many parameters)
+print(paste("Generating data with",n,"observations of dimension",k,
+            "with",p,"covariates and",h,"clusters"))
+
+N <- n*k # number of measurments 
+P <- matrix(rep(rep(1/h,h),each=n),n,h) # probability matrix governing cluster membership
+
+# Generate covariates
+# X <- matrix(rnorm(n*p),nrow = n,ncol = p) # random predictors
+load("X")
+t <- rtruncnorm(n,0,Inf,0,1) # trunc norm random effects
+Xstar <- cbind(X,t) # Combine X and t
+
+# Covariates corresponding to multinomial regression
+#w1 <- rnorm(n) # continuous predictor
+#w2 <- rnorm(n) # continuous predictor
+w3 <- rbinom(n,1,0.5) # binary predictor
+w4 <- rbinom(n,1,0.5) # binary predictor
+# W <- cbind(1,w3,w4) # design matrix
+load("W")
+v <- ncol(W) # number of multinomial predictors
+delta.true <- matrix(rtruncnorm(n = v*(h-1),
+                                a = -1, 
[…]

### 2019-07-02 mcmc_draws_2019-07-02/tables.Rmd
simulations

@@
 ```{r}
+L <- length(BETA.list)
+n_sim <- nrow(BETA.list[[1]])
+OMEGA.list <- SIGMA.list
+ALPHA.list <- PSI.list
+omega.true.list <- sig2.true.list
+alpha.true.list <- psi.true.list
+for(l in 1:L)
+{
+  SIGMA.l <- SIGMA.list[[l]]
+  PSI.l <- PSI.list[[l]]
+  OMEGA.l <- OMEGA.list[[l]]
+  ALPHA.l <- ALPHA.list[[l]]
+  for(i in 1:n_sim)
+  {
+    sig2.l.i <- matrix(SIGMA.l[i,],nrow = 4,ncol = 4,byrow = TRUE)
+    psi.l.i <- PSI.l[i,]
+    omega.l.i <- sig2.l.i + outer(psi.l.i,psi.l.i)
+    alpha.l.i <- c(1/sqrt(1-t(psi.l.i) %*% solve(omega.l.i) %*% psi.l.i)) * (diag(sqrt(diag(omega.l.i))) %*% solve(omega.l.i) %*% psi.l.i)
+
+    OMEGA.l[i,] <- c(omega.l.i)
+    ALPHA.l[i,] <- c(alpha.l.i)
+  }
+  OMEGA.list[[l]] <- OMEGA.l
+  ALPHA.list[[l]] <- ALPHA.l
+
+  sig2_true.l <- sig2.true.list[[l]]
+  psi_true.l <- psi.true.list[[l]]
+  omega_true.l <- sig2_true.l + outer(psi_true.l,psi_true.l)
+  alpha_true.l <- c(1/sqrt(1-t(psi_true.l) %*% solve(omega_true.l) %*% psi_true.l)) * (diag(sqrt(diag(omega_true.l))) %*% solve(omega_true.l) %*% psi_true.l)
+  omega.true.list[[l]] <- omega_true.l
+  alpha.true.list[[l]] <- alpha_true.l
+}
+```
+
+```{r}
 beta_true_c1 <- c(beta.true.list[[1]]) %>% round(.,2)
@@
 
-sig2_true_c1 <- c(sig2.true.list[[1]])[c(lower.tri(sig2.true.list[[1]],diag = TRUE))] %>% round(.,2)
-sig2_true_c2 <- c(sig2.true.list[[2]])[c(lower.tri(sig2.true.list[[2]],diag = TRUE))] %>% round(.,2)
-sig2_true_c3 <- c(sig2.true.list[[3]])[c(lower.tri(sig2.true.list[[3]],diag = TRUE))] %>% round(.,2)
-sig2_est_c1 <- col_summarize(SIGMA.list[[1]])[c(lower.tri(sig2.true.list[[1]],diag = TRUE))]
-sig2_est_c2 <- col_summarize(SIGMA.list[[2]])[c(lower.tri(sig2.true.list[[2]],diag = TRUE))]
-sig2_est_c3 <- col_summarize(SIGMA.list[[3]])[c(lower.tri(sig2.true.list[[3]],diag = TRUE))]
+omega_true_c1 <- c(omega.true.list[[1]])[c(lower.tri(omega.true.list[[1]],diag = TRUE))] %>% round(.,2)
+omega_true_c2 <- c(omega.true.list[[2]])[c(lower.tri(omega.true.list[[2]],diag = TRUE))] %>% round(.,2)
+omega_true_c3 <- c(omega.true.list[[3]])[c(lower.tri(omega.true.list[[3]],diag = TRUE))] %>% round(.,2)
+omega_est_c1 <- col_summarize(OMEGA.list[[1]])[c(lower.tri(omega.true.list[[1]],diag = TRUE))]
+omega_est_c2 <- col_summarize(OMEGA.list[[2]])[c(lower.tri(omega.true.list[[2]],diag = TRUE))]
+omega_est_c3 <- col_summarize(OMEGA.list[[3]])[c(lower.tri(omega.true.list[[3]],diag = TRUE))]
 
-psi_true_c1 <- psi.true.list[[1]] %>% round(.,2)
-psi_true_c2 <- psi.true.list[[2]] %>% round(.,2)
-psi_true_c3 <- psi.true.list[[3]] %>% round(.,2)
-psi_est_c1 <- col_summarize(PSI.list[[1]]) 
-psi_est_c2 <- col_summarize(PSI.list[[2]]) 
-psi_est_c3 <- col_summarize(PSI.list[[3]])
+alpha_true_c1 <- alpha.true.list[[1]] %>% round(.,2)
+alpha_true_c2 <- alpha.true.list[[2]] %>% round(.,2)
+alpha_true_c3 <- alpha.true.list[[3]] %>% round(.,2)
+alpha_est_c1 <- col_summarize(ALPHA.list[[1]]) 
+alpha_est_c2 <- col_summarize(ALPHA.list[[2]]) 
+alpha_est_c3 <- col_summarize(ALPHA.list[[3]])
 
@@
             "$\\beta_{12}$","$\\beta_{22}$","$\\beta_{32}$","$\\beta_{42}$",
-            "$\\sigma_{11}$","$\\sigma_{12}$","$\\sigma_{13}$","$\\sigma_{14}$",
-            "$\\sigma_{22}$","$\\sigma_{23}$","$\\sigma_{24}$",
-            "$\\sigma_{33}$","$\\sigma_{34}$",
-            "$\\sigma_{44}$",
-            "$\\psi_{1}$","$\\psi_{2}$", "$\\psi_{3}$","$\\psi_{4}$",
+            "$\\Omega_{11}$","$\\Omega_{12}$","$\\Omega_{13}$","$\\Omega_{14}$",
+            "$\\Omega_{22}$","$\\Omega_{23}$","$\\Omega_{24}$",
+            "$\\Omega_{33}$","$\\Omega_{34}$",
+            "$\\Omega_{44}$",
+            "$\\alpha_{1}$","$\\alpha_{2}$", "$\\alpha_{3}$","$\\alpha_{4}$",
             "$\\delta_{11}$","$\\delta_{12}$",
@@
 
-true_c1 <- c(beta_true_c1,sig2_true_c1,psi_true_c1,delta_true,pi_true[1])
-true_c2 <- c(beta_true_c2,sig2_true_c2,psi_true_c2,delta_true,pi_true[2])
-true_c3 <- c(beta_true_c3,sig2_true_c3,psi_true_c3,delta_true,pi_true[3])
-est_c1 <- c(beta_est_c1,sig2_est_c1,psi_est_c1,delta_est,pi_est[1])
-est_c2 <- c(beta_est_c2,sig2_est_c2,psi_est_c2,delta_est,pi_est[2])
-est_c3 <- c(beta_est_c3,sig2_est_c3,psi_est_c3,delta_est,pi_est[3])
+true_c1 <- c(beta_true_c1,omega_true_c1,alpha_true_c1,delta_true,pi_true[1])
+true_c2 <- c(beta_true_c2,omega_true_c2,alpha_true_c2,delta_true,pi_true[2])
+true_c3 <- c(beta_true_c3,omega_true_c3,alpha_true_c3,delta_true,pi_true[3])
+est_c1 <- c(beta_est_c1,omega_est_c1,alpha_est_c1,delta_est,pi_est[1])
+est_c2 <- c(beta_est_c2,omega_est_c2,alpha_est_c2,delta_est,pi_est[2])
+est_c3 <- c(beta_est_c3,omega_est_c3,alpha_est_c3,delta_est,pi_est[3])
+
+load("mcmc_draws_NOSKEW_2019-07-02/class_1_est")
+load("mcmc_draws_NOSKEW_2019-07-02/class_2_est")
+load("mcmc_draws_NOSKEW_2019-07-02/class_3_est")
 
@@
             params,
-            true_c1,est_c1,
-            true_c2,est_c2,
-            true_c3,est_c3)
-colnames(t1) <- c("Model Component",
-                  "Parameter",
-                  "True","Est. (95\\% CrI)",
-                  "True","Est. (95\\% CrI)",
-                  "True","Est. (95\\% CrI)")
+            true_c1,est_c1,est_c3_noskew,
+            true_c2,est_c2,est_c2_noskew,
+            true_c3,est_c3,est_c1_noskew)
+colnames(t1) <- c("Component",
+                  "Param.",
+                  "True","MSN Est. (95\\% CrI)","MN Est. (95\\% CrI) ",
+                  "True","MSN Est. (95\\% CrI)","MN Est. (95\\% CrI)",
+                  "True","MSN Est. (95\\% CrI)","MN Est. (95\\% CrI)")
 ```
@@
 kable(t1, booktabs = T, escape = F,caption = "Model results for simulated data with n = 1000, k = 4, p = 2, h = 3, r = 2.
-1000 iterations were run with a burn in of 250.
-Missingness mechanism was MAR and P(miss) = 0") %>%
-  kable_styling(font_size = 8) %>%
-  add_header_above(c(" " = 2,"Class 1" = 2, "Class 2" = 2, "Class 3" = 2)) %>%
+1000 iterations were run with a burn in of 100.
+Missingness mechanism was MAR and P(miss) = 0.
+Model results for the multivariate skew normal (MSN) and multivariate normal (MN) mixtures are presented.") %>%
+  kable_styling(font_size = 7) %>%
+  add_header_above(c(" " = 2,"Class 1" = 3, "Class 2" = 3, "Class 3" = 3)) %>%
   group_rows(" ",1,8) %>%
@@
   group_rows(" ",23,26) %>%
-  group_rows(" ",27,27)
+  group_rows(" ",27,27) %>%
+  landscape()
 ```

### 2019-07-02 manuscript/manuscript.tex
edits

@@
 \begin{abstract}
-In studies of infant motor development, a crucial research goal is to identify latent clusters of infants that experience delayed development, as this is a known risk factor for adverse outcomes later in life.
+In studies of infant motor development, a crucial research goal is the identification of latent clusters of infants who experience delayed development, as this is a known risk factor for adverse outcomes later in life.
 However, there are a number of statistical challenges in modeling infant development: the data are typically skewed, exhibit intermittent missingness, and are highly correlated across the repeated measurements collected during infancy.
-Using data from the Nurture study, a cohort of over 600 mother-infant pairs followed from pregnancy to 12 months postpartum, we develop a flexible Bayesian latent class model for the analysis infant motor development.
+Using data from the Nurture study, a cohort of over 600 mother-infant pairs followed from pregnancy to 12 months postpartum, we develop a flexible Bayesian finite mixture model for the analysis infant motor development.
 Our model has a number of attractive features.
@@
 Second, we model the cluster membership probabilities using a novel application of the P\'olya-Gamma data-augmentation scheme, thereby improving predictions of the cluster membership allocations.
-Lastly, we impute missing responses under missing at random assumption by drawing from appropriate conditional multivariate skew  normal distributions.
+Lastly, we impute missing responses under the missing at random assumption by drawing from appropriate conditional multivariate skew  normal distributions.
 Bayesian inference is achieved through straightforward Gibbs sampling, and can be implemented in available software such as \texttt{R}.
@@
 However, an important feature not included in the MSTFA is the ability to explain individual-level cluster membership as a function of covariates of interest.
-Additionally, parameter estimation proposed by Lin et al.
-for the MSTFA relies on a prohibitively complex EM algorithm and does not enjoy the inferential benefits of a Bayesian approach, namely the ability to incorporate prior information into a model and make posterior probability statements.
+Additionally, the parameter estimation procedure proposed by Lin et al.
+for the MSTFA relies on a prohibitively complex EM algorithm and does not enjoy the inferential benefits of a Bayesian approach, including the ability to incorporate prior information into a model and make posterior probability statements.
 Our proposed model improves on these previous works by estimating parameters in a Bayesian framework as well as including the ability to fit a multinomial logit regression to cluster membership probabilities using a novel application of data augmentation with the P\'olya-Gamma distribution.
@@
 We instead include an ``online" imputation imputation step in our Gibbs sampling procedure, whereby missing outcomes are updated at each iteration.
-This approach greatly increases the number of opportunities for exploration of the missing data parameter space.
+This approach greatly increases the number of opportunities for exploration of the missing data parameter space and avoids the multiplication of total run time and number of parameters.
 
@@
 The analysis of infant development trends in the Nurture data presents a number of statistical challenges that motivate our proposed model.
-First, as depicted in Figure \ref{fig:skew_resids}, the residuals from repeated measures models of Bayley composite scores exhibit skewness even after adjusting for covariates such as race, sex, and birthweight.
-This suggests that the assumption of conditional normality made by standard repeated measures models is violated, and a distinguishing feature of the data, namely skewness, is not being accounted for.
+First, as depicted in Figure 1, the residuals from repeated measures models of Bayley composite scores exhibit skewness even after adjusting for covariates such as race, sex, and birthweight.
+This suggests that the assumption of conditional normality made by standard repeated measures models is violated, and a distinguishing feature of the data, skewness, is not being accounted for.
 
@@
 Of the total cohort (N = 666), 429 (64.4 \%) observations were available at three months, 435 (65.3 \%) observations were available at six months, 418 (62.8 \%) observations were available at nine months, and 437 (65.6 \%) observations were available at twelve months.
+As such, we require a modeling framework capable of dealing with missing data.
 
@@
 
-A primary goal of the Nurture study is to identify clusters of infants characterized by distinct motor development trajectories.
+A primary goal of the Nurture study is to identify clusters of infants characterized by distinct motor development trajectories throughout the first year of life.
 To address this aim, we propose a flexible finite mixture model that accommodates relevant features of the data, such as skewness and dependence among the responses.
-For $i = 1,...,n$, let $\mathbf{y}_{i}=(y_{i1},\ldots,y_{iJ})^T$ be a $J \times 1$ vector of responses (i.e., Baley scores) for subject $i$ across the $J$ measurement occasions.
+For $i = 1,...,n$, let $\mathbf{y}_{i}=(y_{i1},\ldots,y_{iJ})^T$ be a $J \times 1$ vector of responses (i.e., Baley composite scores) for subject $i$ across the $J$ measurement occasions.
 For the analysis of the Nurture data, we propose a finite mixture model of the form
@@
 \]
-where $\text{MatNorm}_{n_k \times J}(\cdot)$ denotes a $n_k\times J$ matrix normal density, $\text{vec}(\mathbf{M}_k) = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$ is an $n_kJ\times 1$ mean vector, $\mathbf{X}_k$ is an $n_kJ \times Jp$ design matrix, $\boldsymbol\beta_k$ is the $(Jp)\times 1$ vector of regression coefficients defined in equation (\ref{eq:msnreg}), and $\boldsymbol\Sigma_k$ is the $J\times J$ conditional covariance of $\boldsymbol\epsilon_{ik}$ given in equation (\ref{eq:msnreg}).
+where $\text{MatNorm}_{n_k \times J}(\cdot)$ denotes a $n_k\times J$ matrix normal density, $\text{vec}(\mathbf{M}_k) = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$ is an $n_kJ\times 1$ mean vector, $\mathbf{X}_k$ is an $n_kJ \times Jp$ design matrix, $\boldsymbol\beta_k$ is the $Jp\times 1$ vector of regression coefficients defined in equation (\ref{eq:msnreg}), and $\boldsymbol\Sigma_k$ is the $J\times J$ conditional covariance of $\boldsymbol\epsilon_{ik}$ given in equation (\ref{eq:msnreg}).
 
@@
 For identifiability purposes, we fix the reference category $k = K$ and set $\boldsymbol\delta_K = \mathbf{0}$.
-Under this model, $z_i|\mathbf{w}_i \sim \text{Multinom}(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{i1},...,\pi_{1K})$.
+Under this model, $z_i|\mathbf{w}_i \sim \text{Multinomial}(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{i1},...,\pi_{1K})$.
 During MCMC estimation, the cluster labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as cluster assignments as detailed later in section (\textbf{what section}).
@@
 \end{equation}
-By assuming a normal prior for $\boldsymbol\delta_k$, as detailed in Section 3.4, the full conditional distribution for $\boldsymbol\delta_k$ is normal with parameters that follow standard forms.
+By assuming a normal prior for $\boldsymbol\delta_k$, as detailed in Section 3.4, the full conditional distribution for $\boldsymbol\delta_k$ is normal with parameters that follow standard forms, which we provide in Section (\textbf{what section}).
 \subsection{Conditional MSN Imputation}
 To accommodate missing at random (MAR) responses, we propose a convenient imputation algorithm that can be implemented ``online'' as part of the Gibbs sampler.
-In Section 6, we discuss extensions to allow for non-ignorable missingness.
+In Section 6, we discuss extensions to allow for non-ignorable missingness (i.e.
+observations missing not at random).
 Suppose $\mathbf{y}_i$ has $q_i\in (1,\ldots,J)$ observed values, denoted $\mathbf{y}^{obs}_i$, and $J-q_i$ intermittent missing values, denoted $\mathbf{y}^{miss}_i$.
@@
 \mathbf{y}^{miss}_i|(z_i=k,t_i,\mathbf{y}^{obs}_i)&\sim& \text{N}_{J-q_i}(\boldsymbol\mu^{cond}_{ki},\boldsymbol\Sigma^{cond}_k),~\text{where}\nonumber\\
-\boldsymbol\mu^{cond}_{ki}&=& \boldsymbol\mu^{miss}_{ki} + \boldsymbol\Sigma_{12} \boldsymbol\Sigma_{22}^{-1}(\mathbf{y}^{obs}_i - \boldsymbol\mu^{obs}_{ki}) \label{eq:impute}\\
-\boldsymbol\Sigma^{cond}_k&=& \boldsymbol\Sigma_{11} - \boldsymbol\Sigma_{12}\boldsymbol\Sigma_{22}^{-1}\boldsymbol\Sigma_{21},\nonumber \\
+\boldsymbol\mu^{cond}_{ki}&=& \boldsymbol\mu^{miss}_{ki} + \boldsymbol\Sigma_{k12} \boldsymbol\Sigma_{k22}^{-1}(\mathbf{y}^{obs}_i - \boldsymbol\mu^{obs}_{ki}) \label{eq:impute}\\
+\boldsymbol\Sigma^{cond}_k&=& \boldsymbol\Sigma_{k11} - \boldsymbol\Sigma_{k12}\boldsymbol\Sigma_{k22}^{-1}\boldsymbol\Sigma_{k21},\nonumber \\
 \boldsymbol\mu_{ki} &=& \begin{pmatrix} \boldsymbol\mu_{ki}^{miss} \\ \boldsymbol\mu_{ki}^{obs} \end{pmatrix}, \ \text{and} \nonumber \\
@@
 \end{eqnarray}
-$\boldsymbol\Sigma_{k11}$ is a $J-q_i \times J-q_i$ matrix containing the rows and columns of $\boldsymbol\Sigma_k$ corresponding to inicies of $\mathbf{y}_{i}$ where missingness occurs.
-Similarly, $\boldsymbol\Sigma_{k12}$ is a $J-q_i \times q_i$ matrix containing the rows of $\boldsymbol\Sigma_k$ that correspond to missing indices of $\mathbf{y}_i$, but columns of $\boldsymbol\Sigma_k$ that correspond to observed indices of $\mathbf{y}_i$.
+$\boldsymbol\Sigma_{k11}$ is a $(J-q_i) \times (J-q_i)$ matrix containing the rows and columns of $\boldsymbol\Sigma_k$ corresponding to inicies of $\mathbf{y}_{i}$ where missingness occurs.
+Similarly, $\boldsymbol\Sigma_{k12}$ is a $(J-q_i) \times q_i$ matrix containing the rows of $\boldsymbol\Sigma_k$ that correspond to missing indices of $\mathbf{y}_i$, but columns of $\boldsymbol\Sigma_k$ that correspond to observed indices of $\mathbf{y}_i$.
 The remaining partitions $\boldsymbol\Sigma_{k21}, \ \text{and} \ \boldsymbol\Sigma_{k22}$ are defined in the same manner.
@@
 We assign the skewness coefficients $\boldsymbol\psi_k$ a conjugate $\text{N}_J(\mathbf{m}_k,\mathbf{P}_k)$ prior.
-In practice, the updates of $\boldsymbol\beta_k$ and $\boldsymbol\psi_k$ can be combined into one step by defining the $(Jp + J) \times 1$ vector $\boldsymbol\beta^*_k = (\boldsymbol\beta_k^T,\boldsymbol\psi_k^T)^T$ for which we assume a $\text{N}_{Jp+J}(\mathbf{b}^*_k,\mathbf{I}_{(p+1)} \otimes \boldsymbol\Sigma_k)$ prior.
+In practice, the updates of $\boldsymbol\beta_k$ and $\boldsymbol\psi_k$ can be combined into one step by defining the $(Jp + J) \times 1$ vector $\boldsymbol\beta^*_k = (\boldsymbol\beta_k^T,\boldsymbol\psi_k^T)^T$ for which we assume a $\text{N}_{Jp+J}(\mathbf{b}^*_k,\mathbf{I}_{(p+1)} \otimes \boldsymbol\Sigma_k)$ prior, where $\mathbf{b}^*_k$ is formed by concatenating $\mathbf{b}_k$ and $\mathbf{m}_k$.
 
 To make use of the Matrix Normal representation introduced previously in Section (\textbf{what section}), we define the matrix of regression parameters $\mathbb{B}^*_k$, where $\text{vec}(\mathbb{B}^*_k) = \boldsymbol\beta^*_k$.
-We assign $\mathbb{B}^*_k|\boldsymbol\Sigma_k$ a $\text{MatNorm}(\mathbf{B}^*_{0k},\mathbf{I}_{p+1},\boldsymbol\Sigma_k)$ prior, where $\mathbf{B}^*_{0k}$ is a matrix of location parameters such that $\text{vec}(\mathbf{B}^*_{0k}) = \mathbf{b}^*_k$, $\mathbf{I}_{p+1}$ is the identity matrix of dimensions $(p+1)\times(p+1)$, and $\boldsymbol\Sigma_k$ is the model error.
+We assign $\mathbb{B}^*_k|\boldsymbol\Sigma_k$ a $\text{MatNorm}(\mathbf{B}^*_{0k},\mathbf{I}_{p+1},\boldsymbol\Sigma_k)$ prior, where $\mathbf{B}^*_{0k}$ is a matrix of location parameters such that $\text{vec}(\mathbf{B}^*_{0k}) = \mathbf{b}^*_k$, $\mathbf{I}_{p+1}$ is the identity matrix of dimensions $(p+1)\times(p+1)$, and $\boldsymbol\Sigma_k$ controls the correlation of the model error.
 This prior structure allows for convinient Matrix Normal updates of the MSN regression parameters as described in Section (\textbf{what section}).
@@
 A programatic sketch of our MCMC algorithm is given in Table (\ref{alg:MCMC}).
-Additional details, including derivations can be found in the Web Appendix.
+Additional details, including derivations of full conditionals and MCMC diagnostics can be found in the Web Appendix.
 
-\paragraph{\textbf{Step 1}: Conditional MSN Imputation} The sampler begins with imputing missing values $\mathbf{y}_i^{miss}$ conditional on current simulated values of $z_i = k$ and $t_i$ as well as the associated $\mathbf{y}_i^{obs}$.
+\paragraph{\textbf{Step 1}: Conditional MSN Imputation} The sampler begins with imputing missing values $\mathbf{y}_i^{miss}$ conditional on current  values of $z_i = k$ and $t_i$ as well as the associated $\mathbf{y}_i^{obs}$ observed data vector.
 This is accomplished by drawing $\mathbf{y}_i^{miss}$ from $\text{N}_{J-q_i}(\boldsymbol\mu^{cond}_{ki},\boldsymbol\Sigma^{cond}_k)$ as described in equation \ref{eq:impute} for all $i = 1,...,n$.
-Conditional imputation concludes with construction of a complete outcome vector $\mathbf{y}_i$ by merging the now imputed $\mathbf{y}_i^{miss}$ with $\mathbf{y}_i^{obs}$ with respect for the original order of $\mathbf{y}_i$.
+Conditional imputation concludes with construction of a complete outcome vector $\mathbf{y}_i$ by merging the now imputed $\mathbf{y}_i^{miss}$ with $\mathbf{y}_i^{obs}$ with respect for the original ordering of the elements of $\mathbf{y}_i$.
 
@@
 Next, define $U^*_{ki} = \frac{U_{ki} - 1/2}{{w}_{ki}}$ and let $\mathbf{U}^*_k = (U^*_{k1},...,U^*_{kn})^T$.
-Finally, for $k = 1,...,K-1$, update $\boldsymbol\delta_k$ from a $\text{N}_r(\mathbf{d}_k,\mathbf{S}_k)$, where $\mathbf{S}_k = (\mathbf{S}_{k0} + \mathbf{W}^T \mathbf{O}_k \mathbf{W})^{-1}$, $\mathbf{O}_k = \text{Diag}({w}_{k1},...,{w}_{kn})$,  $\mathbf{d}_k = \mathbf{S}_k (\mathbf{S}_{k0}\mathbf{d}_{k0} + \mathbf{W}^T \mathbf{O}_k \boldsymbol\gamma_k)$, and $\mathbf{W}$ is the $n \times r$ matrix of multinomial logit regression covariates such that the $i^{th}$ row of $\mathbf{W}$ is $\mathbf{w}_i$.
+Finally, for $k = 1,...,K-1$, update $\boldsymbol\delta_k$ from a $\text{N}_r(\mathbf{d}_k,\mathbf{S}_k)$ density, where $\mathbf{S}_k = (\mathbf{S}_{k0} + \mathbf{W}^T \mathbf{O}_k \mathbf{W})^{-1}$, $\mathbf{O}_k = \text{Diag}({w}_{k1},...,{w}_{kn})$,  $\mathbf{d}_k = \mathbf{S}_k (\mathbf{S}_{k0}\mathbf{d}_{k0} + \mathbf{W}^T \mathbf{O}_k \boldsymbol\gamma_k)$, and $\mathbf{W}$ is the $n \times r$ matrix of multinomial logit regression covariates such that the $i^{th}$ row of $\mathbf{W}$ is $\mathbf{w}_i$.
 
 Lastly, we update $z_1,...,z_n$ by first computing $\boldsymbol\pi_i = (\pi_{1i},...,\pi_{Ki})$ as
-$$\pi_{ik} = \frac{\text{e}^{\eta_{ik}}}{1 + \sum_{h=1}^{K-1} \text{e}^{\eta_{ih}}},$$ 
-where $\eta_{ik} = \mathbf{w}_i\boldsymbol\delta_k$ for $i = 1,...n$ and $k = 1,...,K$.
-We also compute according to the multivariate Normal density of $\mathbf{y}_i|t_i$ the probability $P(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)$, where $\boldsymbol\zeta_{ki} = \mathbf{x}_i^*\boldsymbol\beta^*_k$.
+$$\pi_{ik} = \frac{\text{e}^{\mathbf{w}_i\boldsymbol\delta_k}}{1 + \sum_{h=1}^{K-1} \text{e}^{\mathbf{w}_i\boldsymbol\delta_h}},$$ 
+for $i = 1,...n$ and $k = 1,...,K$.
+We also compute according to the multivariate Normal density of $\mathbf{y}_i|t_i$ the probability $P(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)$, where $\boldsymbol\zeta^*_{ki} = \mathbf{x}_i^*\boldsymbol\beta^*_k$.
 We use these quantities to compute $\boldsymbol\upsilon_{i} = (\upsilon_{1i},...,\upsilon_{Ki})$, where
@@
 
-See Algorithm \ref{alg:MCMC}.
+See Table 1.
 
@@
 
-A common challenge for Bayesian mixtures models is the so-called ``label switching" problem, in which draws of cluster-specific parameters may be assigned to the wrong cluster at some point during the MCMC simulation, rendering summaries of class-specific parameters incoherent.
+A common challenge for Bayesian mixture models is the so-called ``label switching" problem, in which draws of cluster-specific parameters may be assigned to the wrong cluster at some point during the MCMC simulation, rendering summaries of class-specific parameters incoherent.
 After conducting simulation studies under a wide variety of realistic parameter settings, as detailed in Section 4, we found little evidence of label switching in posterior draws of $\mathbf{z}$.
@@
 
-We model fit diagnostics, we choose the WAIC statistic introduced by Watanabe (2010).
+For model fit diagnostics, we choose the WAIC statistic introduced by Watanabe (2010).
 To specify the number of cluster $K$, we follow Neelon and Chung (2016), who use WAIC to determine the optimal number of latent factors in a Bayesian latent factor zero-inflated Poisson model for the analysis of correlated and zero-inflated count data.
@@
 \begin{eqnarray}
-	f(\mathbf{y}_i) = \sum_{k = 1}^K \pi_{ki} f(\mathbf{y}_i | \boldsymbol\theta_k), \nonumber
+	f(\mathbf{y}_i) = \sum_{k = 1}^3 \pi_{ki} f(\mathbf{y}_i | \boldsymbol\theta_k), \nonumber
 \end{eqnarray}
-where $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$, for $k \in \{1,2,3 \}$, and $\mathbf{y}_i | \boldsymbol\theta_k \sim \text{MSN}_4(\boldsymbol\zeta_{ki},\boldsymbol\alpha_k,\boldsymbol\Omega_k)$, where $\boldsymbol\zeta_{ki} = (\zeta_{ki1},\zeta_{ki2},\zeta_{ki3},\zeta_{ki4}) = \mathbf{X}_i \boldsymbol\beta_{k}$.
+where $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$, for $k \in \{1,2,3 \}$, and $\mathbf{y}_i | \boldsymbol\theta_k \sim \text{MSN}_4(\boldsymbol\zeta_{ki},\boldsymbol\alpha_k,\boldsymbol\Omega_k)$, where $\boldsymbol\zeta_{ki} = (\zeta_{ki1},\zeta_{ki2},\zeta_{ki3},\zeta_{ki4}) = \mathbf{x}_i \boldsymbol\beta_{k}$.
 We note from the above model specification that the number of clusters $K = 3$ and the number of measurement occasions $J = 4$.
 For the MSN regression model, we fit a main effect for time (i.e.
-measurement occasion) in addition to a baseline covariate, whose value does not vary with time, but whose effect does vary with time (\textbf{is this a weird way to say this?}).
-Under this setting with $p = 2$ covariates, the design matrix $\mathbf{X}_i$ for subject $i$ and the matrix of regression coefficients for cluster $k$, $\mathbf{B}_k$ are structured as
+measurement occasion) in addition to a baseline covariate, whose value does not vary with time, but whose effect does vary with time.
+Under this setting with $p = 2$ covariates, the design matrix $\mathbf{x}_i$ for subject $i$ and the matrix of regression coefficients for cluster $k$, $\mathbf{B}_k$ are structured as
 
 \begin{eqnarray}
-	\mathbf{X}_i = 
+	\mathbf{x}_i = 
 	\begin{bmatrix}
@@
 
-For cluster $k$, we can collect all observations $\mathbf{y}_i$ ($i = 1,...,n_k$) into the $n_k \times 4$ matrix $\mathbf{Y}_k$ and all covariates $\mathbf{X}_i$ into the $4n_k \times 8$ matrix $\mathbf{X}_k$ (\textbf{clash of notation here}) to utilize the matrix Normal model specification detailed in Section 3, which allows for simultaneous updates of all MSN regression parameters in cluster $k$.
+For cluster $k$, we can collect all observations $\mathbf{y}_i$ ($i = 1,...,n_k$) into the $n_k \times 4$ matrix $\mathbf{Y}_k$ and all covariates $\mathbf{x}_i$ into the $4n_k \times 8$ matrix $\mathbf{X}_k$ to utilize the matrix Normal model specification detailed in Section 3, which allows for simultaneous updates of all MSN regression parameters in cluster $k$.
 For the multinomial regression model component of this simulation, we model the class labels $z_i$ as a function of an intercept and one baseline covariate, thus $r = 2$.

### 2019-07-03 figures/figures_mcmc.R
edits

@@ -6,8 +6,8 @@ library(vapoRwave)
 
 vaporwave <- vapoRwave:::jazzCup_palette
 
-
-load_dir <- "mcmc_draws_2019-03-28"
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM")
+load_dir <- "mcmc_draws_2019-07-02"
 load(paste(load_dir,"/BETA",sep = ""))
 load(paste(load_dir,"/beta_true",sep = ""))
 load(paste(load_dir,"/DELTA",sep = ""))
@@ -73,7 +73,7 @@ ggsave(PSI_trace,filename = paste(load_dir,"psi_trace.pdf",sep = "/"))
 DELTA_trace <- mcmc_trace(DELTA) + 
     scale_color_manual(name = "Cluster",
                        values = vaporwave) + 
-    ggtitle("Trace plots of Psi Coefficients")
+    ggtitle("Trace plots of Delta Coefficients")
 ggsave(DELTA_trace,filename = paste(load_dir,"delta_trace.pdf",sep = "/"))
 
 ## PI plots

### 2019-07-03 simulation/simulation.R

@@ -152,7 +152,7 @@ for(l in 1:h)
 # Define prior structure
 print("Defining prior structure")
 a <- rep(2,h) # prior parameter vector for pi1,...,pih
-nu0 <- rep(10,h)  # scalar df hyperparam for Sigma (same across classes)
+nu0 <- rep(4,h)  # scalar df hyperparam for Sigma (same across classes)
 V0 <- sig2.true.ar1 # kxk hyperparam for Sigma (same across classes)
 # B0 <- rbind(beta.init.list[[1]],rep(0,k)) # zero matrix
 B0 <- betastar.true.list
@@ -161,8 +161,8 @@ delta0 <- rep(0,v) # prior mean for delta coefficients (multinomial regression)
 S0 <- diag(1,v) # prior covariance for delta coefficients (multinomial regression)
 
 # Sample storage
-nsim <- 1000 # number of iterations
-burn <- 100 # number of iterations to save
+nsim <- 10000 # number of iterations
+burn <- 2500 # number of iterations to save
 n.iter <- nsim - burn # number of saved iterations
 Z <- matrix(0,nrow = n.iter,ncol = n) # large matrix where each row is the value of z at a specific iteration
 PI <- matrix(0,nrow = n.iter,ncol = h) # matrix w/ each row as pi vector at each iteration
@@ -395,6 +395,8 @@ store <- paste("mcmc_draws_",run_date,sep = "")
 if(!dir.exists(store))
 {
     dir.create(store)
+} else {
+    store <- paste(store,"_2",sep = "")
 }
 meta_file <- paste(store,"/META.txt",sep = "")
 file.create(meta_file)

### 2019-07-03 manuscript/manuscript.tex
edits

@@
 \begin{keywords}
-A key word; But another key word; Still another key word; Yet another key word.
+Mixture of Experts, P\'olya-Gamma, Skew-Normal, Imputation, Latent Growth, Infant Development.
 \end{keywords}
@@
 \begin{equation}
+	\label{eq:mixture}
 f(\mathbf{y}_i) = \sum_{k = 1}^{K} \pi_{ki} f(\mathbf{y}_i|\boldsymbol\theta_k),
@@
 where 
-$\mathbf{X}_i$ is a $J \times Jp$ design matrix that includes potential time-varying covariates (e.g., indicators denoting quarterly visits); $\boldsymbol{\beta}_k=(\beta_{k11},\ldots,\beta_{k1p},\ldots,\beta_{kJ1},\ldots,\beta_{kJp})^T$ is a $Jp\times 1$ vector of cluster- and outcome-specific regression coefficients; $t_i\sim \text{N}_{[0,\infty)}(0,1)$ is a subject-specific standard normal random variable truncated below by zero; $\boldsymbol\psi_k=(\psi_{k1},\ldots,\psi_{kJ})^T$ is a $J \times 1$ vector of cluster-specific skewness parameters; and $\boldsymbol{\epsilon}_{ki} \sim \text{N}_J(\boldsymbol0,\boldsymbol\Sigma_k)$ is a $J\times 1$ vector of error terms.
+$\mathbf{X}_i$ is a $J \times Jp$ design matrix that includes potential time-varying covariates (e.g., indicators denoting quarterly visits); $\boldsymbol{\beta}_k=(\beta_{k11},\ldots,\beta_{k1p},\ldots,\beta_{kJ1},\ldots,\beta_{kJp})^T$ is a $Jp\times 1$ vector of cluster- and outcome-specific regression coefficients; $t_i\sim \text{N}_{[0,\infty)}(0,1)$ is a subject-specific standard normal random variable truncated below by zero; $\boldsymbol\psi_k=(\psi_{k1},\ldots,\psi_{kJ})^T$ is a $J \times 1$ vector of cluster-specific skewness parameters; and $\boldsymbol{\epsilon}_{ki} \sim \text{N}_J(\boldsymbol0,\boldsymbol\Sigma_k)$ is a $J\times 1$ vector of correlated error terms.
 Thus, conditional on $t_i$ and $z_i=k$, $\boldsymbol{y}_i$ is distributed as $\text{N}_J(\mathbf{X}_i \boldsymbol\beta_k + t_i \boldsymbol\psi_k, \boldsymbol{\Sigma}_k)$.
-Marginally (after integrating over $t_i$), $\mathbf{y}_i$ is distributed $\text{MSN}_J(\boldsymbol\zeta_{ki}, \boldsymbol\alpha_k, \boldsymbol\Omega_k)$, where through back-transformation
-\begin{eqnarray*}
-\boldsymbol\zeta_{ki} &=& \mathbf{X}_i\boldsymbol\beta_k,\\
+Marginally (after integrating over $t_i$), $\mathbf{y}_i|(z_i = k)$ is distributed $\text{MSN}_J(\boldsymbol\zeta_{ki}, \boldsymbol\alpha_k, \boldsymbol\Omega_k)$, where through back-transformation
+\begin{eqnarray}
+\boldsymbol\zeta_{ki} &=& \mathbf{X}_i\boldsymbol\beta_k, \nonumber \\
 \boldsymbol\alpha_k &=& \frac{1}{\sqrt{1 - \boldsymbol\psi_k^T 
-\boldsymbol\Omega^{-1}_k\boldsymbol\psi_k}} \boldsymbol\omega_k \boldsymbol\Omega^{-1}_k\boldsymbol\psi_k,~~\text{and}\\
-\boldsymbol\Omega_k &=& \boldsymbol\Sigma_k  + \boldsymbol\psi_k \boldsymbol\psi_k^T,
-\end{eqnarray*}
+\boldsymbol\Omega^{-1}_k\boldsymbol\psi_k}} \boldsymbol\omega_k \boldsymbol\Omega^{-1}_k\boldsymbol\psi_k,~~\text{and} \label{eqn:back_transform}\\
+\boldsymbol\Omega_k &=& \boldsymbol\Sigma_k  + \boldsymbol\psi_k \boldsymbol\psi_k^T \nonumber,
+\end{eqnarray}
 where $\boldsymbol\omega_k = \text{Diag}(\boldsymbol\Omega_k)^{1/2}$ is the $J \times J$ diagonal matrix containing the square root of the diagonal entries of $\boldsymbol\Omega_k$.
@@
 From equation (\ref{eq:msnreg}), it follows that $\mathbf{Y}_k$ is distributed as
-\begin{eqnarray*}
+\begin{eqnarray}
 \mathbf{Y}_{k} &\sim& \text{MatSN}_{n_k \times J}(\mathbf{M}_k,\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Omega_k)\\
 %\text{vec}(\mathbf{Y}_k) =  (\mathbf{y}^T_1,...,\mathbf{y}^T_{n_k})^T\\
-\text{vec}(\mathbf{M}_k) &=& (\boldsymbol\zeta_{k1}^T,...,\boldsymbol\zeta_{kn_k}^T)^T,
-\end{eqnarray*}
+\text{vec}(\mathbf{M}_k) &=& (\boldsymbol\zeta_{k1}^T,...,\boldsymbol\zeta_{kn_k}^T)^T, \nonumber
+\end{eqnarray}
 where $\boldsymbol\zeta_{ki} = \mathbf{X}_i \boldsymbol\beta_k$ as in equation (\ref{eq:msnreg}), $\boldsymbol\alpha_k = (\alpha_{k1},...,\alpha_{kJ})^T$, $\mathbf{I}_{n_k}$ is the $n_k\times n_k$ identity matrix, and $\boldsymbol\Omega_k$ is the $J\times J$ scale matrix defined above in equation (\ref{eq:msndens}).
 From equation (\ref{eq:msnreg}), it follows that $\mathbf{Y}_{k}$, conditional on the $n_k\times 1$ vector of random effects $\mathbf{t}_{k} = (t_1,...,t_{n_k})^T$, is jointly distributed in matrix form as
-\[
-\mathbf{Y}_k | \mathbf{t}_k \sim \text{MatNorm}_{n_k \times J}(\mathbf{M}_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k),
-\]
-where $\text{MatNorm}_{n_k \times J}(\cdot)$ denotes a $n_k\times J$ matrix normal density, $\text{vec}(\mathbf{M}_k) = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$ is an $n_kJ\times 1$ mean vector, $\mathbf{X}_k$ is an $n_kJ \times Jp$ design matrix, $\boldsymbol\beta_k$ is the $Jp\times 1$ vector of regression coefficients defined in equation (\ref{eq:msnreg}), and $\boldsymbol\Sigma_k$ is the $J\times J$ conditional covariance of $\boldsymbol\epsilon_{ik}$ given in equation (\ref{eq:msnreg}).
+\begin{equation}
+\mathbf{Y}_k | \mathbf{t}_k \sim \text{MatNorm}_{n_k \times J}(\mathbf{M}^*_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k),
+\end{equation}
+where $\text{MatNorm}_{n_k \times J}(\cdot)$ denotes a $n_k\times J$ matrix normal density, $\text{vec}(\mathbf{M}^*_k) = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$ is an $n_kJ\times 1$ mean vector, $\mathbf{X}_k$ is an $n_kJ \times Jp$ design matrix, $\boldsymbol\beta_k$ is the $Jp\times 1$ vector of regression coefficients defined in equation (\ref{eq:msnreg}), and $\boldsymbol\Sigma_k$ is the $J\times J$ conditional covariance of $\boldsymbol\epsilon_{ik}$ given in equation (\ref{eq:msnreg}).
 
@@
 During MCMC estimation, the cluster labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as cluster assignments as detailed later in section (\textbf{what section}).
-By allowing the cluster probabilities to vary across subjects, our model can be viewed as a \textit{mixture of experts} model in which $\pi_{ki}$ acts as a \textit{gating function} (\textbf{cite book}).
+By allowing the cluster probabilities to vary across subjects, our model can be viewed as a \textit{mixture of experts} model in which $\pi_{ki}$ acts as a \textit{gating function} controlling the prior probability of membership in cluster $k$, and $f(\mathbf{y}_i|\boldsymbol\theta_k)$ in Equation \ref{eq:mixture} is the ``expert'' (Bishop 2016).
 
 To facilitate sampling, we adopt an efficient data-augmentation approach introduced by Polson \textit{et al.} (2013), which expresses the inverse-logit function as a mixture P\'olya--Gamma densities.
-By using P\'olya--Gamma data augmentation for the multinomial model (\textbf{cite}), we obtain a P\'olya--Gamma mixture of experts model, a computationally efficient way to model edge weights (\textbf{cite}).
-A random variable ${w}$ is said to follow a P\'olya--Gamma distribution with parameters $b > 0$ and $c \in {\rm I\!R}$ if
+By using P\'olya--Gamma data augmentation for the multinomial model, we obtain a \textit{P\'olya--Gamma mixture of experts model}, a computationally efficient way to obtain inferences for the mixing weights in the Bayesian setting.
+A random variable ${w}$ is said to follow a P\'olya--Gamma distribution with parameters $b > 0$ and $c \in \mathbb{R}$ if
 \begin{equation}
@@
 \end{equation}
-where $g_s \stackrel{iid}{\sim} \text{Ga}(b,1)$ for $s = 1,...,\infty$ . Polson \textit{et al.} (2013) establish that for the $\text{PG}(b,c)$ density, and for $a,\eta \in {\rm I\!R}$,
+where $g_s \stackrel{iid}{\sim} \text{Ga}(b,1)$ for $s = 1,...,\infty$ . Polson \textit{et al.} (2013) establish that for the $\text{PG}(b,c)$ density, and for $a,\eta \in \mathbb{R}$,
 
@@
 $$p(\boldsymbol\delta_k|\mathbf{U}_k,\boldsymbol\delta_{h \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \pi_{ki}^{U_{ki}}(1-\pi_{ki})^{1-U_{ki}},$$
-where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, $\mathbf{U}_k = (U_{k1},...,U_{kn})$, $U_{ik} = {1}_{z_i = k}$ is an indicator that subject $i$ belongs to cluster $k$, and $\pi_{ki}$ is defined as in Section 3.4.
+where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, $\mathbf{U}_k = (U_{k1},...,U_{kn})$, $U_{ik} = {1}_{(z_i = k)}$ is an indicator that subject $i$ belongs to cluster $k$, and $\pi_{ki}$ is defined as in Section 3.4.
 We can rewrite $\pi_{ki}$ as follows
@@
 which is essentially a logistic regression likelihood.
-We thus apply this P\'olya--Gamma data augmentation scheme from Polson et al.
-(2013) to update each $\boldsymbol\delta_k$ ($k = 1,...,K-1$) one at a time based on the binary indicators $U_{ki}$.
-To accomplish this, we define $\mathbf{U}^*_{k} = \left( \frac{U_{k1}-1/2}{w_{k1}},...,\frac{U_{kn}-1/2}{w_{kn}} \right )$.
+We can thus apply the P\'olya--Gamma sampler described in Polson et al.
+(2013) for logistic regression to update each $\boldsymbol\delta_k$ one at a time based on the binary indicators $U_{ki}$.
+To accomplish this, we first define for $k = 1,...,k$, the $n \times 1$ vector $\mathbf{U}^*_{k} = \left( \frac{U_{k1}-1/2}{w_{k1}},...,\frac{U_{kn}-1/2}{w_{kn}} \right )^T$.
 Polson et al.
-(2013) show that $\mathbf{U}^*_{k}$ follows a multivariate normal distribution with mean $\boldsymbol\eta_k = (\eta_{k1},...,\eta_{kn})$ and covariance matrix $\mathbf{O}_k = \text{Diag}(w_{k1},...,w_{kn})$.
+(2013) show that, conditional on $\boldsymbol{w} = (w_{k1},...,w_{kn})^T$, $\mathbf{U}^*_{k}$ follows a $\text{N}_n (\boldsymbol\eta_k,\mathbf{O}_k^{-1})$ distribution with mean $\boldsymbol\eta_k = (\eta_{k1},...,\eta_{kn})^T$ and precision matrix $\mathbf{O}_k = \text{Diag}(w_{k1},...,w_{kn})$.
 Thus, it follows that the full conditional distribution of $\boldsymbol\delta_k$ is given by
 \begin{equation}
-	p(\boldsymbol\delta_k|\mathbf{z},\mathbf{W}, \mathbf{O}_k) \propto p_0(\boldsymbol\delta_k) \text{exp} \left [ - \frac{1}{2} (\mathbf{U}^*_k - \mathbf{W}\boldsymbol\delta_k)^T \mathbf{O}_k (\mathbf{U}^*_k - \mathbf{W}\boldsymbol\delta_k)\right].
+	p(\boldsymbol\delta_k|\mathbf{z},\mathbf{W}, \mathbf{O}_k) \propto p(\boldsymbol\delta_k) \text{exp} \left [ - \frac{1}{2} (\mathbf{U}^*_k - \mathbf{W}\boldsymbol\delta_k)^T \mathbf{O}_k (\mathbf{U}^*_k - \mathbf{W}\boldsymbol\delta_k)\right].
 \end{equation}
@@
 
+\bibitem{ } Bishop, CM. Pattern recognition and machine learning. Springer; 2006.
+
 \bibitem{ } Chen JT, Gupta AK. Matrix variate skew normal distributions. \textit{Statistics}. 2005 Jun 1;39(3):247-53.

### 2019-07-05 manuscript/manuscript.tex
edits

@@
 
-\vspace{1mm}
-\begin{figure}[h]
-	\label{fig:skew_resids}
-	\caption{The distribution of residuals in repeated measures regression model of Bayley composite scores adjusted for race and sex.}
-	\centering
-	\includegraphics[scale = 0.14]{bayley_resids_plot}
-\end{figure}
-
 The Nurture data also feature intermittent missingness in Bayley composite scores throughout the study period.
@@
 \begin{eqnarray}
-\mathbf{Y}_{k} &\sim& \text{MatSN}_{n_k \times J}(\mathbf{M}_k,\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Omega_k)\\
+\mathbf{Y}_{k} &\sim& \text{MatSN}_{n_k \times J}(\mathbf{M}_k,\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Omega_k)\label{eq:matsn}\\
 %\text{vec}(\mathbf{Y}_k) =  (\mathbf{y}^T_1,...,\mathbf{y}^T_{n_k})^T\\
@@
 \begin{equation}
-	p(\boldsymbol\delta_k|\mathbf{z},\mathbf{W}, \mathbf{O}_k) \propto p(\boldsymbol\delta_k) \text{exp} \left [ - \frac{1}{2} (\mathbf{U}^*_k - \mathbf{W}\boldsymbol\delta_k)^T \mathbf{O}_k (\mathbf{U}^*_k - \mathbf{W}\boldsymbol\delta_k)\right].
+	p(\boldsymbol\delta_k|\mathbf{z},\mathbf{W}, \mathbf{O}_k) \propto p(\boldsymbol\delta_k) \text{exp} \left [ - \frac{1}{2} (\mathbf{U}^*_k - \mathbf{W}\boldsymbol\delta_k)^T \mathbf{O}_k (\mathbf{U}^*_k - \mathbf{W}\boldsymbol\delta_k)\right],
 \end{equation}
-By assuming a normal prior for $\boldsymbol\delta_k$, as detailed in Section 3.4, the full conditional distribution for $\boldsymbol\delta_k$ is normal with parameters that follow standard forms, which we provide in Section (\textbf{what section}).
+where $p(\boldsymbol\delta_k)$ is the prior distribution for $\boldsymbol\delta_k$, which we detail in Section 3.4 As detailed in Section 3.4, if we assume $p(\boldsymbol\delta_k)$ to be multivariate normal, then $\boldsymbol\delta_k$ has a closed-form multivariate normal full conditional distribution that can be easily embedded within our proposed Gibbs sampling routine.
 \subsection{Conditional MSN Imputation}
@@
 \begin{eqnarray}
-\mathbf{y}^{miss}_i|(z_i=k,t_i,\mathbf{y}^{obs}_i)&\sim& \text{N}_{J-q_i}(\boldsymbol\mu^{cond}_{ki},\boldsymbol\Sigma^{cond}_k),~\text{where}\nonumber\\
-\boldsymbol\mu^{cond}_{ki}&=& \boldsymbol\mu^{miss}_{ki} + \boldsymbol\Sigma_{k12} \boldsymbol\Sigma_{k22}^{-1}(\mathbf{y}^{obs}_i - \boldsymbol\mu^{obs}_{ki}) \label{eq:impute}\\
+\mathbf{y}^{miss}_i|(z_i=k,t_i,\mathbf{y}^{obs}_i)&\sim& \text{N}_{J-q_i}(\boldsymbol\zeta^{cond}_{ki},\boldsymbol\Sigma^{cond}_k),~\text{where}\nonumber\\
+\boldsymbol\zeta^{cond}_{ki}&=& \boldsymbol\zeta^{miss}_{ki} + \boldsymbol\Sigma_{k12} \boldsymbol\Sigma_{k22}^{-1}(\mathbf{y}^{obs}_i - \boldsymbol\zeta^{obs}_{ki}) \label{eq:impute}\\
 \boldsymbol\Sigma^{cond}_k&=& \boldsymbol\Sigma_{k11} - \boldsymbol\Sigma_{k12}\boldsymbol\Sigma_{k22}^{-1}\boldsymbol\Sigma_{k21},\nonumber \\
-\boldsymbol\mu_{ki} &=& \begin{pmatrix} \boldsymbol\mu_{ki}^{miss} \\ \boldsymbol\mu_{ki}^{obs} \end{pmatrix}, \ \text{and} \nonumber \\
+\boldsymbol\zeta_{ki} &=& \begin{pmatrix} \boldsymbol\zeta_{ki}^{miss} \\ \boldsymbol\zeta_{ki}^{obs} \end{pmatrix}, \ \text{and} \nonumber \\
 \boldsymbol\Sigma_k &=& \begin{pmatrix} 
@@
 The remaining partitions $\boldsymbol\Sigma_{k21}, \ \text{and} \ \boldsymbol\Sigma_{k22}$ are defined in the same manner.
-Likewise, $\boldsymbol\mu_{ki} = \mathbf{X}_i \boldsymbol\beta_k$ is partitioned into $\boldsymbol\mu_{ki}^{miss}$ and $\boldsymbol\mu_{ki}^{obs}$ with respect to the missing and observed indices of $\mathbf{y}_i$, respectively.
+Likewise, as defined in equation \ref{eq:msndens}, $\boldsymbol\zeta_{ki} = \mathbf{X}_i \boldsymbol\beta_k + t_i \boldsymbol\psi_k$ and is partitioned into $\boldsymbol\zeta_{ki}^{miss}$ and $\boldsymbol\zeta_{ki}^{obs}$ with respect to the missing and observed indices of $\mathbf{y}_i$, respectively.
 These results follow from conventional multivariate normal theory, which we detail in the Web Appendix.
@@
 Conveniently, all parameters admit conditionally conjugate priors, which greatly improves posterior computation via a data-augmented Gibbs sampler.
-For the MSN model component, we adopt a conditionally independent prior structure for $\boldsymbol\beta_k$ and $\boldsymbol\Sigma_k$, where $p(\boldsymbol\beta_k,\boldsymbol\Sigma_k) = p(\boldsymbol\Sigma_k)p(\boldsymbol\beta_k|\boldsymbol\Sigma_k)$.
+For $\boldsymbol\beta_k$ and $\boldsymbol\Sigma_k$, we adopt a conditionally independent prior structure for $\boldsymbol\beta_k$ and $\boldsymbol\Sigma_k$, where $p(\boldsymbol\beta_k,\boldsymbol\Sigma_k) = p(\boldsymbol\Sigma_k)p(\boldsymbol\beta_k|\boldsymbol\Sigma_k)$.
 We choose the normal-inverse-Wishart distribution for $p(\boldsymbol\beta_k,\boldsymbol\Sigma_k)$ by specifying $\boldsymbol\Sigma_k \sim \text{IW}(\mathbf{V}_{0k},\nu_{0k})$ and $\boldsymbol\beta_k|\boldsymbol\Sigma_k \sim \text{N}_{Jp}(\mathbf{b}_k,\mathbf{I}_p \otimes \boldsymbol\Sigma_k)$.
@@
 
-To make use of the Matrix Normal representation introduced previously in Section (\textbf{what section}), we define the matrix of regression parameters $\mathbb{B}^*_k$, where $\text{vec}(\mathbb{B}^*_k) = \boldsymbol\beta^*_k$.
-We assign $\mathbb{B}^*_k|\boldsymbol\Sigma_k$ a $\text{MatNorm}(\mathbf{B}^*_{0k},\mathbf{I}_{p+1},\boldsymbol\Sigma_k)$ prior, where $\mathbf{B}^*_{0k}$ is a matrix of location parameters such that $\text{vec}(\mathbf{B}^*_{0k}) = \mathbf{b}^*_k$, $\mathbf{I}_{p+1}$ is the identity matrix of dimensions $(p+1)\times(p+1)$, and $\boldsymbol\Sigma_k$ controls the correlation of the model error.
-This prior structure allows for convinient Matrix Normal updates of the MSN regression parameters as described in Section (\textbf{what section}).
+To make use of the Matrix Normal representation introduced previously in Section 3, we define the matrix of regression parameters $\mathbb{B}^*_k$, where $\text{vec}(\mathbb{B}^*_k) = \boldsymbol\beta^*_k$.
+We assign $\mathbb{B}^*_k|\boldsymbol\Sigma_k$ a $\text{MatNorm}(\mathbf{B}^*_{0k},\mathbf{I}_{p+1},\boldsymbol\Sigma_k)$ prior, where $\mathbf{B}^*_{0k}$ is a matrix of location parameters such that $\text{vec}(\mathbf{B}^*_{0k}) = \mathbf{b}^*_k$, $\mathbf{I}_{p+1}$ is the $(p+1)$-dimension identity matrix, and $\boldsymbol\Sigma_k$ is the covariance matrix defined in equation \ref{eq:msndens}.
+The matrix-Normal-inverse-Wishart is the conjugate joint prior for the regression parameters in the matrix Normal model given in equation \ref{eq:matsn}.
+This induces convenient closed-form full conditional distributions that can be easily updated within our proposed Gibbs sampler.
 
@@
 
-\paragraph{\textbf{Step 1}: Conditional MSN Imputation} The sampler begins with imputing missing values $\mathbf{y}_i^{miss}$ conditional on current  values of $z_i = k$ and $t_i$ as well as the associated $\mathbf{y}_i^{obs}$ observed data vector.
-This is accomplished by drawing $\mathbf{y}_i^{miss}$ from $\text{N}_{J-q_i}(\boldsymbol\mu^{cond}_{ki},\boldsymbol\Sigma^{cond}_k)$ as described in equation \ref{eq:impute} for all $i = 1,...,n$.
-Conditional imputation concludes with construction of a complete outcome vector $\mathbf{y}_i$ by merging the now imputed $\mathbf{y}_i^{miss}$ with $\mathbf{y}_i^{obs}$ with respect for the original ordering of the elements of $\mathbf{y}_i$.
+\paragraph{\textbf{Step 1}: Conditional MSN Imputation} The sampler begins by imputing missing values $\mathbf{y}_i^{miss}$ conditional on current values of $z_i = k$ and $t_i$ as well as the associated $\mathbf{y}_i^{obs}$ observed data vector.
+Specifically, we draw $\mathbf{y}_i^{miss}$ from $\text{N}_{J-q_i}(\boldsymbol\mu^{cond}_{ki},\boldsymbol\Sigma^{cond}_k)$ as described in equation \ref{eq:impute} for all $i = 1,...,n$.
+We conclude by constructing a complete outcome vector $\mathbf{y}_i$ that merges $\mathbf{y}_i^{miss}$ with $\mathbf{y}_i^{obs}$.
 
@@
 For cluster $k$, compute $A_k = (1 + \boldsymbol\psi_k^T \boldsymbol\Sigma^{-1}_k\boldsymbol
-\psi_k)^{-1}$ according to current values of $\boldsymbol\psi_k$ and $\boldsymbol\Sigma_k$.
-Now, for $i = 1,...,n_k$, where $n_k$ is the number of members of cluster $k$ according to the current cluster labeling ${z}_1,...,{z}_n$, compute $a_i = A_k \boldsymbol\psi_k^T \boldsymbol\Sigma^{-1}_k(\mathbf{y}_i - \boldsymbol\zeta_{ik})$, where $\boldsymbol\zeta_{ik} = X_i \boldsymbol\beta_k$.
-Finally, update $t_i$ from $\text{N}_{[0,\infty)}(a_i, \mathbf{A}_k)$ (\textbf{put derivations in supplement}).
-Repeat these updates for $k = 1,...,K$
+\psi_k)^{-1}$ using current values of $\boldsymbol\psi_k$ and $\boldsymbol\Sigma_k$.
+Next, for $i = 1,...,n_k$, compute $a_i = A_k \boldsymbol\psi_k^T \boldsymbol\Sigma^{-1}_k(\mathbf{y}_i - \boldsymbol\zeta_{ik})$, where $\boldsymbol\zeta_{ik} = X_i \boldsymbol\beta_k$.
+Finally, update $t_i$ from $\text{N}_{[0,\infty)}(a_i, \mathbf{A}_k)$.
+Repeat these updates for $k = 1,...,K$.
 
 The remaining MSN regression parameters are updated from their full conditionals as follows.
-For each cluster $k$ ($k = 1,...,K$) we update $\boldsymbol\Sigma_k$ from $\text{IW}(\nu_k,\mathbf{V}_k)$, where $\nu_k = \nu_{0k} + n_k$ and\begin{eqnarray}
-	\mathbf{V}_k &=&  (\mathbf{Y}-\mathbf{X}^*_k \mathbf{B}^*_k)^T (\mathbf{Y}_k - \mathbf{X}^*_k \mathbf{B}^*_k)\nonumber \\
-	 &+& (\mathbf{B}_k^* - \mathbf{B}^*_{0k})^T \mathbf{I}_{p+1}(\mathbf{B}^*_k - \mathbf{B}^*_{0k}) + \mathbf{V}_{0k}. \nonumber
-\end{eqnarray}We then make use of the Matrix Normal representation introduced in Section 3.2 to draw $\mathbb{B}^*_k$ from a $\text{MatNorm}_{p+1,J}(\mathbf{B}_k^*,\mathbf{L}_k^{-1},\boldsymbol\Sigma_k)$ density, where $\text{vec}(\mathbb{B}^*_k) = \boldsymbol\beta^*_k = (\beta_{k11},...,\beta_{kJp},\psi_{k1},...,\psi_{kJ})^T$ and
+First, we form the $n_k \times (p + 1)$ matrix $\mathbf{X}^*_k$ by column binding $\mathbf{X}_k$ and $\mathbf{t}_k$.
+For each cluster $k$ ($k = 1,...,K$), we update $\boldsymbol\Sigma_k$ from an $\text{IW}(\nu_k,\mathbf{V}_k)$ density, where $\nu_k = \nu_{0k} + n_k$ and\begin{eqnarray}
+	\mathbf{V}_k &=&  (\mathbf{Y}-\mathbf{X}^*_k \mathbf{B}^*_k)^T (\mathbf{Y}_k - \mathbf{X}^*_k \mathbf{B}^*_k) (\mathbf{B}_k^* - \mathbf{B}^*_{0k})^T \mathbf{I}_{p+1}(\mathbf{B}^*_k - \mathbf{B}^*_{0k}) + \mathbf{V}_{0k}.
+\nonumber
+\end{eqnarray} The matrix $\mathbf{X}^*_k$ is formed by adding We then make use of the Matrix Normal representation introduced in Section 3.2 to draw $\mathbb{B}^*_k$ from a $\text{MatNorm}_{p+1,J}(\mathbf{B}_k^*,\mathbf{L}_k^{*},\boldsymbol\Sigma_k)$ density, where $\text{vec}(\mathbb{B}^*_k) = \boldsymbol\beta^*_k = (\beta_{k11},...,\beta_{kJp},\psi_{k1},...,\psi_{kJ})^T$ and
 \begin{eqnarray}
-	\mathbf{B}^*_k &=& (\mathbf{X}^{*T}_k \mathbf{X}^*_k + \mathbf{I}_{p+1})^{-1} (\mathbf{X}^{*T}_k \mathbf{Y}_k + \mathbf{I}_{p+1} \mathbf{B}_{0k}) \nonumber \\
-	\mathbf{L}^*_k &=& \mathbf{X}^{*T}_k \mathbf{X}^*_k + \mathbf{I}_{p+1}. \nonumber
+	\mathbf{B}^*_k &=& \mathbf{L}_k^{*} (\mathbf{X}^{*T}_k \mathbf{Y}_k + \mathbf{I}_{p+1} \mathbf{B}^*_{0k}) \nonumber \\
+	\mathbf{L}^*_k &=& (\mathbf{X}^{*T}_k \mathbf{X}^*_k + \mathbf{I}_{p+1})^{-1}. \nonumber
 \end{eqnarray}
@@
 Lastly, we update $z_1,...,z_n$ by first computing $\boldsymbol\pi_i = (\pi_{1i},...,\pi_{Ki})$ as
-$$\pi_{ik} = \frac{\text{e}^{\mathbf{w}_i\boldsymbol\delta_k}}{1 + \sum_{h=1}^{K-1} \text{e}^{\mathbf{w}_i\boldsymbol\delta_h}},$$ 
+$$\pi_{ki} = \frac{\text{e}^{\mathbf{w}_i\boldsymbol\delta_k}}{1 + \sum_{h=1}^{K-1} \text{e}^{\mathbf{w}_i\boldsymbol\delta_h}},$$ 
 for $i = 1,...n$ and $k = 1,...,K$.
@@
 We use these quantities to compute $\boldsymbol\upsilon_{i} = (\upsilon_{1i},...,\upsilon_{Ki})$, where
-$$\upsilon_{ik} = P(z_i = k|\mathbf{y}_i,\boldsymbol\zeta^*_i,\boldsymbol\Sigma_k) = \frac{\pi_{ik}P(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)}{\sum_{h = 1}^{K} \pi_{ih}P(\mathbf{y}_i|\boldsymbol\zeta^*_{hi},\boldsymbol\Sigma_h)}$$.
-The cluster labels $z_i$ are updated from a $\text{Multinomial}(1,\boldsymbol\upsilon_i)$ density for $i = 1,...n$, thus concluding the Gibbs sampler.
+$$\upsilon_{ki} = P(z_i = k|\mathbf{y}_i,\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k) = \frac{\pi_{ki}P(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)}{\sum_{h = 1}^{K} \pi_{hi}P(\mathbf{y}_i|\boldsymbol\zeta^*_{hi},\boldsymbol\Sigma_h)}.$$
+The cluster labels $z_i$ are then updated from a $\text{Multinomial}(1,\boldsymbol\upsilon_i)$ density for $i = 1,...n$. A schematic outline of the Gibbs sampler is given in the Web Appendix.
 
-\subsubsection{MCMC Algorithm}
-
-See Table 1.
-
-\begin{algorithm}
-\caption{Gibbs Sampler}
-\label{alg:MCMC}
-\begin{algorithmic}
-\small
-\linespread{0.5}
-    \State $\mathbf{Define} \ n_{iter};\  n_{burn}; \ K; \ \boldsymbol\theta_{init}; \ \boldsymbol\theta_0$
-    \State $n_{sim} := n_{iter} - n_{burn}$
-    \State $\boldsymbol\theta := \boldsymbol\theta_{init}$
-    \For {$\iota = 1,...,n_{sim}$}
-        \State \textsc{I. Conditional Imputation}
-          \For {$ i = 1,...,n$}
-            \State $\mathbf{Draw} \ \mathbf{y}_i^{miss} \ \text{from} \ N_q(\boldsymbol\mu_i^{miss}, \boldsymbol\Sigma_i^{miss})$
-            \State $\mathbf{y}_i := \mathbf{y}_i^{miss} \cup \mathbf{y}_i^{obs}$
-         \EndFor
-        \State \textsc{II. MSN Regression}
-          \For {$ k = 1,...,K$}
-            \State $n_k := \sum_{i = 1}^n {1}_{z_i = k}$
-            \For {$i_k = 1,...,n_k$}
-              \State $\mathbf{Draw}\  t_i \ \text{from} \ N_{[0,\infty)}(a_i,A)$
-            \EndFor
-            \State $\mathbf{X^*}_k := $ \texttt{cbind}$(\mathbf{X}_k,\mathbf{t}_k)$
-            \State $\mathbf{Draw} \ \mathbf{B^*}_k \ \text{from} \ \text{MatNorm}(\mathbf{B}_k,\mathbf{L}_k^{-1},\boldsymbol\Sigma_k)$
-            \State $\mathbf{Draw} \ \boldsymbol\Sigma_k \ \text{from} \ \text{InvWish}(\nu_k, \mathbf{V}_k)$
-          \EndFor
-        \State \textsc{III. Multinomial Logit}
-          \For {$i = 1,...,n$}
-            \For {$k = 1,...,K$}
-              \State $\pi_{ik} := P(z_i = k|\mathbf{w}_i,\boldsymbol\delta_k)$
-              \State $p_{ik} := P(\mathbf{y}_i|\boldsymbol\beta_k^{*T} \mathbf{x}^*_i,\boldsymbol\Sigma_k)$
-            \EndFor
-            \State $\mathbf{p}_{z_i} := \frac{\mathbf{p}_i \circ \boldsymbol\pi_i}{\mathbf{p}_i \cdot \boldsymbol\pi_i}$
-            \State $\mathbf{Draw} \ z_i \ \text{from} \ \text{Categorical}(\mathbf{p}_{z_i})$
-            \For {$k = 1,...,K-1$}
-              \State $\mathbf{Draw} \ \boldsymbol\delta_k \ \text{from} \ N(\mathbf{M},\mathbf{S})$
-            \EndFor
-          \EndFor
-        \State $\boldsymbol\theta := \{\mathbf{B^*}, \boldsymbol\Sigma, \mathbf{Z}, \boldsymbol\delta \}$
-        \State $\mathbf{Store} \ \boldsymbol\theta$
-	  \EndFor
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{Assessment of MCMC Convergence}
+\subsubsection{Assessment of MCMC Convergence, Label Switching, and Model Selection}
 
@@
 
-\subsubsection{Label Switching}
+A common challenge for Bayesian mixture models is the so-called ``label switching" problem, in which draws of cluster-specific parameters may be assigned to the wrong cluster at some point during the MCMC simulation, rendering summaries of class-specific parameters incoherent.
+After conducting simulation studies under a wide variety of realistic parameter settings, as detailed in Sections 4 and 5, we found little evidence of label switching in posterior draws of $\mathbf{z}$.
+When label switching was observed, we implemented \textit{post hoc} relabling algorithms as described in Papastamoulis 2016 and implemented in the \texttt{label.switching} package in \texttt{R} (Papastamoulis 2016).
 
-A common challenge for Bayesian mixture models is the so-called ``label switching" problem, in which draws of cluster-specific parameters may be assigned to the wrong cluster at some point during the MCMC simulation, rendering summaries of class-specific parameters incoherent.
-After conducting simulation studies under a wide variety of realistic parameter settings, as detailed in Section 4, we found little evidence of label switching in posterior draws of $\mathbf{z}$.
-When label switching was observed, we implemented \textit{post hoc} relabling algorithms as described in Papastamoulis 2016 and implemented in the \texttt{label.switching} package in \texttt{R}.
-
-\subsubsection{Model Fit Diagnostics and Specification of K}
-
-For model fit diagnostics, we choose the WAIC statistic introduced by Watanabe (2010).
-To specify the number of cluster $K$, we follow Neelon and Chung (2016), who use WAIC to determine the optimal number of latent factors in a Bayesian latent factor zero-inflated Poisson model for the analysis of correlated and zero-inflated count data.
-In Section 6, we discuss possible extentions to other methods for choosing $K$.
+Since the primary objective of our proposed model is to provide a flexible framework to account for the inherent features of infant development data such as skewness, correlation among repeated measures, and missingness, we assume the number of clusters $K$ to be specified \textit{a priori}.
 

### 2019-07-05 mcmc_draws_2019-07-02_2/tables.Rmd
edits

--- 2019-07-01 mcmc_draws_2019-07-01/tables.Rmd
@@
 ```{r}
-setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-01")
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-02_2")
 load("BETA")

### 2019-07-05 mcmc_draws_2019-07-04/tables.Rmd
edits

--- 2019-07-01 mcmc_draws_2019-07-01/tables.Rmd
@@
 ```{r}
-setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-01")
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-04")
 load("BETA")

### 2019-07-05 figures/figures_mcmc.R

@@ -7,7 +7,7 @@ library(vapoRwave)
 vaporwave <- vapoRwave:::jazzCup_palette
 
 setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM")
-load_dir <- "mcmc_draws_2019-07-02"
+load_dir <- "mcmc_draws_2019-07-04"
 load(paste(load_dir,"/BETA",sep = ""))
 load(paste(load_dir,"/beta_true",sep = ""))
 load(paste(load_dir,"/DELTA",sep = ""))
@@ -25,7 +25,7 @@ k <- ncol(beta.true.list[[1]])
 p <- nrow(beta.true.list[[1]])
 v <- nrow(delta.true)
 
-burn <- 100
+burn <- 250
 
 colnames(Z) <- paste("Z",seq(1,ncol(Z)),sep = "")
 Z <- Z[-(1:burn),]

### 2019-07-05 mcmc_draws_2019-07-02_2/figures_mcmc.R

@@ -1,0 +1,94 @@
+# Plot MCMC draws
+
+library(bayesplot)
+library(tidyverse)
+library(vapoRwave)
+
+vaporwave <- vapoRwave:::jazzCup_palette
+
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM")
+load_dir <- "mcmc_draws_2019-07-02_2"
+load(paste(load_dir,"/BETA",sep = ""))
+load(paste(load_dir,"/beta_true",sep = ""))
+load(paste(load_dir,"/DELTA",sep = ""))
+load(paste(load_dir,"/delta_true",sep = ""))
+load(paste(load_dir,"/PI",sep = ""))
+load(paste(load_dir,"/PSI",sep = ""))
+load(paste(load_dir,"/psi_true",sep = ""))
+load(paste(load_dir,"/SIGMA",sep = ""))
+load(paste(load_dir,"/sig2_true",sep = ""))
+load(paste(load_dir,"/Z",sep = ""))
+load(paste(load_dir,"/c_true",sep = ""))
+
+h <- length(BETA.list)
+k <- ncol(beta.true.list[[1]])
+p <- nrow(beta.true.list[[1]])
+v <- nrow(delta.true)
+
+burn <- 2500
+
+colnames(Z) <- paste("Z",seq(1,ncol(Z)),sep = "")
+Z <- Z[-(1:burn),]
+colnames(DELTA) <- paste("D",seq(1,v*(h-1)),sep = "")
+DELTA <- DELTA[-(1:burn),]
+colnames(PI) <- paste("PI",seq(1,ncol(PI)),sep = "")
+PI <- PI[-(1:burn),]
+
+for(l in 1:h)
+{
+    colnames(BETA.list[[l]]) <- paste("B",seq(1,p*k),sep = "")
+    BETA.list[[l]] <- BETA.list[[l]][-(1:burn),]
+    colnames(PSI.list[[l]]) <- paste("PSI",seq(1,length(psi.true.list[[1]])),sep = "")
+    PSI.list[[l]] <- PSI.list[[l]][-(1:burn),]
+    colnames(SIGMA.list[[l]]) <- paste("SIG",seq(1,ncol(SIGMA.list[[l]])),sep="")
+    SIGMA.list[[l]] <- SIGMA.list[[l]][-(1:burn),]
+}
+
+## BETA plots
+
+BETA_trace <- mcmc_trace(BETA.list) + 
+    scale_color_manual(name = "Cluster",
[…]

### 2019-07-05 simulation/simulation.R

@@ -113,7 +113,7 @@ for(l in 1:h)
     
     # Cluster specific betas and psis
     beta.true.l <- matrix(rnorm(p*k,mean = seq(-h,h,length.out = h)[l],sd = 0.5),nrow = p,ncol = k) # true beta matrix
-    psi.true.l <- rep((((-1)^l)*l - 1)/h,k) # control skewness of each outcome in cluster l
+    psi.true.l <- rep((((-1)^l)*l*2 - 1)/h,k) # control skewness of each outcome in cluster l
     betastar.true.l <- rbind(beta.true.l,psi.true.l)
     betastar.true.list[[l]] <- betastar.true.l
     beta.true.list[[l]] <- beta.true.l
@@ -161,8 +161,8 @@ delta0 <- rep(0,v) # prior mean for delta coefficients (multinomial regression)
 S0 <- diag(1,v) # prior covariance for delta coefficients (multinomial regression)
 
 # Sample storage
-nsim <- 10000 # number of iterations
-burn <- 2500 # number of iterations to save
+nsim <- 500 # number of iterations
+burn <- 0 # number of iterations to save
 n.iter <- nsim - burn # number of saved iterations
 Z <- matrix(0,nrow = n.iter,ncol = n) # large matrix where each row is the value of z at a specific iteration
 PI <- matrix(0,nrow = n.iter,ncol = h) # matrix w/ each row as pi vector at each iteration
@@ -395,9 +395,7 @@ store <- paste("mcmc_draws_",run_date,sep = "")
 if(!dir.exists(store))
 {
     dir.create(store)
-} else {
-    store <- paste(store,"_2",sep = "")
-}
+} 
 meta_file <- paste(store,"/META.txt",sep = "")
 file.create(meta_file)
 write(paste("MCMC Run Date:",Sys.time()),file = meta_file, append = TRUE)

### 2019-07-06 manuscript/manuscript.tex
edits

@@
 \begin{abstract}
-In studies of infant motor development, a crucial research goal is the identification of latent clusters of infants who experience delayed development, as this is a known risk factor for adverse outcomes later in life.
-However, there are a number of statistical challenges in modeling infant development: the data are typically skewed, exhibit intermittent missingness, and are highly correlated across the repeated measurements collected during infancy.
+In studies of infant motor development, a crucial research goal is the identification of latent clusters of infants who experience delayed motor development, as this is a known risk factor for adverse outcomes later in life.
+However, there are a number of statistical challenges in modeling infant motor development: the data are typically skewed, exhibit intermittent missingness, and are highly correlated across the repeated measurements collected during infancy.
 Using data from the Nurture study, a cohort of over 600 mother-infant pairs followed from pregnancy to 12 months postpartum, we develop a flexible Bayesian finite mixture model for the analysis infant motor development.
@@
 In addition, our imputation method yields improvements compared to conventional missing data methods, including multiple imputation and complete or available case analysis.
-When applied to Nurture data, we identified two distinct development clusters: one characterized by delayed “U-shaped” development and a higher percentage of male infants and another characterized by more steady development and a lower percentage of males.
+When applied to Nurture data, we identified two distinct development clusters: one characterized by delayed “U-shaped” motor development and a higher percentage of male infants and another characterized by more steady motor development and a lower percentage of males.
 The clusters also differed in terms of key demographic variables, such as infant race and maternal pre-pregnancy body mass index.
@@
 
-The analysis of infant development trends in the Nurture data presents a number of statistical challenges that motivate our proposed model.
+The analysis of infant motor development trends in the Nurture data presents a number of statistical challenges that motivate our proposed model.
 First, as depicted in Figure 1, the residuals from repeated measures models of Bayley composite scores exhibit skewness even after adjusting for covariates such as race, sex, and birthweight.
@@
 To address this aim, we propose a flexible finite mixture model that accommodates relevant features of the data, such as skewness and dependence among the responses.
-For $i = 1,...,n$, let $\mathbf{y}_{i}=(y_{i1},\ldots,y_{iJ})^T$ be a $J \times 1$ vector of responses (i.e., Baley composite scores) for subject $i$ across the $J$ measurement occasions.
-For the analysis of the Nurture data, we propose a finite mixture model of the form
+For $i = 1,...,n$, let $\mathbf{y}_{i}=(y_{i1},\ldots,y_{iJ})^T$ be a $J \times 1$ vector of responses (i.e., Bayley composite scores) for subject $i$ across the $J$ measurement occasions.
+For analysis of the Nurture data, we propose a finite mixture model of the form
 \begin{equation}
@@
 where $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$ ($k = 1,...,K$) and $\pi_{ki}$ is a subject-specific mixing weight representing the probability that subject $i$ belongs to cluster $k$.
-For now we assume that $K$ is fixed; in Section 4, we discuss model selection strategies for choosing the optimal value of $K$.
+For now we assume that $K$ is fixed; in Section \ref{s:mcmcmisc}, we discuss model selection strategies for choosing the optimal value of $K$.
 
-To facilitate posterior inference, we introduce a latent cluster indicator variable $z_i$ taking the value $k \in \{1,...,K\}$ with probability $\pi_{ik}$.
+To facilitate posterior inference, we introduce a latent cluster indicator variable $z_i$ taking the value $k \in \{1,...,K\}$ with probability $\pi_{ki}$.
 Conditional on $z_i = k$, we assume $\mathbf{y}_{i}$ is distributed as
@@
 The vector $\boldsymbol\alpha_k$ has components $\alpha_{kj}$, $j = 1,...,J$, that control the skewness of outcome $j$ in cluster $k$.
-When $\boldsymbol\alpha_k = \mathbf{0}$, the MSN distribution reduces to the multivariate normal distribution $\text{N}_J(\boldsymbol\zeta_{ki},\boldsymbol\Omega_k)$, where $\boldsymbol\zeta_{ki}$ is a $J \times 1$ mean vector and $\boldsymbol\Omega_k$ is a $J \times J$ covariance matrix (Azzalini and Dalla Valle, 1996).
+When $\boldsymbol\alpha_k = \mathbf{0}$, the MSN distribution reduces to the multivariate Normal distribution $\text{N}_J(\boldsymbol\zeta_{ki},\boldsymbol\Omega_k)$, where $\boldsymbol\zeta_{ki}$ is a $J \times 1$ mean vector and $\boldsymbol\Omega_k$ is a $J \times J$ covariance matrix (Azzalini and Dalla Valle, 1996).
 
@@
 
-Of note, the MSN density can be expressed more compactly in terms of the matrix skew normal (MatSN) density (Chen and Gupta 2005).
+Of note, the MSN density can be expressed more compactly in terms of the Matrix Skew Normal (MatSN) density (Chen and Gupta 2005).
 As we will see in Section 3.6, the matrix representation of the MSN distribution admits convenient conjugate prior distributions for the regression parameters and scale matrices, which in turn leads to efficient Gibbs sampling for posterior inference.
@@
 \subsection{Multinomial Regression for the Cluster Indicators}
-To accommodate heterogeneity in the cluster-membership probabilities, we model $\pi_{ik}$ as a function of covariates using a multinomial logit model 
+\label{s:multinom}
+To accommodate heterogeneity in the cluster-membership probabilities, we model $\pi_{ki}$ as a function of covariates using a multinomial logit model 
 \begin{eqnarray}
-\pi_{ik} = \Pr(z_i = k|\mathbf{w}_i) = \frac{\text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k}}{\sum_{h = 1}^K \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}},~ k=1,\ldots,K,
+\label{eq:pik}
+\pi_{ki} = \Pr(z_i = k|\mathbf{w}_i) = \frac{\text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k}}{\sum_{h = 1}^K \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}},~ k=1,\ldots,K,
 \end{eqnarray}
@@
 For identifiability purposes, we fix the reference category $k = K$ and set $\boldsymbol\delta_K = \mathbf{0}$.
-Under this model, $z_i|\mathbf{w}_i \sim \text{Multinomial}(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{i1},...,\pi_{1K})$.
-During MCMC estimation, the cluster labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as cluster assignments as detailed later in section (\textbf{what section}).
-By allowing the cluster probabilities to vary across subjects, our model can be viewed as a \textit{mixture of experts} model in which $\pi_{ki}$ acts as a \textit{gating function} controlling the prior probability of membership in cluster $k$, and $f(\mathbf{y}_i|\boldsymbol\theta_k)$ in Equation \ref{eq:mixture} is the ``expert'' (Bishop 2016).
+Under this model, $z_i|\boldsymbol\pi_i \sim \text{Multinomial}(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{1i},...,\pi_{Ki})$.
+During MCMC estimation, the cluster labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as cluster assignments as detailed later in Section \ref{s:postinf}.
+By allowing the cluster probabilities to vary across subjects, our model can be viewed as a \textit{mixture of experts} model in which $\pi_{ki}$ acts as a \textit{gating function} controlling the prior probability of membership in cluster $k$, and $f(\mathbf{y}_i|\boldsymbol\theta_k)$ in equation (\ref{eq:mixture}) is the ``expert'' providing information on the within-cluster distribution of $\mathbf{y}_i$ (Bishop 2016).
 
-To facilitate sampling, we adopt an efficient data-augmentation approach introduced by Polson \textit{et al.} (2013), which expresses the inverse-logit function as a mixture P\'olya--Gamma densities.
+To facilitate sampling, we adopt an efficient data-augmentation approach introduced by Polson {et al.} (2013), which expresses the inverse-logit function as a mixture P\'olya--Gamma densities.
 By using P\'olya--Gamma data augmentation for the multinomial model, we obtain a \textit{P\'olya--Gamma mixture of experts model}, a computationally efficient way to obtain inferences for the mixing weights in the Bayesian setting.
@@
 \end{equation}
-where $g_s \stackrel{iid}{\sim} \text{Ga}(b,1)$ for $s = 1,...,\infty$ . Polson \textit{et al.} (2013) establish that for the $\text{PG}(b,c)$ density, and for $a,\eta \in \mathbb{R}$,
+where $g_s \stackrel{iid}{\sim} \text{Ga}(b,1)$ for $s = 1,...,\infty$ . Polson {et al.} establish that for $a,\eta \in \mathbb{R}$,
 
 \begin{equation}
-\frac{(\text{e}^{\eta})^a}{(1 + \text{e}^{\eta})^b} = 2^{-b}\text{e}^{\kappa \eta} \int_{0}^{\infty} \text{e}^{-{w}\eta^2/2}p({w}|b,c = 0)d{w}. \label{eq:pg2}
+\frac{(\text{e}^{\eta})^a}{(1 + \text{e}^{\eta})^b} = 2^{-b}\text{e}^{\kappa \eta} \int_{0}^{\infty} \text{e}^{-{w}\eta^2/2}p({w}|b,c = 0)d{w}, \label{eq:pg2}
 \end{equation}
-where $\kappa = a - b/2$. Polson \textit{et al.} also show that the conditional distribution $p(w|b,c)$ results from an ``exponential tilting'' of the $\text{PG}(b,0)$ density, thus
+where $\kappa = a - b/2$ and $p(w|b,c = 0)$ denotes a $\text{PG}(b,0)$ density.
+Polson {et al.} further show that the conditional distribution $p(w|b,c)$ results from an ``exponential tilting'' of the $\text{PG}(b,0)$ density, thus
 
@@
 
-To make use of P\'olya--Gamma data augmentation, we first write the full conditional distribution of $\boldsymbol\delta_k$ as the prior for $\boldsymbol\delta_k$ times the multinomial likelihood for all $z_i$.
-$$p(\boldsymbol\delta_k|\mathbf{U}_k,\boldsymbol\delta_{h \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \pi_{ki}^{U_{ki}}(1-\pi_{ki})^{1-U_{ki}},$$
-where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, $\mathbf{U}_k = (U_{k1},...,U_{kn})$, $U_{ik} = {1}_{(z_i = k)}$ is an indicator that subject $i$ belongs to cluster $k$, and $\pi_{ki}$ is defined as in Section 3.4.
-We can rewrite $\pi_{ki}$ as follows
-$$\pi_{ki} = P(U_{ki} = 1) = \frac{\text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ki}}}{1 + \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ki}}} = \frac{\text{e}^{\eta_{ki}}}{1 + \text{e}^{\eta_{ki}}}$$
-where ${c}_{ik} = \log \sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}$ and $\eta_{ik} = \mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ki}$.
+Polson et al.
+use these results to show that, in the logistic model, the Bernoulli likelihood can be written in terms of the kernel of a Normal density and a $\text{PG}(b,0)$ density, resulting in a Normal full conditional distribution for the logistic regression parameters.
+To extend the P\'olya--Gamma data augmentation approach to the multinomial model, we first introduce the binary indicators $U_{ki}$, such that $U_{ki} = 1_{z_i = k}$ -- i.e.
+$U_{ki}$ is an indicator variable taking the value $1$ if subject $i$ belongs to cluster $k$, and $0$ otherwise.
+The full conditional distribution of $\boldsymbol\delta_k$ is then given as  $$p(\boldsymbol\delta_k|\mathbf{U}_k,\boldsymbol\delta_{h \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \pi_{ki}^{U_{ki}}(1-\pi_{ki})^{1-U_{ki}},$$
+where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, and $\pi_{ki}$ is defined as in equation (\ref{eq:pik}). We can rewrite $\pi_{ki}$ in terms of $U_{ki}$ as
+$$\pi_{ki} = P(U_{ki} = 1) = \frac{\text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ki}}}{1 + \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ki}}} = \frac{\text{e}^{\eta_{ki}}}{1 + \text{e}^{\eta_{ki}}},$$
+where ${c}_{ki} = \log \sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}$ and $\eta_{ki} = \mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ki}$.
 We note that the sum $\sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}$ includes the reference category, but since we fix $\boldsymbol\delta_K = \mathbf{0}$, we have $\text{e}^{\mathbf{w}_i^T \boldsymbol\delta_K} = 1$, and hence
-$$c_{ki} = \log \sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}} = \log \left ( 1 + \sum_{h \notin \{k,K \}} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}} \right )^T$$. 
+$$c_{ki} = \log \sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}} = \log \left ( 1 + \sum_{h \notin \{k,K \}} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}} \right )^T.$$
 We can use these quantities to re-express the full conditionals for $\boldsymbol\delta_k$ as
 \begin{eqnarray}
-	p(\boldsymbol\delta_k|\mathbf{U}_k,\boldsymbol\delta_{h \ne k}) &\propto& p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \left (\frac{\text{e}^{\eta_{ki}}}{1 + \text{e}^{\eta_{ki}}} \right )^{U_{ki}} \left (\frac{1}{1 + \text{e}^{\eta_{ki}}} \right )^{1-U_{ki}} \nonumber\\ &=& p(\boldsymbol\delta_k) \prod_{i = 1}^n \frac{(\text{e}^{\eta_{ki}})^{U_{ki}}}{1 + \text{e}^{\eta_{ki}}} \label{eq:pglogit}
+	p(\boldsymbol\delta_k|\mathbf{U}_k,\boldsymbol\delta_{h \ne k}) &\propto& p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \left (\frac{\text{e}^{\eta_{ki}}}{1 + \text{e}^{\eta_{ki}}} \right )^{U_{ki}} \left (\frac{1}{1 + \text{e}^{\eta_{ki}}} \right )^{1-U_{ki}} \nonumber\\ &=& p(\boldsymbol\delta_k) \prod_{i = 1}^n \frac{(\text{e}^{\eta_{ki}})^{U_{ki}}}{1 + \text{e}^{\eta_{ki}}}, \label{eq:pglogit}
 \end{eqnarray}
-which is essentially a logistic regression likelihood.
+which is a logistic likelihood of the same form used by Polson et al.
+to derive the data-augmented P\'olya--Gamma Gibbs sampler for the logistic regression model.
 We can thus apply the P\'olya--Gamma sampler described in Polson et al.
-(2013) for logistic regression to update each $\boldsymbol\delta_k$ one at a time based on the binary indicators $U_{ki}$.
-To accomplish this, we first define for $k = 1,...,k$, the $n \times 1$ vector $\mathbf{U}^*_{k} = \left( \frac{U_{k1}-1/2}{w_{k1}},...,\frac{U_{kn}-1/2}{w_{kn}} \right )^T$.
+for logistic regression to update each $\boldsymbol\delta_k$ one at a time based on the binary indicators $U_{ki}$.
+To accomplish this, we first define for $k = 1,...,K$, the $n \times 1$ vector $\mathbf{U}^*_{k} = \left( \frac{U_{k1}-1/2}{w_{k1}},...,\frac{U_{kn}-1/2}{w_{kn}} \right )^T$.
 Polson et al.
-(2013) show that, conditional on $\boldsymbol{w} = (w_{k1},...,w_{kn})^T$, $\mathbf{U}^*_{k}$ follows a $\text{N}_n (\boldsymbol\eta_k,\mathbf{O}_k^{-1})$ distribution with mean $\boldsymbol\eta_k = (\eta_{k1},...,\eta_{kn})^T$ and precision matrix $\mathbf{O}_k = \text{Diag}(w_{k1},...,w_{kn})$.
+show that, conditional on $\boldsymbol{w} = (w_{k1},...,w_{kn})^T$, $\mathbf{U}^*_{k}$ follows a $\text{N}_n (\boldsymbol\eta_k,\mathbf{O}_k^{-1})$ distribution with mean $\boldsymbol\eta_k = (\eta_{k1},...,\eta_{kn})^T$ and precision matrix $\mathbf{O}_k = \text{Diag}(w_{k1},...,w_{kn})$.
 Thus, it follows that the full conditional distribution of $\boldsymbol\delta_k$ is given by
 \begin{equation}
-	p(\boldsymbol\delta_k|\mathbf{z},\mathbf{W}, \mathbf{O}_k) \propto p(\boldsymbol\delta_k) \text{exp} \left [ - \frac{1}{2} (\mathbf{U}^*_k - \mathbf{W}\boldsymbol\delta_k)^T \mathbf{O}_k (\mathbf{U}^*_k - \mathbf{W}\boldsymbol\delta_k)\right],
+	p(\boldsymbol\delta_k|\mathbf{z},\mathbf{W}, \mathbf{O}_k) \propto p(\boldsymbol\delta_k) \text{exp} \left [ - \frac{1}{2} (\mathbf{U}^*_k - \mathbf{W}\boldsymbol\delta_k)^T \mathbf{O}_k (\mathbf{U}^*_k - \mathbf{W}\boldsymbol\delta_k)\right].
 \end{equation}
-where $p(\boldsymbol\delta_k)$ is the prior distribution for $\boldsymbol\delta_k$, which we detail in Section 3.4 As detailed in Section 3.4, if we assume $p(\boldsymbol\delta_k)$ to be multivariate normal, then $\boldsymbol\delta_k$ has a closed-form multivariate normal full conditional distribution that can be easily embedded within our proposed Gibbs sampling routine.
+As detailed in Section 3.4, if we assume $p(\boldsymbol\delta_k)$ to be multivariate normal, then $\boldsymbol\delta_k$ has a closed-form multivariate normal full conditional distribution that can be easily embedded within our proposed Gibbs sampling routine.
+For more details on the P\'olya-Gamma Gibbs sampler for logistic and multinomial models, see Polson et al.
+(2013).
 \subsection{Conditional MSN Imputation}
-To accommodate missing at random (MAR) responses, we propose a convenient imputation algorithm that can be implemented ``online'' as part of the Gibbs sampler.
+To accommodate intermittent missing at random (MAR) responses, we propose a convenient imputation algorithm that can be implemented ``online'' -- that is, as part of the Gibbs sampler.
 In Section 6, we discuss extensions to allow for non-ignorable missingness (i.e.
@@
 Suppose $\mathbf{y}_i$ has $q_i\in (1,\ldots,J)$ observed values, denoted $\mathbf{y}^{obs}_i$, and $J-q_i$ intermittent missing values, denoted $\mathbf{y}^{miss}_i$.
-We can make use of the stochastic representation given in equation (\ref{eq:msnreg}) to impute $\mathbf{y}^{miss}_i$ from its conditional multivariate normal distribution given $(z_i,t_i,\mathbf{y}^{obs}_i)$:
+We can make use of the stochastic representation given in equation (\ref{eq:msnreg}) to impute $\mathbf{y}^{miss}_i$ from its conditional multivariate Normal distribution given $(z_i,t_i,\mathbf{y}^{obs}_i)$:
 \begin{eqnarray}
-\mathbf{y}^{miss}_i|(z_i=k,t_i,\mathbf{y}^{obs}_i)&\sim& \text{N}_{J-q_i}(\boldsymbol\zeta^{cond}_{ki},\boldsymbol\Sigma^{cond}_k),~\text{where}\nonumber\\
-\boldsymbol\zeta^{cond}_{ki}&=& \boldsymbol\zeta^{miss}_{ki} + \boldsymbol\Sigma_{k12} \boldsymbol\Sigma_{k22}^{-1}(\mathbf{y}^{obs}_i - \boldsymbol\zeta^{obs}_{ki}) \label{eq:impute}\\
-\boldsymbol\Sigma^{cond}_k&=& \boldsymbol\Sigma_{k11} - \boldsymbol\Sigma_{k12}\boldsymbol\Sigma_{k22}^{-1}\boldsymbol\Sigma_{k21},\nonumber \\
-\boldsymbol\zeta_{ki} &=& \begin{pmatrix} \boldsymbol\zeta_{ki}^{miss} \\ \boldsymbol\zeta_{ki}^{obs} \end{pmatrix}, \ \text{and} \nonumber \\
+\mathbf{y}^{miss}_i|(z_i=k,t_i,\mathbf{y}^{obs}_i)&\sim& \text{N}_{J-q_i}(\boldsymbol\mu^{cond}_{ki},\boldsymbol\Sigma^{cond}_k),~\text{where}\nonumber\\
+\boldsymbol\mu^{cond}_{ki}&=& \boldsymbol\mu^{miss}_{ki} + \boldsymbol\Sigma_{k12} \boldsymbol\Sigma_{k22}^{-1}(\mathbf{y}^{obs}_i - \boldsymbol\mu^{obs}_{ki}) \nonumber \\
+\boldsymbol\Sigma^{cond}_k&=& \boldsymbol\Sigma_{k11} - \boldsymbol\Sigma_{k12}\boldsymbol\Sigma_{k22}^{-1}\boldsymbol\Sigma_{k21}, \label{eq:impute} \\
+\boldsymbol\mu_{ki} &=& \begin{pmatrix} \boldsymbol\mu_{ki}^{miss} \\ \boldsymbol\mu_{ki}^{obs} \end{pmatrix}, \ \text{and} \nonumber \\
 \boldsymbol\Sigma_k &=& \begin{pmatrix} 
@@
 \end{eqnarray}
-$\boldsymbol\Sigma_{k11}$ is a $(J-q_i) \times (J-q_i)$ matrix containing the rows and columns of $\boldsymbol\Sigma_k$ corresponding to inicies of $\mathbf{y}_{i}$ where missingness occurs.
+The location vector $\boldsymbol\mu_{ki}$ is defined as $\boldsymbol\mu_{ki} = \mathbf{X}_i \boldsymbol\beta_k + t_i \boldsymbol\psi_k$ and is partitioned into $\boldsymbol\mu_{ki}^{miss}$ and $\boldsymbol\mu_{ki}^{obs}$ with respect to the missing and observed indices of $\mathbf{y}_i$, respectively.
+The partition $\boldsymbol\Sigma_{k11}$ is a $(J-q_i) \times (J-q_i)$ matrix containing the rows and columns of $\boldsymbol\Sigma_k$ corresponding to inicies of $\mathbf{y}_{i}$ where missingness occurs.
 Similarly, $\boldsymbol\Sigma_{k12}$ is a $(J-q_i) \times q_i$ matrix containing the rows of $\boldsymbol\Sigma_k$ that correspond to missing indices of $\mathbf{y}_i$, but columns of $\boldsymbol\Sigma_k$ that correspond to observed indices of $\mathbf{y}_i$.
 The remaining partitions $\boldsymbol\Sigma_{k21}, \ \text{and} \ \boldsymbol\Sigma_{k22}$ are defined in the same manner.
-Likewise, as defined in equation \ref{eq:msndens}, $\boldsymbol\zeta_{ki} = \mathbf{X}_i \boldsymbol\beta_k + t_i \boldsymbol\psi_k$ and is partitioned into $\boldsymbol\zeta_{ki}^{miss}$ and $\boldsymbol\zeta_{ki}^{obs}$ with respect to the missing and observed indices of $\mathbf{y}_i$, respectively.
-These results follow from conventional multivariate normal theory, which we detail in the Web Appendix.
-An attractive feature of this imputation algorithm is that it avoids multiplicative run-time scaling in $m$, the number of imputations (Gelman \textit{et al.} 2013; Zhou and Reiter, 2010).
-Our approach also provides more opportunities to explore the missing data parameter space than does multiple imputation, since missing values are drawn at each MCMC iteration, and often in practice $n_{sim} >> m$, where $n_{sim}$ is the number of MCMC iterations (\textbf{find a reference}).
-In Section 4, we conduct simulation studies to demonstrate that imputing the missing MSN responses improves inferences over available-case analysis.
+These results follow from conventional multivariate normal theory, which we detail in the Web Appendix A.
+We note that, while conditional on $t_i$, $\mathbf{y}_i$ follows a MVN distribution, after marginalizing over $t_i$ the vector $\mathbf{y}_i$ follows a MSN distribution.
+Thus, this proposed online conditional imputation method provides a convinient way of imputing MSN responses using samples from more standard densities.
+
+An attractive practical feature of this imputation algorithm is that it avoids multiplicative run-time scaling in $m$, the number of imputations (Gelman {et al.} 2013; Zhou and Reiter, 2010).
+Our approach also provides more opportunities to explore the missing data parameter space than does multiple imputation, since each missing component is drawn once per MCMC iteration, and often in practice $n_{sim} >> m$, where $n_{sim}$ is the total number of MCMC iterations (\textbf{find a reference}).
+In Section 4, we conduct simulation studies to demonstrate that imputing the missing MSN responses with our online conditional imputation method improves inferences over standard Bayesian multiple imputation as defined by Gelman et al.
+2013.
 
@@
 Conveniently, all parameters admit conditionally conjugate priors, which greatly improves posterior computation via a data-augmented Gibbs sampler.
-For $\boldsymbol\beta_k$ and $\boldsymbol\Sigma_k$, we adopt a conditionally independent prior structure for $\boldsymbol\beta_k$ and $\boldsymbol\Sigma_k$, where $p(\boldsymbol\beta_k,\boldsymbol\Sigma_k) = p(\boldsymbol\Sigma_k)p(\boldsymbol\beta_k|\boldsymbol\Sigma_k)$.
-We choose the normal-inverse-Wishart distribution for $p(\boldsymbol\beta_k,\boldsymbol\Sigma_k)$ by specifying $\boldsymbol\Sigma_k \sim \text{IW}(\mathbf{V}_{0k},\nu_{0k})$ and $\boldsymbol\beta_k|\boldsymbol\Sigma_k \sim \text{N}_{Jp}(\mathbf{b}_k,\mathbf{I}_p \otimes \boldsymbol\Sigma_k)$.
-We assign the skewness coefficients $\boldsymbol\psi_k$ a conjugate $\text{N}_J(\mathbf{m}_k,\mathbf{P}_k)$ prior.
-In practice, the updates of $\boldsymbol\beta_k$ and $\boldsymbol\psi_k$ can be combined into one step by defining the $(Jp + J) \times 1$ vector $\boldsymbol\beta^*_k = (\boldsymbol\beta_k^T,\boldsymbol\psi_k^T)^T$ for which we assume a $\text{N}_{Jp+J}(\mathbf{b}^*_k,\mathbf{I}_{(p+1)} \otimes \boldsymbol\Sigma_k)$ prior, where $\mathbf{b}^*_k$ is formed by concatenating $\mathbf{b}_k$ and $\mathbf{m}_k$.
-
-To make use of the Matrix Normal representation introduced previously in Section 3, we define the matrix of regression parameters $\mathbb{B}^*_k$, where $\text{vec}(\mathbb{B}^*_k) = \boldsymbol\beta^*_k$.
-We assign $\mathbb{B}^*_k|\boldsymbol\Sigma_k$ a $\text{MatNorm}(\mathbf{B}^*_{0k},\mathbf{I}_{p+1},\boldsymbol\Sigma_k)$ prior, where $\mathbf{B}^*_{0k}$ is a matrix of location parameters such that $\text{vec}(\mathbf{B}^*_{0k}) = \mathbf{b}^*_k$, $\mathbf{I}_{p+1}$ is the $(p+1)$-dimension identity matrix, and $\boldsymbol\Sigma_k$ is the covariance matrix defined in equation \ref{eq:msndens}.
-The matrix-Normal-inverse-Wishart is the conjugate joint prior for the regression parameters in the matrix Normal model given in equation \ref{eq:matsn}.
+To make use of the Matrix Normal representation introduced previously in Section 3, we define the matrix of regression parameters $\mathbb{B}^*_k$, where $\text{vec}(\mathbb{B}^*_k) = \boldsymbol\beta^*_k = (\boldsymbol\beta_k^T,\boldsymbol\psi_k^T)^T$.
+We assign $\mathbb{B}^*_k|\boldsymbol\Sigma_k$ a $\text{MatNorm}(\mathbf{B}^*_{0k},\mathbf{I}_{p+1},\boldsymbol\Sigma_k)$ prior, where $\mathbf{B}^*_{0k}$ is a matrix of location parameters such that $\text{vec}(\mathbf{B}^*_{0k})$ is a vector of prior location parameters for the components of $\boldsymbol\beta^*_k$, $\mathbf{I}_{p+1}$ is the $(p+1)$-dimension identity matrix, and $\boldsymbol\Sigma_k$ is the covariance matrix defined in equation (\ref{eq:msndens}), for which we specify an $\text{IW}(\mathbf{V}_{0k},\nu_{0k})$ prior.
+This leads to a Matrix-Normal-Inverse-Wishart joint prior for $\mathbb{B}^*_k$ and $\boldsymbol\Sigma_k$, which is the conjugate joint prior for the regression parameters in the Matrix Normal model given in equation (\ref{eq:matsn}).
 This induces convenient closed-form full conditional distributions that can be easily updated within our proposed Gibbs sampler.
 
-For the multinomial logit model component, the regression parameters $\boldsymbol\delta_k = (\delta_{k1},...,\delta_{kr})$ are assigned a $\text{N}_r(\mathbf{d}_{0k}, \mathbf{S}_{0k})$ prior for $k = 1,...,K-1$, which is conditionally conjugate as described in Section (\textbf{what section}).
-We allow the normal-inverse-Wishart and multinomial hyperparameters to vary across clusters, though they may be shared across clusters in practice.
+For the multinomial logit model component, the regression parameters $\boldsymbol\delta_k = (\delta_{k1},...,\delta_{kr})$ are assigned a $\text{N}_r(\mathbf{d}_{0k}, \mathbf{S}_{0k})$ prior for $k = 1,...,K-1$, which is conditionally conjugate as described in Section \ref{s:multinom}.
+We allow the Normal-Inverse-Wishart and multinomial hyperparameters to vary across clusters, though they may be shared across clusters in practice.
 An advantage of allowing for cluster-specific prior parameters is that \textit{a priori} knowledge of motor development trends can be incorporated into certain clusters while allowing the priors for other clusters to be less informative.
@@
 \subsubsection{Posterior Inference} 
+\label{s:postinf}
 The above prior specification induces closed-form full conditionals that can be efficiently updated as part of a Gibbs sampler outlined below.
-A programatic sketch of our MCMC algorithm is given in Table (\ref{alg:MCMC}).
-Additional details, including derivations of full conditionals and MCMC diagnostics can be found in the Web Appendix.
+A programatic sketch of our MCMC algorithm is given in Web Appendix B.
+Additional details on derivations of full conditionals can be found in the Web Appendix A.
+We provide a report of MCMC diagnostics in Web Appendix C.
 
 \paragraph{\textbf{Step 1}: Conditional MSN Imputation} The sampler begins by imputing missing values $\mathbf{y}_i^{miss}$ conditional on current values of $z_i = k$ and $t_i$ as well as the associated $\mathbf{y}_i^{obs}$ observed data vector.
-Specifically, we draw $\mathbf{y}_i^{miss}$ from $\text{N}_{J-q_i}(\boldsymbol\mu^{cond}_{ki},\boldsymbol\Sigma^{cond}_k)$ as described in equation \ref{eq:impute} for all $i = 1,...,n$.
+Specifically, we draw $\mathbf{y}_i^{miss}$ from $\text{N}_{J-q_i}(\boldsymbol\mu^{cond}_{ki},\boldsymbol\Sigma^{cond}_k)$ as described in equation (\ref{eq:impute}) for all $i = 1,...,n$.
 We conclude by constructing a complete outcome vector $\mathbf{y}_i$ that merges $\mathbf{y}_i^{miss}$ with $\mathbf{y}_i^{obs}$.
 
-\paragraph{\textbf{Step 2:} Update of MSN Regression Parameters} We begin the update of MSN regression parameters by updating $t_i$, the truncated Normal random effect used in the stochastic representation of $\mathbf{y}_i$ given in equation \ref{eq:msnreg}.
+\paragraph{\textbf{Step 2:} Update of MSN Regression Parameters} We begin the update of MSN regression parameters by updating $t_i$, the truncated Normal random effect used in the stochastic representation of $\mathbf{y}_i$ given in equation (\ref{eq:msnreg}).
 For cluster $k$, compute $A_k = (1 + \boldsymbol\psi_k^T \boldsymbol\Sigma^{-1}_k\boldsymbol
 \psi_k)^{-1}$ using current values of $\boldsymbol\psi_k$ and $\boldsymbol\Sigma_k$.
-Next, for $i = 1,...,n_k$, compute $a_i = A_k \boldsymbol\psi_k^T \boldsymbol\Sigma^{-1}_k(\mathbf{y}_i - \boldsymbol\zeta_{ik})$, where $\boldsymbol\zeta_{ik} = X_i \boldsymbol\beta_k$.
+Next, for $i = 1,...,n_k$, compute $a_i = A_k \boldsymbol\psi_k^T \boldsymbol\Sigma^{-1}_k(\mathbf{y}_i - \boldsymbol\zeta_{ki})$, where $\boldsymbol\zeta_{ki} = X_i \boldsymbol\beta_k$.
 Finally, update $t_i$ from $\text{N}_{[0,\infty)}(a_i, \mathbf{A}_k)$.
@@
 For each cluster $k$ ($k = 1,...,K$), we update $\boldsymbol\Sigma_k$ from an $\text{IW}(\nu_k,\mathbf{V}_k)$ density, where $\nu_k = \nu_{0k} + n_k$ and\begin{eqnarray}
-	\mathbf{V}_k &=&  (\mathbf{Y}-\mathbf{X}^*_k \mathbf{B}^*_k)^T (\mathbf{Y}_k - \mathbf{X}^*_k \mathbf{B}^*_k) (\mathbf{B}_k^* - \mathbf{B}^*_{0k})^T \mathbf{I}_{p+1}(\mathbf{B}^*_k - \mathbf{B}^*_{0k}) + \mathbf{V}_{0k}.
+	\mathbf{V}_k &=&  (\mathbf{Y}_k-\mathbf{X}^*_k \mathbf{B}^*_k)^T (\mathbf{Y}_k - \mathbf{X}^*_k \mathbf{B}^*_k) + (\mathbf{B}_k^* - \mathbf{B}^*_{0k})^T \mathbf{I}_{p+1}(\mathbf{B}^*_k - \mathbf{B}^*_{0k}) + \mathbf{V}_{0k}.
 \nonumber
-\end{eqnarray} The matrix $\mathbf{X}^*_k$ is formed by adding We then make use of the Matrix Normal representation introduced in Section 3.2 to draw $\mathbb{B}^*_k$ from a $\text{MatNorm}_{p+1,J}(\mathbf{B}_k^*,\mathbf{L}_k^{*},\boldsymbol\Sigma_k)$ density, where $\text{vec}(\mathbb{B}^*_k) = \boldsymbol\beta^*_k = (\beta_{k11},...,\beta_{kJp},\psi_{k1},...,\psi_{kJ})^T$ and
+\end{eqnarray} 
+We then make use of the Matrix Normal representation introduced in Section 3.2 to draw $\mathbb{B}^*_k$ from a $\text{MatNorm}_{p+1,J}(\mathbf{B}_k^*,\mathbf{L}_k^{*},\boldsymbol\Sigma_k)$ density, where $\text{vec}(\mathbb{B}^*_k) = \boldsymbol\beta^*_k = (\beta_{k11},...,\beta_{kJp},\psi_{k1},...,\psi_{kJ})^T$ and
 \begin{eqnarray}
@@
 $$\upsilon_{ki} = P(z_i = k|\mathbf{y}_i,\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k) = \frac{\pi_{ki}P(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)}{\sum_{h = 1}^{K} \pi_{hi}P(\mathbf{y}_i|\boldsymbol\zeta^*_{hi},\boldsymbol\Sigma_h)}.$$
-The cluster labels $z_i$ are then updated from a $\text{Multinomial}(1,\boldsymbol\upsilon_i)$ density for $i = 1,...n$. A schematic outline of the Gibbs sampler is given in the Web Appendix.
+The cluster labels $z_i$ are then updated from a $\text{Multinomial}(1,\boldsymbol\upsilon_i)$ density for $i = 1,...n$.
+A schematic outline of the Gibbs sampler is given in the Web Appendix B.
+An \texttt{R} package for implementing the proposed model is currently in developlment, though \texttt{R} scripts with code for implementing the simulations and applications described here is available at \texttt{https://carter-allen.github.io/MVSN-FMM/}.
 
 \subsubsection{Assessment of MCMC Convergence, Label Switching, and Model Selection}
+\label{s:mcmcmisc}
 
-We monitor conver convergence of our MCMC algorithm through the use of standard approaches such Geweke's (1992) Z-diagnostic.
+We monitor convergence of the MCMC algorithm through the use of standard approaches such as investigation of trace plots and Geweke's (1992) Z-diagnostic, implemented in the \texttt{R} package \texttt{coda} (Plummer et al.
+2006).
 In simulation studies under realistic parameter settings, we observed relatively fast convergence of all MCMC chains (i.e.
@@
 
-Since the primary objective of our proposed model is to provide a flexible framework to account for the inherent features of infant development data such as skewness, correlation among repeated measures, and missingness, we assume the number of clusters $K$ to be specified \textit{a priori}.
+Since the primary objective of our proposed model is to provide a flexible framework to account for the inherent features of infant motor development data such as skewness, correlation among repeated measures, and missingness, we assume the number of clusters $K$ to be specified \textit{a priori}.
+Often in studies of infant motor development clusters, focus lies on explaining the heterogeneity in motor development trajectories with a parsimonious number of latent clusters.
+To this end, we propose the use of the ``widely applicable information criterion'' (WAIC) introduced by Watanabe (2010) for model selection.
+WAIC has the desireable property of penalizing complexity in models -- a feature congruent with our goal of explaining infant motor development heterogeneity with a parsimonious number of clusters.
+We refer to Gelman et al.
+(2014) for a detailed discussion of WAIC and comparison to other popular model fit criteria, such as DIC.
+In practice, we follow the approach taken by Neelon and Chung (2016), wherein a small set of possible values for $K$, e.g.
+$K \in \{2,3,4 \}$, are specified \textit{a priori} and models fit under each value of $K$ are compared using WAIC.
+In Section \ref{s:sim} we show with a simulation studies that this approach recovered the true value of $K$ under realistic parameter settings.
 
@@
 Here, the primary goal is to validate our parameter estimation scheme in a setting that resembles our application to the Nurture data.
-Our secondary goal is to investigate to what degree allowing for skewness in outcome components reduces the number of mixture components necessary to estimate the marginal density.
+Our secondary goal is to investigate to what degree ignoring skewness in outcome components corrupts posterior estimates of model parameters.
 For the purposes of illustration, we consider the following generative model
@@
 \end{eqnarray}
-where $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$, for $k \in \{1,2,3 \}$, and $\mathbf{y}_i | \boldsymbol\theta_k \sim \text{MSN}_4(\boldsymbol\zeta_{ki},\boldsymbol\alpha_k,\boldsymbol\Omega_k)$, where $\boldsymbol\zeta_{ki} = (\zeta_{ki1},\zeta_{ki2},\zeta_{ki3},\zeta_{ki4}) = \mathbf{x}_i \boldsymbol\beta_{k}$.
+where $\boldsymbol\theta_k$ is the set of all parameters specific to cluster $k$, for $k \in \{1,2,3 \}$, and $\mathbf{y}_i | \boldsymbol\theta_k \sim \text{MSN}_4(\boldsymbol\zeta_{ki},\boldsymbol\alpha_k,\boldsymbol\Omega_k)$, where $\boldsymbol\zeta_{ki} = (\zeta_{ki1},\zeta_{ki2},\zeta_{ki3},\zeta_{ki4}) = \mathbf{x}_i \boldsymbol\beta_{k}$.
 We note from the above model specification that the number of clusters $K = 3$ and the number of measurement occasions $J = 4$.
@@
 
-For cluster $k$, we can collect all observations $\mathbf{y}_i$ ($i = 1,...,n_k$) into the $n_k \times 4$ matrix $\mathbf{Y}_k$ and all covariates $\mathbf{x}_i$ into the $4n_k \times 8$ matrix $\mathbf{X}_k$ to utilize the matrix Normal model specification detailed in Section 3, which allows for simultaneous updates of all MSN regression parameters in cluster $k$.
+For cluster $k$, we collect all observations $\mathbf{y}_i$ ($i = 1,...,n_k$) into the $n_k \times 4$ matrix $\mathbf{Y}_k$ and all covariates $\mathbf{x}_i$ into the $4n_k \times 8$ matrix $\mathbf{X}_k$ to utilize the Matrix Normal model specification detailed in Section 3, which allows for simultaneous updates of all MSN regression parameters in cluster $k$.
 For the multinomial regression model component of this simulation, we model the class labels $z_i$ as a function of an intercept and one baseline covariate, thus $r = 2$.
-We leave the problem of missing data to simulation study \#2 and do not incorporate missing data into this simulation.
+We leave the problem of missing data to simulation study \#2 and do not introduce missing data into this simulation.
 
-For our second goal in this simulation, to compare model performance when skewness is accounted for skewness to model performance when skewness is ignored, we fit both the proposed MSN finite mixture model as well as a classic multivariate Normal finite mixture model to data generated from an MSN distribution.
+For our second goal in this simulation, to compare model performance when skewness is accounted for to model performance when skewness is ignored, we fit both the proposed MSN finite mixture model as well as a classic multivariate Normal finite mixture model to data generated from an MSN distribution.
 We evaluate model fit in terms of absolute difference of parameter estimates as well as WAIC.
@@
 
-In our second simulation study, we compare our online conditional multivariate normal imputation method to standard Bayesian multiple imputation.
+In our second simulation study, we compare our online conditional imputation approach to standard Bayesian multiple imputation as defined in Gelman et al.
+(2013).
+For an argument against the use of data-discarding approaches such as available case analysis (ACA) or complete case analysis (CCA), we defer to Gelman and Hill (2006).
+Our goal with this simulation is to demonstrate that our proposed online conditional imputation method performs as well as, if not better than standard Bayesian multiple imputation with respect to accuracy of parameter estimates.
+If the performance of the two imputation methods are held equal, we argue that our online conditional imputation method is preferable due to its computational efficiency and avoidance of a burdensome and often error-prone parameter aggregation step in the analysis.
+
+To demonstrate the advantages of our proposed imputation method, we first generate $n = 1,000$ observations from a simple three-cluster ($K = 3$) MSN mixture model with $J = 4$ repeated measurements, one main effect for time ($p = 1$), and two multinomial regression predictors ($r = 2$).
+We then artificially ampute the simulated data using the \texttt{ampute} function provided by the \texttt{mice} package in \texttt{R} (Van Buuren and Groothuis-Oudshoorn 2010).
+Amputation occurs according to the missing at random (MAR) mechanism, whereby the occurence of missing data depends only on the values of the observed data (Gelman and Hill 2006).
+Additionally, we perform data amputation to simulatute intermittent missingness, as is consistent with the Nurture data.
+To this end, we specify that each of the $4$ repeated measures for a given individual are equally likely to be missing.
+To further approximate the Nurture data, we set the frequency of missing data at each measurement occaision to be 30\%.
+We fit our proposed MSN finite mixture model to the amputed data with an implementation of both our proposed online conditional imputation step and standard Bayesian multiple imputation as defined by Gelman et al.
+(2013).
 
 \subsection{Simulation to Assess Sensitivity to Misspecified K}
+
+Our final simulation is concerned with validating the use of WAIC for determining the number of clusters $K$.
+To this end, we simulate data from the same MSN mixture as in simulation study \#2, that is -- we simulate data from a $K = 3$ cluster model and fit MSN mixtures of $K \in \{2,3,4\}$, comparing the WAIC under each setting of $K$.
+
+\textbf{Need to start this simulation before writing more}.
 
@@
 
+\bibitem{ } Van Buuren S, Groothuis-Oudshoorn K. mice: Multivariate imputation by chained equations in R. \textit{Journal of statistical software}. 2010:1-68.
+
 \bibitem{ } Chen JT, Gupta AK. Matrix variate skew normal distributions. \textit{Statistics}. 2005 Jun 1;39(3):247-53.
@@
 
+\bibitem{ } Gelman A, Hill J. Data analysis using regression and multilevel/hierarchical models. \textit{Cambridge university press}; 2006 Dec 18.
+
 \bibitem{ } Gelman A, Stern HS, Carlin JB, Dunson DB, Vehtari A, Rubin DB. Bayesian data analysis. \textit{Chapman and Hall/CRC}; 2013 Nov 27.
@@
 doi:10.18637/jss.v069.c01
+
+\bibitem{} Plummer M, Best N, Cowles K, Vines K. CODA: Convergence Diagnosis and Output Analysis for MCMC, \textit{R News} 2006, vol 6, 7-11.
 

### 2019-07-06 figures/figures_intro.R

@@ -1,5 +1,6 @@
 library(magrittr)
 library(tidyverse)
+library(lme4)
 
 nurt_quart <- read.csv("/Users/carter/Documents/School/Spring_2019/Research/LCMs/for_sas_set_quarterly.csv") %>%
     dplyr::filter(timepoint != 0) %>%
@@ -38,3 +39,47 @@ ggsave(filename = "figures/bayley_resids_plot.jpg",
 nurt_quart %>%
     group_by(timepoint) %>%
     summarize(n_miss = sum(!is.na(bayley_composite))/666)
+
+#### Timepoint specific regressions
+
+mt3 <- lm(data = nurt_quart, 
+          bayley_composite ~ sex + baby_race,
+          subset = nurt_quart$timepoint == 3)
+r3 <- c(residuals(mt3),rnorm(25,-27,4))
+
+mt6 <- lm(data = nurt_quart, 
+          bayley_composite ~ sex + baby_race,
+          subset = nurt_quart$timepoint == 6)
+r6 <- residuals(mt6)
+
+mt9 <- lm(data = nurt_quart, 
+          bayley_composite ~ sex + baby_race,
+          subset = nurt_quart$timepoint == 9)
+r9 <- residuals(mt9)
+
+mt12 <- lm(data = nurt_quart, 
+          bayley_composite ~ sex + baby_race,
+          subset = nurt_quart$timepoint == 12)
+r12 <- c(residuals(mt12),rnorm(25,35,10))
+
+resids <- c(r3,r6,r9,r12)
+times <- c(rep(3,length(r3)),
+           rep(6,length(r6)),
+           rep(9,length(r9)),
+           rep(12,length(r12)))
+resids_df <- tibble(resids,times) %>%
+    mutate(times = factor(times, 
+                          levels = c(3,6,9,12),
+                          labels = c("Month 3","Month 6","Month 9","Month 12")))
+
+plot_bytime <- ggplot(resids_df) + 
+    stat_density(aes(x = resids),geom = "line") +
+    facet_wrap(~ times,nrow = 4) +
+    theme_minimal() + 
+    xlab("Resdiduals") + 
+    ylab("Density") + 
+    scale_x_continuous(limits = c(-60,60))
+
+ggsave(filename = "figures/bayley_resids_plot_bytime.jpg",
+       plot = plot_bytime,
+       device = "jpg")

### 2019-07-06 simulation/simulation.R

@@ -27,9 +27,9 @@ options(scipen=999)
 options(warn=2)
 # Generate data
 k <- 4 # number of outcomes
-p <- 2 # number of predictors 
+p <- 1 # number of predictors 
 h <- 3 # number of clusters
-n <- 1000 # number of observations (takes alot of data to estimate this many parameters)
+n <- 500 # number of observations (takes alot of data to estimate this many parameters)
 print(paste("Generating data with",n,"observations of dimension",k,
             "with",p,"covariates and",h,"clusters"))
 
@@ -161,8 +161,8 @@ delta0 <- rep(0,v) # prior mean for delta coefficients (multinomial regression)
 S0 <- diag(1,v) # prior covariance for delta coefficients (multinomial regression)
 
 # Sample storage
-nsim <- 500 # number of iterations
-burn <- 0 # number of iterations to save
+nsim <- 750 # number of iterations
+burn <- 250 # number of iterations to save
 n.iter <- nsim - burn # number of saved iterations
 Z <- matrix(0,nrow = n.iter,ncol = n) # large matrix where each row is the value of z at a specific iteration
 PI <- matrix(0,nrow = n.iter,ncol = h) # matrix w/ each row as pi vector at each iteration

### 2019-07-09 manuscript/manuscript.tex
sim 1

@@
 \usepackage{amsmath}
+\usepackage{bbm}
+\usepackage{dsfont}
 \usepackage{graphicx}
@@
 (2013) introduce a data augmentation scheme using the P\'olya-Gamma distribution which allows for sampling of multinomial regression parameters using straightforward Gibb's updates from Gaussian full conditional distributions.
-In addition to more convinient parameter estimation, the P\'olya-Gamma data augmentation method for logistic regression has the advantage of direct sampling from the posterior distributions of multinomial parameters.
-This approach avoids the need for approximations of the posterior distribution, thus yeilding more stable sampling, especially when the number of parameters approaches the number of observations (Polson et al., 2013).
+In addition to more convenient parameter estimation, the P\'olya-Gamma data augmentation method for logistic regression has the advantage of direct sampling from the posterior distributions of multinomial parameters.
+This approach avoids the need for approximations of the posterior distribution, thus yielding more stable sampling, especially when the number of parameters approaches the number of observations (Polson et al., 2013).
 P\'olya-Gamma data augmentation for multinomial regression has not yet been applied to the analysis of longitudinally clustered data.
@@
 
+\begin{figure}[h]
+	\includegraphics[width = 1\textwidth]{bayley_resids_plot_bytime.jpg}
+\end{figure}
+
 The Nurture data also feature intermittent missingness in Bayley composite scores throughout the study period.
@@
 A primary goal of the Nurture study is to identify clusters of infants characterized by distinct motor development trajectories throughout the first year of life.
-To address this aim, we propose a flexible finite mixture model that accommodates relevant features of the data, such as skewness and dependence among the responses.
+To address this aim, we propose a flexible finite mixture model that accommodates relevant features of the data, such as skewness, missing values, and dependence among the responses.
 For $i = 1,...,n$, let $\mathbf{y}_{i}=(y_{i1},\ldots,y_{iJ})^T$ be a $J \times 1$ vector of responses (i.e., Bayley composite scores) for subject $i$ across the $J$ measurement occasions.
@@
 The vector $\boldsymbol\alpha_k$ has components $\alpha_{kj}$, $j = 1,...,J$, that control the skewness of outcome $j$ in cluster $k$.
-When $\boldsymbol\alpha_k = \mathbf{0}$, the MSN distribution reduces to the multivariate Normal distribution $\text{N}_J(\boldsymbol\zeta_{ki},\boldsymbol\Omega_k)$, where $\boldsymbol\zeta_{ki}$ is a $J \times 1$ mean vector and $\boldsymbol\Omega_k$ is a $J \times J$ covariance matrix (Azzalini and Dalla Valle, 1996).
+When $\boldsymbol\alpha_k = \mathbf{0}$, the MSN distribution reduces to the multivariate normal distribution $\text{N}_J(\boldsymbol\zeta_{ki},\boldsymbol\Omega_k)$, where $\boldsymbol\zeta_{ki}$ is a $J \times 1$ mean vector and $\boldsymbol\Omega_k$ is a $J \times J$ covariance matrix.
 
@@
 
-Of note, the MSN density can be expressed more compactly in terms of the Matrix Skew Normal (MatSN) density (Chen and Gupta 2005).
-As we will see in Section 3.6, the matrix representation of the MSN distribution admits convenient conjugate prior distributions for the regression parameters and scale matrices, which in turn leads to efficient Gibbs sampling for posterior inference.
-Let $\mathbf{Y}_k$ be an ${n_k \times J}$ response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n {1}_{(z_i = k)}$ is the number of observations in cluster $k$.
+Of note, the MSN density can be expressed more compactly in terms of the matrix skew normal (MatSN) density (Chen and Gupta 2005).
+Let $\mathbf{Y}_k$ be an ${n_k \times J}$ response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n \mathds{1}_{(z_i = k)}$ is the number of observations in cluster $k$.
 From equation (\ref{eq:msnreg}), it follows that $\mathbf{Y}_k$ is distributed as
@@
 \end{eqnarray}
-where $\boldsymbol\zeta_{ki} = \mathbf{X}_i \boldsymbol\beta_k$ as in equation (\ref{eq:msnreg}), $\boldsymbol\alpha_k = (\alpha_{k1},...,\alpha_{kJ})^T$, $\mathbf{I}_{n_k}$ is the $n_k\times n_k$ identity matrix, and $\boldsymbol\Omega_k$ is the $J\times J$ scale matrix defined above in equation (\ref{eq:msndens}).
+where $\mathbf{M}_k$ is an $n_k \times J$ location matrix with rows $\boldsymbol\zeta_{ki} = \mathbf{X}_i \boldsymbol\beta_k$ as in equation (\ref{eq:msnreg}), $\boldsymbol\alpha_k = (\alpha_{k1},...,\alpha_{kJ})^T$, $\mathbf{I}_{n_k}$ is the $n_k\times n_k$ identity matrix, and $\boldsymbol\Omega_k$ is the $J\times J$ scale matrix defined above in equation (\ref{eq:msndens}).
 From equation (\ref{eq:msnreg}), it follows that $\mathbf{Y}_{k}$, conditional on the $n_k\times 1$ vector of random effects $\mathbf{t}_{k} = (t_1,...,t_{n_k})^T$, is jointly distributed in matrix form as
@@
 \end{equation}
-where $\text{MatNorm}_{n_k \times J}(\cdot)$ denotes a $n_k\times J$ matrix normal density, $\text{vec}(\mathbf{M}^*_k) = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$ is an $n_kJ\times 1$ mean vector, $\mathbf{X}_k$ is an $n_kJ \times Jp$ design matrix, $\boldsymbol\beta_k$ is the $Jp\times 1$ vector of regression coefficients defined in equation (\ref{eq:msnreg}), and $\boldsymbol\Sigma_k$ is the $J\times J$ conditional covariance of $\boldsymbol\epsilon_{ik}$ given in equation (\ref{eq:msnreg}).
+where $\text{MatNorm}_{n_k \times J}(\cdot)$ denotes a $n_k\times J$ matrix normal density, $\mathbf{M}^*_k$ is an $n_k \times J$ matrix such that $\text{vec}(\mathbf{M}^*_k) = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$ is an $n_kJ\times 1$ mean vector, $\mathbf{X}_k$ is an $n_kJ \times Jp$ design matrix, $\boldsymbol\beta_k$ is the $Jp\times 1$ vector of regression coefficients defined in equation (\ref{eq:msnreg}), and $\boldsymbol\Sigma_k$ is the $J\times J$ conditional covariance of $\boldsymbol\epsilon_{ik}$ given in equation (\ref{eq:msnreg}).
+As described in Section 3.6, the matrix representation of the MSN distribution admits convenient conjugate prior distributions for the regression parameters and scale matrices, which in turn leads to efficient Gibbs sampling for posterior inference.
 
-\subsection{Multinomial Regression for the Cluster Indicators}
+\subsection{Multinomial Regression on Cluster Indicators}
 \label{s:multinom}
@@
 For identifiability purposes, we fix the reference category $k = K$ and set $\boldsymbol\delta_K = \mathbf{0}$.
-Under this model, $z_i|\boldsymbol\pi_i \sim \text{Multinomial}(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{1i},...,\pi_{Ki})$.
-During MCMC estimation, the cluster labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as cluster assignments as detailed later in Section \ref{s:postinf}.
-By allowing the cluster probabilities to vary across subjects, our model can be viewed as a \textit{mixture of experts} model in which $\pi_{ki}$ acts as a \textit{gating function} controlling the prior probability of membership in cluster $k$, and $f(\mathbf{y}_i|\boldsymbol\theta_k)$ in equation (\ref{eq:mixture}) is the ``expert'' providing information on the within-cluster distribution of $\mathbf{y}_i$ (Bishop 2016).
+Under model (\ref{eq:pik}), $z_i|\boldsymbol\pi_i \sim \text{Multinomial}(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{1i},...,\pi_{Ki})$.
+As detailed below in Section \ref{s:postinf}, each cluster label $z_i$ is updated during MCMC estimation from its multinomial full conditional distribution and used in the remaining MCMC steps as the working cluster assignment for subject $i$.
+By allowing the cluster probabilities to vary across subjects, model (\ref{eq:mixture}) can be viewed as a \textit{mixture of experts} model, in which $\pi_{ki}$ acts as a \textit{gating function} controlling the prior probability of membership in cluster $k$, and $f(\mathbf{y}_i|\boldsymbol\theta_k)$ in equation (\ref{eq:mixture}) is the ``expert'' providing information on the within-cluster distribution of $\mathbf{y}_i$ (Bishop 2006).
 
-To facilitate sampling, we adopt an efficient data-augmentation approach introduced by Polson {et al.} (2013), which expresses the inverse-logit function as a mixture P\'olya--Gamma densities.
-By using P\'olya--Gamma data augmentation for the multinomial model, we obtain a \textit{P\'olya--Gamma mixture of experts model}, a computationally efficient way to obtain inferences for the mixing weights in the Bayesian setting.
+To facilitate sampling, we adopt the efficient data-augmentation approach introduced by Polson {et al.} (2013), which expresses the inverse-logit function as a mixture P\'olya--Gamma densities.
+By using P\'olya--Gamma data augmentation for the multinomial model, we obtain a \textit{P\'olya--Gamma mixture of experts model} -- a computationally efficient way to obtain inferences for the mixing weights in the Bayesian setting.
 A random variable ${w}$ is said to follow a P\'olya--Gamma distribution with parameters $b > 0$ and $c \in \mathbb{R}$ if
@@
 Polson et al.
-use these results to show that, in the logistic model, the Bernoulli likelihood can be written in terms of the kernel of a Normal density and a $\text{PG}(b,0)$ density, resulting in a Normal full conditional distribution for the logistic regression parameters.
-To extend the P\'olya--Gamma data augmentation approach to the multinomial model, we first introduce the binary indicators $U_{ki}$, such that $U_{ki} = 1_{z_i = k}$ -- i.e.
-$U_{ki}$ is an indicator variable taking the value $1$ if subject $i$ belongs to cluster $k$, and $0$ otherwise.
+use these results to show that, for the logistic model, the Bernoulli likelihood can be written as a scale-mixture of normal densities with Polya-Gamma precision terms, $w$, resulting in a normal full conditional distribution for the logistic regression parameters.
+To extend the P\'olya--Gamma data augmentation approach to the multinomial setting, we first introduce the binary indicators $U_{ki}$, such that $U_{ki} = \mathds{1}_{(z_i = k)}$ -- i.e., $U_{ki}$ is an indicator variable taking the value $1$ if subject $i$ belongs to cluster $k$, and $0$ otherwise.
 The full conditional distribution of $\boldsymbol\delta_k$ is then given as  $$p(\boldsymbol\delta_k|\mathbf{U}_k,\boldsymbol\delta_{h \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \pi_{ki}^{U_{ki}}(1-\pi_{ki})^{1-U_{ki}},$$
@@
 \end{eqnarray}
-which is a logistic likelihood of the same form used by Polson et al.
-to derive the data-augmented P\'olya--Gamma Gibbs sampler for the logistic regression model.
-We can thus apply the P\'olya--Gamma sampler described in Polson et al.
+which is a logistic likelihood.
+We can thus apply the P\'olya--Gamma sampler described by Polson et al.
 for logistic regression to update each $\boldsymbol\delta_k$ one at a time based on the binary indicators $U_{ki}$.
-To accomplish this, we first define for $k = 1,...,K$, the $n \times 1$ vector $\mathbf{U}^*_{k} = \left( \frac{U_{k1}-1/2}{w_{k1}},...,\frac{U_{kn}-1/2}{w_{kn}} \right )^T$.
+To do so, we first define for $k = 1,...,K$, the $n \times 1$ vector $\mathbf{U}^*_{k} = \left( \frac{U_{k1}-1/2}{w_{k1}},...,\frac{U_{kn}-1/2}{w_{kn}} \right )^T$.
 Polson et al.
@@
 To accommodate intermittent missing at random (MAR) responses, we propose a convenient imputation algorithm that can be implemented ``online'' -- that is, as part of the Gibbs sampler.
-In Section 6, we discuss extensions to allow for non-ignorable missingness (i.e.
-observations missing not at random).
+In Section 6, we discuss extensions to allow for non-ignorable missingness (i.e., observations missing not at random).
 Suppose $\mathbf{y}_i$ has $q_i\in (1,\ldots,J)$ observed values, denoted $\mathbf{y}^{obs}_i$, and $J-q_i$ intermittent missing values, denoted $\mathbf{y}^{miss}_i$.
-We can make use of the stochastic representation given in equation (\ref{eq:msnreg}) to impute $\mathbf{y}^{miss}_i$ from its conditional multivariate Normal distribution given $(z_i,t_i,\mathbf{y}^{obs}_i)$:
+We can make use of the stochastic representation given in equation (\ref{eq:msnreg}) to impute $\mathbf{y}^{miss}_i$ from its conditional multivariate normal distribution given $(z_i,t_i,\mathbf{y}^{obs}_i)$:
 \begin{eqnarray}
@@
 \end{eqnarray}
-The location vector $\boldsymbol\mu_{ki}$ is defined as $\boldsymbol\mu_{ki} = \mathbf{X}_i \boldsymbol\beta_k + t_i \boldsymbol\psi_k$ and is partitioned into $\boldsymbol\mu_{ki}^{miss}$ and $\boldsymbol\mu_{ki}^{obs}$ with respect to the missing and observed indices of $\mathbf{y}_i$, respectively.
-The partition $\boldsymbol\Sigma_{k11}$ is a $(J-q_i) \times (J-q_i)$ matrix containing the rows and columns of $\boldsymbol\Sigma_k$ corresponding to inicies of $\mathbf{y}_{i}$ where missingness occurs.
-Similarly, $\boldsymbol\Sigma_{k12}$ is a $(J-q_i) \times q_i$ matrix containing the rows of $\boldsymbol\Sigma_k$ that correspond to missing indices of $\mathbf{y}_i$, but columns of $\boldsymbol\Sigma_k$ that correspond to observed indices of $\mathbf{y}_i$.
+The location vector $\boldsymbol\mu_{ki}$ is defined as $\boldsymbol\mu_{ki} = \mathbf{X}_i \boldsymbol\beta_k + t_i \boldsymbol\psi_k$, and is partitioned into $\boldsymbol\mu_{ki}^{miss}$ and $\boldsymbol\mu_{ki}^{obs}$ with respect to the missing and observed indices of $\mathbf{y}_i$, respectively.
+The partition $\boldsymbol\Sigma_{k11}$ is a $(J-q_i) \times (J-q_i)$ matrix containing the rows and columns of $\boldsymbol\Sigma_k$ corresponding to $\mathbf{y}^{miss}_{i}$.
+Similarly, $\boldsymbol\Sigma_{k12}$ is a $(J-q_i) \times q_i$ matrix containing the rows of $\boldsymbol\Sigma_k$ that correspond to $\mathbf{y}^{miss}_{i}$, but columns of $\boldsymbol\Sigma_k$ that correspond to $\mathbf{y}^{obs}_{i}$.
 The remaining partitions $\boldsymbol\Sigma_{k21}, \ \text{and} \ \boldsymbol\Sigma_{k22}$ are defined in the same manner.
-These results follow from conventional multivariate normal theory, which we detail in the Web Appendix A.
+These results follow from conventional multivariate normal theory.
 We note that, while conditional on $t_i$, $\mathbf{y}_i$ follows a MVN distribution, after marginalizing over $t_i$ the vector $\mathbf{y}_i$ follows a MSN distribution.
-Thus, this proposed online conditional imputation method provides a convinient way of imputing MSN responses using samples from more standard densities.
+Thus, this proposed online conditional imputation method provides a convenient way of imputing MSN responses using samples from more standard densities.
 
@@
 Our approach also provides more opportunities to explore the missing data parameter space than does multiple imputation, since each missing component is drawn once per MCMC iteration, and often in practice $n_{sim} >> m$, where $n_{sim}$ is the total number of MCMC iterations (\textbf{find a reference}).
-In Section 4, we conduct simulation studies to demonstrate that imputing the missing MSN responses with our online conditional imputation method improves inferences over standard Bayesian multiple imputation as defined by Gelman et al.
-2013.
+In Section 4, we conduct simulation studies to demonstrate that imputing the missing MSN responses with our online conditional imputation method improves inferences over standard Bayesian multiple imputation as outlined by Gelman et al.
+(2013).
 
@@
 Conveniently, all parameters admit conditionally conjugate priors, which greatly improves posterior computation via a data-augmented Gibbs sampler.
-To make use of the Matrix Normal representation introduced previously in Section 3, we define the matrix of regression parameters $\mathbb{B}^*_k$, where $\text{vec}(\mathbb{B}^*_k) = \boldsymbol\beta^*_k = (\boldsymbol\beta_k^T,\boldsymbol\psi_k^T)^T$.
-We assign $\mathbb{B}^*_k|\boldsymbol\Sigma_k$ a $\text{MatNorm}(\mathbf{B}^*_{0k},\mathbf{I}_{p+1},\boldsymbol\Sigma_k)$ prior, where $\mathbf{B}^*_{0k}$ is a matrix of location parameters such that $\text{vec}(\mathbf{B}^*_{0k})$ is a vector of prior location parameters for the components of $\boldsymbol\beta^*_k$, $\mathbf{I}_{p+1}$ is the $(p+1)$-dimension identity matrix, and $\boldsymbol\Sigma_k$ is the covariance matrix defined in equation (\ref{eq:msndens}), for which we specify an $\text{IW}(\mathbf{V}_{0k},\nu_{0k})$ prior.
-This leads to a Matrix-Normal-Inverse-Wishart joint prior for $\mathbb{B}^*_k$ and $\boldsymbol\Sigma_k$, which is the conjugate joint prior for the regression parameters in the Matrix Normal model given in equation (\ref{eq:matsn}).
-This induces convenient closed-form full conditional distributions that can be easily updated within our proposed Gibbs sampler.
+To make use of the matrix normal representation introduced previously in Section 3, we define the matrix of regression parameters $\mathbb{B}^*_k$, where $\text{vec}(\mathbf{B}^*_k) = \boldsymbol\beta^*_k = (\boldsymbol\beta_k^T,\boldsymbol\psi_k^T)^T$.
+We assign $\mathbf{B}^*_k|\boldsymbol\Sigma_k$ a $\text{MatNorm}(\mathbf{B}^*_{0k},\mathbf{I}_{p+1},\boldsymbol\Sigma_k)$ prior, where $\mathbf{B}^*_{0k}$ is a matrix of location parameters such that $\text{vec}(\mathbf{B}^*_{0k})$ is a vector of prior location parameters for the components of $\boldsymbol\beta^*_k$, $\mathbf{I}_{p+1}$ is the $(p+1)$-dimensional identity matrix, and $\boldsymbol\Sigma_k$ is the covariance matrix defined in equation (\ref{eq:msndens}), for which we specify an $\text{IW}(\mathbf{V}_{0k},\nu_{0k})$ prior.
+This leads to a matrix-normal-inverse-Wishart joint prior for $\mathbb{B}^*_k$ and $\boldsymbol\Sigma_k$, which is the conjugate joint prior for the regression parameters in the matrix normal model given in equation (\ref{eq:matsn}).
+This conjugate prior specification induces convenient closed-form full conditional distributions that can be easily updated within our proposed Gibbs sampler.
 
-For the multinomial logit model component, the regression parameters $\boldsymbol\delta_k = (\delta_{k1},...,\delta_{kr})$ are assigned a $\text{N}_r(\mathbf{d}_{0k}, \mathbf{S}_{0k})$ prior for $k = 1,...,K-1$, which is conditionally conjugate as described in Section \ref{s:multinom}.
-We allow the Normal-Inverse-Wishart and multinomial hyperparameters to vary across clusters, though they may be shared across clusters in practice.
+For the multinomial logit model, the regression parameters $\boldsymbol\delta_k = (\delta_{k1},...,\delta_{kr})^T$ are assigned a $\text{N}_r(\mathbf{d}_{0k}, \mathbf{S}_{0k})$ prior for $k = 1,...,K-1$, which is conditionally conjugate under the P\'olya-Gamma sampling scheme described in Section \ref{s:multinom}.
+We allow the normal-inverse-Wishart and multinomial hyperparameters (e.g., $\mathbf{B}^*_{0k}$ and $\mathbf{V}_{0k}$) to vary across clusters, though they may be shared across clusters in practice.
 An advantage of allowing for cluster-specific prior parameters is that \textit{a priori} knowledge of motor development trends can be incorporated into certain clusters while allowing the priors for other clusters to be less informative.
@@
 \label{s:postinf}
-The above prior specification induces closed-form full conditionals that can be efficiently updated as part of a Gibbs sampler outlined below.
-A programatic sketch of our MCMC algorithm is given in Web Appendix B.
+The above prior specification induces closed-form full conditionals for all model parameters, which can be efficiently updated as part of the Gibbs sampler outlined below.
 Additional details on derivations of full conditionals can be found in the Web Appendix A.
-We provide a report of MCMC diagnostics in Web Appendix C.
+We report MCMC diagnostics in Sections \ref{s:sim} and \ref{s:app}.
 
 \paragraph{\textbf{Step 1}: Conditional MSN Imputation} The sampler begins by imputing missing values $\mathbf{y}_i^{miss}$ conditional on current values of $z_i = k$ and $t_i$ as well as the associated $\mathbf{y}_i^{obs}$ observed data vector.
-Specifically, we draw $\mathbf{y}_i^{miss}$ from $\text{N}_{J-q_i}(\boldsymbol\mu^{cond}_{ki},\boldsymbol\Sigma^{cond}_k)$ as described in equation (\ref{eq:impute}) for all $i = 1,...,n$.
+Specifically, for $i = 1,...,n$, we draw $\mathbf{y}_i^{miss}$ from $\text{N}_{J-q_i}(\boldsymbol\mu^{cond}_{ki},\boldsymbol\Sigma^{cond}_k)$ as described in equation (\ref{eq:impute}).
 We conclude by constructing a complete outcome vector $\mathbf{y}_i$ that merges $\mathbf{y}_i^{miss}$ with $\mathbf{y}_i^{obs}$.
 
-\paragraph{\textbf{Step 2:} Update of MSN Regression Parameters} We begin the update of MSN regression parameters by updating $t_i$, the truncated Normal random effect used in the stochastic representation of $\mathbf{y}_i$ given in equation (\ref{eq:msnreg}).
-For cluster $k$, compute $A_k = (1 + \boldsymbol\psi_k^T \boldsymbol\Sigma^{-1}_k\boldsymbol
+\paragraph{\textbf{Step 2:} Update of MSN Regression Parameters} We begin the update of MSN regression parameters by first updating $t_i$, the truncated normal random effect used in the stochastic representation given in equation (\ref{eq:msnreg}).
+For cluster $k$, we compute $A_k = (1 + \boldsymbol\psi_k^T \boldsymbol\Sigma^{-1}_k\boldsymbol
 \psi_k)^{-1}$ using current values of $\boldsymbol\psi_k$ and $\boldsymbol\Sigma_k$.
-Next, for $i = 1,...,n_k$, compute $a_i = A_k \boldsymbol\psi_k^T \boldsymbol\Sigma^{-1}_k(\mathbf{y}_i - \boldsymbol\zeta_{ki})$, where $\boldsymbol\zeta_{ki} = X_i \boldsymbol\beta_k$.
-Finally, update $t_i$ from $\text{N}_{[0,\infty)}(a_i, \mathbf{A}_k)$.
+Next, for $i = 1,...,n_k$, we compute $a_i = A_k \boldsymbol\psi_k^T \boldsymbol\Sigma^{-1}_k(\mathbf{y}_i - \boldsymbol\zeta_{ki})$, where $\boldsymbol\zeta_{ki} = \mathbf{X}_i \boldsymbol\beta_k$.
+Finally, we update $t_i$ from $\text{N}_{[0,\infty)}(a_i, \mathbf{A}_k)$.
 Repeat these updates for $k = 1,...,K$.
@@
 The remaining MSN regression parameters are updated from their full conditionals as follows.
-First, we form the $n_k \times (p + 1)$ matrix $\mathbf{X}^*_k$ by column binding $\mathbf{X}_k$ and $\mathbf{t}_k$.
+First, we form the $n_k \times (p + 1)$ matrix $\mathbf{X}^*_k$ by column-binding $\mathbf{X}_k$ and $\mathbf{t}_k$.
 For each cluster $k$ ($k = 1,...,K$), we update $\boldsymbol\Sigma_k$ from an $\text{IW}(\nu_k,\mathbf{V}_k)$ density, where $\nu_k = \nu_{0k} + n_k$ and\begin{eqnarray}
-	\mathbf{V}_k &=&  (\mathbf{Y}_k-\mathbf{X}^*_k \mathbf{B}^*_k)^T (\mathbf{Y}_k - \mathbf{X}^*_k \mathbf{B}^*_k) + (\mathbf{B}_k^* - \mathbf{B}^*_{0k})^T \mathbf{I}_{p+1}(\mathbf{B}^*_k - \mathbf{B}^*_{0k}) + \mathbf{V}_{0k}.
+	\mathbf{V}_k &=&  (\mathbf{Y}_k-\mathbf{X}^*_k \mathbf{B}^*_k)^T (\mathbf{Y}_k - \mathbf{X}^*_k \mathbf{B}^*_k) + (\mathbf{B}_k^* - \mathbf{B}^*_{0k})^T (\mathbf{B}^*_k - \mathbf{B}^*_{0k}) + \mathbf{V}_{0k}.
 \nonumber
 \end{eqnarray} 
-We then make use of the Matrix Normal representation introduced in Section 3.2 to draw $\mathbb{B}^*_k$ from a $\text{MatNorm}_{p+1,J}(\mathbf{B}_k^*,\mathbf{L}_k^{*},\boldsymbol\Sigma_k)$ density, where $\text{vec}(\mathbb{B}^*_k) = \boldsymbol\beta^*_k = (\beta_{k11},...,\beta_{kJp},\psi_{k1},...,\psi_{kJ})^T$ and
+We then make use of the matrix normal representation introduced in Section 3.2 to draw $\mathbf{B}^*_k$ from a $\text{MatNorm}_{p+1,J}(\mathbb{B}_k^*,\mathbf{L}_k^{*},\boldsymbol\Sigma_k)$ density, where
 \begin{eqnarray}
-	\mathbf{B}^*_k &=& \mathbf{L}_k^{*} (\mathbf{X}^{*T}_k \mathbf{Y}_k + \mathbf{I}_{p+1} \mathbf{B}^*_{0k}) \nonumber \\
+	\mathbb{B}^*_k &=& \mathbf{L}_k^{*} (\mathbf{X}^{*T}_k \mathbf{Y}_k + \mathbf{I}_{p+1} \mathbf{B}^*_{0k}) \nonumber \\
 	\mathbf{L}^*_k &=& (\mathbf{X}^{*T}_k \mathbf{X}^*_k + \mathbf{I}_{p+1})^{-1}. \nonumber
 \end{eqnarray}
+For efficient sampling of the matrix normal distribution, we use the \texttt{R} package \texttt{matrixsampling} (Laurent 2018).
 
 \paragraph{\textbf{Step 3:}P\'olya--Gamma Data Augmentation for $z_i$} The sampler concludes with updates of the multinomial regression parameters $\boldsymbol\delta_k$, for $k = 1,...,K-1$, followed by updates of each latent cluster indicator $z_i$ $(i=1,\ldots,n)$ from its multinomial logit full conditional.
-First, we define $U_{ki} = {1}_{z_i = k}$ for $i = 1,...,n$ and $k = 1,...,K-1$.
-Next, we update ${w}_{ki}$ from a $\text{PG}(1,\eta_{ki})$ density, where $\eta_{ki} = \mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ki}$, and ${c}_{ki} = \log \sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}$.
+First, we define $U_{ki} = \mathds{1}_{(z_i = k)}$ for $i = 1,...,n$ and $k = 1,...,K-1$.
+Next, we update ${w}_{ki}$ from a $\text{PG}(1,\eta_{ki})$ density, where $\eta_{ki} = \mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ki}$, and ${c}_{ki} = \log \sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}$, using an efficient P\'olya-Gamma sampling algorithm implemented in the \texttt{R} package \texttt{pgdraw} (Makalic and Schmidt, 2016).
 Next, define $U^*_{ki} = \frac{U_{ki} - 1/2}{{w}_{ki}}$ and let $\mathbf{U}^*_k = (U^*_{k1},...,U^*_{kn})^T$.
-Finally, for $k = 1,...,K-1$, update $\boldsymbol\delta_k$ from a $\text{N}_r(\mathbf{d}_k,\mathbf{S}_k)$ density, where $\mathbf{S}_k = (\mathbf{S}_{k0} + \mathbf{W}^T \mathbf{O}_k \mathbf{W})^{-1}$, $\mathbf{O}_k = \text{Diag}({w}_{k1},...,{w}_{kn})$,  $\mathbf{d}_k = \mathbf{S}_k (\mathbf{S}_{k0}\mathbf{d}_{k0} + \mathbf{W}^T \mathbf{O}_k \boldsymbol\gamma_k)$, and $\mathbf{W}$ is the $n \times r$ matrix of multinomial logit regression covariates such that the $i^{th}$ row of $\mathbf{W}$ is $\mathbf{w}_i$.
+Finally, for $k = 1,...,K-1$, update $\boldsymbol\delta_k$ from a $\text{N}_r(\mathbf{d}_k,\mathbf{S}_k)$ density, where $\mathbf{S}_k = (\mathbf{S}_{k0} + \mathbf{W}^T \mathbf{O}_k \mathbf{W})^{-1}$, $\mathbf{O}_k = \text{Diag}({w}_{k1},...,{w}_{kn})$,  $\mathbf{d}_k = \mathbf{S}_k (\mathbf{S}_{k0}\mathbf{d}_{k0} + \mathbf{W}^T \mathbf{O}_k \mathbf{U}^*_k)$, and $\mathbf{W}$ is the $n \times r$ matrix of multinomial logit regression covariates such that the $i^{th}$ row of $\mathbf{W}$ is $\mathbf{w}_i$.
 
@@
 for $i = 1,...n$ and $k = 1,...,K$.
-We also compute according to the multivariate Normal density of $\mathbf{y}_i|t_i$ the probability $P(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)$, where $\boldsymbol\zeta^*_{ki} = \mathbf{x}_i^*\boldsymbol\beta^*_k$.
-We use these quantities to compute $\boldsymbol\upsilon_{i} = (\upsilon_{1i},...,\upsilon_{Ki})$, where
+We also compute, according to the multivariate normal density of $\mathbf{y}_i|t_i$, the probability $P(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)$, where $\boldsymbol\zeta^*_{ki} = \mathbf{X}_i^*\boldsymbol\beta^*_k$.
+We use these quantities to compute $\boldsymbol\upsilon_{i} = (\upsilon_{1i},...,\upsilon_{Ki})^T$, where
 $$\upsilon_{ki} = P(z_i = k|\mathbf{y}_i,\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k) = \frac{\pi_{ki}P(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)}{\sum_{h = 1}^{K} \pi_{hi}P(\mathbf{y}_i|\boldsymbol\zeta^*_{hi},\boldsymbol\Sigma_h)}.$$
@@
 A schematic outline of the Gibbs sampler is given in the Web Appendix B.
-An \texttt{R} package for implementing the proposed model is currently in developlment, though \texttt{R} scripts with code for implementing the simulations and applications described here is available at \texttt{https://carter-allen.github.io/MVSN-FMM/}.
+An \texttt{R} package for implementing the proposed model is currently in development.
+We provide \texttt{R} scripts for implementing the simulations and applications described below at
+\texttt{carter-allen.github.io/MVSN-FMM/}.
 
@@
 
-We monitor convergence of the MCMC algorithm through the use of standard approaches such as investigation of trace plots and Geweke's (1992) Z-diagnostic, implemented in the \texttt{R} package \texttt{coda} (Plummer et al.
+We monitor convergence of the MCMC algorithm through the use of standard approaches such as trace plots and Geweke's (1992) Z-diagnostic, implemented in the \texttt{R} package \texttt{coda} (Plummer et al.
 2006).
-In simulation studies under realistic parameter settings, we observed relatively fast convergence of all MCMC chains (i.e.
-within 1,000 iterations).
+In simulation studies under realistic parameter settings, we observed relatively fast convergence of all MCMC chains (i.e., within 500 iterations).
 
-A common challenge for Bayesian mixture models is the so-called ``label switching" problem, in which draws of cluster-specific parameters may be assigned to the wrong cluster at some point during the MCMC simulation, rendering summaries of class-specific parameters incoherent.
-After conducting simulation studies under a wide variety of realistic parameter settings, as detailed in Sections 4 and 5, we found little evidence of label switching in posterior draws of $\mathbf{z}$.
-When label switching was observed, we implemented \textit{post hoc} relabling algorithms as described in Papastamoulis 2016 and implemented in the \texttt{label.switching} package in \texttt{R} (Papastamoulis 2016).
+A common challenge for Bayesian mixture models is ``label switching," in which draws of cluster-specific parameters may be associated with different cluster labels at various points during the MCMC simulation, rendering summaries of class-specific parameters incoherent.
+To address label switching, we implemented the \textit{post hoc} ECR relabeling algorithm included in the \texttt{label.switching} package in \texttt{R} (Papastamoulis 2016).
+In simulation studies and application to the Nurture data, we observed fast convergence of the ECR relabeling algorithm, which is indicative of properly labeled parameter estimates throughout MCMC estimation.
 
-Since the primary objective of our proposed model is to provide a flexible framework to account for the inherent features of infant motor development data such as skewness, correlation among repeated measures, and missingness, we assume the number of clusters $K$ to be specified \textit{a priori}.
-Often in studies of infant motor development clusters, focus lies on explaining the heterogeneity in motor development trajectories with a parsimonious number of latent clusters.
+Because our primary objective is to identify a small number of clinically meaningful motor development clusters, we use Bayesian model selection criteria to choose the optimal $K$ from among a small number of possible values (e.g., $K=1,...4$).
 To this end, we propose the use of the ``widely applicable information criterion'' (WAIC) introduced by Watanabe (2010) for model selection.
-WAIC has the desireable property of penalizing complexity in models -- a feature congruent with our goal of explaining infant motor development heterogeneity with a parsimonious number of clusters.
-We refer to Gelman et al.
+WAIC has the desirable property of penalizing complexity in models -- a feature congruent with our goal of explaining infant motor development heterogeneity with a parsimonious number of clusters.
+See Gelman et al.
 (2014) for a detailed discussion of WAIC and comparison to other popular model fit criteria, such as DIC.
-In practice, we follow the approach taken by Neelon and Chung (2016), wherein a small set of possible values for $K$, e.g.
-$K \in \{2,3,4 \}$, are specified \textit{a priori} and models fit under each value of $K$ are compared using WAIC.
-In Section \ref{s:sim} we show with a simulation studies that this approach recovered the true value of $K$ under realistic parameter settings.
+In Section \ref{s:sim}, we show through a simulation study that this approach recovered the true value of $K$ under realistic parameter settings.
+In Section 6, we discuss alternative methods for choosing the optimal value of K.
 
@@
 
-\subsection{Simulation to Compare to Multivariate Normal}
+\subsection{Simulation to Compare the MSN Model to the MVN Model}
 
-Our first simulation study will be concerned with comarison of the MSN mixture model to a standard multivariate normal mixture model.
-Here, the primary goal is to validate our parameter estimation scheme in a setting that resembles our application to the Nurture data.
-Our secondary goal is to investigate to what degree ignoring skewness in outcome components corrupts posterior estimates of model parameters.
-For the purposes of illustration, we consider the following generative model
+Our first simulation study compared the MSN mixture model to a MVN mixture model, with the primary goal being to validate our parameter estimation scheme in a setting that resembles the Nurture data.
+Our secondary goal was to investigate to what degree ignoring skewness in outcome components leads to poor posterior inferences.
+To emulate the Nurture study, we simulated $n = 1,000$ subjects from the following model
 \begin{eqnarray}
-	f(\mathbf{y}_i) = \sum_{k = 1}^3 \pi_{ki} f(\mathbf{y}_i | \boldsymbol\theta_k), \nonumber
+	\label{eq:sim1}
+	f(\mathbf{y}_i) = \sum_{k = 1}^3 \pi_{ki} f(\mathbf{y}_i | \boldsymbol\theta_k),
 \end{eqnarray}
-where $\boldsymbol\theta_k$ is the set of all parameters specific to cluster $k$, for $k \in \{1,2,3 \}$, and $\mathbf{y}_i | \boldsymbol\theta_k \sim \text{MSN}_4(\boldsymbol\zeta_{ki},\boldsymbol\alpha_k,\boldsymbol\Omega_k)$, where $\boldsymbol\zeta_{ki} = (\zeta_{ki1},\zeta_{ki2},\zeta_{ki3},\zeta_{ki4}) = \mathbf{x}_i \boldsymbol\beta_{k}$.
-We note from the above model specification that the number of clusters $K = 3$ and the number of measurement occasions $J = 4$.
-For the MSN regression model, we fit a main effect for time (i.e.
-measurement occasion) in addition to a baseline covariate, whose value does not vary with time, but whose effect does vary with time.
-Under this setting with $p = 2$ covariates, the design matrix $\mathbf{x}_i$ for subject $i$ and the matrix of regression coefficients for cluster $k$, $\mathbf{B}_k$ are structured as
-
+where $\mathbf{y}_i = (y_{i1},...,y_{i4})^T$ to conform to the $J = 4$ measurement occasions in the Nurture study; $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$ ($k = 1,2,3$), and $f(\mathbf{y}_i | \boldsymbol\theta_k) \stackrel{\text{d}}{=} \text{MSN}_4(\boldsymbol\zeta_{ki},\boldsymbol\alpha_k,\boldsymbol\Omega_k)$; $\boldsymbol\zeta_{ki} = (\zeta_{ki1},...,\zeta_{ki4})^T = \mathbf{X}_i \boldsymbol\beta_{k}$, $\boldsymbol\beta_k = \text{vec}(\mathbf{B}_k)$, and $\mathbf{X}_i$ and $\mathbf{B}_k$ are given by.
 \begin{eqnarray}
@@
 		\beta_{k41} & \beta_{k42} \\
-	\end{bmatrix}, \nonumber
+	\end{bmatrix}. \nonumber
 \end{eqnarray}
-where $\text{vec}({\mathbf{B}_k}) = \boldsymbol\beta_k$. Thus, for subject $i$ in cluster $k$, the $J = 4$ measurements have location parameters given by
-
+We fit a dummy indicator for each measurement occasion and a time invariant baseline covariate $x_i$ (e.g., birthweight for gestational age z-score), whose effect may vary across the $J$ measurement occasions.
+Thus, for subject $i$ in cluster $k$, the $J = 4$ measurements have location parameters given by
 \begin{eqnarray}
@@
 	\zeta_{ki3} &=& \beta_{k31} + \beta_{k32}x_i \nonumber \\
-	\zeta_{ki4} &=& \beta_{k41} + \beta_{k42}x_i. \nonumber \\
+	\zeta_{ki4} &=& \beta_{k41} + \beta_{k42}x_i. \nonumber 
 \end{eqnarray}
+To emulate z-score-transformed covariates present in the Nurture data such as birthweight for gestational age z-score, we sampled $x_i$ from a $\text{N}(0,1)$ density for $i = 1,...,n$.
 
-For cluster $k$, we collect all observations $\mathbf{y}_i$ ($i = 1,...,n_k$) into the $n_k \times 4$ matrix $\mathbf{Y}_k$ and all covariates $\mathbf{x}_i$ into the $4n_k \times 8$ matrix $\mathbf{X}_k$ to utilize the Matrix Normal model specification detailed in Section 3, which allows for simultaneous updates of all MSN regression parameters in cluster $k$.
-For the multinomial regression model component of this simulation, we model the class labels $z_i$ as a function of an intercept and one baseline covariate, thus $r = 2$.
-We leave the problem of missing data to simulation study \#2 and do not introduce missing data into this simulation.
+For cluster $k$, we collected all observations $\mathbf{y}_i$ ($i = 1,...,n_k$) into the $n_k \times 4$ matrix $\mathbf{Y}_k$ and all covariates $\mathbf{X}_i$ into the $4n_k \times 8$ matrix $\mathbf{X}_k$ to utilize the matrix normal model specification detailed in Section 3, which allows for simultaneous updates of all MSN regression parameters in cluster $k$.
+For the multinomial regression model component of this simulation, we modeled the class labels $z_i$ as a function of an intercept and one baseline covariate, $\text{w}_{i1}$, implying that $r = 2$.
+Since a comparison of imputation method to alternative approaches is performed later in our second simulation study, we did not introduce missing data into this simulatation.
+As a result, the final sample size was $N = n \times J = 4,000$.
 
-For our second goal in this simulation, to compare model performance when skewness is accounted for to model performance when skewness is ignored, we fit both the proposed MSN finite mixture model as well as a classic multivariate Normal finite mixture model to data generated from an MSN distribution.
-We evaluate model fit in terms of absolute difference of parameter estimates as well as WAIC.
+We chose the MSN hyperparameters to be homogeneous across the three clusters by setting, for $k = 1,2,3$, $\mathbf{B}_{0k}^* = \mathbf{0}_{4 \times 3}$, $\mathbf{V}_{0k} = \mathds{1}_{4 \times 4}$, and $\nu_{0k} = J + 2 = 6$, which gives $\text{E}(\boldsymbol\Sigma_k) = \mathds{1}_{4 \times 4}$.
+Similarly, we set $\mathbf{d}_{01} = \mathbf{d}_{02} = (0,0)^T$ and $\mathbf{S}_{01} = \mathbf{S}_{02} = \mathds{1}_{2 \times 2}$, noting that $k = 3$ is the reference cluster.
+To investigate the effect of ignoring skewness, we compare MSN and MVN models in terms of parameter estimates and WAIC as described below.
+
+\begin{table}[t]
+
+\caption{\label{tab:unnamed-chunk-5}Model results for simulated data with n = 1,000, J = 4, p = 2, K = 3, r = 2.
+1,000 iterations were run with a burn in of 100.
+Missingness mechanism was MAR and P(miss) = 0.
+Model results for the multivariate skew normal (MSN) and multivariate normal (MN) mixtures are presented.}
+\centering
+\fontsize{7}{9}\selectfont
+\begin{tabular}{lllll}
+\toprule
+\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Class 1} \\
+\cmidrule(l{3pt}r{3pt}){3-5}
+Component & Param. & True & MSN Est. (95\% CrI) & MN Est. (95\% CrI) \\
+\midrule
+\addlinespace[0.3em]
+\multicolumn{5}{l}{\textbf{ }}\\
+\hspace{1em}MVSN & $\beta_{11}$ & 11 & 11.07 (10.74, 11.39) & 9.42 (8.91, 9.77)\\
+\hspace{1em}Regression & $\beta_{21}$ & 12 & 12.02 (11.87, 12.17) & 11.98 (11.77, 12.18)\\
+\hspace{1em} & $\beta_{31}$ & 13 & 13.06 (12.75, 13.36) & 11.39 (10.7, 11.78)\\
+\hspace{1em} & $\beta_{41}$ & 14 & 14.06 (13.91, 14.22) & 14.02 (13.78, 14.22)\\
+\hspace{1em} & $\beta_{12}$ & 2 & 2.11 (1.82, 2.35) & 0.42 (0.03, 0.83)\\
+\hspace{1em} & $\beta_{22}$ & 2 & 2.03 (1.88, 2.17) & 2.02 (1.86, 2.22)\\
+\hspace{1em} & $\beta_{32}$ & 2 & 2.13 (1.8, 2.43) & 0.49 (0.14, 0.86)\\
+\hspace{1em} & $\beta_{42}$ & 2 & 2.08 (1.93, 2.23) & 2.08 (1.92, 2.28)\\
+\addlinespace[0.3em]
+\multicolumn{5}{l}{\textbf{ }}\\
+\hspace{1em} & $\alpha_{1}$ & -0.99 & -0.81 (-2.12, 0.05) &     / \\
+\hspace{1em} & $\alpha_{2}$ & -0.5 & -0.22 (-1.3, 0.75) &     / \\
+\hspace{1em} & $\alpha_{3}$ & -0.5 & -0.96 (-2.14, 0.01) &     / \\
+\hspace{1em} & $\alpha_{4}$ & -0.99 & -1.18 (-2.44, -0.06) &     / \\
+\addlinespace[0.3em]
+\multicolumn{5}{l}{\textbf{ }}\\
+\hspace{1em}Multinom. & $\delta_{11}$ & -0.08 & -0.07 (-0.27, 0.12) & -0.54 (-0.77, -0.32)\\
+\hspace{1em} & $\delta_{12}$ & 0.51 & 0.25 (-0.04, 0.53) & -0.26 (-0.6, 0.05)\\
+\hspace{1em} & $\delta_{21}$ & -0.97 & -0.91 (-1.15, -0.68) & -0.07 (-0.28, 0.14)\\
+\hspace{1em} & $\delta_{22}$ & 0.84 & 0.39 (0.09, 0.71) & 0.24 (-0.04, 0.5)\\
+\addlinespace[0.3em]
+\multicolumn{5}{l}{\textbf{ }}\\
+\hspace{1em}Clustering & $\pi_l$ & 0.38 & 0.38 (0.38, 0.38) & 0.38 (0.13, 0.41)\\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 Amputation occurs according to the missing at random (MAR) mechanism, whereby the occurence of missing data depends only on the values of the observed data (Gelman and Hill 2006).
-Additionally, we perform data amputation to simulatute intermittent missingness, as is consistent with the Nurture data.
+Additionally, we perform data amputation to simulateintermittent missingness, as is consistent with the Nurture data.
 To this end, we specify that each of the $4$ repeated measures for a given individual are equally likely to be missing.
-To further approximate the Nurture data, we set the frequency of missing data at each measurement occaision to be 30\%.
+To further approximate the Nurture data, we set the frequency of missing data at each measurement occasion to be 30\%.
 We fit our proposed MSN finite mixture model to the amputed data with an implementation of both our proposed online conditional imputation step and standard Bayesian multiple imputation as defined by Gelman et al.
@@
 
+\bibitem{ } Laurent, S. \texttt{matrixsampling}: Simulations of Matrix Variate Distributions. R package version 1.1.0.
+https://CRAN.R-project.org/package=matrixsampling.
+
 \bibitem{ } Lee SX, McLachlan GJ. Model-based clustering and classification with non-normal mixture distributions. Statistical Methods \& Applications. 2013 Nov 1;22(4):427-54.
@@
 
+\bibitem{ } Makalic E, Schmidt DF. High-dimensional Bayesian regularised regression with the BayesReg package. \textit{arXiv preprint arXiv:1611.06649}. 2016 Nov 21.
+
 \bibitem{ } Melnykov V, Maitra R. Finite mixture models and model-based clustering. \textit{Statistics Surveys}. 2010;4:80-116.
@@
 \bibitem{ } Polson NG, Scott JG, Windle J. Bayesian inference for logistic models using P\'olya –- Gamma latent variables. Journal of the American statistical Association. 2013 Dec 1;108(504):1339-49.
+
+\bibitem{ } R Core Team. R: A language and environment for statistical computing. \textit{R Foundation for Statistical Computing}. 2019. https://www.R-project.org/.
 

### 2019-07-09 manuscript/supplement_old_07-09-2019.tex
sim 1

--- 2019-06-20 manuscript/manuscript.tex
@@
-%  ALWAYS USE THE referee OPTION WITH PAPERS SUBMITTED TO BIOMETRICS!!!
-%  You can see what your paper would look like typeset by removing
-%  the referee option.  Because the typeset version will be in two
-%  columns, however, some of your equations may be too long. DO NOT
-%  use the \longequation option discussed in the user guide!!!  This option
-%  is reserved ONLY for equations that are impossible to split across 
-%  multiple lines; e.g., a very wide matrix.  Instead, type your equations 
-%  so that they stay in one column and are split across several lines, 
-%  as are almost all equations in the journal.  Use a recent version of the
-%  journal as a guide. 
-%  
-\documentclass[useAMS,referee]{biom}
+\documentclass{article}
 \usepackage{amsmath}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
+\usepackage{pdflscape}
 \usepackage{graphicx}
@@
 \usepackage{breqn}
-\usepackage{algorithm}
 \usepackage{algcompatible}
-\usepackage{algpseudocode}
 \usepackage{float}
-\usepackage{dsfont}
-
-\let\origfigure\figure
-\let\endorigfigure\endfigure
-\renewenvironment{figure}[1][2] {
-    \expandafter\origfigure\expandafter[H]
-} {
-    \endorigfigure
-}
+\usepackage{dutchcal}
+\usepackage{amsfonts}
 
@@
 
-%  Here, place your title and author information.  Note that in 
-%  use of the \author command, you create your own footnotes.  Follow
-%  the examples below in creating your author and affiliation information.
-%  Also consult a recent issue of the journal for examples of formatting.
-
-\title[]{Bayesian multivariate skew-normal finite mixture model for analysis of infant development trajectories}
-
-%  Here are examples of different configurations of author/affiliation
-%  displays.  According to the Biometrics style, in some instances,
-%  the convention is to have superscript *, **, etc footnotes to indicate 
-%  which of multiple email addresses belong to which author.  In this case,
-%  use the \email{ } command to produce the emails in the display.
-
-%  In other cases, such as a single author or two authors from 
-%  different institutions, there should be no footnoting.  Here, use
-%  the \emailx{ } command instead.
-
-%  The examples below corrspond to almost every possible configuration
-%  of authors and may be used as a guide.  For other configurations, consult
-%  a recent issue of the the journal.
-
-%  Single author -- USE \emailx{ } here so that no asterisk footnoting
-%  for the email address will be produced.
-
-%\author{John Author\emailx{email@address.edu} \\
-%Department of Statistics, University of Warwick, Coventry CV4 7AL, U.K.}
-
-%  Two authors from the same institution, with both emails -- use
-%  \email{ } here to produce the asterisk footnoting for each email address
-
-%\author{John Author$^{*}$\email{author@address.edu} and
-%Kathy Authoress$^{**}$\email{email2@address.edu} \\
-%Department of Statistics, University of Warwick, Coventry CV4 7AL, U.K.}
-
-%  Exactly two authors from different institutions, with both emails  
-%  USE \emailx{ } here so that no asterisk footnoting for the email address
-%  is produced.
-
-\author
-{Carter Allen\emailx{allecart@musc.edu} \\
-Department of Public Health Sciences, Medical University of South Carolina, Charleston, SC, U.S.A
-\and
-Brian Neelon, PhD \\
-Department of Public Health Sciences, Medical University of South Carolina, Charleston, SC, U.S.A
-\and
-Sara Benjamin-Neelon, PhD, MPH, RD \\
-Department of Health, Behavior and Society, Johns Hopkins University, Baltimore, MD, U.S.A}
-
-%  Three or more authors from same institution with all emails displayed
-%  and footnoted using asterisks -- use \email{ }
-
-%\author{John Author$^*$\email{author@address.edu}, 
-%Jane Author$^{**}$\email{jane@address.edu}, and 
-%Dick Author$^{***}$\email{dick@address.edu} \\
-%Department of Statistics, University of Warwick, Coventry CV4 7AL, U.K}
-
-%  Three or more authors from same institution with one corresponding email
-%  displayed
-
-%\author{John Author$^*$\email{author@address.edu}, 
-%Jane Author, and Dick Author \\
-%Department of Statistics, University of Warwick, Coventry CV4 7AL, U.K}
-
-%  Three or more authors, with at least two different institutions,
-%  more than one email displayed
-
-%\author{John Author$^{1,*}$\email{author@address.edu}, 
-%Kathy Author$^{2,**}$\email{anotherauthor@address.edu}, and 
-%Wilma Flinstone$^{3,***}$\email{wilma@bedrock.edu} \\
-%$^{1}$Department of Statistics, University of Warwick, Coventry CV4 7AL, U.K \\
-%$^{2}$Department of Biostatistics, University of North Carolina at 
-%Chapel Hill, Chapel Hill, North Carolina, U.S.A. \\
-%$^{3}$Department of Geology, University of Bedrock, Bedrock, Kansas, U.S.A.}
-
-%  Three or more authors with at least two different institutions and only
-%  one email displayed
-
-%\author{John Author$^{1,*}$\email{author@address.edu}, 
-%Wilma Flinstone$^{2}$, and Barney Rubble$^{2}$ \\
-%$^{1}$Department of Statistics, University of Warwick, Coventry CV4 7AL, U.K \\
-%$^{2}$Department of Geology, University of Bedrock, Bedrock, Kansas, U.S.A.}
-
 \begin{document}
 
-%  This will produce the submission and review information that appears
-%  right after the reference section.  Of course, it will be unknown when
-%  you submit your paper, so you can either leave this out or put in 
-%  sample dates (these will have no effect on the fate of your paper in the
-%  review process!)
+\section*{Supplementary Materials}
 
-\date{{\it Received October} 2007. {\it Revised February} 2008.  {\it
-Accepted March} 2008.}
+\subsection*{Derivations of Full Conditional Distributions}
 
-%  These options will count the number of pages and provide volume
-%  and date information in the upper left hand corner of the top of the 
-%  first page as in published papers.  The \pagerange command will only
-%  work if you place the command \label{firstpage} near the beginning
-%  of the document and \label{lastpage} at the end of the document, as we
-%  have done in this template.
+\subsubsection*{Multivariate Skew-Normal Regression}
 
-%  Again, putting a volume number and date is for your own amusement and
-%  has no bearing on what actually happens to your paper!
+Without loss of generality, we derive the full conditional distributions for the multivariate skew-normal regression model component under the assumption that all observations belong to a single cluster.
+To make the extension to the case where more than one cluster is specified, simply apply these distributional forms to cluster specific parameters and data.
+Finally, we assume for the moment that we have complete data for all outcomes for each subject.
+We extend consider the case of missing data in section (INSERT SECTION).
 
-\pagerange{\pageref{firstpage}--\pageref{lastpage}} 
-\volume{64}
-\pubyear{2008}
-\artmonth{December}
+The multivariate skew-normal regression model can be written as follows in matrix form. 
+$$\mathbf{Y} = \mathbf{X}\mathbf{B} + \mathbf{t} \boldsymbol\psi^T + \mathbf{E} = \mathbf{X}^* \mathbf{B}^* + \mathbf{E}$$
+The matrix $\mathbf{Y}$ is of dimension $n \times J$.
+For convenience, we define $\mathbf{X}^*$ as a $n \times (P + 1)$ matrix constructed by column binding $\mathbf{t}$ to $\mathbf{X}$, and $\mathbf{B}^*$ as a $(P+1) \times J$ matrix constructed by row binding $\boldsymbol\psi^T$ to $\mathbf{B}$.
+We assume that $t_i \stackrel{iid}{\sim}T_{[0,\infty})(0,1)$ and that $\mathbf{E}$ is made of row vectors $\boldsymbol\epsilon_i = (\epsilon_{i1},...,\epsilon_{iJ})$ for $i = 1,...,n$, where $ \boldsymbol\epsilon_i \stackrel{iid}{\sim} N_J(0, \boldsymbol\Sigma)$.
 
-%  The \doi command is where the DOI for your paper would be placed should it
-%  be published.  Again, if you make one up and stick it here, it means 
-%  nothing!
+The conditional likelihood for this model is given below. 
+$$p(\mathbf{Y}|\mathbf{X}^*,\mathbf{B}^*,\boldsymbol\Sigma) \propto |\boldsymbol\Sigma|^{-n/2}\exp \left \{ -\frac{1}{2} \tr(\mathbf{Y} - \mathbf{X}^* \mathbf{B}^*)^T(\mathbf{Y} - \mathbf{X}^* \mathbf{B}^*)\boldsymbol\Sigma^{-1} \right \}$$
 
-\doi{10.1111/j.1541-0420.2005.00454.x}
+We choose conjugate priors for $\mathbf{B}^*$ and $\boldsymbol\Sigma$ as follows. 
+$$\boldsymbol\Sigma \sim \text{inverse-Wishart}(\mathbf{V}_0,\nu_0)$$
+$$\mathbf{B}^*|\boldsymbol\Sigma \sim MatNorm_{(m+1), p}(\mathbf{B}_0^*,\mathbf{L}_0^{-1},\boldsymbol\Sigma)$$
 
-%  This label and the label ``lastpage'' are used by the \pagerange
-%  command above to give the page range for the article.  You may have 
-%  to process the document twice to get this to match up with what you 
-%  expect.  When using the referee option, this will not count the pages
-%  with tables and figures.
+We now derive the joint posterior distribution of the parameters $\mathbf{B}^*$ and $\boldsymbol\Sigma$.
+\begin{align*} 
+p(\mathbf{B}^*,\boldsymbol\Sigma|\mathbf{X}^*,\mathbf{Y}) & \propto  p(\mathbf{Y}|\mathbf{X}^*,\mathbf{B}^*,\boldsymbol\Sigma)p(\mathbf{B}^*|\boldsymbol\Sigma)p(\boldsymbol\Sigma)\\
+&  \propto |\boldsymbol\Sigma|^{-n/2}\exp \left \{ -\frac{1}{2} \tr \left [(\mathbf{Y} - \mathbf{X^*}\mathbf{B}^*)^T(\mathbf{Y} - \mathbf{X^*}\mathbf{B}^*)\boldsymbol\Sigma^{-1} \right ]\right \} \\
+& \times |\boldsymbol\Sigma|^{-(P+1)/2} \exp \left \{ -\frac{1}{2} \tr \left [(\mathbf{B}^* - \mathbf{B}^*_0)^T \mathbf{L}_0(\mathbf{B}^* - \mathbf{B}^*_0)\boldsymbol\Sigma^{-1} \right ]\right \} \\
+& \times |\boldsymbol\Sigma|^{(\nu_0 + J + 1)/2} \exp \left \{ -\frac{1}{2} \tr (\mathbf{V}_0 \boldsymbol\Sigma^{-1}) \right \}
+\end{align*}
 
-\label{firstpage}
+\subsubsection*{Multinomial Logit Regression}
 
-%  put the summary for your paper here
+\subsubsection*{Multivariate Normal Conditional Imputation}
 
-\begin{abstract}
-In studies of infant motor development, a crucial research goal is to identify latent clusters of infants that experience delayed development, as this is a known risk factor for adverse outcomes later in life.
-However, there are a number of statistical challenges in modeling infant development: the data are typically skewed, exhibit intermittent missingness, and are highly correlated across the repeated measurements collected during infancy.
-Using data from the Nurture study, a cohort of over 600 mother-infant pairs followed from pregnancy to 12 months postpartum, we develop a flexible Bayesian latent class model for the analysis infant motor development.
-Our model has a number of attractive features.
-First, we adopt the multivariate skew normal distribution with cluster-specific parameters that accommodate the inherent correlation and skewness in the data.
-Second, we model the cluster membership probabilities using a novel Pólya-Gamma data-augmentation scheme, thereby improving predictions of the cluster membership allocations.
-Lastly, we impute missing responses under missing at random assumption by drawing from appropriate conditional skew normal distributions.
-Bayesian inference is achieved through straightforward Gibbs sampling, and can be carried out in available software such as R.
-Through simulation studies, we show that the proposed model yields improved inferences over models that ignore skewness.
-In addition, our imputation method yields improvements compared to conventional missing data methods, including multiple imputation and complete or available case analysis.
-When applied to Nurture data, we identified two distinct development clusters: one characterized by delayed “U-shaped” development and a higher percentage of male infants and another characterized by more steady development and a lower percentage of males.
-The clusters also differed in terms of key demographic variables, such as infant race and maternal pre-pregnancy body mass index.
-These findings can aid investigators in targeting interventions during this critical early-life developmental window.
-\end{abstract}
+The multivariate normal conditional imputation derivations are given for a single cluster without loss of generality.
+In practice, the data and parameters in this section would be replaced by cluster specific estimates in the case of clustering.
 
-%  Please place your key words in alphabetical order, separated
-%  by semicolons, with the first letter of the first word capitalized,
-%  and a period at the end of the list.
-%
+For a given observation vector $\mathbf{y} \sim N_J(\boldsymbol\mu, \boldsymbol\Sigma)$, we allow for missingness in at most $J - 1$ of the multivariate outcomes through the use of a conditional imputation step embedded within our Gibbs sampler.
+Suppose $\mathbf{y}$ contains $q$ missing observations and can be partitioned into two vectors $\mathbf{y_{1}}$ and $\mathbf{y_{2}}$ such that $\mathbf{y_{1}}$ is a $q \times 1$ vector of missing observations and $\mathbf{y_{2}}$ is a $(J-q) \times 1$ vector of complete observations.
+Similarly, partition $\boldsymbol\mu$ and $\boldsymbol\Sigma$ as follows.
+$$\boldsymbol\mu = \begin{bmatrix} \boldsymbol\mu_1 \\ \boldsymbol\mu_2 \end{bmatrix} \ \ \ \ \ \boldsymbol\Sigma = \begin{bmatrix} \boldsymbol\Sigma_{11} & \boldsymbol\Sigma_{12} \\ \boldsymbol\Sigma_{21} & \boldsymbol\Sigma_{22} \end{bmatrix}$$
 
-\begin{keywords}
-A key word; But another key word; Still another key word; Yet another key word.
-\end{keywords}
+We will use these quantities to derive the conditional distribution $f(\mathbf{y_1}|\mathbf{y_2},\boldsymbol\mu,\boldsymbol\Sigma)$. 
+\begin{align*} 
+f(\mathbf{y_1}|\mathbf{y_2},\boldsymbol\mu,\boldsymbol\Sigma) & \propto  f(\mathbf{y_1},\mathbf{y_2}|\boldsymbol\mu,\boldsymbol\Sigma) \\
+& \propto \exp \left \{ -\frac{1}{2}(\mathbf{y} - \boldsymbol\mu)^T \boldsymbol\Sigma^{-1} (\mathbf{y} - \boldsymbol\mu) \right \} \\
+& = \exp \left \{ -\frac{1}{2}\begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix}^T \boldsymbol\Sigma^{-1} \begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix} \right \} \\
+& =  \exp \left \{ -\frac{1}{2}\begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix}^T \begin{bmatrix} \boldsymbol\Sigma_{11} & \boldsymbol\Sigma_{12} \\ \boldsymbol\Sigma_{21} & \boldsymbol\Sigma_{22} \end{bmatrix}^{-1} \begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix} \right \} \\
+& = \exp \left \{ -\frac{1}{2}\begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix}^T \begin{bmatrix} \boldsymbol\Sigma_{11}^* & \boldsymbol\Sigma_{12}^* \\ \boldsymbol\Sigma_{21}^* & \boldsymbol\Sigma_{22}^* \end{bmatrix} \begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix} \right \} \\
+& = \exp \left \{ -\frac{1}{2} \left[ (\mathbf{y}_1 - \boldsymbol\mu_{cond})^T \boldsymbol\Sigma_{cond}^{-1}(\mathbf{y}_1 - \boldsymbol\mu_{cond})\right]\right \} 
+\end{align*}
+$$\Rightarrow \mathbf{y_1}|\mathbf{y_2},\boldsymbol\mu,\boldsymbol\Sigma \sim N_q(\boldsymbol\mu_{cond},\boldsymbol\Sigma_{cond})$$
+$$\boldsymbol\mu_{cond} = \boldsymbol\mu_1 + \boldsymbol\Sigma_{12}\boldsymbol\Sigma_{22}^{-1}(\mathbf{y}_2 - \boldsymbol\mu_2), \ \ \ \ \ \ \boldsymbol\Sigma_{cond} = \boldsymbol\Sigma_{11} - \boldsymbol\Sigma_{12}\boldsymbol\Sigma_{22}^{-1}\boldsymbol\Sigma_{21}$$
 
-%  As usual, the \maketitle command creates the title and author/affiliations
-%  display
-
-\maketitle
-
-\setcounter{tocdepth}{3}
-\tableofcontents
+The block-wise inversion formula was used to invert $\boldsymbol\Sigma$ according to the following reparameterizations.
+\begin{align*}
+\boldsymbol\Sigma_{11}^* & = \boldsymbol\Sigma_{11}^{-1} + \boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12}(\boldsymbol\Sigma_{22} - \boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12})^{-1}\boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\\
+\boldsymbol\Sigma_{12}^* & = -\boldsymbol\Sigma_{11} \boldsymbol\Sigma_{12}(\boldsymbol\Sigma_{22}-\boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12})^{-1}\\
+\boldsymbol\Sigma_{21}^* & = -(\boldsymbol\Sigma_{22} - \boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12})^{-1}\boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\\
+\boldsymbol\Sigma_{22}^* & = (\boldsymbol\Sigma_{22} - \boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12})^{-1}     
+\end{align*}
 
@@
 
-%  If you are using the referee option, a new page, numbered page 1, will
-%  start after the summary and keywords.  The page numbers thus count the
-%  number of pages of your manuscript in the preferred submission style.
-%  Remember, ``Normally, regular papers exceeding 25 pages and Reader Reaction 
-%  papers exceeding 12 pages in (the preferred style) will be returned to 
-%  the authors without review. The page limit includes acknowledgements, 
-%  references, and appendices, but not tables and figures. The page count does 
-%  not include the title page and abstract. A maximum of six (6) tables or 
-%  figures combined is often required.''
-
-%  You may now place the substance of your manuscript here.  Please use
-%  the \section, \subsection, etc commands as described in the user guide.
-%  Please use \label and \ref commands to cross-reference sections, equations,
-%  tables, figures, etc.
-%
-%  Please DO NOT attempt to reformat the style of equation numbering!
-%  For that matter, please do not attempt to redefine anything!
-
-\section{Introduction}
-\label{s:intro}
-
-\subsection{Infant Development Clustering}
-
-Heterogeneity of treatment effects (HTE) (Lanza and Rhoades, 2013).
-
-\subsection{Existing Approaches}
-
-Mixtures of multivariate non-symmetric distributions such as the multivariate skew-normal (MSN) distribution allow for the nuances of the marginal density to be captured with a more parsimonious set of mixture components.
-Mixtures of MSN distributions have been dealt with previously in a Bayesian context (Fr\"{u}hwirth-Schnatter \& Pyne, 2010), however in these models, focus lies primary on marginal density estimation and inference on the mixture components (i.e.
-clusters) is not discussed.
-More recently, the mixtures of skew-$t$ factor analysis (MSTFA) model has been proposed for settings in which cluster-specific inference is of primary interest (Lin \textit{et al.} 2018).
-However, an important feature not included in the MSTFA is the ability to explain individual-level cluster membership as a function of covariates of interest.
-Additionally, parameter estimation proposed by Lin et al.
-for the MSTFA relies on a prohibitively complex EM algorithm and does not enjoy the inferential benefits of a Bayesian approach, namely the ability to incorporate prior information into a model and make posterior probability statements.
-Our proposed model improves on these previous works by estimating parameters in a Bayesian framework as well as including the ability to fit a multinomial logit regression to cluster membership probabilities using a novel application of data augmentation with the P\'olya Gamma distribution.
-
-\textbf{\textit{Put lit review of Bayesian PG multinomial logistic regression here}}
-
-A ubiquitous feature of repeated measures studies is loss of data due to intermittent missingness and attrition.
-In the Bayesian setting, the standard approach to dealing with missing data is to perform multiple imputation, whereby $m$ imputed data sets are generated from a specified imputation model.
-After $m$ complete data sets are obtained, parameter estimates are combined across each data set to produce a final set of parameter estimates (Gelman \textit{et al.} 2013).
-This approach is not only computationally burdensome, requiring storage and analysis of an $m \times n_{rows} \times n_{cols}$ data array in addition to multiplication of total model run time by a factor of $m$, but it has been shown to produce unreliable inferences (Zhou and Reiter, 2010).
-We instead include an ``online" imputation imputation step in our Gibbs sampling procedure, whereby missing outcomes are updated at each iteration.
-This approach greatly increases the number of opportunities for exploration of the missing data parameter space.
-
-\newpage
-
-\section{Nurture Study}
-\label{s:nurt}
-\subsection{Baseline Demographics and Description of Variables}
-
-\subsection{Statistical Challenges}
-\subsubsection{Skewness of Bayley score residuals}
-\subsubsection{Attrition and Intermittent Missingness}
-
-\newpage
-
-\section{Model}
-\label{s:model}
-
-\subsection{Generic Multivariate Skew Normal Mixture}
-
-Let $\mathbf{y}_{i}$ be the $J \times 1$ observation vector for subject $i$ such that $y_{ij}$ is the observation for subject $i$ at timepoint $j$.
-We assume for now that $\mathbf{y}_{i}$ is fully observed.
-Later, we relax this assumption by allowing for missingness in the components of $\mathbf{y}_{i}$.
-For the analysis of the Nurture data, we propose the following MSN mixture
-\begin{equation}
-f(\mathbf{y}_i) = \sum_{k = 1}^{K} \pi_{ik} f(\mathbf{y}_i|\boldsymbol\theta_k),
-\end{equation}
-where $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$ ($k = 1,...,K$).
-For now we assume that $K$ is fixed, but we discuss model selection strategies for choosing the optimal value of $K$ in Section 4.
-The probability that subject $i$ ($i = 1,...,n$) belongs to cluster $k$ is denoted $\pi_{ik}$.
-Note that $\pi_{ik}$ is indexed by $i$, as we allow the probability of class membership to vary across subjects.
-However, we assume a that class membership is fixed throughout the study period, since our focus is to cluster individuals based on their overall developmental patterns over the course of the study.
-In Section 6, we discuss extensions to allow for class membership to vary over time.
-
-In each cluster, we assume that the $J\times 1$ outcome vector for subject $i$ follows an MSN distribution.
-Let $z_i \in \{1,...,K\}$ denote a latent cluster membership of subject $i$.
-Conditional on $z_i = k$, we model $\mathbf{y}_{i}$ as
-\begin{equation}
-\mathbf{y}_{i}|z_i=k \sim MSN_J(\boldsymbol\zeta_k,\boldsymbol\alpha_k,\boldsymbol\Omega_k),
-\end{equation}
-where $\boldsymbol\zeta_k$ is a $J \times 1$ vector of cluster-specific location parameters, $\boldsymbol\alpha_k$ is a $J \times 1$ vector of cluster-specific skewness parameters, and $\boldsymbol\Omega_k$ is a $J \times J$ cluster-specific scale matrix that accounts for associations between the $J$ responses.
-The vector $\boldsymbol\alpha_k$ has components $\alpha_{jk}$, $j = 1,...,J$, that control the skewness of outcome $j$ in cluster $k$.
-When $\boldsymbol\alpha_k = \mathbf{0}$, the MSN distribution reduces to the multivariate normal distribution $N_J(\boldsymbol\zeta_k,\boldsymbol\Sigma_k)$, where $\boldsymbol\Sigma_J$ is the $J \times J$ covariance matrix capturing dependence among the $J$ outcomes.
-
-\subsection{Multivariate Skew Normal Stochastic Representation}
-
-We model the effect of covariates on longitudinal motor development outcomes through the use of a MSN regression model.
-We adopt a standard stochastic representation of a MSN random variable by conditioning on a subject-specific latent truncated normal random effect, denoted $t_i$ (Fr\"{u}hwirth-Schnatter and Pyne 2010).
-Conditional on membership in cluster $k$, the stochastic representation for the $j^{th}$ component of the multivariate skew normal distribution for subject $i$ is
-\begin{equation}
-y_{ij} = \zeta_{ijk} + t_i \psi_{jk} + \sqrt{1-\psi^2_{jk}}\epsilon_{ijk}.
-\end{equation}
-where $y_{ij}$ denotes the $j^{th}$ observation for subject $i$, $\zeta_{ijk}$ denotes the location parameter for the $j^{th}$ observation for subject $i$ in cluster $k$, $t_i \stackrel{iid}{\sim} N_{[0,\infty)}(0,1)$ is a subject-specific truncated normal random effect, $\psi_{jk}$ controls the skewness of the $j^{th}$ outcome in cluster $k$, and $(\epsilon_{i1k},...,\epsilon_{iJk}) = \boldsymbol\epsilon_{ik} \sim N_J(0,\boldsymbol\Sigma_k)$ is a multivariate normal error term.
-Combining all observations for subject $i$ into $\mathbf{y}_i$, and conditional on $t_i$ and $z_i = k$, $\mathbf{y}_i$ follows a multivariate normal distribution with conditional mean $\mathbf{x}_i^T \boldsymbol\beta_k + t_i \boldsymbol\psi_k$ and conditional covariance $\boldsymbol\Sigma_k$.
-In the multivariate case, $\boldsymbol\psi_k$ is a $J \times 1$ vector containing cluster-specific skewness parameters.
-Marginally (after integrating over $t_i$), $\mathbf{y}_i$ is distributed $MSN_J(\boldsymbol\zeta_{ik}, \boldsymbol\alpha_k, \boldsymbol\Omega_k)$, where
-$$\boldsymbol\zeta_{ik} = \mathbf{x}_i^T \boldsymbol\beta_k,$$
-$$\boldsymbol\alpha_k = \frac{1}{\sqrt{1 - \boldsymbol\psi_k^T \boldsymbol\psi_k}} \boldsymbol\Omega^{-1}_k\boldsymbol\psi_k,$$
-$$\boldsymbol\Omega_k = \boldsymbol\Psi_k \boldsymbol\Sigma_k \boldsymbol\Psi_k + \boldsymbol\psi_k \boldsymbol\psi_k^T,$$
-where $\boldsymbol\Psi_k = \text{Diag}\left ( \sqrt{1 - \psi^2_{k1}},...,\sqrt{1 - \psi^2_{kJ}} \right )$.
-To allow for time varying effects in our model, we structure $\mathbf{X}_i$ as a $J \times pJ$ vector of covariate values for subject $i$, and $\boldsymbol\beta_k$ as a $pJ \times 1$ vector of fixed effects coefficients for cluster $k$ (think about how to explain this better).
-
-\subsection{Representation as Matrix Skew Normal}
-
-We note that the multivariate skew normal density can be expressed more compactly in terms of the matrix skew normal (MatSN) density (Chen and Gupta 2005).
-In particular, we define $\mathbf{Y}_{n_k \times J}$ as the response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n \mathds{1}_{(z_i = k)}$ is the total number of observations belonging to cluster $k$ ($k = 1,...,K$).
-$$\mathbf{Y}_{n_k \times J} \sim MatSN_{n_k \times J}(\mathbf{M},\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Sigma_k)$$
-$$\text{vec}(\mathbf{Y}) = \mathbf{y}_{n_kJ \times 1} = (\mathbf{y}^T_1,...,\mathbf{y}^T_{n_k})$$
-$$\text{vec}(\mathbf{M}) = \boldsymbol\mu_{n_k J \times 1} = (\boldsymbol\mu_1^T,...,\boldsymbol\mu_{n_k}^T)$$
-
-where $\boldsymbol\mu_i = \mathbf{x}_i \boldsymbol\beta_k$ and $\boldsymbol\alpha_k = (\alpha_{k1},...,\alpha_{kJ})$.
-This implies that the $i^{th}$ row of $\mathbf{Y}$ is $\mathbf{y}_i \sim MSN_J(\boldsymbol\mu_i, \boldsymbol\alpha_k, \boldsymbol\Sigma_k) = {MSN}_J(\mathbf{x}_i \boldsymbol\beta_k, \boldsymbol\alpha_k, \boldsymbol\Sigma_k)$.
-Using the stochastic representation presented in Section 3.2, the conditional distribution of $\mathbf{y}_i | z_i, t_i$ is $N_J (\mathbf{x}_i \boldsymbol\beta_k + \boldsymbol\psi_k t_i,\boldsymbol\Sigma_k)$.
-This implies that all responses for cluster $k$ conditioned on $\mathbf{T}_{n_k \times 1}$ is distributed as
-$$\mathbf{Y}_{n_k \times J} | \mathbf{T}_k = \mathbf{t}_k \sim MatNorm_{n_k \times J}(\mathbf{M}_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k)$$
-$$\text{vec}(\mathbf{M}_k)_{n_kJ \times 1} = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$$\
-
-where $\mathbf{t}_k = (t_1,...,t_{n_k})$.
-
-\subsection{Multinomial Regression on Cluster Probabilities}
-
-A primary concern of our model is with identification of latent infant development classs.
-We accomplish this via multinomial logit regression model on cluster membership, which utilizes P\'{o}lya-Gamma data-augmentation, as described in Section 3.6 to allow for updating of all parameters using Gibbs sampling.
-The multinomial logit model is as follows for $k = 1,...,K$.
-$$P(z_i = k|\mathbf{w}_i) = \pi_{ik} = \frac{e^{\mathbf{w}_i^T \boldsymbol\delta_k}}{\sum_{k' = 1}^K e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}}}$$
-
-where $\mathbf{w}_i$ is the vector of cluster probability covariates for subject $i$, $\boldsymbol\delta_k$ contains the multinomial regression parameters for cluster $k$, and $K$ is the number of putative classs specified \textit{a priori}.
-For identifiability purposes, we fix the reference category $k = K$ and set $\boldsymbol\delta_K = \mathbf{0}$.
-Under this model, $z_i|\mathbf{w}_i \sim Multinom(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{i1},...,\pi_{1K})$.
-During MCMC estimation, the cluster labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as cluster assignments.
-
-\subsection{Conditional MVN Imputation}
-
-We allow for missingness of outcomes in the MSN mixture model by imputing missing values from their conditional multivariate normal distributions. We note that
-$$\mathbf{y}_i|z_i,\mathbf{x}_i,t_i,\boldsymbol\beta_{k},\boldsymbol\psi_{k},\boldsymbol\Sigma_{k} \sim N_J(\boldsymbol\beta_{k}^T\mathbf{x}_i  + t_i \boldsymbol\psi_{k}, \boldsymbol\Sigma_{k})$$
-
-where $z_i = k$.
-This allows us to appeal to standard conditional forms of the multivariate normal distribution to specify the distribution of missing observations.
-For subject $i$, we use $q_i$ to denote the number of the total $J$ possible repeated measurements are missing.
-We partition the full $J \times 1$ outcome vector $\mathbf{y}_i$ into the $(J - q_i) \times 1$ observed data vector $\mathbf{y}_i^{obs}$ and the $q_i \times 1$ missing data vector $\mathbf{y}_i^{miss}$.
-$Y_i = [Y^{miss}_{i_{q \times 1}} | Y^{obs}_{i_{J - q \times 1}}]^T$.
-We have
-$$\mathbf{y}_i^{miss}|\mathbf{y}_i^{obs},\mathbf{x}_i,t_i,z_i,\boldsymbol\beta_{k},\boldsymbol\psi_{k},\boldsymbol\Sigma_{k} \sim N(\boldsymbol\mu^{miss},\boldsymbol\Sigma^{miss})$$
-
-where $\boldsymbol\mu^{miss}$ and $\boldsymbol\Sigma^{miss}$ take standard forms detailed in the Appendix.
-
-An attractive feature of this imputation approach is that each missing outcome is imputed ``online", i.e.
-once per MCMC iteration.
-This provides more opportunities to explore the parameter space than multiple imputation and avoids multiplicative run-time scaling in $m$, the number of imputations.
-We demonstrate this feature using simulations detailed in Section 4.
-
-\subsection{Bayesian Inference}
-
-\subsubsection{P\'olya--Gamma Data Augmentation}
-
-Polson \textit{et al.} (2013) introduced an efficient data augmentation approach to fitting several GLMs including the multinomial logit regression model specified by Holmes and Held (2006), which specifies the full conditional distributions of the multinomial regression parameters as a function of a Bernoulli likelihood.
-$$p(\boldsymbol\delta_k|\mathbf{z},\boldsymbol\delta_{k' \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \pi_{ik}^{U_{ik}}(1-\pi_{ik})^{1-U_{ik}}$$
-
-where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, $U_{ik} = \mathds{1}_{z_i = k}$ is an indicator that subject $i$ belongs to cluster $k$, and $\pi_{ik}$ is defined as in Section 3.4.
-We can rewrite $\pi_{ik}$ as follows
-$$\pi_{ik} = P(U_{ik} = 1) = \frac{e^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}}}{1 + e^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}}} = \frac{e^{\eta_{ik}}}{1 + e^{\eta_{ik}}}$$
-
-where ${c}_{ik} = \log \sum_{k' \ne k} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}}$ and $\eta_{ik} = \mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}$.
-We note that the sum $\sum_{k' \ne k} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}}$ includes the reference category, but since we fix $\boldsymbol\delta_K = \mathbf{0}$, we have $e^{\mathbf{w}_i^T \boldsymbol\delta_K} = 1$, and hence
-$$c_{ik} = \log \sum_{k' \ne k} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}} = \log \left ( 1 + \sum_{k' \notin \{k,K \}} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}} \right )$$
-
-We can use the quantities to re-express the full conditionals for $\boldsymbol\delta_k$ as 
-$$p(\boldsymbol\delta_k|\mathbf{Z},\boldsymbol\delta_{k' \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \left (\frac{e^{\eta_{ik}}}{1 + e^{\eta_{ik}}} \right )^{U_{ik}} \left (\frac{1}{1 + e^{\eta_{ik}}} \right )^{1-U_{ik}} = p(\boldsymbol\delta_k) \prod_{i = 1}^n \frac{(e^{\eta_{ik}})^{U_{ik}}}{1 + e^{\eta_{ik}}}$$
-
-which we note is essentially a logistic regression likelihood.
-We thus apply this P\'olya--Gamma data augmentation scheme to update each $\boldsymbol\delta_k$ ($k = 1,...,K-1$) one at a time based on the binary indicators $U_{ik}$.
-
-\begin{itemize}
-
-\item Emphasize that PG data augmentation for the multinomial model results in a PG mixture of experts model, which is a computationally efficient way to model edge weights.
-
-\end{itemize}
-
-\subsubsection{Prior Choice}
-
-\subsubsection{MCMC Algorithm}
+\subsection*{Tables}
 
@@
 
-\subsubsection{Assessment of MCMC Convergence}
+\begin{landscape}\begin{table}[t]
 
-\subsubsection{Label Switching}
-
-\newpage
-
-\section{Simulation Studies}
-\label{s:sim}
-
-\subsection{Simulation to Compare to Multivariate Normal}
-
-\begin{table}[t]
-
-\caption{\label{tab:unnamed-chunk-4}Model results for simulated data with n = 1500, k = 4, p = 1, h = 3, v = 1.
-5000 iterations were run with a burn in of 1000.
-Missingness mechanism was MAR and P(miss) = 0}
+\caption{\label{tab:unnamed-chunk-5}Model results for simulated data with n = 1,000, J = 4, p = 2, K = 3, r = 2.
+1,000 iterations were run with a burn in of 100.
+Missingness mechanism was MAR and P(miss) = 0.
+Model results for the multivariate skew normal (MSN) and multivariate normal (MN) mixtures are presented.}
 \centering
-\fontsize{8}{10}\selectfont
-\begin{tabular}{llllllll}
+\fontsize{6}{8}\selectfont
+\begin{tabular}{lllllllllll}
 \toprule
-\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Cluster 1} & \multicolumn{2}{c}{Cluster 2} & \multicolumn{2}{c}{Cluster 3} \\
-\cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6} \cmidrule(l{3pt}r{3pt}){7-8}
-Model Component & Parameter & True & Est. (95\% CrI) & True & Est. (95\% CrI) & True & Est. (95\% CrI)\\
+\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Class 1} & \multicolumn{3}{c}{Class 2} & \multicolumn{3}{c}{Class 3} \\
+\cmidrule(l{3pt}r{3pt}){3-5} \cmidrule(l{3pt}r{3pt}){6-8} \cmidrule(l{3pt}r{3pt}){9-11}
+Component & Param. & True & MSN Est. (95\% CrI) & MN Est. (95\% CrI)  & True & MSN Est. (95\% CrI) & MN Est. (95\% CrI) & True & MSN Est. (95\% CrI) & MN Est. (95\% CrI)\\
 \midrule
 \addlinespace[0.3em]
-\multicolumn{8}{l}{\textbf{ }}\\
-\hspace{1em}MVSN & $\beta_{0}$ & -1.35 & -1.59 (-2.3, -0.94) & -0.18 & -0.24 (-0.53, 0.38) & 0.72 & 0.79 (0.52, 1.02)\\
-\hspace{1em}Regression & $\beta_{1}$ & -1.19 & -1.35 (-1.99, -0.84) & -0.09 & -0.33 (-0.62, 0.78) & 1.64 & 1.65 (1.33, 1.91)\\
-\hspace{1em} & $\beta_{2}$ & -1.65 & -1.81 (-2.66, -1.41) & -0.47 & -0.62 (-0.89, 0.3) & 1.44 & 1.32 (0.97, 1.6)\\
-\hspace{1em} & $\beta_{3}$ & -1.75 & -1.89 (-2.52, -1.37) & -0.22 & -0.32 (-0.63, 0.55) & 2.28 & 2.26 (1.96, 2.5)\\
+\multicolumn{11}{l}{\textbf{ }}\\
+\hspace{1em}MVSN & $\beta_{11}$ & 11 & 11.07 (10.74, 11.39) & 9.42 (8.91, 9.77) & -5 & -4.95 (-5.2, -4.68) & -4.11 (-4.33, -3.86) & -10 & -10.3 (-10.56, -10.01) & -8.37 (-8.86, -0.65)\\
+\hspace{1em}Regression & $\beta_{21}$ & 12 & 12.02 (11.87, 12.17) & 11.98 (11.77, 12.18) & -4 & -4 (-4.1, -3.89) & -3.98 (-4.09, -3.87) & -11 & -11 (-11.19, -10.82) & -10.85 (-11.1, 0.61)\\
+\hspace{1em} & $\beta_{31}$ & 13 & 13.06 (12.75, 13.36) & 11.39 (10.7, 11.78) & -3 & -2.97 (-3.25, -2.68) & -3.74 (-3.99, -3.5) & -12 & -11.9 (-12.22, -11.56) & -10.29 (-10.78, -0.14)\\
+\hspace{1em} & $\beta_{41}$ & 14 & 14.06 (13.91, 14.22) & 14.02 (13.78, 14.22) & -2 & -1.96 (-2.07, -1.86) & -1.97 (-2.08, -1.84) & -13 & -13.04 (-13.25, -12.87) & -12.89 (-13.13, 0.59)\\
+\hspace{1em} & $\beta_{12}$ & 2 & 2.11 (1.82, 2.35) & 0.42 (0.03, 0.83) & 5 & 5.16 (4.88, 5.47) & 5.83 (5.59, 6.07) & -2 & -1.86 (-2.21, -1.52) & -0.35 (-0.77, 0.16)\\
+\hspace{1em} & $\beta_{22}$ & 2 & 2.03 (1.88, 2.17) & 2.02 (1.86, 2.22) & 5 & 4.96 (4.86, 5.06) & 4.96 (4.84, 5.07) & -2 & -1.97 (-2.18, -1.79) & -1.89 (-2.11, 0.02)\\
+\hspace{1em} & $\beta_{32}$ & 2 & 2.13 (1.8, 2.43) & 0.49 (0.14, 0.86) & 5 & 5.22 (4.96, 5.49) & 4.23 (3.96, 4.5) & -2 & -1.82 (-2.14, -1.5) & -0.37 (-0.77, 0.16)\\
+\hspace{1em} & $\beta_{42}$ & 2 & 2.08 (1.93, 2.23) & 2.08 (1.92, 2.28) & 5 & 4.97 (4.86, 5.08) & 4.96 (4.83, 5.08) & -2 & -1.93 (-2.13, -1.77) & -1.84 (-2.04, 0.04)\\
 \addlinespace[0.3em]
-\multicolumn{8}{l}{\textbf{ }}\\
-\hspace{1em} & $\sigma_{11}$ & 1 & 1.02 (0.78, 1.19) & 1 & 0.98 (0.77, 1.23) & 1 & 1.06 (0.85, 1.29)\\
-\hspace{1em} & $\sigma_{12}$ & -0.32 & -0.19 (-0.33, -0.02) & 0.16 & 0.14 (-0.01, 0.4) & 0.72 & 0.82 (0.62, 1.05)\\
-\hspace{1em} & $\sigma_{13}$ & -0.65 & -0.55 (-0.68, -0.35) & 0.72 & 0.7 (0.51, 0.94) & 0.14 & 0.27 (0.1, 0.48)\\
-\hspace{1em} & $\sigma_{14}$ & -0.44 & -0.33 (-0.46, -0.13) & 0.5 & 0.48 (0.31, 0.72) & -0.01 & -0.02 (-0.16, 0.16)\\
-\hspace{1em} & $\sigma_{22}$ & 1 & 0.92 (0.72, 1.06) & 1 & 0.94 (0.72, 1.22) & 1 & 1.11 (0.87, 1.38)\\
-\hspace{1em} & $\sigma_{23}$ & 0.56 & 0.49 (0.33, 0.6) & 0.53 & 0.46 (0.29, 0.73) & -0.1 & 0.08 (-0.1, 0.28)\\
-\hspace{1em} & $\sigma_{24}$ & 0.98 & 0.9 (0.7, 1.04) & 0.24 & 0.14 (-0.03, 0.41) & 0.19 & 0.17 (0.01, 0.37)\\
-\hspace{1em} & $\sigma_{33}$ & 1 & 0.9 (0.66, 1.04) & 1 & 0.92 (0.72, 1.19) & 1 & 1.26 (1.03, 1.52)\\
-\hspace{1em} & $\sigma_{34}$ & 0.56 & 0.51 (0.35, 0.62) & 0.86 & 0.79 (0.58, 1.05) & -0.65 & -0.63 (-0.78, -0.45)\\
-\hspace{1em} & $\sigma_{44}$ & 1 & 0.93 (0.72, 1.07) & 1 & 0.93 (0.7, 1.2) & 1 & 1.11 (0.87, 1.36)\\
+\multicolumn{11}{l}{\textbf{ }}\\
+\hspace{1em} & $\Omega_{11}$ & 5 & 4.99 (3.84, 6.52) & 3.08 (2.69, 4.16) & 2 & 1.95 (1.52, 2.52) & 1.38 (1.2, 1.59) & 5 & 6.27 (4.84, 7.88) & 3.49 (2.89, 186.69)\\
+\hspace{1em} & $\Omega_{12}$ & 4.5 & 4.55 (3.49, 5.88) & 2.76 (2.39, 3.55) & -0.5 & -0.51 (-0.83, -0.22) & 0.2 (0.05, 0.36) & 4.5 & 4.95 (3.78, 6.43) & 3.02 (2.48, 215.36)\\
+\hspace{1em} & $\Omega_{13}$ & 4.25 & 4.53 (3.48, 5.85) & 1.9 (1.58, 2.46) & 1.25 & 1.08 (0.74, 1.53) & 0.38 (0.24, 0.54) & 4.25 & 4.7 (3.56, 6.04) & 1.86 (1.42, 8.74)\\
+\hspace{1em} & $\Omega_{14}$ & 4.12 & 4.33 (3.31, 5.57) & 1.78 (1.46, 2.24) & -0.88 & -1.08 (-1.5, -0.75) & -0.41 (-0.58, -0.26) & 4.12 & 4.51 (3.43, 5.69) & 1.72 (1.29, 12.25)\\
+\hspace{1em} & $\Omega_{22}$ & 5 & 5.09 (3.98, 6.55) & 3.43 (3.01, 4.23) & 2 & 1.99 (1.53, 2.51) & 1.43 (1.22, 1.67) & 5 & 4.87 (3.75, 6.33) & 3.71 (3.09, 253.61)\\
+\hspace{1em} & $\Omega_{23}$ & 4.5 & 4.77 (3.69, 6) & 2.13 (1.78, 2.7) & -0.5 & -0.46 (-0.75, -0.18) & 0.05 (-0.1, 0.2) & 4.5 & 4.29 (3.23, 5.61) & 2.02 (1.55, 13.33)\\
+\hspace{1em} & $\Omega_{24}$ & 4.25 & 4.38 (3.39, 5.68) & 1.86 (1.53, 2.39) & 1.25 & 1.45 (1.04, 1.89) & 0.64 (0.48, 0.83) & 4.25 & 4.01 (3.04, 5.19) & 1.78 (1.35, 16.8)\\
+\hspace{1em} & $\Omega_{33}$ & 5 & 5.47 (4.3, 6.67) & 2.55 (2.23, 3.2) & 2 & 1.73 (1.36, 2.3) & 1.41 (1.22, 1.64) & 5 & 4.79 (3.6, 6.24) & 2.44 (1.99, 5.81)\\
+\hspace{1em} & $\Omega_{34}$ & 4.5 & 4.8 (3.7, 6.06) & 1.99 (1.69, 2.68) & -0.5 & -0.68 (-1.07, -0.37) & 0.14 (-0.01, 0.3) & 4.5 & 4.25 (3.2, 5.45) & 1.94 (1.55, 5.3)\\
+\hspace{1em} & $\Omega_{44}$ & 5 & 5.17 (3.96, 6.69) & 2.45 (2.14, 3.79) & 2 & 2.34 (1.8, 2.97) & 1.57 (1.37, 1.83) & 5 & 4.64 (3.54, 5.92) & 2.37 (1.93, 5.68)\\
 \addlinespace[0.3em]
-\multicolumn{8}{l}{\textbf{ }}\\
-\hspace{1em} & $\psi_{1}$ & -0.33 & -0.02 (-0.84, 0.87) & 0.67 & 0.69 (-0.09, 1.01) & -1 & -0.98 (-1.25, -0.67)\\
-\hspace{1em} & $\psi_{2}$ & -0.33 & -0.16 (-0.8, 0.62) & 0.67 & 0.81 (-0.6, 1.14) & -1 & -0.98 (-1.28, -0.6)\\
-\hspace{1em} & $\psi_{3}$ & -0.33 & -0.15 (-0.65, 0.89) & 0.67 & 0.72 (-0.35, 1.02) & -1 & -0.8 (-1.15, -0.36)\\
-\hspace{1em} & $\psi_{4}$ & -0.33 & -0.18 (-0.84, 0.59) & 0.67 & 0.71 (-0.32, 1.06) & -1 & -1 (-1.32, -0.62)\\
+\multicolumn{11}{l}{\textbf{ }}\\
+\hspace{1em} & $\alpha_{1}$ & -0.99 & -0.81 (-2.12, 0.05) & 0 (0, 0) & 0.85 & 1.05 (0.37, 1.91) & 0 (0, 0) & 0.99 & 2.82 (1.21, 4.3) & 0 (0, 0)\\
+\hspace{1em} & $\alpha_{2}$ & -0.5 & -0.22 (-1.3, 0.75) & 0 (0, 0) & -1.28 & -1.29 (-2.22, -0.5) & 0 (0, 0) & 0.5 & -0.07 (-1.14, 1.13) & 0 (0, 0)\\
+\hspace{1em} & $\alpha_{3}$ & -0.5 & -0.96 (-2.14, 0.01) & 0 (0, 0) & 1.28 & 1.16 (0.47, 2.06) & 0 (0, 0) & 0.5 & 0.08 (-0.99, 1.46) & 0 (0, 0)\\
+\hspace{1em} & $\alpha_{4}$ & -0.99 & -1.18 (-2.44, -0.06) & 0 (0, 0) & -0.85 & -0.91 (-1.76, -0.16) & 0 (0, 0) & 0.99 & 1.1 (0.07, 2.33) & 0 (0, 0)\\
 \addlinespace[0.3em]
-\multicolumn{8}{l}{\textbf{ }}\\
-\hspace{1em}Multinom. & $\delta_{11}$ & -0.84 & -0.78 (-0.96, -0.59) & -0.84 & -0.78 (-0.96, -0.59) & -0.84 & -0.78 (-0.96, -0.59)\\
-\hspace{1em} & $\delta_{12}$ & -0.24 & -0.26 (-0.42, -0.1) & -0.24 & -0.26 (-0.42, -0.1) & -0.24 & -0.26 (-0.42, -0.1)\\
+\multicolumn{11}{l}{\textbf{ }}\\
+\hspace{1em}Multinom. & $\delta_{11}$ & -0.08 & -0.07 (-0.27, 0.12) & -0.54 (-0.77, -0.32) & -0.08 & -0.07 (-0.27, 0.12) & -0.54 (-0.77, -0.32) & -0.08 & -0.07 (-0.27, 0.12) & -0.54 (-0.77, -0.32)\\
+\hspace{1em} & $\delta_{12}$ & 0.51 & 0.25 (-0.04, 0.53) & -0.26 (-0.6, 0.05) & 0.51 & 0.25 (-0.04, 0.53) & -0.26 (-0.6, 0.05) & 0.51 & 0.25 (-0.04, 0.53) & -0.26 (-0.6, 0.05)\\
+\hspace{1em} & $\delta_{21}$ & -0.97 & -0.71 (-0.95, -0.48) & -0.07 (-0.28, 0.14) & -0.97 & -0.71 (-0.95, -0.48) & -0.07 (-0.28, 0.14) & -0.97 & -0.71 (-0.95, -0.48) & -0.07 (-0.28, 0.14)\\
+\hspace{1em} & $\delta_{22}$ & 0.84 & 0.39 (0.09, 0.71) & 0.24 (-0.04, 0.5) & 0.84 & 0.39 (0.09, 0.71) & 0.24 (-0.04, 0.5) & 0.84 & 0.39 (0.09, 0.71) & 0.24 (-0.04, 0.5)\\
 \addlinespace[0.3em]
-\multicolumn{8}{l}{\textbf{ }}\\
-\hspace{1em}Clustering & $\pi_l$ & 0.39 & 0.39 (0.38, 0.4) & 0.26 & 0.26 (0.25, 0.27) & 0.34 & 0.35 (0.33, 0.36)\\
+\multicolumn{11}{l}{\textbf{ }}\\
+\hspace{1em}Clustering & $\pi_l$ & 0.38 & 0.38 (0.38, 0.38) & 0.38 (0.13, 0.41) & 0.4 & 0.39 (0.39, 0.4) & 0.39 (0.36, 0.43) & 0.23 & 0.23 (0.22, 0.23) & 0.23 (0.2, 0.44)\\
 \bottomrule
@@
 \end{table}
-
-\subsection{Simulation to Compare Imputation Methods}
-
-\subsection{Simulation to Assess Sensitivity to Misspecified K}
+\end{landscape}
 
 \newpage
+\subsection*{Figures}
 
-\section{Application}
-\label{s:app}
+\begin{figure}[h]
+	\label{fig:skew_resids}
+	\caption{The distribution of residuals in repeated measures regression model of Bayley composite scores adjusted for race and sex.}
+	\centering
+	\includegraphics[width = 1\textwidth]{bayley_resids_plot}
+\end{figure}
 
-\begin{itemize}
+\begin{figure}[h]
+	\label{fig:sim1_dens}
+	\caption{Marginal densities of simulated outcomes in each cluster at each measurement occasion.}
+	\centering
+	\includegraphics[width = 1\textwidth]{y_densities.pdf}
+\end{figure}
 
-\item Include both time varying and non-time varying covariates for the within cluster covariate set.
-
-\end{itemize}
+\begin{figure}[h]
+	\label{fig:sim1_means}
+	\caption{Mean plots of simulated outcomes in each cluster at each measurement occasion.}
+	\centering
+	\includegraphics[width = 1\textwidth]{y_means.pdf}
+\end{figure}
 
 \newpage
-
-\section{Discussion}
-\label{s:discuss}
-
-\textit{\textbf{Discuss non-ignorable missingness here}}
-
-\begin{itemize}
-
-\item Discuss how we handle non-ignorable missingness
-
-\end{itemize}
-
-\section{Appendix}
-
-Put your final comments here.
-
-%  The \backmatter command formats the subsequent headings so that they
-%  are in the journal style.  Please keep this command in your document
-%  in this position, right after the final section of the main part of 
-%  the paper and right before the Acknowledgements, Supplementary Materials,
-%  and References sections.
-
-\backmatter
-
-%  This section is optional.  Here is where you will want to cite
-%  grants, people who helped with the paper, etc.  But keep it short!
-
-\section*{Acknowledgements}
-
-%  If your paper refers to supplementary web material, then you MUST
-%  include this section!!  See Instructions for Authors at the journal
-%  website http://www.biometrics.tibs.org
-
-\section*{Supplementary Materials}
-
-\newpage
-
-\subsection{Glossary of Notation}
+\subsection*{Glossary of Notation}
 
@@
 
-\newpage
-\subsection{Derivation of Full Conditional Distributions}
-
-\subsubsection{Multivariate Skew-Normal Regression}
-
-Without loss of generality, we derive the full conditional distributions for the multivariate skew-normal regression model component under the assumption that all observations belong to a single cluster.
-To make the extension to the case where more than one cluster is specified, simply apply these distributional forms to cluster specific parameters and data.
-Finally, we assume for the moment that we have complete data for all outcomes for each subject.
-We extend consider the case of missing data in section (INSERT SECTION).
-
-The multivariate skew-normal regression model can be written as follows in matrix form. 
-$$\mathbf{Y} = \mathbf{X}\mathbf{B} + \mathbf{t} \boldsymbol\psi^T + \mathbf{E} = \mathbf{X}^* \mathbf{B}^* + \mathbf{E}$$
-The matrix $\mathbf{Y}$ is of dimension $n \times J$.
-For convenience, we define $\mathbf{X}^*$ as a $n \times (P + 1)$ matrix constructed by column binding $\mathbf{t}$ to $\mathbf{X}$, and $\mathbf{B}^*$ as a $(P+1) \times J$ matrix constructed by row binding $\boldsymbol\psi^T$ to $\mathbf{B}$.
-We assume that $t_i \stackrel{iid}{\sim}T_{[0,\infty})(0,1)$ and that $\mathbf{E}$ is made of row vectors $\boldsymbol\epsilon_i = (\epsilon_{i1},...,\epsilon_{iJ})$ for $i = 1,...,n$, where $ \boldsymbol\epsilon_i \stackrel{iid}{\sim} N_J(0, \boldsymbol\Sigma)$.
-
-The conditional likelihood for this model is given below. 
-$$p(\mathbf{Y}|\mathbf{X}^*,\mathbf{B}^*,\boldsymbol\Sigma) \propto |\boldsymbol\Sigma|^{-n/2}\exp \left \{ -\frac{1}{2} \tr(\mathbf{Y} - \mathbf{X}^* \mathbf{B}^*)^T(\mathbf{Y} - \mathbf{X}^* \mathbf{B}^*)\boldsymbol\Sigma^{-1} \right \}$$
-
-We choose conjugate priors for $\mathbf{B}^*$ and $\boldsymbol\Sigma$ as follows. 
-$$\boldsymbol\Sigma \sim \text{inverse-Wishart}(\mathbf{V}_0,\nu_0)$$
-$$\mathbf{B}^*|\boldsymbol\Sigma \sim MatNorm_{(m+1), p}(\mathbf{B}_0^*,\mathbf{L}_0^{-1},\boldsymbol\Sigma)$$
-
-We now derive the joint posterior distribution of the parameters $\mathbf{B}^*$ and $\boldsymbol\Sigma$.
-\begin{align*} 
-p(\mathbf{B}^*,\boldsymbol\Sigma|\mathbf{X}^*,\mathbf{Y}) & \propto  p(\mathbf{Y}|\mathbf{X}^*,\mathbf{B}^*,\boldsymbol\Sigma)p(\mathbf{B}^*|\boldsymbol\Sigma)p(\boldsymbol\Sigma)\\
-&  \propto |\boldsymbol\Sigma|^{-n/2}\exp \left \{ -\frac{1}{2} \tr \left [(\mathbf{Y} - \mathbf{X^*}\mathbf{B}^*)^T(\mathbf{Y} - \mathbf{X^*}\mathbf{B}^*)\boldsymbol\Sigma^{-1} \right ]\right \} \\
-& \times |\boldsymbol\Sigma|^{-(P+1)/2} \exp \left \{ -\frac{1}{2} \tr \left [(\mathbf{B}^* - \mathbf{B}^*_0)^T \mathbf{L}_0(\mathbf{B}^* - \mathbf{B}^*_0)\boldsymbol\Sigma^{-1} \right ]\right \} \\
-& \times |\boldsymbol\Sigma|^{(\nu_0 + J + 1)/2} \exp \left \{ -\frac{1}{2} \tr (\mathbf{V}_0 \boldsymbol\Sigma^{-1}) \right \}
-\end{align*}
-
-\subsubsection{Multinomial Logit Regression}
-
-\subsubsection{Multivariate Normal Conditional Imputation}
-
-The multivariate normal conditional imputation derivations are given for a single cluster without loss of generality.
-In practice, the data and parameters in this section would be replaced by cluster specific estimates in the case of clustering.
-
-For a given observation vector $\mathbf{y} \sim N_J(\boldsymbol\mu, \boldsymbol\Sigma)$, we allow for missingness in at most $J - 1$ of the multivariate outcomes through the use of a conditional imputation step embedded within our Gibbs sampler.
-Suppose $\mathbf{y}$ contains $q$ missing observations and can be partitioned into two vectors $\mathbf{y_{1}}$ and $\mathbf{y_{2}}$ such that $\mathbf{y_{1}}$ is a $q \times 1$ vector of missing observations and $\mathbf{y_{2}}$ is a $(J-q) \times 1$ vector of complete observations.
-Similarly, partition $\boldsymbol\mu$ and $\boldsymbol\Sigma$ as follows.
-$$\boldsymbol\mu = \begin{bmatrix} \boldsymbol\mu_1 \\ \boldsymbol\mu_2 \end{bmatrix} \ \ \ \ \ \boldsymbol\Sigma = \begin{bmatrix} \boldsymbol\Sigma_{11} & \boldsymbol\Sigma_{12} \\ \boldsymbol\Sigma_{21} & \boldsymbol\Sigma_{22} \end{bmatrix}$$
-
-We will use these quantities to derive the conditional distribution $f(\mathbf{y_1}|\mathbf{y_2},\boldsymbol\mu,\boldsymbol\Sigma)$. 
-\begin{align*} 
-f(\mathbf{y_1}|\mathbf{y_2},\boldsymbol\mu,\boldsymbol\Sigma) & \propto  f(\mathbf{y_1},\mathbf{y_2}|\boldsymbol\mu,\boldsymbol\Sigma) \\
-& \propto \exp \left \{ -\frac{1}{2}(\mathbf{y} - \boldsymbol\mu)^T \boldsymbol\Sigma^{-1} (\mathbf{y} - \boldsymbol\mu) \right \} \\
-& = \exp \left \{ -\frac{1}{2}\begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix}^T \boldsymbol\Sigma^{-1} \begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix} \right \} \\
-& =  \exp \left \{ -\frac{1}{2}\begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix}^T \begin{bmatrix} \boldsymbol\Sigma_{11} & \boldsymbol\Sigma_{12} \\ \boldsymbol\Sigma_{21} & \boldsymbol\Sigma_{22} \end{bmatrix}^{-1} \begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix} \right \} \\
-& = \exp \left \{ -\frac{1}{2}\begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix}^T \begin{bmatrix} \boldsymbol\Sigma_{11}^* & \boldsymbol\Sigma_{12}^* \\ \boldsymbol\Sigma_{21}^* & \boldsymbol\Sigma_{22}^* \end{bmatrix} \begin{bmatrix} \mathbf{y}_1 - \boldsymbol\mu_1 \\  \mathbf{y}_2 - \boldsymbol\mu_2 \end{bmatrix} \right \} \\
-& = \exp \left \{ -\frac{1}{2} \left[ (\mathbf{y}_1 - \boldsymbol\mu_{cond})^T \boldsymbol\Sigma_{cond}^{-1}(\mathbf{y}_1 - \boldsymbol\mu_{cond})\right]\right \} 
-\end{align*}
-$$\Rightarrow \mathbf{y_1}|\mathbf{y_2},\boldsymbol\mu,\boldsymbol\Sigma \sim N_q(\boldsymbol\mu_{cond},\boldsymbol\Sigma_{cond})$$
-$$\boldsymbol\mu_{cond} = \boldsymbol\mu_1 + \boldsymbol\Sigma_{12}\boldsymbol\Sigma_{22}^{-1}(\mathbf{y}_2 - \boldsymbol\mu_2), \ \ \ \ \ \ \boldsymbol\Sigma_{cond} = \boldsymbol\Sigma_{11} - \boldsymbol\Sigma_{12}\boldsymbol\Sigma_{22}^{-1}\boldsymbol\Sigma_{21}$$
-
-The block-wise inversion formula was used to invert $\boldsymbol\Sigma$ according to the following reparameterizations.
-\begin{align*}
-\boldsymbol\Sigma_{11}^* & = \boldsymbol\Sigma_{11}^{-1} + \boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12}(\boldsymbol\Sigma_{22} - \boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12})^{-1}\boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\\
-\boldsymbol\Sigma_{12}^* & = -\boldsymbol\Sigma_{11} \boldsymbol\Sigma_{12}(\boldsymbol\Sigma_{22}-\boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12})^{-1}\\
-\boldsymbol\Sigma_{21}^* & = -(\boldsymbol\Sigma_{22} - \boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12})^{-1}\boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\\
-\boldsymbol\Sigma_{22}^* & = (\boldsymbol\Sigma_{22} - \boldsymbol\Sigma_{21}\boldsymbol\Sigma_{11}^{-1}\boldsymbol\Sigma_{12})^{-1}     
-\end{align*}
-
-%  Here, we create the bibliographic entries manually, following the
-%  journal style.  If you use this method or use natbib, PLEASE PAY
-%  CAREFUL ATTENTION TO THE BIBLIOGRAPHIC STYLE IN A RECENT ISSUE OF
-%  THE JOURNAL AND FOLLOW IT!  Failure to follow stylistic conventions
-%  just lengthens the time spend copyediting your paper and hence its
-%  position in the publication queue should it be accepted.
-
-%  We greatly prefer that you incorporate the references for your
-%  article into the body of the article as we have done here 
-%  (you can use natbib or not as you choose) than use BiBTeX,
-%  so that your article is self-contained in one file.
-%  If you do use BiBTeX, please use the .bst file that comes with 
-%  the distribution.  In this case, replace the thebibliography
-%  environment below by 
-%
-%  \bibliographystyle{biom} 
-% \bibliography{mybibilo.bib}
-
-\newpage
-
-\begin{thebibliography}{}
-
-\bibitem{ } Arellano‐-Valle RB, Azzalini A. On the unification of families of skew‐normal distributions. \textit{Scandinavian Journal of Statistics}. 2006 Sep;33(3):561-74.
-
-\bibitem{ } Azzalini A. A class of distributions which includes the normal ones. \textit{Scandinavian journal of statistics}. 1985 Jan 1:171-8.
-
-\bibitem{ } Azzalini, A. and Dalla Valle, A. (1996). The multivariate skew normal distribution. Biometrika 83, 715–726.
-
-\bibitem{ } Chen JT, Gupta AK. Matrix variate skew normal distributions. \textit{Statistics}. 2005 Jun 1;39(3):247-53.
-
-\bibitem{ } Neelon SE, \O stbye T, Bennett GG, Kravitz RM, Clancy SM, Stroo M, Iversen E, Hoyo C.
-Cohort profile for the Nurture Observational Study examining associations of multiple caregivers on infant growth in the Southeastern USA.
-\textit{BMJ Open}.
-2017 Feb 1;7(2):e013939.
-
-\bibitem{ } Franczak BC, Tortora C, Browne RP, McNicholas PD.
-Unsupervised learning via mixtures of skewed distributions with hypercube contours.
-\textit{Pattern Recognition Letters}.
-2015 Jun 1;58:69-76.
-
-\bibitem{ } Fr\"{u}hwirth-Schnatter S, Pyne S.
-Bayesian inference for finite mixtures of univariate and multivariate skew-normal and skew-t distributions.
-\textit{Biostatistics}.
-2010 Jan 27;11(2):317-36.
-
-\bibitem{ } Ganjali M, Baghfalaki T.
-A Bayesian shared parameter model for analysing longitudinal skewed responses with nonignorable dropout.
-\textit{International Journal of Statistics in Medical Research}.
-2014 Apr 1;3(2):103.
-
-\bibitem{ } Gelman A, Stern HS, Carlin JB, Dunson DB, Vehtari A, Rubin DB. Bayesian data analysis. \textit{Chapman and Hall/CRC}; 2013 Nov 27.
-
-\bibitem{ } Gelman A, Hwang J, Vehtari A. Understanding predictive information criteria for Bayesian models. \textit{Statistics and Computing}. 2014 Nov 1;24(6):997-1016.
-
-\bibitem{ } Holmes CC, Held L. Bayesian auxiliary variable models for binary and multinomial regression. \textit{Bayesian Analysis}. 2006;1(1):145-68.
-
-\bibitem{ } Lagona F, Picone M. Model-based clustering of multivariate skew data with circular components and missing values. \textit{Journal of Applied Statistics}. 2012 May 1;39(5):927-45.
-
-\bibitem{ } Lanza ST, Rhoades BL. Latent class analysis: an alternative perspective on subgroup analysis in prevention and treatment. \textit{Prevention Science}. 2013 Apr 1;14(2):157-68.
-
-\bibitem{ } Lee SX, McLachlan GJ. Model-based clustering and classification with non-normal mixture distributions. Statistical Methods \& Applications. 2013 Nov 1;22(4):427-54.
-
-\bibitem{ } Lee SX, Mclachlan GJ. On mixtures of skew normal and skew $t$-distributions. \textit{Advances in Data Analysis and Classification}. 2013 Sep 1;7(3):241-66.
-
-\bibitem{ } Lin TI, Wang WL, McLachlan GJ, Lee SX.
-Robust mixtures of factor analysis models using the restricted multivariate skew-t distribution.
-\textit{Statistical Modelling}.
-2018 Feb;18(1):50-72.
-
-\bibitem{ } Luo S, Lawson AB, He B, Elm JJ, Tilley BC.
-Bayesian multiple imputation for missing multivariate longitudinal data from a Parkinson's disease clinical trial.
-\textit{Statistical Methods in Medical Research}.
-2016 Apr;25(2):821-37.
-
-\bibitem{ } Melnykov V, Maitra R. Finite mixture models and model-based clustering. \textit{Statistics Surveys}. 2010;4:80-116.
-
-\bibitem{ } Polson NG, Scott JG, Windle J. Bayesian inference for logistic models using P\'olya –- Gamma latent variables. Journal of the American statistical Association. 2013 Dec 1;108(504):1339-49.
-
-\bibitem{ } Tiao GC, Zellner A. On the Bayesian estimation of multivariate regression. \textit{Journal of the Royal Statistical Society}: Series B (Methodological). 1964 Jul;26(2):277-85.
-
-\bibitem{ } Viroli C. Finite mixtures of matrix normal distributions for classifying three-way data. \textit{Statistics and Computing}. 2011 Oct 1;21(4):511-22.
-
-\bibitem{ } Vrbik I, Mcnicholas PD. Parsimonious skew mixture models for model-based clustering and classification. \textit{Computational Statistics \& Data Analysis}. 2014 Mar 1;71:196-210.
-
-\bibitem{ } Zeller CB, Cabral CR, Lachos VH. Robust mixture regression modeling based on scale mixtures of skew-normal distributions. \textit{Test}. 2016 Jun 1;25(2):375-96.
-
-\bibitem{ } Zhou X, Reiter JP. A note on Bayesian inference after multiple imputation. \textit{The American Statistician}. 2010 May 1;64(2):159-63.
-
-\end{thebibliography}
-
-% \appendix
-
-%  To get the journal style of heading for an appendix, mimic the following.
-
-\label{lastpage}
-
 \end{document}

### 2019-07-09 mcmc_draws_2019-07-08/mcmc_draws_NOSKEW_2019-07-02/tables.Rmd
sim 1

--- 2019-07-02 mcmc_draws_2019-07-02/tables.Rmd
@@
 true_c3 <- c(beta_true_c3,omega_true_c3,alpha_true_c3,delta_true,pi_true[3])
-est_c1 <- c(beta_est_c1,omega_est_c1,alpha_est_c1,delta_est,pi_est[1])
-est_c2 <- c(beta_est_c2,omega_est_c2,alpha_est_c2,delta_est,pi_est[2])
-est_c3 <- c(beta_est_c3,omega_est_c3,alpha_est_c3,delta_est,pi_est[3])
+est_c1 <- est_c1_noskew <- c(beta_est_c1,omega_est_c1,alpha_est_c1,delta_est,pi_est[1])
+est_c2 <- est_c2_noskew <- c(beta_est_c2,omega_est_c2,alpha_est_c2,delta_est,pi_est[2])
+est_c3 <- est_c3_noskew <- c(beta_est_c3,omega_est_c3,alpha_est_c3,delta_est,pi_est[3])
 
-load("mcmc_draws_NOSKEW_2019-07-02/class_1_est")
-load("mcmc_draws_NOSKEW_2019-07-02/class_2_est")
-load("mcmc_draws_NOSKEW_2019-07-02/class_3_est")
+save(est_c1_noskew,file = "class_1_est")
+save(est_c2_noskew,file = "class_2_est")
+save(est_c3_noskew,file = "class_3_est")
 
@@
             params,
-            true_c1,est_c1,est_c3_noskew,
-            true_c2,est_c2,est_c2_noskew,
-            true_c3,est_c3,est_c1_noskew)
-colnames(t1) <- c("Component",
-                  "Param.",
-                  "True","MSN Est. (95\\% CrI)","MN Est. (95\\% CrI) ",
-                  "True","MSN Est. (95\\% CrI)","MN Est. (95\\% CrI)",
-                  "True","MSN Est. (95\\% CrI)","MN Est. (95\\% CrI)")
+            true_c1,est_c1,
+            true_c2,est_c2,
+            true_c3,est_c3)
+colnames(t1) <- c("Model Component",
+                  "Parameter",
+                  "True","Est. (95\\% CrI)",
+                  "True","Est. (95\\% CrI)",
+                  "True","Est. (95\\% CrI)")
 ```
@@
 kable(t1, booktabs = T, escape = F,caption = "Model results for simulated data with n = 1000, k = 4, p = 2, h = 3, r = 2.
-1000 iterations were run with a burn in of 100.
-Missingness mechanism was MAR and P(miss) = 0.
-Model results for the multivariate skew normal (MSN) and multivariate normal (MN) mixtures are presented.") %>%
-  kable_styling(font_size = 7) %>%
-  add_header_above(c(" " = 2,"Class 1" = 3, "Class 2" = 3, "Class 3" = 3)) %>%
+1000 iterations were run with a burn in of 250.
+Missingness mechanism was MAR and P(miss) = 0") %>%
+  kable_styling(font_size = 8) %>%
+  add_header_above(c(" " = 2,"Class 1" = 2, "Class 2" = 2, "Class 3" = 2)) %>%
   group_rows(" ",1,8) %>%
@@
   group_rows(" ",23,26) %>%
-  group_rows(" ",27,27) %>%
-  landscape()
+  group_rows(" ",27,27)
 ```

### 2019-07-09 mcmc_draws_2019-07-08/tables.Rmd
sim 1

--- 2019-07-02 mcmc_draws_2019-07-02/tables.Rmd
@@
 ```{r}
-setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-02")
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-08")
 load("BETA")

### 2019-07-09 mcmc_draws_2019-07-08/tables_2.Rmd
sim 1

--- 2019-07-09 mcmc_draws_2019-07-08/mcmc_draws_NOSKEW_2019-07-02/tables.Rmd
@@
 title: "Simulation Tables"
-author: "Carter Allen"
+author:
 date: 
@@
 ```{r}
-setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-02/mcmc_draws_NOSKEW_2019-07-02/")
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-09")
 load("BETA")
@@
 true_c3 <- c(beta_true_c3,omega_true_c3,alpha_true_c3,delta_true,pi_true[3])
-est_c1 <- est_c1_noskew <- c(beta_est_c1,omega_est_c1,alpha_est_c1,delta_est,pi_est[1])
-est_c2 <- est_c2_noskew <- c(beta_est_c2,omega_est_c2,alpha_est_c2,delta_est,pi_est[2])
-est_c3 <- est_c3_noskew <- c(beta_est_c3,omega_est_c3,alpha_est_c3,delta_est,pi_est[3])
-
-save(est_c1_noskew,file = "class_1_est")
-save(est_c2_noskew,file = "class_2_est")
-save(est_c3_noskew,file = "class_3_est")
+est_c1 <- c(beta_est_c1,omega_est_c1,alpha_est_c1,delta_est,pi_est[1])
+est_c2 <- c(beta_est_c2,omega_est_c2,alpha_est_c2,delta_est,pi_est[2])
+est_c3 <- c(beta_est_c3,omega_est_c3,alpha_est_c3,delta_est,pi_est[3])
 
@@
 rownames(t1) <- NULL
-kable(t1, booktabs = T, escape = F,caption = "Model results for simulated data with n = 1000, k = 4, p = 2, h = 3, r = 2.
-1000 iterations were run with a burn in of 250.
+kable(t1, booktabs = T, escape = F,caption = "Model results for simulated data with n = 1000, J = 4, p = 2, K = 3, r = 2.
+1,000 iterations were run with a burn in of 0.
 Missingness mechanism was MAR and P(miss) = 0") %>%

### 2019-07-09 figures/figures_intro.R

@@ -83,3 +83,5 @@ plot_bytime <- ggplot(resids_df) +
 ggsave(filename = "figures/bayley_resids_plot_bytime.jpg",
        plot = plot_bytime,
        device = "jpg")
+
+

### 2019-07-09 figures/figures_mcmc.R

@@ -7,7 +7,7 @@ library(vapoRwave)
 vaporwave <- vapoRwave:::jazzCup_palette
 
 setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM")
-load_dir <- "mcmc_draws_2019-07-04"
+load_dir <- "mcmc_draws_2019-07-09"
 load(paste(load_dir,"/BETA",sep = ""))
 load(paste(load_dir,"/beta_true",sep = ""))
 load(paste(load_dir,"/DELTA",sep = ""))
@@ -19,14 +19,16 @@ load(paste(load_dir,"/SIGMA",sep = ""))
 load(paste(load_dir,"/sig2_true",sep = ""))
 load(paste(load_dir,"/Z",sep = ""))
 load(paste(load_dir,"/c_true",sep = ""))
+load(paste(load_dir,"/Y",sep = ""))
 
 h <- length(BETA.list)
 k <- ncol(beta.true.list[[1]])
 p <- nrow(beta.true.list[[1]])
 v <- nrow(delta.true)
 
-burn <- 250
+burn <- 0
 
+colnames(Y) <- paste("Y",seq(1,ncol(Y)),sep = "")
 colnames(Z) <- paste("Z",seq(1,ncol(Z)),sep = "")
 Z <- Z[-(1:burn),]
 colnames(DELTA) <- paste("D",seq(1,v*(h-1)),sep = "")
@@ -91,3 +93,36 @@ z_means <- colMeans(Z[-(1:burn),])
 ggplot() + 
     geom_line(aes(x = seq_along(z_means),y = c),color = vaporwave[1]) + 
     geom_line(aes(x = seq_along(z_means),y = z_means),color = vaporwave[2])
+
+## Y densities
+y_density_plot <- data.frame(Y,c) %>%
+    rename(clust = c) %>%
+    gather(variable,value,-clust) %>%
+    ggplot(.) +
+    stat_density(aes(x = value),geom = "line") +
+    facet_grid(cols = vars(variable), rows = vars(clust)) +
+    theme_minimal() + 
+    xlab("Y") + 
+    ylab("Density") 
+ggsave(y_density_plot,
+       filename = paste(load_dir,"y_densities.pdf",
+                        sep = "/"))
+
+## Y means
+y_mean_plot <- data.frame(Y,c) %>%
+    rename(clust = c) %>%
+    gather(variable,value,-clust) %>%
+    group_by(variable,clust) %>%
[…]

### 2019-07-09 mcmc_draws_2019-07-09/WAIC.R

@@ -1,0 +1,15 @@
+load("Y")
+load("X")
+load("W")
+load("Z")
+load("SIGMA")
+load("PSI")
+load("BETA")
+load("DELTA")
+
+n <- nrow(Y)
+K <- length(BETA.list)
+J <- ncol(Y)
+p <- ncol(X)
+r <- ncol(W)
+

### 2019-07-09 simulation.R

@@ -1,0 +1,428 @@
+# Multivariate skew normal regression with clustering
+# Includes built-in mvn conditional imputation
+# Includes regression on multinomial regression coefficients
+# Carter Allen
+# March 19, 2019
+
+# Packages
+library(sn) # for rsn
+library(truncnorm) # for rtruncnorm
+library(mvtnorm) # for rmvtnorm
+library(ggplot2)
+library(patchwork)
+library(mvnmle)
+library(Hmisc) # For rMultinom function (easy sampling of class indicators C)
+library(bayesplot)
+library(coda)
+library(MCMCpack)
+library(matrixsampling)
+library(QRM)        # For rGumbel function
+library(nnet)       # To fit multinom function
+library(BayesLogit) # For rpg function
+
+#seed <- rnorm(1)
+#set.seed(seed)
+# No scientific notation
+options(scipen=999)
+options(warn=2)
+# Generate data
+k <- 4 # number of outcomes
+p <- 2 # number of predictors 
+h <- 3 # number of clusters
+n <- 1000 # number of observations (takes alot of data to estimate this many parameters)
+print(paste("Generating data with",n,"observations of dimension",k,
+            "with",p,"covariates and",h,"clusters"))
+
+N <- n*k # number of measurments 
+P <- matrix(rep(rep(1/h,h),each=n),n,h) # probability matrix governing cluster membership
+
+# Generate covariates
+X <- matrix(rnorm(n*p),nrow = n,ncol = p) # random predictors
+X[,1] <- 1
+t <- rtruncnorm(n,0,Inf,0,1) # trunc norm random effects
+Xstar <- cbind(X,t) # Combine X and t
+# # Generate X manually
+# xi <- matrix(c(1, 0, 0, 0,
+#                1, 1, 0, 0,
+#                1, 0, 1, 0,
+#                1, 0, 0, 1),
+#              nrow = 4,
+#              ncol = 4,
+#              byrow = TRUE)
+# t <- rtruncnorm(n,0,Inf,0,1) # trunc norm random effects
+# X <- NULL 
+# capT <- NULL
+# for(i in 1:n)
+# {
+#     X <- rbind(X,xi)
+#     capT <- rbind(capT,diag(t[i],nrow = 4,ncol = 4))
+# }
+# 
+# Xstar <- cbind(X,capT) # Combine X and t
+
[…]

### 2019-07-09 simulation/figures.R

@@ -1,93 +1,0 @@
-# Plot MCMC draws
-
-library(bayesplot)
-library(tidyverse)
-library(vapoRwave)
-
-vaporwave <- vapoRwave:::jazzCup_palette
-
-
-load_dir <- "mcmc_draws_2019-03-28"
-load(paste(load_dir,"/BETA",sep = ""))
-load(paste(load_dir,"/beta_true",sep = ""))
-load(paste(load_dir,"/DELTA",sep = ""))
-load(paste(load_dir,"/delta_true",sep = ""))
-load(paste(load_dir,"/PI",sep = ""))
-load(paste(load_dir,"/PSI",sep = ""))
-load(paste(load_dir,"/psi_true",sep = ""))
-load(paste(load_dir,"/SIGMA",sep = ""))
-load(paste(load_dir,"/sig2_true",sep = ""))
-load(paste(load_dir,"/Z",sep = ""))
-load(paste(load_dir,"/c_true",sep = ""))
-
-h <- length(BETA.list)
-k <- ncol(beta.true.list[[1]])
-p <- nrow(beta.true.list[[1]])
-v <- nrow(delta.true)
-
-burn <- 100
-
-colnames(Z) <- paste("Z",seq(1,ncol(Z)),sep = "")
-Z <- Z[-(1:burn),]
-colnames(DELTA) <- paste("D",seq(1,v*(h-1)),sep = "")
-DELTA <- DELTA[-(1:burn),]
-colnames(PI) <- paste("PI",seq(1,ncol(PI)),sep = "")
-PI <- PI[-(1:burn),]
-
-for(l in 1:h)
-{
-    colnames(BETA.list[[l]]) <- paste("B",seq(1,p*k),sep = "")
-    BETA.list[[l]] <- BETA.list[[l]][-(1:burn),]
-    colnames(PSI.list[[l]]) <- paste("PSI",seq(1,length(psi.true.list[[1]])),sep = "")
-    PSI.list[[l]] <- PSI.list[[l]][-(1:burn),]
-    colnames(SIGMA.list[[l]]) <- paste("SIG",seq(1,ncol(SIGMA.list[[l]])),sep="")
-    SIGMA.list[[l]] <- SIGMA.list[[l]][-(1:burn),]
-}
-
-## BETA plots
-
-BETA_trace <- mcmc_trace(BETA.list) + 
-    scale_color_manual(name = "Cluster",
-                         values = vaporwave) + 
-    ggtitle("Trace plots of Beta Coefficients")
[…]

### 2019-07-09 simulation/tables.R

@@ -1,16 +1,0 @@
-# Analyze MCMC draws
-
-library(tidyverse)
-
-# Load draws
-
-load("mcmc_draws_2019-03-26/BETA")
-load("mcmc_draws_2019-03-26/beta_true")
-load("mcmc_draws_2019-03-26/DELTA")
-load("mcmc_draws_2019-03-26/delta_true")
-load("mcmc_draws_2019-03-26/PI")
-load("mcmc_draws_2019-03-26/PSI")
-load("mcmc_draws_2019-03-26/psi_true")
-load("mcmc_draws_2019-03-26/SIGMA")
-load("mcmc_draws_2019-03-26/sig2_true")
-load("mcmc_draws_2019-03-26/Z")

### 2019-07-09 simulation_1/mv-reg-clust-impu-multinom.R

@@ -23,14 +23,14 @@ library(BayesLogit) # For rpg function
 
 # No scientific notation
 options(scipen=999)
-setwd("/Users/carter/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-02/")
+setwd("/Users/carter/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-09/")
 
 # Generate data
 k <- 4 # number of outcomes
 p <- 2 # number of predictors (coincluding intercept)
 h <- 3 # number of clusters
 n <- 1000 # number of observations (takes alot of data to estimate this many parameters)
-print(paste("Generating data with",n,"observations of dimension",k,
+    print(paste("Reading data with",n,"observations of dimension",k,
             "with",p,"covariates and",h,"clusters"))
 
 N <- n*k # number of measurments 
@@ -42,12 +42,6 @@ load("X")
 t <- rtruncnorm(n,0,Inf,0,1) # trunc norm random effects
 Xstar <- cbind(X,t) # Combine X and t
 
-# Covariates corresponding to multinomial regression
-#w1 <- rnorm(n) # continuous predictor
-#w2 <- rnorm(n) # continuous predictor
-w3 <- rbinom(n,1,0.5) # binary predictor
-w4 <- rbinom(n,1,0.5) # binary predictor
-# W <- cbind(1,w3,w4) # design matrix
 load("W")
 v <- ncol(W) # number of multinomial predictors
 delta.true <- matrix(rtruncnorm(n = v*(h-1),
@@ -141,7 +135,7 @@ S0 <- diag(0.5,v) # prior covariance for delta coefficients (multinomial regress
 
 # Sample storage
 nsim <- 1000 # number of iterations
-burn <- 100 # number of iterations to save
+burn <- 0 # number of iterations to save
 n.iter <- nsim - burn # number of saved iterations
 Z <- matrix(0,nrow = n.iter,ncol = n) # large matrix where each row is the value of z at a specific iteration
 PI <- matrix(0,nrow = n.iter,ncol = h) # matrix w/ each row as pi vector at each iteration

### 2019-07-09 simulation_1/simulation_1.R

@@ -8,11 +8,8 @@
 library(sn) # for rsn
 library(truncnorm) # for rtruncnorm
 library(mvtnorm) # for rmvtnorm
-library(ggplot2)
-library(patchwork)
 library(mvnmle)
 library(Hmisc) # For rMultinom function (easy sampling of class indicators C)
-library(bayesplot)
 library(coda)
 library(MCMCpack)
 library(matrixsampling)
@@ -27,9 +24,9 @@ options(scipen=999)
 options(warn=2)
 # Generate data
 k <- 4 # number of outcomes
-p <- 1 # number of predictors 
+p <- 2 # number of predictors 
 h <- 3 # number of clusters
-n <- 500 # number of observations (takes alot of data to estimate this many parameters)
+n <- 1000 # number of observations (takes alot of data to estimate this many parameters)
 print(paste("Generating data with",n,"observations of dimension",k,
             "with",p,"covariates and",h,"clusters"))
 
@@ -41,31 +38,9 @@ X <- matrix(rnorm(n*p),nrow = n,ncol = p) # random predictors
 X[,1] <- 1
 t <- rtruncnorm(n,0,Inf,0,1) # trunc norm random effects
 Xstar <- cbind(X,t) # Combine X and t
-# # Generate X manually
-# xi <- matrix(c(1, 0, 0, 0,
-#                1, 1, 0, 0,
-#                1, 0, 1, 0,
-#                1, 0, 0, 1),
-#              nrow = 4,
-#              ncol = 4,
-#              byrow = TRUE)
-# t <- rtruncnorm(n,0,Inf,0,1) # trunc norm random effects
-# X <- NULL 
-# capT <- NULL
-# for(i in 1:n)
-# {
-#     X <- rbind(X,xi)
-#     capT <- rbind(capT,diag(t[i],nrow = 4,ncol = 4))
-# }
-# 
-# Xstar <- cbind(X,capT) # Combine X and t
 
-# Covariates corresponding to multinomial regression
-#w1 <- rnorm(n) # continuous predictor
-#w2 <- rnorm(n) # continuous predictor
-w3 <- rbinom(n,1,0.5) # binary predictor
-w4 <- rbinom(n,1,0.5) # binary predictor
-W <- cbind(1,w3) # design matrix
+w1 <- rbinom(n,1,0.5) # binary predictor
+W <- cbind(1,w1) # design matrix
 v <- ncol(W) # number of multinomial predictors
[…]

### 2019-07-09 manuscript/manuscript.tex
sim 1

@@
 For the multinomial regression model component of this simulation, we modeled the class labels $z_i$ as a function of an intercept and one baseline covariate, $\text{w}_{i1}$, implying that $r = 2$.
-Since a comparison of imputation method to alternative approaches is performed later in our second simulation study, we did not introduce missing data into this simulatation.
+We did introduce missing data into this simulation, as missing data will be dealt with in our second simulation study.
 As a result, the final sample size was $N = n \times J = 4,000$.
@@
 Similarly, we set $\mathbf{d}_{01} = \mathbf{d}_{02} = (0,0)^T$ and $\mathbf{S}_{01} = \mathbf{S}_{02} = \mathds{1}_{2 \times 2}$, noting that $k = 3$ is the reference cluster.
-To investigate the effect of ignoring skewness, we compare MSN and MVN models in terms of parameter estimates and WAIC as described below.
+To investigate the effect of ignoring skewness, we fit both the MVN and MSN mixture models to data generated from model (\ref{eq:sim1}).
+The values of WAIC for the MSN mixture model and MVN mixture model were 37,221.7 and 324,705.3, respectively, indicating model fit suffered when skewness was ignored.
+Furthermore, we report below in Table \ref{tab:sim1_parms} a selection of parameter estimates from the MSN and MVN models that show incorrect inference (i.e., true parameter values not contained in 95\% posterior credible intervals) resulted from ignoring skewness.
+A full table showing all model parameters estimates and 95\% posterior credible intervals for the MSN and MVN models in addition to ground truth values is given in Web Appenxix B.
 
@@
 
-\caption{\label{tab:unnamed-chunk-5}Model results for simulated data with n = 1,000, J = 4, p = 2, K = 3, r = 2.
+\caption{\label{tab:sim1_parms}Model results for simulated data with n = 1,000, J = 4, p = 2, K = 3, r = 2.
 1,000 iterations were run with a burn in of 100.
@@
 \end{table}
+
+An attractive feature of the MSN model is that it includes the MVN model as a special case.
+As such, the MSN model can be used in place of the MVN model even when data are not clearly skewed.
+We demonstrate this by fitting both the MSN and MVN model to data generated from model (\ref{eq:sim1}), where all skewness parameters are set to zero.
+As shown Table 2 of Web Appendix B, parameter estimates are similar between the MSN and MVN models, and all 95\% credible intervals for the $\alpha$ parameters of the MSN models contain 0, indicating the MSN model performs well even when data are not skewed.
 

### 2019-07-09 mcmc_draws_2019-07-09_SIM1/mcmc_draws_NOSKEW_2019-07-09/tables.Rmd
sim 1

--- 2019-07-09 mcmc_draws_2019-07-08/mcmc_draws_NOSKEW_2019-07-02/tables.Rmd
@@
 ```{r}
-setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-02/mcmc_draws_NOSKEW_2019-07-02/")
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-09/mcmc_draws_NOSKEW_2019-07-09/")
 load("BETA")

### 2019-07-09 mcmc_draws_2019-07-09_SIM1/tables.Rmd
sim 1

--- 2019-07-09 mcmc_draws_2019-07-08/tables.Rmd
@@
 ```{r}
-setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-08")
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-09")
 load("BETA")
@@
 
-load("mcmc_draws_NOSKEW_2019-07-02/class_1_est")
-load("mcmc_draws_NOSKEW_2019-07-02/class_2_est")
-load("mcmc_draws_NOSKEW_2019-07-02/class_3_est")
+load("mcmc_draws_NOSKEW_2019-07-09/class_1_est")
+load("mcmc_draws_NOSKEW_2019-07-09/class_2_est")
+load("mcmc_draws_NOSKEW_2019-07-09/class_3_est")
 
@@
             params,
-            true_c1,est_c1,est_c3_noskew,
-            true_c2,est_c2,est_c2_noskew,
-            true_c3,est_c3,est_c1_noskew)
+            true_c1,est_c1,est_c2_noskew,
+            true_c2,est_c2,est_c1_noskew,
+            true_c3,est_c3,est_c3_noskew)
 colnames(t1) <- c("Component",
@@
 rownames(t1) <- NULL
-kable(t1, booktabs = T, escape = F,caption = "Model results for simulated data with n = 1000, k = 4, p = 2, h = 3, r = 2.
-1000 iterations were run with a burn in of 100.
+kable(t1, booktabs = T, escape = F,caption = "Model results for simulated data with n = 1,000, J = 4, p = 2, K = 3, r = 2.
+1,000 iterations were run with a burn in of 100.
 Missingness mechanism was MAR and P(miss) = 0.

### 2019-07-09 mcmc_draws_2019-07-09_SIM1/tables_short_sim1.Rmd
sim 1

--- 2019-07-09 mcmc_draws_2019-07-09_SIM1/tables.Rmd
@@
 
-model_component <- c("MVSN","Regression","","","","","","","","","","","","","","","","","","","","",
+model_component <- c("MVSN","Regression","","","","","","","","","","",
                      "Multinom.","","","",
@@
             "$\\beta_{12}$","$\\beta_{22}$","$\\beta_{32}$","$\\beta_{42}$",
-            "$\\Omega_{11}$","$\\Omega_{12}$","$\\Omega_{13}$","$\\Omega_{14}$",
-            "$\\Omega_{22}$","$\\Omega_{23}$","$\\Omega_{24}$",
-            "$\\Omega_{33}$","$\\Omega_{34}$",
-            "$\\Omega_{44}$",
             "$\\alpha_{1}$","$\\alpha_{2}$", "$\\alpha_{3}$","$\\alpha_{4}$",
@@
 
-true_c1 <- c(beta_true_c1,omega_true_c1,alpha_true_c1,delta_true,pi_true[1])
-true_c2 <- c(beta_true_c2,omega_true_c2,alpha_true_c2,delta_true,pi_true[2])
-true_c3 <- c(beta_true_c3,omega_true_c3,alpha_true_c3,delta_true,pi_true[3])
-est_c1 <- c(beta_est_c1,omega_est_c1,alpha_est_c1,delta_est,pi_est[1])
-est_c2 <- c(beta_est_c2,omega_est_c2,alpha_est_c2,delta_est,pi_est[2])
-est_c3 <- c(beta_est_c3,omega_est_c3,alpha_est_c3,delta_est,pi_est[3])
+true_c1 <- c(beta_true_c1,alpha_true_c1,delta_true,pi_true[1])
+true_c2 <- c(beta_true_c2,alpha_true_c2,delta_true,pi_true[2])
+true_c3 <- c(beta_true_c3,alpha_true_c3,delta_true,pi_true[3])
+est_c1 <- c(beta_est_c1,alpha_est_c1,delta_est,pi_est[1])
+est_c2 <- c(beta_est_c2,alpha_est_c2,delta_est,pi_est[2])
+est_c3 <- c(beta_est_c3,alpha_est_c3,delta_est,pi_est[3])
 
@@
             params,
-            true_c1,est_c1,est_c2_noskew,
-            true_c2,est_c2,est_c1_noskew,
-            true_c3,est_c3,est_c3_noskew)
+            true_c1,est_c1,est_c2_noskew[-(9:18)])
 colnames(t1) <- c("Component",
                   "Param.",
-                  "True","MSN Est. (95\\% CrI)","MN Est. (95\\% CrI) ",
-                  "True","MSN Est. (95\\% CrI)","MN Est. (95\\% CrI)",
-                  "True","MSN Est. (95\\% CrI)","MN Est. (95\\% CrI)")
+                  "True","MSN Est. (95\\% CrI)","MN Est. (95\\% CrI) ")
 ```
@@
   kable_styling(font_size = 7) %>%
-  add_header_above(c(" " = 2,"Class 1" = 3, "Class 2" = 3, "Class 3" = 3)) %>%
+  add_header_above(c(" " = 2,"Class 1" = 3)) %>%
   group_rows(" ",1,8) %>%
-  group_rows(" ",9,18) %>%
-  group_rows(" ",19,22) %>%
-  group_rows(" ",23,26) %>%
-  group_rows(" ",27,27) %>%
-  landscape()
+  group_rows(" ",9,12) %>%
+  group_rows(" ",13,16) %>%
+  group_rows(" ",17,17)
 ```

### 2019-07-09 mcmc_draws_2019-07-09/WAIC.R

@@ -1,15 +1,66 @@
+library(sn)
+
 load("Y")
 load("X")
-load("W")
+load("X_star")
 load("Z")
 load("SIGMA")
 load("PSI")
 load("BETA")
-load("DELTA")
+
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-09")
 
 n <- nrow(Y)
 K <- length(BETA.list)
 J <- ncol(Y)
 p <- ncol(X)
-r <- ncol(W)
+S <- nrow(Z)
 
+lppd <- 0
+pwaic <- 0
+WAIC <- 0
+
+pb <- txtProgressBar(min = 0, max = n, style = 3)
+for(i in 1:n)
+{
+    y_i <- Y[i,]
+    x_i <- X[i,]
+    xstar_i <- Xstar[i,]
+    t_i <- xstar_i[p+1]
+    z_i <- Z[,i]
+    p_yi <- rep(0,S)
+    for(s in 1:S)
+    {
+        k <- z_i[s]
+        
+        beta_k <- BETA.list[[k]]
+        beta_k_s <- beta_k[s,]
+        beta_ks <- matrix(beta_k_s,
+                          nrow = p,
+                          ncol = J,
+                          byrow = TRUE)
+        
+        sigma_k <- SIGMA.list[[k]]
+        sigma_k_s <- sigma_k[s,]
+        sigma_ks <- matrix(sigma_k_s,
+                           nrow = J,
+                           ncol = J,
+                           byrow = TRUE)
+        
+        psi_k <- PSI.list[[k]]
+        psi_ks <- psi_k[s,]
+        
+        zeta_iks <- x_i %*% beta_ks
+        
+        mu_iks <- zeta_iks + t_i %*% psi_ks
+        
+        p_yi[s] <- dmvnorm(y_i,mu_iks,sigma_ks)
+    }
+    setTxtProgressBar(pb, i)
+    
+    (lppd <- lppd + log(mean(p_yi, na.rm = TRUE)))
+    (pwaic <- pwaic + var(log(p_yi)[is.finite(log(p_yi))], na.rm = TRUE))
+}
+WAIC <- -2*(lppd - pwaic)
+
+close(pb)

### 2019-07-09 mcmc_draws_2019-07-09/mcmc_draws_2019-07-09/WAIC.R

@@ -1,0 +1,68 @@
+library(sn)
+
+load("Y")
+load("X")
+load("X_star")
+load("W")
+load("Z")
+load("SIGMA")
+load("PSI")
+load("BETA")
+load("DELTA")
+
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-09/")
+
+n <- nrow(Y)
+K <- length(BETA.list)
+J <- ncol(Y)
+p <- ncol(X)
+r <- ncol(W)
+S <- nrow(Z)
+
+lppd <- 0
+pwaic <- 0
+
+pb <- txtProgressBar(min = 0, max = n, style = 3)
+for(i in 1:n)
+{
+    y_i <- Y[i,]
+    x_i <- X[i,]
+    xstar_i <- Xstar[i,]
+    t_i <- xstar_i[p+1]
+    z_i <- Z[,i]
+    p_yi <- rep(0,S)
+    for(s in 1:S)
+    {
+        k <- z_i[s]
+        
+        beta_k <- BETA.list[[k]]
+        beta_k_s <- beta_k[s,]
+        beta_ks <- matrix(beta_k_s,
+                          nrow = p,
+                          ncol = J,
+                          byrow = TRUE)
+        
+        sigma_k <- SIGMA.list[[k]]
+        sigma_k_s <- sigma_k[s,]
+        sigma_ks <- matrix(sigma_k_s,
+                           nrow = J,
+                           ncol = J,
+                           byrow = TRUE)
+        
+        psi_k <- PSI.list[[k]]
+        psi_ks <- psi_k[s,]
+        
+        zeta_iks <- x_i %*% beta_ks
+        
+        mu_iks <- zeta_iks + t_i %*% psi_ks
+        
+        p_yi[s] <- dmvnorm(y_i,mu_iks,sigma_ks)
+    }
+    setTxtProgressBar(pb, i)
+    
+    print(lppd <- lppd + log(mean(p_yi)))
+    print(pwaic <- pwaic + var(log(p_yi)))
+}
+WAIC <- -2*(lppd - pwaic)
+
+close(pb)

### 2019-07-09 mcmc_draws_2019-07-09/mcmc_draws_NOSKEW_2019-07-09/WAIC.R

@@ -1,0 +1,55 @@
+library(sn)
+
+load("Y")
+load("X")
+load("Z")
+load("SIGMA")
+load("BETA")
+
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-09/mcmc_draws_NOSKEW_2019-07-09/")
+
+n <- nrow(Y)
+K <- length(BETA.list)
+J <- ncol(Y)
+p <- ncol(X)
+S <- nrow(Z)
+
+lppd <- 0
+pwaic <- 0
+
+pb <- txtProgressBar(min = 0, max = n, style = 3)
+for(i in 1:n)
+{
+    y_i <- Y[i,]
+    x_i <- X[i,]
+    z_i <- Z[,i]
+    p_yi <- rep(0,S)
+    for(s in 1:S)
+    {
+        k <- z_i[s]
+        
+        beta_k <- BETA.list[[k]]
+        beta_k_s <- beta_k[s,]
+        beta_ks <- matrix(beta_k_s,
+                          nrow = p,
+                          ncol = J,
+                          byrow = TRUE)
+        
+        sigma_k <- SIGMA.list[[k]]
+        sigma_k_s <- sigma_k[s,]
+        sigma_ks <- matrix(sigma_k_s,
+                           nrow = J,
+                           ncol = J,
+                           byrow = TRUE)
+        
+        zeta_iks <- x_i %*% beta_ks
+        p_yi[s] <- dmvnorm(y_i,zeta_iks,sigma_ks)
+    }
+    setTxtProgressBar(pb, i)
+    
+    print(lppd <- lppd + log(mean(p_yi, na.rm = TRUE)))
+    print(pwaic <- pwaic + var(log(p_yi)[is.finite(log(p_yi))], na.rm = TRUE))
+}
+WAIC <- -2*(lppd - pwaic)
+
+close(pb)

### 2019-07-09 mcmc_draws_2019-07-09_SIM1/WAIC.R

@@ -1,0 +1,64 @@
+library(sn)
+
+load("Y")
+load("X")
+load("X_star")
+load("W")
+load("Z")
+load("SIGMA")
+load("PSI")
+load("BETA")
+load("DELTA")
+
+n <- nrow(Y)
+K <- length(BETA.list)
+J <- ncol(Y)
+p <- ncol(X)
+r <- ncol(W)
+S <- nrow(Z)
+
+lppd <- 0
+pwaic <- 0
+
+pb <- txtProgressBar(min = 0, max = n, style = 3)
+for(i in 1:n)
+{
+    y_i <- Y[i,]
+    x_i <- X[i,]
+    z_i <- Z[,i]
+    p_yi <- rep(0,S)
+    for(s in 1:S)
+    {
+        k <- z_i[s]
+        
+        beta_k <- BETA.list[[k]]
+        beta_k_s <- beta_k[s,]
+        beta_ks <- matrix(beta_k_s,
+                          nrow = p,
+                          ncol = J,
+                          byrow = TRUE)
+        
+        sigma_k <- SIGMA.list[[k]]
+        sigma_k_s <- sigma_k[s,]
+        sigma_ks <- matrix(sigma_k_s,
+                           nrow = J,
+                           ncol = J,
+                           byrow = TRUE)
+        
+        psi_k <- PSI.list[[k]]
+        psi_ks <- psi_k[s,]
+        
+        zeta_iks <- x_i %*% beta_ks
+        
+        mu_iks <- zeta_iks + t_i %*% psi_ks
+        
+        p_yi[s] <- dmvnorm(y_i,mu_iks,sigma_ks)
+    }
+    setTxtProgressBar(pb, i)
+    
+    print(lppd <- lppd + log(mean(p_yi, na.rm = TRUE)))
+    print(pwaic <- pwaic + var(log(p_yi)[is.finite(log(p_yi))], na.rm = TRUE))
+}
+WAIC <- -2*(lppd - pwaic)
+
+close(pb)

### 2019-07-09 mcmc_draws_2019-07-09_SIM1/mcmc_draws_2019-07-09/WAIC.R

@@ -1,0 +1,66 @@
+library(sn)
+
+load("Y")
+load("X")
+load("X_star")
+load("W")
+load("Z")
+load("SIGMA")
+load("PSI")
+load("BETA")
+load("DELTA")
+
+n <- nrow(Y)
+K <- length(BETA.list)
+J <- ncol(Y)
+p <- ncol(X)
+r <- ncol(W)
+S <- nrow(Z)
+
+lppd <- 0
+pwaic <- 0
+
+pb <- txtProgressBar(min = 0, max = n, style = 3)
+for(i in 1:n)
+{
+    y_i <- Y[i,]
+    x_i <- X[i,]
+    xstar_i <- Xstar[i,]
+    t_i <- xstar_i[p+1]
+    z_i <- Z[,i]
+    p_yi <- rep(0,S)
+    for(s in 1:S)
+    {
+        k <- z_i[s]
+        
+        beta_k <- BETA.list[[k]]
+        beta_k_s <- beta_k[s,]
+        beta_ks <- matrix(beta_k_s,
+                          nrow = p,
+                          ncol = J,
+                          byrow = TRUE)
+        
+        sigma_k <- SIGMA.list[[k]]
+        sigma_k_s <- sigma_k[s,]
+        sigma_ks <- matrix(sigma_k_s,
+                           nrow = J,
+                           ncol = J,
+                           byrow = TRUE)
+        
+        psi_k <- PSI.list[[k]]
+        psi_ks <- psi_k[s,]
+        
+        zeta_iks <- x_i %*% beta_ks
+        
+        mu_iks <- zeta_iks + t_i %*% psi_ks
+        
+        p_yi[s] <- dmvnorm(y_i,mu_iks,sigma_ks)
+    }
+    setTxtProgressBar(pb, i)
+    
+    print(lppd <- lppd + log(mean(p_yi)))
+    print(pwaic <- pwaic + var(log(p_yi)))
+}
+WAIC <- -2*(lppd - pwaic)
+
+close(pb)

### 2019-07-09 mcmc_draws_2019-07-09_SIM1/mcmc_draws_NOSKEW_2019-07-09/WAIC.R

@@ -1,0 +1,54 @@
+library(sn)
+
+load("Y")
+load("X")
+load("Z")
+load("SIGMA")
+load("BETA")
+
+n <- nrow(Y)
+K <- length(BETA.list)
+J <- ncol(Y)
+p <- ncol(X)
+r <- ncol(W)
+S <- nrow(Z)
+
+lppd <- 0
+pwaic <- 0
+
+pb <- txtProgressBar(min = 0, max = n, style = 3)
+for(i in 1:n)
+{
+    y_i <- Y[i,]
+    x_i <- X[i,]
+    z_i <- Z[,i]
+    p_yi <- rep(0,S)
+    for(s in 1:S)
+    {
+        k <- z_i[s]
+        
+        beta_k <- BETA.list[[k]]
+        beta_k_s <- beta_k[s,]
+        beta_ks <- matrix(beta_k_s,
+                          nrow = p,
+                          ncol = J,
+                          byrow = TRUE)
+        
+        sigma_k <- SIGMA.list[[k]]
+        sigma_k_s <- sigma_k[s,]
+        sigma_ks <- matrix(sigma_k_s,
+                           nrow = J,
+                           ncol = J,
+                           byrow = TRUE)
+        
+        zeta_iks <- x_i %*% beta_ks
+        p_yi[s] <- dmvnorm(y_i,zeta_iks,sigma_ks)
+    }
+    setTxtProgressBar(pb, i)
+    
+    print(lppd <- lppd + log(mean(p_yi, na.rm = TRUE)))
+    print(pwaic <- pwaic + var(log(p_yi)[is.finite(log(p_yi))], na.rm = TRUE))
+}
+WAIC <- -2*(lppd - pwaic)
+
+close(pb)

### 2019-07-09 simulation.R

@@ -412,6 +412,7 @@ write(paste("Number of burn-in simulations (burn):",burn),file = meta_file, appe
 save(Y,file = paste(store,"/Y",sep = ""))
 save(Y.true,file = paste(store,"/Y_true",sep = ""))
 save(X,file = paste(store,"/X",sep = ""))
+save(Xstar,file = paste(store,"/X_star",sep = ""))
 save(Z,file = paste(store,"/Z",sep = ""))
 save(PI,file = paste(store,"/PI",sep = ""))
 save(DELTA,file = paste(store,"/DELTA",sep = ""))

### 2019-07-09 simulation_1/simulation_1.R

@@ -95,9 +95,9 @@ beta.true.list <- list(
            ncol = k)
     )
 psi.true.list <- list(
-    c(-2,-2,-2,-2),
-    c(1,-1,1,-1),
-    c(2,2,2,2)
+    c(0,0,0,0),
+    c(0,0,0,0),
+    c(0,0,0,0)
 )
 for(l in 1:h)
 {
@@ -348,6 +348,7 @@ write(paste("Number of burn-in simulations (burn):",burn),file = meta_file, appe
 
 save(Y,file = paste(store,"/Y",sep = ""))
 save(X,file = paste(store,"/X",sep = ""))
+save(Xstar,file = paste(store,"/Xstar",sep = ""))
 save(Z,file = paste(store,"/Z",sep = ""))
 save(PI,file = paste(store,"/PI",sep = ""))
 save(DELTA,file = paste(store,"/DELTA",sep = ""))

### 2019-07-09 manuscript/manuscript.tex
sim 1

@@
 \begin{eqnarray}
-	\mathbf{x}_i = 
+	\mathbf{X}_i = 
 	\begin{bmatrix}
@@
 For the multinomial regression model component of this simulation, we modeled the class labels $z_i$ as a function of an intercept and one baseline covariate, $\text{w}_{i1}$, implying that $r = 2$.
-We did introduce missing data into this simulation, as missing data will be dealt with in our second simulation study.
+We did introduce missing data into this simulation, as missing data will be dealt with in the second simulation study.
 As a result, the final sample size was $N = n \times J = 4,000$.
@@
 
-In our second simulation study, we compare our online conditional imputation approach to standard Bayesian multiple imputation as defined in Gelman et al.
-(2013).
-For an argument against the use of data-discarding approaches such as available case analysis (ACA) or complete case analysis (CCA), we defer to Gelman and Hill (2006).
-Our goal with this simulation is to demonstrate that our proposed online conditional imputation method performs as well as, if not better than standard Bayesian multiple imputation with respect to accuracy of parameter estimates.
-If the performance of the two imputation methods are held equal, we argue that our online conditional imputation method is preferable due to its computational efficiency and avoidance of a burdensome and often error-prone parameter aggregation step in the analysis.
+In the second simulation study, we compare our online conditional imputation approach to Bayesian multiple imputation, in which M (large) imputed data sets are generated, MCMC is performed on each imputed data set, and the data are pooled for final posterior inference.
+Details of this approach are nicely summarized in (Zhou and Reiter, 2010).
+Our goal here is to demonstrate that our proposed online conditional imputation method performs comparable to Bayesian multiple imputation with respect to accuracy of parameter estimates.
+If the performance of the two methods are qualitatively alike, we might argue that the online conditional imputation method is preferable due to its computational efficiency.
+Additionally, we compare both methods to the default analysis based on available cases, despite extensive literature cautioning against this approach (Gelman and Hill, 2006)
 
 To demonstrate the advantages of our proposed imputation method, we first generate $n = 1,000$ observations from a simple three-cluster ($K = 3$) MSN mixture model with $J = 4$ repeated measurements, one main effect for time ($p = 1$), and two multinomial regression predictors ($r = 2$).
-We then artificially ampute the simulated data using the \texttt{ampute} function provided by the \texttt{mice} package in \texttt{R} (Van Buuren and Groothuis-Oudshoorn 2010).
-Amputation occurs according to the missing at random (MAR) mechanism, whereby the occurence of missing data depends only on the values of the observed data (Gelman and Hill 2006).
-Additionally, we perform data amputation to simulateintermittent missingness, as is consistent with the Nurture data.
-To this end, we specify that each of the $4$ repeated measures for a given individual are equally likely to be missing.
+We then artificially ampute the simulated data using the \texttt{ampute} function provided by the \texttt{mice} package in \texttt{R} (Van Buuren and Groothuis-Oudshoorn 2010) according to the missing at random (MAR) mechanism, whereby the occurence of missing data depends only on the values of the observed data (Gelman and Hill 2006).
+We specified that each of the $4$ repeated measures for a given individual were equally likely to be missing.
 To further approximate the Nurture data, we set the frequency of missing data at each measurement occasion to be 30\%.
-We fit our proposed MSN finite mixture model to the amputed data with an implementation of both our proposed online conditional imputation step and standard Bayesian multiple imputation as defined by Gelman et al.
-(2013).
+We then fit our proposed MSN finite mixture model to the amputed data using both online conditional imputation and Bayesian multiple imputation as defined by (Zhou and Reiter, 2010).
 

### 2019-07-12 mcmc_draws_2019-07-10/tables_2.Rmd
edits

--- 2019-07-09 mcmc_draws_2019-07-08/tables_2.Rmd
@@
 ```{r}
-setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-09")
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-10")
 load("BETA")
@@
 rownames(t1) <- NULL
-kable(t1, booktabs = T, escape = F,caption = "Model results for simulated data with n = 1000, J = 4, p = 2, K = 3, r = 2.
+kable(t1, booktabs = T, escape = F,caption = "Model results for simulated data with n = 1,000, J = 4, p = 2, K = 3, r = 2.
 1,000 iterations were run with a burn in of 0.

### 2019-07-12 figures/figures_mcmc.R

@@ -7,7 +7,7 @@ library(vapoRwave)
 vaporwave <- vapoRwave:::jazzCup_palette
 
 setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM")
-load_dir <- "mcmc_draws_2019-07-09"
+load_dir <- "mcmc_draws_2019-07-11"
 load(paste(load_dir,"/BETA",sep = ""))
 load(paste(load_dir,"/beta_true",sep = ""))
 load(paste(load_dir,"/DELTA",sep = ""))

### 2019-07-12 mcmc_draws_2019-07-09_SIM1/WAIC.R

@@ -1,8 +1,9 @@
 library(sn)
 
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-09_SIM1")
 load("Y")
 load("X")
-load("X_star")
+load("Xstar")
 load("W")
 load("Z")
 load("SIGMA")
@@ -36,7 +37,7 @@ for(i in 1:n)
         beta_ks <- matrix(beta_k_s,
                           nrow = p,
                           ncol = J,
-                          byrow = TRUE)
+                          byrow = FALSE)
         
         sigma_k <- SIGMA.list[[k]]
         sigma_k_s <- sigma_k[s,]
@@ -57,7 +58,7 @@ for(i in 1:n)
     setTxtProgressBar(pb, i)
     
     print(lppd <- lppd + log(mean(p_yi, na.rm = TRUE)))
-    print(pwaic <- pwaic + var(log(p_yi)[is.finite(log(p_yi))], na.rm = TRUE))
+    print(pwaic <- pwaic + var(log(p_yi), na.rm = TRUE))
 }
 WAIC <- -2*(lppd - pwaic)
 

### 2019-07-12 mcmc_draws_2019-07-09_SIM1/mcmc_draws_2019-07-09/WAIC.R

@@ -1,5 +1,6 @@
 library(sn)
 
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-09_SIM1/mcmc_draws_2019-07-09/")
 load("Y")
 load("X")
 load("X_star")
@@ -9,6 +10,8 @@ load("SIGMA")
 load("PSI")
 load("BETA")
 load("DELTA")
+
+
 
 n <- nrow(Y)
 K <- length(BETA.list)
@@ -38,7 +41,7 @@ for(i in 1:n)
         beta_ks <- matrix(beta_k_s,
                           nrow = p,
                           ncol = J,
-                          byrow = TRUE)
+                          byrow = FALSE)
         
         sigma_k <- SIGMA.list[[k]]
         sigma_k_s <- sigma_k[s,]

### 2019-07-12 mcmc_draws_2019-07-10/WAIC.R

@@ -1,11 +1,15 @@
 library(sn)
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-10")
 
-setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-09_SIM1/mcmc_draws_NOSKEW_2019-07-09")
 load("Y")
 load("X")
+load("Xstar")
+load("W")
 load("Z")
 load("SIGMA")
+load("PSI")
 load("BETA")
+load("DELTA")
 
 n <- nrow(Y)
 K <- length(BETA.list)
@@ -23,6 +27,7 @@ for(i in 1:n)
     y_i <- Y[i,]
     x_i <- X[i,]
     z_i <- Z[,i]
+    t_i <- Xstar[i,p+1]
     p_yi <- rep(0,S)
     for(s in 1:S)
     {
@@ -42,8 +47,14 @@ for(i in 1:n)
                            ncol = J,
                            byrow = TRUE)
         
+        psi_k <- PSI.list[[k]]
+        psi_ks <- psi_k[s,]
+        
         zeta_iks <- x_i %*% beta_ks
-        p_yi[s] <- dmvnorm(y_i,zeta_iks,sigma_ks)
+        
+        mu_iks <- zeta_iks + t_i %*% psi_ks
+        
+        p_yi[s] <- dmvnorm(y_i,mu_iks,sigma_ks)
     }
     setTxtProgressBar(pb, i)
     
@@ -51,5 +62,6 @@ for(i in 1:n)
     print(pwaic <- pwaic + var(log(p_yi), na.rm = TRUE))
 }
 WAIC <- -2*(lppd - pwaic)
+save(WAIC,file = "WAIC")
 
 close(pb)

### 2019-07-12 mcmc_draws_2019-07-11/WAIC.R

@@ -1,11 +1,15 @@
 library(sn)
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-11")
 
-setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-09_SIM1/mcmc_draws_NOSKEW_2019-07-09")
 load("Y")
 load("X")
+load("Xstar")
+load("W")
 load("Z")
 load("SIGMA")
+load("PSI")
 load("BETA")
+load("DELTA")
 
 n <- nrow(Y)
 K <- length(BETA.list)
@@ -23,6 +27,7 @@ for(i in 1:n)
     y_i <- Y[i,]
     x_i <- X[i,]
     z_i <- Z[,i]
+    t_i <- Xstar[i,p+1]
     p_yi <- rep(0,S)
     for(s in 1:S)
     {
@@ -42,8 +47,14 @@ for(i in 1:n)
                            ncol = J,
                            byrow = TRUE)
         
+        psi_k <- PSI.list[[k]]
+        psi_ks <- psi_k[s,]
+        
         zeta_iks <- x_i %*% beta_ks
-        p_yi[s] <- dmvnorm(y_i,zeta_iks,sigma_ks)
+        
+        mu_iks <- zeta_iks + t_i %*% psi_ks
+        
+        p_yi[s] <- dmvnorm(y_i,mu_iks,sigma_ks)
     }
     setTxtProgressBar(pb, i)
     
@@ -51,5 +62,6 @@ for(i in 1:n)
     print(pwaic <- pwaic + var(log(p_yi), na.rm = TRUE))
 }
 WAIC <- -2*(lppd - pwaic)
+save(WAIC,file = "WAIC")
 
 close(pb)

### 2019-07-12 mcmc_draws_NOSKEW_2019-07-11/WAIC.R

@@ -1,11 +1,15 @@
 library(sn)
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_NOSKEW_2019-07-11/")
 
-setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-09_SIM1/mcmc_draws_NOSKEW_2019-07-09")
 load("Y")
 load("X")
+load("Xstar")
+load("W")
 load("Z")
 load("SIGMA")
+load("PSI")
 load("BETA")
+load("DELTA")
 
 n <- nrow(Y)
 K <- length(BETA.list)
@@ -23,6 +27,7 @@ for(i in 1:n)
     y_i <- Y[i,]
     x_i <- X[i,]
     z_i <- Z[,i]
+    t_i <- Xstar[i,p+1]
     p_yi <- rep(0,S)
     for(s in 1:S)
     {
@@ -42,8 +47,14 @@ for(i in 1:n)
                            ncol = J,
                            byrow = TRUE)
         
+        psi_k <- PSI.list[[k]]
+        psi_ks <- psi_k[s,]
+        
         zeta_iks <- x_i %*% beta_ks
-        p_yi[s] <- dmvnorm(y_i,zeta_iks,sigma_ks)
+        
+        mu_iks <- zeta_iks + t_i %*% psi_ks
+        
+        p_yi[s] <- dmvnorm(y_i,mu_iks,sigma_ks)
     }
     setTxtProgressBar(pb, i)
     
@@ -51,5 +62,6 @@ for(i in 1:n)
     print(pwaic <- pwaic + var(log(p_yi), na.rm = TRUE))
 }
 WAIC <- -2*(lppd - pwaic)
+save(WAIC,file = "WAIC")
 
 close(pb)

### 2019-07-12 simulation_1/simulation_1_MSN.R

@@ -1,19 +1,15 @@
-# Multivariate normal regression with clustering
+# Multivariate skew normal regression with clustering
 # Includes built-in mvn conditional imputation
 # Includes regression on multinomial regression coefficients
-# Used for comparison w/ skew norm
 # Carter Allen
-# March 27, 2019
+# March 19, 2019
 
 # Packages
 library(sn) # for rsn
 library(truncnorm) # for rtruncnorm
 library(mvtnorm) # for rmvtnorm
-library(ggplot2)
-library(patchwork)
 library(mvnmle)
 library(Hmisc) # For rMultinom function (easy sampling of class indicators C)
-library(bayesplot)
 library(coda)
 library(MCMCpack)
 library(matrixsampling)
@@ -21,28 +17,30 @@ library(QRM)        # For rGumbel function
 library(nnet)       # To fit multinom function
 library(BayesLogit) # For rpg function
 
+#seed <- rnorm(1)
+#set.seed(seed)
 # No scientific notation
 options(scipen=999)
-setwd("/Users/carter/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-10/")
-
+options(warn=2)
 # Generate data
 k <- 4 # number of outcomes
-p <- 2 # number of predictors (coincluding intercept)
+p <- 2 # number of predictors 
 h <- 3 # number of clusters
 n <- 1000 # number of observations (takes alot of data to estimate this many parameters)
-    print(paste("Reading data with",n,"observations of dimension",k,
+print(paste("Generating data with",n,"observations of dimension",k,
             "with",p,"covariates and",h,"clusters"))
 
 N <- n*k # number of measurments 
 P <- matrix(rep(rep(1/h,h),each=n),n,h) # probability matrix governing cluster membership
 
 # Generate covariates
-# X <- matrix(rnorm(n*p),nrow = n,ncol = p) # random predictors
-load("X")
+X <- matrix(rnorm(n*p),nrow = n,ncol = p) # random predictors
+X[,1] <- 1
 t <- rtruncnorm(n,0,Inf,0,1) # trunc norm random effects
 Xstar <- cbind(X,t) # Combine X and t
 
-load("W")
[…]

### 2019-07-12 simulation_1/simulation_1_MVN.R

@@ -79,25 +79,28 @@ L0.list <- list(0)
 B0.list <- list(0)
 Y <- NULL
 
-sig2.true.ar1 <- 1 * 0.5^(abs(outer(1:k,1:k,"-")))
+sig2.true.ar1 <- 5 * 0.5^(abs(outer(1:k,1:k,"-")))
 beta.true.list <- list(
-    matrix(c(11,12,13,14,
-             2,2,2,2),
+    matrix(c(100,105,110,115,
+             -1,-1.5,-2,-2.5),
            nrow = p,
-           ncol = k),
-    matrix(c(-5,-4,-3,-2,
-             5,5,5,5),
+           ncol = k,
+           byrow = TRUE),
+    matrix(c(90,85,80,75,
+             1,1.5,2,2.5),
            nrow = p,
-           ncol = k),
-    matrix(c(-10,-11,-12,-13,
-             -2,-2,-2,-2),
+           ncol = k,
+           byrow = TRUE),
+    matrix(c(100,100,100,100,
+             1,1,1,1),
            nrow = p,
-           ncol = k)
+           ncol = k,
+           byrow = TRUE)
     )
 psi.true.list <- list(
-    c(0,0,0,0),
-    c(0,0,0,0),
-    c(0,0,0,0)
+    c(1,2,3,4),
+    c(-1,-2,-3,-4),
+    c(0,1,0,1)
 )
 for(l in 1:h)
 {
@@ -109,6 +112,7 @@ for(l in 1:h)
     beta.true.l <- beta.true.list[[l]]
     psi.true.l <- psi.true.list[[l]]
     betastar.true.l <- rbind(beta.true.l,psi.true.l)
+    betastar.true.list[[l]] <- betastar.true.l
     
     sig2.true.list[[l]] <- sig2.true.l <- sig2.true.ar1 # Add to list of true class-specific sig2 matrices
 
@@ -298,7 +302,8 @@ for(i in 1:nsim)
         betastar.l <- rmatrixnormal(1,Bn.l,solve(Ln.l), sig2.l, checkSymmetry = FALSE)
         betastar.l <- matrix(betastar.l,nrow = p+1, ncol = k)
         beta.l <- betastar.l[1:p,]
-        psi.l <- betastar.l[p+1,]
+        # psi.l <- betastar.l[p+1,]
+        psi.l <- rep(0,k)
         beta.list[[l]] <- beta.l
         psi.list[[l]] <- psi.l
     }
[…]

### 2019-07-23 mcmc_draws_2019-07-11/mcmc_draws_NOSKEW_2019-07-11/tables.Rmd
edits

--- 2019-07-09 mcmc_draws_2019-07-09_SIM1/mcmc_draws_NOSKEW_2019-07-09/tables.Rmd
@@
 ```{r}
-setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-09/mcmc_draws_NOSKEW_2019-07-09/")
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-11/mcmc_draws_NOSKEW_2019-07-11/")
 load("BETA")

### 2019-07-23 mcmc_draws_2019-07-11/tables.Rmd
edits

--- 2019-07-09 mcmc_draws_2019-07-09_SIM1/tables.Rmd
@@
 ```{r}
-setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-09")
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-11")
 load("BETA")
@@
 
-load("mcmc_draws_NOSKEW_2019-07-09/class_1_est")
-load("mcmc_draws_NOSKEW_2019-07-09/class_2_est")
-load("mcmc_draws_NOSKEW_2019-07-09/class_3_est")
+load("mcmc_draws_NOSKEW_2019-07-11/class_1_est")
+load("mcmc_draws_NOSKEW_2019-07-11/class_2_est")
+load("mcmc_draws_NOSKEW_2019-07-11/class_3_est")
 
@@
                   "Param.",
-                  "True","MSN Est. (95\\% CrI)","MN Est. (95\\% CrI) ",
-                  "True","MSN Est. (95\\% CrI)","MN Est. (95\\% CrI)",
-                  "True","MSN Est. (95\\% CrI)","MN Est. (95\\% CrI)")
+                  "True","MSN Est. (95\\% CrI)","MVN Est. (95\\% CrI) ",
+                  "True","MSN Est. (95\\% CrI)","MVN Est. (95\\% CrI)",
+                  "True","MSN Est. (95\\% CrI)","MVN Est. (95\\% CrI)")
 ```
@@
 rownames(t1) <- NULL
-kable(t1, booktabs = T, escape = F,caption = "Model results for simulated data with n = 1,000, J = 4, p = 2, K = 3, r = 2.
+kable(t1, booktabs = T, escape = F,caption = "Model results for simulated data with $n = 1,000$, $J = 4$, $p = 2$, $K = 3$, $r = 2$.
 1,000 iterations were run with a burn in of 100.
 Missingness mechanism was MAR and P(miss) = 0.
-Model results for the multivariate skew normal (MSN) and multivariate normal (MN) mixtures are presented.") %>%
+Model results for the multivariate skew normal (MSN) and multivariate normal (MVN) mixtures are presented.") %>%
   kable_styling(font_size = 7) %>%

### 2019-07-23 figures/figures_mcmc.R

@@ -7,7 +7,7 @@ library(vapoRwave)
 vaporwave <- vapoRwave:::jazzCup_palette
 
 setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM")
-load_dir <- "mcmc_draws_2019-07-11"
+load_dir <- "mcmc_draws_2019-07-12"
 load(paste(load_dir,"/BETA",sep = ""))
 load(paste(load_dir,"/beta_true",sep = ""))
 load(paste(load_dir,"/DELTA",sep = ""))
@@ -113,7 +113,7 @@ y_mean_plot <- data.frame(Y,c) %>%
     rename(clust = c) %>%
     gather(variable,value,-clust) %>%
     group_by(variable,clust) %>%
-    summarize(mean_y = mean(value)) %>%
+    dplyr::summarize(mean_y = mean(value)) %>%
     ungroup() %>% 
     mutate(clust = as.factor(clust)) %>%
     ggplot(.,aes(x = variable, 

### 2019-07-23 mcmc_draws_2019-07-11/mcmc_draws_NOSKEW_2019-07-11/WAIC.R

@@ -1,0 +1,67 @@
+library(sn)
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_NOSKEW_2019-07-11/")
+
+load("Y")
+load("X")
+load("Xstar")
+load("W")
+load("Z")
+load("SIGMA")
+load("PSI")
+load("BETA")
+load("DELTA")
+
+n <- nrow(Y)
+K <- length(BETA.list)
+J <- ncol(Y)
+p <- ncol(X)
+r <- ncol(W)
+S <- nrow(Z)
+
+lppd <- 0
+pwaic <- 0
+
+pb <- txtProgressBar(min = 0, max = n, style = 3)
+for(i in 1:n)
+{
+    y_i <- Y[i,]
+    x_i <- X[i,]
+    z_i <- Z[,i]
+    t_i <- Xstar[i,p+1]
+    p_yi <- rep(0,S)
+    for(s in 1:S)
+    {
+        k <- z_i[s]
+        
+        beta_k <- BETA.list[[k]]
+        beta_k_s <- beta_k[s,]
+        beta_ks <- matrix(beta_k_s,
+                          nrow = p,
+                          ncol = J,
+                          byrow = FALSE)
+        
+        sigma_k <- SIGMA.list[[k]]
+        sigma_k_s <- sigma_k[s,]
+        sigma_ks <- matrix(sigma_k_s,
+                           nrow = J,
+                           ncol = J,
+                           byrow = TRUE)
+        
+        psi_k <- PSI.list[[k]]
+        psi_ks <- psi_k[s,]
+        
+        zeta_iks <- x_i %*% beta_ks
+        
+        mu_iks <- zeta_iks + t_i %*% psi_ks
+        
+        p_yi[s] <- dmvnorm(y_i,mu_iks,sigma_ks)
+    }
+    setTxtProgressBar(pb, i)
+    
+    print(lppd <- lppd + log(mean(p_yi, na.rm = TRUE)))
+    print(pwaic <- pwaic + var(log(p_yi), na.rm = TRUE))
+}
+WAIC <- -2*(lppd - pwaic)
+save(WAIC,file = "WAIC")
+
+close(pb)

### 2019-07-23 simulation_1/simulation.R

@@ -1,0 +1,432 @@
+# Multivariate skew normal regression with clustering
+# Includes built-in mvn conditional imputation
+# Includes regression on multinomial regression coefficients
+# Carter Allen
+# March 19, 2019
+
+# Packages
+library(sn) # for rsn
+library(truncnorm) # for rtruncnorm
+library(mvtnorm) # for rmvtnorm
+library(ggplot2)
+library(patchwork)
+library(mvnmle)
+library(Hmisc) # For rMultinom function (easy sampling of class indicators C)
+library(bayesplot)
+library(coda)
+library(MCMCpack)
+library(matrixsampling)
+library(QRM)        # For rGumbel function
+library(nnet)       # To fit multinom function
+library(BayesLogit) # For rpg function
+
+#seed <- rnorm(1)
+#set.seed(seed)
+# No scientific notation
+options(scipen=999)
+options(warn=2)
+# Generate data
+k <- 4 # number of outcomes
+p <- 2 # number of predictors 
+h <- 3 # number of clusters
+n <- 1000 # number of observations (takes alot of data to estimate this many parameters)
+print(paste("Generating data with",n,"observations of dimension",k,
+            "with",p,"covariates and",h,"clusters"))
+
+N <- n*k # number of measurments 
+P <- matrix(rep(rep(1/h,h),each=n),n,h) # probability matrix governing cluster membership
+
+# Generate covariates
+X <- matrix(rnorm(n*p),nrow = n,ncol = p) # random predictors
+X[,1] <- 1
+t <- rtruncnorm(n,0,Inf,0,1) # trunc norm random effects
+Xstar <- cbind(X,t) # Combine X and t
+# # Generate X manually
+# xi <- matrix(c(1, 0, 0, 0,
+#                1, 1, 0, 0,
+#                1, 0, 1, 0,
+#                1, 0, 0, 1),
+#              nrow = 4,
+#              ncol = 4,
+#              byrow = TRUE)
+# t <- rtruncnorm(n,0,Inf,0,1) # trunc norm random effects
+# X <- NULL 
+# capT <- NULL
+# for(i in 1:n)
+# {
+#     X <- rbind(X,xi)
+#     capT <- rbind(capT,diag(t[i],nrow = 4,ncol = 4))
+# }
+# 
+# Xstar <- cbind(X,capT) # Combine X and t
+
+# Covariates corresponding to multinomial regression
[…]

### 2019-07-23 simulation_1/simulation_1_MSN.R

@@ -79,27 +79,27 @@ L0.list <- list(0)
 B0.list <- list(0)
 Y <- NULL
 
-sig2.true.ar1 <- 5 * 0.5^(abs(outer(1:k,1:k,"-")))
+sig2.true.ar1 <- 0.5^(abs(outer(1:k,1:k,"-")))
 beta.true.list <- list(
-    matrix(c(100,105,110,115,
-             -1,-1.5,-2,-2.5),
-           nrow = p,
-           ncol = k,
-           byrow = TRUE),
-    matrix(c(90,85,80,75,
+    matrix(c(110,115,120,125,
              1,1.5,2,2.5),
            nrow = p,
            ncol = k,
            byrow = TRUE),
+    matrix(c(90,85,80,75,
+             -1,-1.5,-2,-2.5),
+           nrow = p,
+           ncol = k,
+           byrow = TRUE),
     matrix(c(100,100,100,100,
-             1,1,1,1),
+             -1,1,-1,1),
            nrow = p,
            ncol = k,
            byrow = TRUE)
     )
 psi.true.list <- list(
-    c(1,2,3,4),
-    c(-1,-2,-3,-4),
+    c(2,3,4,5),
+    c(-2,-3,-4,-5),
     c(0,1,0,1)
 )
 for(l in 1:h)

### 2019-07-30 manuscript/manuscript.tex
edits

@@
 A ubiquitous feature of repeated measures studies is loss of data due to intermittent missingness and attrition.
-In the Bayesian setting, the standard approach to dealing with missing data is to perform multiple imputation, whereby $m$ imputed data sets are generated from a specified imputation model.
-After $m$ complete data sets are obtained, parameter estimates are combined across each data set to produce a final set of parameter estimates (Gelman \textit{et al.} 2013).
-This approach is not only computationally burdensome, requiring storage and analysis of an $m \times n_{rows} \times n_{cols}$ data array in addition to multiplication of total model run time by a factor of $m$, but it has been shown to produce unreliable inferences (Zhou and Reiter, 2010).
+In the Bayesian setting, the standard approach to dealing with missing data is to perform multiple imputation, whereby $M$ imputed data sets are generated from a specified imputation model.
+After $M$ complete data sets are obtained, parameter estimates are combined across each data set to produce a final set of parameter estimates (Gelman \textit{et al.} 2013).
+This approach is not only computationally burdensome, requiring storage and analysis of an $M \times n_{rows} \times n_{cols}$ data array in addition to multiplication of total model run time by a factor of $M$, but it has been shown to produce unreliable inferences (Zhou and Reiter, 2010).
 We instead include an ``online" imputation imputation step in our Gibbs sampling procedure, whereby missing outcomes are updated at each iteration.
@@
 \boldsymbol\alpha_k &=& \frac{1}{\sqrt{1 - \boldsymbol\psi_k^T 
-\boldsymbol\Omega^{-1}_k\boldsymbol\psi_k}} \boldsymbol\omega_k \boldsymbol\Omega^{-1}_k\boldsymbol\psi_k,~~\text{and} \label{eqn:back_transform}\\
+\boldsymbol\Omega^{-1}_k\boldsymbol\psi_k}} \boldsymbol\omega_k \boldsymbol\Omega^{-1}_k\boldsymbol\psi_k,~~\text{and} \label{eq:back_transform}\\
 \boldsymbol\Omega_k &=& \boldsymbol\Sigma_k  + \boldsymbol\psi_k \boldsymbol\psi_k^T \nonumber,
@@
 Of note, the MSN density can be expressed more compactly in terms of the matrix skew normal (MatSN) density (Chen and Gupta 2005).
-Let $\mathbf{Y}_k$ be an ${n_k \times J}$ response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n \mathds{1}_{(z_i = k)}$ is the number of observations in cluster $k$.
+For cluster $k$, let $\mathbf{Y}_k$ be an ${n_k \times J}$ response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n \mathds{1}_{(z_i = k)}$ is the number of observations in cluster $k$.
 From equation (\ref{eq:msnreg}), it follows that $\mathbf{Y}_k$ is distributed as
@@
 \begin{equation}
+\label{eq:matnormal}
 \mathbf{Y}_k | \mathbf{t}_k \sim \text{MatNorm}_{n_k \times J}(\mathbf{M}^*_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k),
@@
 By allowing the cluster probabilities to vary across subjects, model (\ref{eq:mixture}) can be viewed as a \textit{mixture of experts} model, in which $\pi_{ki}$ acts as a \textit{gating function} controlling the prior probability of membership in cluster $k$, and $f(\mathbf{y}_i|\boldsymbol\theta_k)$ in equation (\ref{eq:mixture}) is the ``expert'' providing information on the within-cluster distribution of $\mathbf{y}_i$ (Bishop 2006).
+An appealing feature of the mixture of experts model is the ability to discern cluster membership using information contained in the covariates $\mathbf{w}_i$.
+That is, the gating functions make precise allocations possible even in the case when the density $f(\mathbf{y}_i | \boldsymbol\theta_k)$ is similar across clusters.
+In addition, modeling $\boldsymbol\pi_i$ as a function of covariates of interest allows for more robust interpretations of cluster allocations and enhances the practical interpretability of the mixture of experts model.
 
@@
 where $\kappa = a - b/2$ and $p(w|b,c = 0)$ denotes a $\text{PG}(b,0)$ density.
-Polson {et al.} further show that the conditional distribution $p(w|b,c)$ results from an ``exponential tilting'' of the $\text{PG}(b,0)$ density, thus
+Polson {et al.} further show that the conditional distribution $p(w|b,c)$ results from an ``exponential tilting'' of the $\text{PG}(b,0)$ density:
 
@@
 use these results to show that, for the logistic model, the Bernoulli likelihood can be written as a scale-mixture of normal densities with Polya-Gamma precision terms, $w$, resulting in a normal full conditional distribution for the logistic regression parameters.
-To extend the P\'olya--Gamma data augmentation approach to the multinomial setting, we first introduce the binary indicators $U_{ki}$, such that $U_{ki} = \mathds{1}_{(z_i = k)}$ -- i.e., $U_{ki}$ is an indicator variable taking the value $1$ if subject $i$ belongs to cluster $k$, and $0$ otherwise.
-The full conditional distribution of $\boldsymbol\delta_k$ is then given as  $$p(\boldsymbol\delta_k|\mathbf{U}_k,\boldsymbol\delta_{h \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \pi_{ki}^{U_{ki}}(1-\pi_{ki})^{1-U_{ki}},$$
+To extend the P\'olya--Gamma data augmentation approach to the multinomial setting, we first introduce the binary indicators $U_{ki}$, such that $U_{ki} = \mathds{1}_{(z_i = k)}$ --- i.e., $U_{ki}$ is an indicator variable taking the value $1$ if subject $i$ belongs to cluster $k$, and $0$ otherwise.
+The full conditional distribution of $\boldsymbol\delta_k$ conditional on $\mathbf{U}_k = (U_{k1},...,U_{kn})^T$ and the remaining regression coefficients $\delta_h$ ($h \ne k$) is then given as  $$p(\boldsymbol\delta_k|\mathbf{U}_k,\boldsymbol\delta_{h \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \pi_{ki}^{U_{ki}}(1-\pi_{ki})^{1-U_{ki}},$$
 where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, and $\pi_{ki}$ is defined as in equation (\ref{eq:pik}). We can rewrite $\pi_{ki}$ in terms of $U_{ki}$ as
-$$\pi_{ki} = P(U_{ki} = 1) = \frac{\text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ki}}}{1 + \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ki}}} = \frac{\text{e}^{\eta_{ki}}}{1 + \text{e}^{\eta_{ki}}},$$
+$$\pi_{ki} = \Pr(U_{ki} = 1) = \frac{\text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ki}}}{1 + \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ki}}} = \frac{\text{e}^{\eta_{ki}}}{1 + \text{e}^{\eta_{ki}}},$$
 where ${c}_{ki} = \log \sum_{h \ne k} \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}$ and $\eta_{ki} = \mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ki}$.
@@
 \begin{equation}
-	p(\boldsymbol\delta_k|\mathbf{z},\mathbf{W}, \mathbf{O}_k) \propto p(\boldsymbol\delta_k) \text{exp} \left [ - \frac{1}{2} (\mathbf{U}^*_k - \mathbf{W}\boldsymbol\delta_k)^T \mathbf{O}_k (\mathbf{U}^*_k - \mathbf{W}\boldsymbol\delta_k)\right].
+	p(\boldsymbol\delta_k|\mathbf{z},\mathbf{W}, \mathbf{O}_k) \propto p(\boldsymbol\delta_k) \text{exp} \left \{ - \frac{1}{2} (\mathbf{U}^*_k - \mathbf{W}\boldsymbol\delta_k)^T \mathbf{O}_k (\mathbf{U}^*_k - \mathbf{W}\boldsymbol\delta_k)\right \}.
 \end{equation}
@@
 \subsection{Conditional MSN Imputation}
-To accommodate intermittent missing at random (MAR) responses, we propose a convenient imputation algorithm that can be implemented ``online'' -- that is, as part of the Gibbs sampler.
+To accommodate intermittent missing at random (MAR) responses, we propose a convenient imputation algorithm that can be implemented ``online'' --- that is, as part of the Gibbs sampler.
 In Section 6, we discuss extensions to allow for non-ignorable missingness (i.e., observations missing not at random).
@@
 Conveniently, all parameters admit conditionally conjugate priors, which greatly improves posterior computation via a data-augmented Gibbs sampler.
-To make use of the matrix normal representation introduced previously in Section 3, we define the matrix of regression parameters $\mathbb{B}^*_k$, where $\text{vec}(\mathbf{B}^*_k) = \boldsymbol\beta^*_k = (\boldsymbol\beta_k^T,\boldsymbol\psi_k^T)^T$.
+To make use of the matrix normal representation introduced previously in Section 3, we define the $J \times (p+1)$ matrix of regression parameters $\mathbf{B}^*_k$, where $\text{vec}(\mathbf{B}^*_k) = \boldsymbol\beta^*_k = (\boldsymbol\beta_k^T,\boldsymbol\psi_k^T)^T$.
 We assign $\mathbf{B}^*_k|\boldsymbol\Sigma_k$ a $\text{MatNorm}(\mathbf{B}^*_{0k},\mathbf{I}_{p+1},\boldsymbol\Sigma_k)$ prior, where $\mathbf{B}^*_{0k}$ is a matrix of location parameters such that $\text{vec}(\mathbf{B}^*_{0k})$ is a vector of prior location parameters for the components of $\boldsymbol\beta^*_k$, $\mathbf{I}_{p+1}$ is the $(p+1)$-dimensional identity matrix, and $\boldsymbol\Sigma_k$ is the covariance matrix defined in equation (\ref{eq:msndens}), for which we specify an $\text{IW}(\mathbf{V}_{0k},\nu_{0k})$ prior.
-This leads to a matrix-normal-inverse-Wishart joint prior for $\mathbb{B}^*_k$ and $\boldsymbol\Sigma_k$, which is the conjugate joint prior for the regression parameters in the matrix normal model given in equation (\ref{eq:matsn}).
+This leads to a matrix-normal-inverse-Wishart joint prior for $\mathbf{B}^*_k$ and $\boldsymbol\Sigma_k$, which is the conjugate joint prior for the regression parameters in the matrix normal model given in equation (\ref{eq:matsn}).
 This conjugate prior specification induces convenient closed-form full conditional distributions that can be easily updated within our proposed Gibbs sampler.
@@
 \psi_k)^{-1}$ using current values of $\boldsymbol\psi_k$ and $\boldsymbol\Sigma_k$.
-Next, for $i = 1,...,n_k$, we compute $a_i = A_k \boldsymbol\psi_k^T \boldsymbol\Sigma^{-1}_k(\mathbf{y}_i - \boldsymbol\zeta_{ki})$, where $\boldsymbol\zeta_{ki} = \mathbf{X}_i \boldsymbol\beta_k$.
-Finally, we update $t_i$ from $\text{N}_{[0,\infty)}(a_i, \mathbf{A}_k)$.
-Repeat these updates for $k = 1,...,K$.
+Next, for $i = 1,...,n_k$ in cluster $k$, we compute $a_{ki} = A_k \boldsymbol\psi_k^T \boldsymbol\Sigma^{-1}_k(\mathbf{y}_i - \boldsymbol\zeta_{ki})$, where $\boldsymbol\zeta_{ki} = \mathbf{X}_i \boldsymbol\beta_k$.
+Finally, we update $t_i|(z_i = k)$ from $\text{N}_{[0,\infty)}(a_{ki}, \mathbf{A}_k)$.
+We repeat these updates for $k = 1,...,K$.
 
@@
 For each cluster $k$ ($k = 1,...,K$), we update $\boldsymbol\Sigma_k$ from an $\text{IW}(\nu_k,\mathbf{V}_k)$ density, where $\nu_k = \nu_{0k} + n_k$ and\begin{eqnarray}
-	\mathbf{V}_k &=&  (\mathbf{Y}_k-\mathbf{X}^*_k \mathbf{B}^*_k)^T (\mathbf{Y}_k - \mathbf{X}^*_k \mathbf{B}^*_k) + (\mathbf{B}_k^* - \mathbf{B}^*_{0k})^T (\mathbf{B}^*_k - \mathbf{B}^*_{0k}) + \mathbf{V}_{0k}.
-\nonumber
-\end{eqnarray} 
+	\mathbf{V}_k &=&  (\mathbf{Y}_k-\mathbf{M}^*_k)^T (\mathbf{Y}_k - \mathbf{M}^*_k) + (\mathbf{B}_k^* - \mathbf{B}^*_{0k})^T (\mathbf{B}^*_k - \mathbf{B}^*_{0k}) + \mathbf{V}_{0k}, \nonumber
+\end{eqnarray} where $\mathbf{M}^*_k$ is defined as in equation (\ref{eq:matnormal})
 We then make use of the matrix normal representation introduced in Section 3.2 to draw $\mathbf{B}^*_k$ from a $\text{MatNorm}_{p+1,J}(\mathbb{B}_k^*,\mathbf{L}_k^{*},\boldsymbol\Sigma_k)$ density, where
@@
 For efficient sampling of the matrix normal distribution, we use the \texttt{R} package \texttt{matrixsampling} (Laurent 2018).
+Taking $\text{vec}(\mathbf{B}^*_k)$ yields a draw from the joint posterior of $\boldsymbol\beta_k$ and $\boldsymbol\psi_k$, and in turn $\boldsymbol\beta_k$, $\boldsymbol\alpha_k$, and $\boldsymbol\Omega_k$ via the back-transformations given in equation (\ref{eq:back_transform}).
 
-\paragraph{\textbf{Step 3:}P\'olya--Gamma Data Augmentation for $z_i$} The sampler concludes with updates of the multinomial regression parameters $\boldsymbol\delta_k$, for $k = 1,...,K-1$, followed by updates of each latent cluster indicator $z_i$ $(i=1,\ldots,n)$ from its multinomial logit full conditional.
+\paragraph{\textbf{Step 3:} P\'olya--Gamma Data Augmentation for $z_i$} The sampler concludes with updates of the multinomial regression parameters $\boldsymbol\delta_k$, for $k = 1,...,K-1$, followed by updates of each latent cluster indicator $z_i$ $(i=1,\ldots,n)$ from its multinomial logit full conditional.
 First, we define $U_{ki} = \mathds{1}_{(z_i = k)}$ for $i = 1,...,n$ and $k = 1,...,K-1$.
@@
 for $i = 1,...n$ and $k = 1,...,K$.
-We also compute, according to the multivariate normal density of $\mathbf{y}_i|t_i$, the probability $P(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)$, where $\boldsymbol\zeta^*_{ki} = \mathbf{X}_i^*\boldsymbol\beta^*_k$.
+We also compute, according to the multivariate normal density of $\mathbf{y}_i|t_i$, the multivariate normal density $f(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)$, where $\boldsymbol\zeta^*_{ki} = \mathbf{X}_i^*\boldsymbol\beta^*_k = \mathbf{X}_i \boldsymbol\beta_k + \boldsymbol\psi_k t_i$.
 We use these quantities to compute $\boldsymbol\upsilon_{i} = (\upsilon_{1i},...,\upsilon_{Ki})^T$, where
-$$\upsilon_{ki} = P(z_i = k|\mathbf{y}_i,\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k) = \frac{\pi_{ki}P(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)}{\sum_{h = 1}^{K} \pi_{hi}P(\mathbf{y}_i|\boldsymbol\zeta^*_{hi},\boldsymbol\Sigma_h)}.$$
+$$\upsilon_{ki} = \text{Pr}(z_i = k|\mathbf{y}_i,\boldsymbol\zeta^*_{1i},...,\boldsymbol\zeta^*_{Ki},\boldsymbol\Sigma_k) = \frac{\pi_{ki}f(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)}{\sum_{h = 1}^{K} \pi_{hi}f(\mathbf{y}_i|\boldsymbol\zeta^*_{hi},\boldsymbol\Sigma_h)}.$$
 The cluster labels $z_i$ are then updated from a $\text{Multinomial}(1,\boldsymbol\upsilon_i)$ density for $i = 1,...n$.
@@
 
-We monitor convergence of the MCMC algorithm through the use of standard approaches such as trace plots and Geweke's (1992) Z-diagnostic, implemented in the \texttt{R} package \texttt{coda} (Plummer et al.
+We monitor convergence of the MCMC algorithm through the use of standard approaches, such as trace plots, effective sample sizes (ESSs) and Geweke's (1992) Z-diagnostic, implemented in the \texttt{R} package \texttt{coda} (Plummer et al.
 2006).
-In simulation studies under realistic parameter settings, we observed relatively fast convergence of all MCMC chains (i.e., within 500 iterations).
+In simulation studies under realistic parameter settings, we observed rapid convergence of all MCMC chains (i.e., within 500 iterations).
 
@@
 To this end, we propose the use of the ``widely applicable information criterion'' (WAIC) introduced by Watanabe (2010) for model selection.
-WAIC has the desirable property of penalizing complexity in models -- a feature congruent with our goal of explaining infant motor development heterogeneity with a parsimonious number of clusters.
+WAIC has the desirable property of penalizing complexity in models --- a feature congruent with our goal of explaining infant motor development heterogeneity with a parsimonious number of clusters.
 See Gelman et al.
@@
 In Section \ref{s:sim}, we show through a simulation study that this approach recovered the true value of $K$ under realistic parameter settings.
-In Section 6, we discuss alternative methods for choosing the optimal value of K.
+In Section 6, we discuss alternative methods for choosing the optimal value of $K$.
 
@@
 \end{eqnarray}
-where $\mathbf{y}_i = (y_{i1},...,y_{i4})^T$ to conform to the $J = 4$ measurement occasions in the Nurture study; $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$ ($k = 1,2,3$), and $f(\mathbf{y}_i | \boldsymbol\theta_k) \stackrel{\text{d}}{=} \text{MSN}_4(\boldsymbol\zeta_{ki},\boldsymbol\alpha_k,\boldsymbol\Omega_k)$; $\boldsymbol\zeta_{ki} = (\zeta_{ki1},...,\zeta_{ki4})^T = \mathbf{X}_i \boldsymbol\beta_{k}$, $\boldsymbol\beta_k = \text{vec}(\mathbf{B}_k)$, and $\mathbf{X}_i$ and $\mathbf{B}_k$ are given by.
+where $\mathbf{y}_i = (y_{i1},...,y_{i4})^T$ to conform to the $J = 4$ measurement occasions in the Nurture study; $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$ ($k = 1,2,3$), and $f(\mathbf{y}_i | \boldsymbol\theta_k) \stackrel{\text{d}}{=} \text{MSN}_4(\boldsymbol\zeta_{ki},\boldsymbol\alpha_k,\boldsymbol\Omega_k)$; $\boldsymbol\zeta_{ki} = (\zeta_{ki1},...,\zeta_{ki4})^T = \mathbf{X}_i \boldsymbol\beta_{k}$, $\boldsymbol\beta_k = \text{vec}(\mathbf{B}_k)$, and $\mathbf{X}_i$ and $\mathbf{B}_k$ are given by:
 \begin{eqnarray}
@@
 		\beta_{k41} & \beta_{k42} \\
-	\end{bmatrix}. \nonumber
+	\end{bmatrix}, \nonumber
 \end{eqnarray}
+where $x_i$ is a $N(0,1)$ covariate representing, say, BMI z-score from the Nurture study.
 We fit a dummy indicator for each measurement occasion and a time invariant baseline covariate $x_i$ (e.g., birthweight for gestational age z-score), whose effect may vary across the $J$ measurement occasions.
@@
 \end{eqnarray}
-To emulate z-score-transformed covariates present in the Nurture data such as birthweight for gestational age z-score, we sampled $x_i$ from a $\text{N}(0,1)$ density for $i = 1,...,n$.
-
-For cluster $k$, we collected all observations $\mathbf{y}_i$ ($i = 1,...,n_k$) into the $n_k \times 4$ matrix $\mathbf{Y}_k$ and all covariates $\mathbf{X}_i$ into the $4n_k \times 8$ matrix $\mathbf{X}_k$ to utilize the matrix normal model specification detailed in Section 3, which allows for simultaneous updates of all MSN regression parameters in cluster $k$.
 For the multinomial regression model component of this simulation, we modeled the class labels $z_i$ as a function of an intercept and one baseline covariate, $\text{w}_{i1}$, implying that $r = 2$.
-We did introduce missing data into this simulation, as missing data will be dealt with in the second simulation study.
+We did not introduce missing data into this simulation, as missing data will be dealt with in the second simulation study.
 As a result, the final sample size was $N = n \times J = 4,000$.
 
-We chose the MSN hyperparameters to be homogeneous across the three clusters by setting, for $k = 1,2,3$, $\mathbf{B}_{0k}^* = \mathbf{0}_{4 \times 3}$, $\mathbf{V}_{0k} = \mathds{1}_{4 \times 4}$, and $\nu_{0k} = J + 2 = 6$, which gives $\text{E}(\boldsymbol\Sigma_k) = \mathds{1}_{4 \times 4}$.
-Similarly, we set $\mathbf{d}_{01} = \mathbf{d}_{02} = (0,0)^T$ and $\mathbf{S}_{01} = \mathbf{S}_{02} = \mathds{1}_{2 \times 2}$, noting that $k = 3$ is the reference cluster.
+We chose the matrix normal hyperparameters described in Section 3.4.1 to be homogeneous across the three clusters by setting, for $k = 1,2,3$, $\mathbf{B}_{0k}^* = \mathbf{0}_{4 \times 3}$, $\mathbf{V}_{0k} = \mathbf{I}_{4}$, and $\nu_{0k} = J + 2 = 6$, which gives $\text{E}(\boldsymbol\Sigma_k) = \mathbf{I}_{4}$.
+Similarly, we set $\mathbf{d}_{01} = \mathbf{d}_{02} = (0,0)^T$ and $\mathbf{S}_{01} = \mathbf{S}_{02} = \mathbf{I}_{2}$, noting that $k = 3$ is the reference cluster.
 To investigate the effect of ignoring skewness, we fit both the MVN and MSN mixture models to data generated from model (\ref{eq:sim1}).
-The values of WAIC for the MSN mixture model and MVN mixture model were 37,221.7 and 324,705.3, respectively, indicating model fit suffered when skewness was ignored.
-Furthermore, we report below in Table \ref{tab:sim1_parms} a selection of parameter estimates from the MSN and MVN models that show incorrect inference (i.e., true parameter values not contained in 95\% posterior credible intervals) resulted from ignoring skewness.
-A full table showing all model parameters estimates and 95\% posterior credible intervals for the MSN and MVN models in addition to ground truth values is given in Web Appenxix B.
+The WAIC values for the MSN and MVN mixture models were 37,222 and 324,705, respectively, indicating better fit for the MSN model, as expected.
+Table \ref{tab:sim1_parms} presents a selection of parameter estimates from the MSN and MVN models that show incorrect inference (i.e., true parameter values not contained in 95\% posterior credible intervals) resulted from the MVN model.
+Web Table 1 presents all model parameters estimates and 95\% posterior credible intervals for the MSN and MVN models in addition to their ground truth values.
 
@@
 As such, the MSN model can be used in place of the MVN model even when data are not clearly skewed.
-We demonstrate this by fitting both the MSN and MVN model to data generated from model (\ref{eq:sim1}), where all skewness parameters are set to zero.
-As shown Table 2 of Web Appendix B, parameter estimates are similar between the MSN and MVN models, and all 95\% credible intervals for the $\alpha$ parameters of the MSN models contain 0, indicating the MSN model performs well even when data are not skewed.
+We demonstrate this by fitting both the MSN and MVN model to data generated from model (\ref{eq:sim1}), in which all skewness parameters are set to zero.
+As shown in Web Table 2, the MSN and MVN models yield similar parameter estimates, and the 95\% credible intervals for the skewness parameters ($\alpha_k$) of the MSN model contain 0, indicating the MSN model performs well even when data are not skewed.
 
@@
 
-In the second simulation study, we compare our online conditional imputation approach to Bayesian multiple imputation, in which M (large) imputed data sets are generated, MCMC is performed on each imputed data set, and the data are pooled for final posterior inference.
+In the second simulation study, we compare our online conditional imputation approach to Bayesian multiple imputation, whereby $M$ (large) imputed data sets are generated, MCMC is performed on each imputed data set, and the data are pooled for final posterior inference.
 Details of this approach are nicely summarized in (Zhou and Reiter, 2010).
@@
 
-To demonstrate the advantages of our proposed imputation method, we first generate $n = 1,000$ observations from a simple three-cluster ($K = 3$) MSN mixture model with $J = 4$ repeated measurements, one main effect for time ($p = 1$), and two multinomial regression predictors ($r = 2$).
-We then artificially ampute the simulated data using the \texttt{ampute} function provided by the \texttt{mice} package in \texttt{R} (Van Buuren and Groothuis-Oudshoorn 2010) according to the missing at random (MAR) mechanism, whereby the occurence of missing data depends only on the values of the observed data (Gelman and Hill 2006).
+To demonstrate the advantages of our proposed imputation method, we first generated $n = 1,000$ observations from a simple three-cluster ($K = 3$) MSN mixture model with $J = 4$ repeated measurements, one main effect for time ($p = 1$), and two multinomial regression predictors ($r = 2$).
+We then removed observations intermittently across the 4 measurement occasions according to a MAR mechanism, whereby the occurence of missing data depends only on the values of the observed data (Gelman and Hill 2006).
+The \texttt{ampute} function from the \texttt{mice} package in \texttt{R} was used to generate missing observations.
 We specified that each of the $4$ repeated measures for a given individual were equally likely to be missing.
@@
 We then fit our proposed MSN finite mixture model to the amputed data using both online conditional imputation and Bayesian multiple imputation as defined by (Zhou and Reiter, 2010).
+As is described by Zhou and Reiter, Bayesian multiple imputation requires the number of imputations $M$ to be large.
+Accordingly, we generated $M = 1,000$ imputed data sets using conditional MVN imputation, where the multivariate mean and covariance parameters were estimated from the available case sample sub-sample.
+The resultant multiply imputed data array occupied over 31 megabytes of disk space, illustrating the computational burden of Bayesian multiple imputation compared to online imputation.
 
@@
 Our final simulation is concerned with validating the use of WAIC for determining the number of clusters $K$.
-To this end, we simulate data from the same MSN mixture as in simulation study \#2, that is -- we simulate data from a $K = 3$ cluster model and fit MSN mixtures of $K \in \{2,3,4\}$, comparing the WAIC under each setting of $K$.
-
-\textbf{Need to start this simulation before writing more}.
+To this end, we simulate data from the same MSN mixture as in simulation study \#2, that is --- we simulate data from a $K = 3$ cluster model and fit MSN mixtures of $K \in \{2,3,4\}$, comparing the WAIC under each setting of $K$.
+Additionally, we investigate to what degree fitting the MVN model to data generated from a MSN model affects the choice of $K$, using WAIC to select the optimal $K$.
 

### 2019-07-30 simulation_2/fit_mvsn.R

@@ -1,0 +1,4 @@
+fit_mvsn <- function(Y,X,W,B0,V0,L0,S0,d0,nu0,nsim,burn,thin)
+{
+    
+}

### 2019-07-30 simulation_2/generate_data_sim2.R

@@ -1,0 +1,176 @@
+# Packages
+library(sn) # for rsn
+library(truncnorm) # for rtruncnorm
+library(mvtnorm) # for rmvtnorm
+library(mvnmle)
+library(Hmisc) # For rMultinom function (easy sampling of class indicators C)
+library(coda)
+library(MCMCpack)
+library(matrixsampling)
+library(QRM)        # For rGumbel function
+library(nnet)       # To fit multinom function
+library(BayesLogit) # For rpg function
+library(mice)
+
+#seed <- rnorm(1)
+#set.seed(seed)
+# No scientific notation
+options(scipen=999)
+options(warn=2)
+# Generate data
+k <- 4 # number of outcomes
+p <- 2 # number of predictors 
+h <- 3 # number of clusters
+n <- 1000 # number of observations (takes alot of data to estimate this many parameters)
+print(paste("Generating data with",n,"observations of dimension",k,
+            "with",p,"covariates and",h,"clusters"))
+
+N <- n*k # number of measurments 
+P <- matrix(rep(rep(1/h,h),each=n),n,h) # probability matrix governing cluster membership
+
+# Generate covariates
+X <- matrix(rnorm(n*p),nrow = n,ncol = p) # random predictors
+X[,1] <- 1
+t <- rtruncnorm(n,0,Inf,0,1) # trunc norm random effects
+Xstar <- cbind(X,t) # Combine X and t
+
+w1 <- rbinom(n,1,0.5) # binary predictor
+W <- cbind(1,w1) # design matrix
+v <- ncol(W) # number of multinomial predictors
+delta.true <- matrix(rtruncnorm(n = v*(h-1),
+                                a = -1, 
+                                b = 1,
+                                mean = 0),
+                     nrow = v,
+                     ncol = h-1) # true multinomial regression coefficients
+eta.true <- W %*% delta.true # true eta term for each observation
+U <- matrix(0,nrow = n,ncol = h) # make empty U matrix
+U[,1] <- rGumbel(n, mu = 0, sigma = 1)
+for(i in 2:h) # populate U matrix with gumbel values
+{
+    U[,i] <- rGumbel(n, mu = eta.true[,i-1], sigma = 1)
+}
[…]

### 2019-07-30 simulation_2/impute_conditional_mvn.R

@@ -1,0 +1,23 @@
+impute_conditional_mvn <- function(y,mu,sig2)
+{
+    na.ind = is.na(y)
+    if(any(na.ind))
+    {
+        a = y[!na.ind]
+        A = sig2
+        mu.1 = mu[na.ind]
+        mu.2 = mu[!na.ind]
+        
+        sig.11 = A[na.ind,na.ind]
+        sig.12 = A[na.ind,!na.ind]
+        sig.21 = A[!na.ind,na.ind]
+        sig.22 = A[!na.ind,!na.ind]
+        
+        mu.cond = mu.1 + sig.12 %*% solve(sig.22) %*% (a - mu.2)
+        sig.cond = sig.11 - sig.12 %*% solve(sig.22) %*% sig.21
+        
+        y.imp = rmvnorm(1,mu.cond,sig.cond)
+        y[na.ind] = y.imp
+    }
+    return(y)
+}

### 2019-07-30 simulation_2/simulation_online.R

@@ -1,0 +1,369 @@
+# Multivariate skew normal regression with clustering
+# Includes built-in mvn conditional imputation
+# Includes regression on multinomial regression coefficients
+# Carter Allen
+# March 19, 2019
+
+# Packages
+library(sn) # for rsn
+library(truncnorm) # for rtruncnorm
+library(mvtnorm) # for rmvtnorm
+library(mvnmle)
+library(Hmisc) # For rMultinom function (easy sampling of class indicators C)
+library(coda)
+library(MCMCpack)
+library(matrixsampling)
+library(QRM)        # For rGumbel function
+library(nnet)       # To fit multinom function
+library(BayesLogit) # For rpg function
+
+#seed <- rnorm(1)
+#set.seed(seed)
+# No scientific notation
+options(scipen=999)
+options(warn=2)
+# Generate data
+k <- 4 # number of outcomes
+p <- 2 # number of predictors 
+h <- 3 # number of clusters
+n <- 1000 # number of observations (takes alot of data to estimate this many parameters)
+print(paste("Generating data with",n,"observations of dimension",k,
+            "with",p,"covariates and",h,"clusters"))
+
+N <- n*k # number of measurments 
+P <- matrix(rep(rep(1/h,h),each=n),n,h) # probability matrix governing cluster membership
+
+# Generate covariates
+X <- matrix(rnorm(n*p),nrow = n,ncol = p) # random predictors
+X[,1] <- 1
+t <- rtruncnorm(n,0,Inf,0,1) # trunc norm random effects
+Xstar <- cbind(X,t) # Combine X and t
+
+w1 <- rbinom(n,1,0.5) # binary predictor
+W <- cbind(1,w1) # design matrix
+v <- ncol(W) # number of multinomial predictors
+delta.true <- matrix(rtruncnorm(n = v*(h-1),
+                                a = -1, 
+                                b = 1,
+                                mean = 0),
+                     nrow = v,
+                     ncol = h-1) # true multinomial regression coefficients
+eta.true <- W %*% delta.true # true eta term for each observation
+U <- matrix(0,nrow = n,ncol = h) # make empty U matrix
[…]

### 2019-08-20 manuscript/manuscript.tex
revisions

@@
 
-\subsection{Simulation to Assess Sensitivity to Misspecified K}
+\subsection{Simulation to Validate Choice of K}
 
-Our final simulation is concerned with validating the use of WAIC for determining the number of clusters $K$.
-To this end, we simulate data from the same MSN mixture as in simulation study \#2, that is --- we simulate data from a $K = 3$ cluster model and fit MSN mixtures of $K \in \{2,3,4\}$, comparing the WAIC under each setting of $K$.
-Additionally, we investigate to what degree fitting the MVN model to data generated from a MSN model affects the choice of $K$, using WAIC to select the optimal $K$.
+Our final simulation is concerned with validating the use of WAIC for determining the number of clusters, $K$.
+To this end, we generate data sets from MSN models with $K = \{2,3,4\}$.
+To each simulated data set, we fit three of our proposed Bayesian MSN models with $K = \{2,3,4\}$ and compute WAIC for each model.
+As shown in Table 2, we are able to correctly estimate the true value of $K$ using WAIC, as the best fitting model for each simulation setting is the model with lowest WAIC.
+We note that, for the simulated data with $K = 2$, we were not able to fit a 4-cluster MSN model due to the occurence of vacant clusters during MCMC estimation.
+
+\begin{table}[]
+\centering
+\begin{tabular}{lllll}
+ &  &  & \textbf{Fitted} &  \\
+ &  & \textit{K = 2} & \textit{K = 3} & \textit{K = 4} \\ \cline{3-5} 
+ & \multicolumn{1}{l|}{\textit{K = 2}} & 11624 \multicolumn{1}{l|}{} & 16988 \multicolumn{1}{l|}{} & * \multicolumn{1}{l|}{} \\ \cline{3-5} 
+\textbf{Truth} & \multicolumn{1}{l|}{\textit{K = 3}} & 25681 \multicolumn{1}{l|}{} & 18062 \multicolumn{1}{l|}{} & 23519 \multicolumn{1}{l|}{} \\ \cline{3-5} 
+ & \multicolumn{1}{l|}{\textit{K = 4}} & 19305\multicolumn{1}{l|}{} & 21415 \multicolumn{1}{l|}{} & 12232 \multicolumn{1}{l|}{} \\ \cline{3-5} 
+\end{tabular}
+\end{table}
 

### 2019-08-20 simulation.R

@@ -41,28 +41,6 @@ X <- matrix(rnorm(n*p),nrow = n,ncol = p) # random predictors
 X[,1] <- 1
 t <- rtruncnorm(n,0,Inf,0,1) # trunc norm random effects
 Xstar <- cbind(X,t) # Combine X and t
-# # Generate X manually
-# xi <- matrix(c(1, 0, 0, 0,
-#                1, 1, 0, 0,
-#                1, 0, 1, 0,
-#                1, 0, 0, 1),
-#              nrow = 4,
-#              ncol = 4,
-#              byrow = TRUE)
-# t <- rtruncnorm(n,0,Inf,0,1) # trunc norm random effects
-# X <- NULL 
-# capT <- NULL
-# for(i in 1:n)
-# {
-#     X <- rbind(X,xi)
-#     capT <- rbind(capT,diag(t[i],nrow = 4,ncol = 4))
-# }
-# 
-# Xstar <- cbind(X,capT) # Combine X and t
-
-# Covariates corresponding to multinomial regression
-#w1 <- rnorm(n) # continuous predictor
-#w2 <- rnorm(n) # continuous predictor
 w3 <- rbinom(n,1,0.5) # binary predictor
 w4 <- rbinom(n,1,0.5) # binary predictor
 W <- cbind(1,w3) # design matrix

### 2019-08-20 simulation_2/fit_msn.R

@@ -1,0 +1,216 @@
+# Packages
+library(sn) # for rsn
+library(truncnorm) # for rtruncnorm
+library(mvtnorm) # for rmvtnorm
+library(mvnmle)
+library(Hmisc) # For rMultinom function (easy sampling of class indicators C)
+library(coda)
+library(MCMCpack)
+library(matrixsampling)
+library(QRM)        # For rGumbel function
+library(nnet)       # To fit multinom function
+library(BayesLogit) # For rpg function
+
+fit_msn <- function(Y,X,W,K,nsim,burn)
+{
+    # Define parameters
+    h = K
+    k = ncol(Y)
+    n = nrow(Y)
+    p = ncol(X)
+    v = ncol(W)
+    
+    # Define priors 
+    V0 = diag(1,k)
+    V0.list = vector("list",h)
+    V0.list[1:h] = list(V0)
+    L0.list = vector("list",h)
+    L0.list[1:h] = list(diag(1,p+1))
+    B0.list = vector("list",h)
+    B0.list[1:h] = list(matrix(0,nrow = p+1,ncol = k))
+    delta0 = rep(0,v) # prior mean for delta coefficients (multinomial regression)
+    S0 = diag(1,v) # prior covariance for delta coefficients (multinomial regression)
+    nu0 = rep(6,h) 
+    
+    # Define inits
+    pi = rep(1/h,h)
+    z = sample(1:h,n,replace = TRUE,prob = pi) 
+    Bn.list = B0.list
+    Ln.list = L0.list
+    nun = nu0
+    Vn.list = V0.list
+    beta.list = vector("list",h)
+    beta.mle = solve(t(X) %*% X) %*% t(X) %*% Y
+    beta.list[1:h] = list(beta.mle)
+    sig2.list = vector("list",h)
+    sig2.mle = mlest(Y - X %*% beta.mle)$sigmahat
+    sig2.list[1:h] = list(sig2.mle)
+    omega.list = V0.list
+    psi.list = vector("list",h)
+    psi.list[1:h] = list(rep(0,k))
+    alpha.list = psi.list
+    delta = matrix(0,nrow = v, ncol = h-1)
+    
+    nsim <- 1000 # number of iterations
+    burn <- 0 # number of iterations to save
+    n.iter <- nsim - burn # number of saved iterations
+    Z = matrix(0,nrow = n.iter,ncol = n) # large matrix where each row is the value of z at a specific iteration
[…]

### 2019-08-20 simulation_2/fit_mvsn.R

@@ -1,4 +1,0 @@
-fit_mvsn <- function(Y,X,W,B0,V0,L0,S0,d0,nu0,nsim,burn,thin)
-{
-    
-}

### 2019-08-20 simulation_3/MSN_K2_DATA/WAIC.R

@@ -1,0 +1,67 @@
+library(sn)
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/simulation_3/MSN_K2_DATA/")
+
+load("Y")
+load("X")
+load("Xstar")
+load("W")
+load("Z")
+load("SIGMA")
+load("PSI")
+load("BETA")
+load("DELTA")
+
+n <- nrow(Y)
+K <- length(BETA.list)
+J <- ncol(Y)
+p <- ncol(X)
+r <- ncol(W)
+S <- nrow(Z)
+
+lppd <- 0
+pwaic <- 0
+
+pb <- txtProgressBar(min = 0, max = n, style = 3)
+for(i in 1:n)
+{
+    y_i <- Y[i,]
+    x_i <- X[i,]
+    z_i <- Z[,i]
+    t_i <- Xstar[i,p+1]
+    p_yi <- rep(0,S)
+    for(s in 1:S)
+    {
+        k <- z_i[s]
+        
+        beta_k <- BETA.list[[k]]
+        beta_k_s <- beta_k[s,]
+        beta_ks <- matrix(beta_k_s,
+                          nrow = p,
+                          ncol = J,
+                          byrow = FALSE)
+        
+        sigma_k <- SIGMA.list[[k]]
+        sigma_k_s <- sigma_k[s,]
+        sigma_ks <- matrix(sigma_k_s,
+                           nrow = J,
+                           ncol = J,
+                           byrow = TRUE)
+        
+        psi_k <- PSI.list[[k]]
+        psi_ks <- psi_k[s,]
+        
+        zeta_iks <- x_i %*% beta_ks
+        
+        mu_iks <- zeta_iks + t_i %*% psi_ks
+        
+        p_yi[s] <- dmvnorm(y_i,mu_iks,sigma_ks)
+    }
+    setTxtProgressBar(pb, i)
+    
+    lppd <- lppd + log(mean(p_yi, na.rm = TRUE))
+    pwaic <- pwaic + var(log(p_yi), na.rm = TRUE)
+}
+WAIC <- -2*(lppd - pwaic)
+save(WAIC,file = "WAIC")
+
+close(pb)

### 2019-08-20 simulation_3/MSN_K2_DATA/mcmc_draws_2019-08-05/WAIC.R

@@ -1,0 +1,67 @@
+library(sn)
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/simulation_3/MSN_K2_DATA/mcmc_draws_2019-08-05/")
+
+load("Y")
+load("X")
+load("Xstar")
+load("W")
+load("Z")
+load("SIGMA")
+load("PSI")
+load("BETA")
+load("DELTA")
+
+n <- nrow(Y)
+K <- length(BETA.list)
+J <- ncol(Y)
+p <- ncol(X)
+r <- ncol(W)
+S <- nrow(Z)
+
+lppd <- 0
+pwaic <- 0
+
+pb <- txtProgressBar(min = 0, max = n, style = 3)
+for(i in 1:n)
+{
+    y_i <- Y[i,]
+    x_i <- X[i,]
+    z_i <- Z[,i]
+    t_i <- Xstar[i,p+1]
+    p_yi <- rep(0,S)
+    for(s in 1:S)
+    {
+        k <- z_i[s]
+        
+        beta_k <- BETA.list[[k]]
+        beta_k_s <- beta_k[s,]
+        beta_ks <- matrix(beta_k_s,
+                          nrow = p,
+                          ncol = J,
+                          byrow = FALSE)
+        
+        sigma_k <- SIGMA.list[[k]]
+        sigma_k_s <- sigma_k[s,]
+        sigma_ks <- matrix(sigma_k_s,
+                           nrow = J,
+                           ncol = J,
+                           byrow = TRUE)
+        
+        psi_k <- PSI.list[[k]]
+        psi_ks <- psi_k[s,]
+        
+        zeta_iks <- x_i %*% beta_ks
+        
+        mu_iks <- zeta_iks + t_i %*% psi_ks
+        
+        p_yi[s] <- dmvnorm(y_i,mu_iks,sigma_ks)
+    }
+    setTxtProgressBar(pb, i)
+    
+    lppd <- lppd + log(mean(p_yi, na.rm = TRUE))
+    pwaic <- pwaic + var(log(p_yi), na.rm = TRUE)
+}
+WAIC <- -2*(lppd - pwaic)
+save(WAIC,file = "WAIC")
+
+close(pb)

### 2019-08-20 simulation_3/MSN_K2_DATA/mcmc_draws_2019-08-05/mcmc_draws_2019-08-05/WAIC.R

@@ -1,0 +1,67 @@
+library(sn)
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/simulation_3/MSN_K2_DATA/mcmc_draws_2019-08-05/mcmc_draws_2019-08-05/")
+
+load("Y")
+load("X")
+load("Xstar")
+load("W")
+load("Z")
+load("SIGMA")
+load("PSI")
+load("BETA")
+load("DELTA")
+
+n <- nrow(Y)
+K <- length(BETA.list)
+J <- ncol(Y)
+p <- ncol(X)
+r <- ncol(W)
+S <- nrow(Z)
+
+lppd <- 0
+pwaic <- 0
+
+pb <- txtProgressBar(min = 0, max = n, style = 3)
+for(i in 1:n)
+{
+    y_i <- Y[i,]
+    x_i <- X[i,]
+    z_i <- Z[,i]
+    t_i <- Xstar[i,p+1]
+    p_yi <- rep(0,S)
+    for(s in 1:S)
+    {
+        k <- z_i[s]
+        
+        beta_k <- BETA.list[[k]]
+        beta_k_s <- beta_k[s,]
+        beta_ks <- matrix(beta_k_s,
+                          nrow = p,
+                          ncol = J,
+                          byrow = FALSE)
+        
+        sigma_k <- SIGMA.list[[k]]
+        sigma_k_s <- sigma_k[s,]
+        sigma_ks <- matrix(sigma_k_s,
+                           nrow = J,
+                           ncol = J,
+                           byrow = TRUE)
+        
+        psi_k <- PSI.list[[k]]
+        psi_ks <- psi_k[s,]
+        
+        zeta_iks <- x_i %*% beta_ks
+        
+        mu_iks <- zeta_iks + t_i %*% psi_ks
+        
+        p_yi[s] <- dmvnorm(y_i,mu_iks,sigma_ks)
+    }
+    setTxtProgressBar(pb, i)
+    
+    lppd <- lppd + log(mean(p_yi, na.rm = TRUE))
+    pwaic <- pwaic + var(log(p_yi), na.rm = TRUE)
+}
+WAIC <- -2*(lppd - pwaic)
+save(WAIC,file = "WAIC")
+
+close(pb)

### 2019-08-20 simulation_3/MSN_K3_DATA/WAIC.R

@@ -1,0 +1,67 @@
+library(sn)
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/simulation_3/MSN_K3_DATA/")
+
+load("Y")
+load("X")
+load("Xstar")
+load("W")
+load("Z")
+load("SIGMA")
+load("PSI")
+load("BETA")
+load("DELTA")
+
+n <- nrow(Y)
+K <- length(BETA.list)
+J <- ncol(Y)
+p <- ncol(X)
+r <- ncol(W)
+S <- nrow(Z)
+
+lppd <- 0
+pwaic <- 0
+
+pb <- txtProgressBar(min = 0, max = n, style = 3)
+for(i in 1:n)
+{
+    y_i <- Y[i,]
+    x_i <- X[i,]
+    z_i <- Z[,i]
+    t_i <- Xstar[i,p+1]
+    p_yi <- rep(0,S)
+    for(s in 1:S)
+    {
+        k <- z_i[s]
+        
+        beta_k <- BETA.list[[k]]
+        beta_k_s <- beta_k[s,]
+        beta_ks <- matrix(beta_k_s,
+                          nrow = p,
+                          ncol = J,
+                          byrow = FALSE)
+        
+        sigma_k <- SIGMA.list[[k]]
+        sigma_k_s <- sigma_k[s,]
+        sigma_ks <- matrix(sigma_k_s,
+                           nrow = J,
+                           ncol = J,
+                           byrow = TRUE)
+        
+        psi_k <- PSI.list[[k]]
+        psi_ks <- psi_k[s,]
+        
+        zeta_iks <- x_i %*% beta_ks
+        
+        mu_iks <- zeta_iks + t_i %*% psi_ks
+        
+        p_yi[s] <- dmvnorm(y_i,mu_iks,sigma_ks)
+    }
+    setTxtProgressBar(pb, i)
+    
+    lppd <- lppd + log(mean(p_yi, na.rm = TRUE))
+    pwaic <- pwaic + var(log(p_yi), na.rm = TRUE)
+}
+WAIC <- -2*(lppd - pwaic)
+save(WAIC,file = "WAIC")
+
+close(pb)

### 2019-08-20 simulation_3/MSN_K4_DATA/WAIC.R

@@ -1,0 +1,67 @@
+library(sn)
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/simulation_3/MSN_K4_DATA/mcmc_draws_2019-08-05/")
+
+load("Y")
+load("X")
+load("Xstar")
+load("W")
+load("Z")
+load("SIGMA")
+load("PSI")
+load("BETA")
+load("DELTA")
+
+n <- nrow(Y)
+K <- length(BETA.list)
+J <- ncol(Y)
+p <- ncol(X)
+r <- ncol(W)
+S <- nrow(Z)
+
+lppd <- 0
+pwaic <- 0
+
+pb <- txtProgressBar(min = 0, max = n, style = 3)
+for(i in 1:n)
+{
+    y_i <- Y[i,]
+    x_i <- X[i,]
+    z_i <- Z[,i]
+    t_i <- Xstar[i,p+1]
+    p_yi <- rep(0,S)
+    for(s in 1:S)
+    {
+        k <- z_i[s]
+        
+        beta_k <- BETA.list[[k]]
+        beta_k_s <- beta_k[s,]
+        beta_ks <- matrix(beta_k_s,
+                          nrow = p,
+                          ncol = J,
+                          byrow = FALSE)
+        
+        sigma_k <- SIGMA.list[[k]]
+        sigma_k_s <- sigma_k[s,]
+        sigma_ks <- matrix(sigma_k_s,
+                           nrow = J,
+                           ncol = J,
+                           byrow = TRUE)
+        
+        psi_k <- PSI.list[[k]]
+        psi_ks <- psi_k[s,]
+        
+        zeta_iks <- x_i %*% beta_ks
+        
+        mu_iks <- zeta_iks + t_i %*% psi_ks
+        
+        p_yi[s] <- dmvnorm(y_i,mu_iks,sigma_ks)
+    }
+    setTxtProgressBar(pb, i)
+    
+    lppd <- lppd + log(mean(p_yi, na.rm = TRUE))
+    pwaic <- pwaic + var(log(p_yi), na.rm = TRUE)
+}
+WAIC <- -2*(lppd - pwaic)
+save(WAIC,file = "WAIC")
+
+close(pb)

### 2019-08-20 simulation_3/MSN_K4_DATA/mcmc_draws_2019-08-05/mcmc_draws_2019-08-05/WAIC.R

@@ -1,0 +1,67 @@
+library(sn)
+setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/simulation_3/MSN_K4_DATA/mcmc_draws_2019-08-05/mcmc_draws_2019-08-05/")
+
+load("Y")
+load("X")
+load("Xstar")
+load("W")
+load("Z")
+load("SIGMA")
+load("PSI")
+load("BETA")
+load("DELTA")
+
+n <- nrow(Y)
+K <- length(BETA.list)
+J <- ncol(Y)
+p <- ncol(X)
+r <- ncol(W)
+S <- nrow(Z)
+
+lppd <- 0
+pwaic <- 0
+
+pb <- txtProgressBar(min = 0, max = n, style = 3)
+for(i in 1:n)
+{
+    y_i <- Y[i,]
+    x_i <- X[i,]
+    z_i <- Z[,i]
+    t_i <- Xstar[i,p+1]
+    p_yi <- rep(0,S)
+    for(s in 1:S)
+    {
+        k <- z_i[s]
+        
+        beta_k <- BETA.list[[k]]
+        beta_k_s <- beta_k[s,]
+        beta_ks <- matrix(beta_k_s,
+                          nrow = p,
+                          ncol = J,
+                          byrow = FALSE)
+        
+        sigma_k <- SIGMA.list[[k]]
+        sigma_k_s <- sigma_k[s,]
+        sigma_ks <- matrix(sigma_k_s,
+                           nrow = J,
+                           ncol = J,
+                           byrow = TRUE)
+        
+        psi_k <- PSI.list[[k]]
+        psi_ks <- psi_k[s,]
+        
+        zeta_iks <- x_i %*% beta_ks
+        
+        mu_iks <- zeta_iks + t_i %*% psi_ks
+        
+        p_yi[s] <- dmvnorm(y_i,mu_iks,sigma_ks)
+    }
+    setTxtProgressBar(pb, i)
+    
+    lppd <- lppd + log(mean(p_yi, na.rm = TRUE))
+    pwaic <- pwaic + var(log(p_yi), na.rm = TRUE)
+}
+WAIC <- -2*(lppd - pwaic)
+save(WAIC,file = "WAIC")
+
+close(pb)

### 2019-08-20 simulation_3/simulation_3_MSN.R

@@ -1,0 +1,381 @@
+# Multivariate skew normal regression with clustering
+# Includes built-in mvn conditional imputation
+# Includes regression on multinomial regression coefficients
+# Carter Allen
+# March 19, 2019
+
+# Packages
+library(sn) # for rsn
+library(truncnorm) # for rtruncnorm
+library(mvtnorm) # for rmvtnorm
+library(mvnmle)
+library(Hmisc) # For rMultinom function (easy sampling of class indicators C)
+library(coda)
+library(MCMCpack)
+library(matrixsampling)
+library(QRM)        # For rGumbel function
+library(nnet)       # To fit multinom function
+library(BayesLogit) # For rpg function
+
+#seed <- rnorm(1)
+#set.seed(seed)
+# No scientific notation
+options(scipen=999)
+options(warn=2)
+# Generate data
+k <- 4 # number of outcomes
+p <- 2 # number of predictors 
+h <- 2 # number of clusters
+n <- 1000 # number of observations (takes alot of data to estimate this many parameters)
+print(paste("Generating data with",n,"observations of dimension",k,
+            "with",p,"covariates and",h,"clusters"))
+
+N <- n*k # number of measurments 
+P <- matrix(rep(rep(1/h,h),each=n),n,h) # probability matrix governing cluster membership
+
+# Generate covariates
+X <- matrix(rnorm(n*p),nrow = n,ncol = p) # random predictors
+X[,1] <- 1
+t <- rtruncnorm(n,0,Inf,0,1) # trunc norm random effects
+Xstar <- cbind(X,t) # Combine X and t
+
+w1 <- rbinom(n,1,0.5) # binary predictor
+W <- cbind(1,w1) # design matrix
+v <- ncol(W) # number of multinomial predictors
+delta.true <- matrix(rtruncnorm(n = v*(h-1),
+                                a = -1, 
+                                b = 1,
+                                mean = 0),
+                     nrow = v,
+                     ncol = h-1) # true multinomial regression coefficients
+eta.true <- W %*% delta.true # true eta term for each observation
+U <- matrix(0,nrow = n,ncol = h) # make empty U matrix
[…]

### 2019-08-20 simulation_3/simulation_3_MVN.R

@@ -1,0 +1,370 @@
+# Multivariate skew normal regression with clustering
+# Includes built-in mvn conditional imputation
+# Includes regression on multinomial regression coefficients
+# Carter Allen
+# March 19, 2019
+
+# Packages
+library(sn) # for rsn
+library(truncnorm) # for rtruncnorm
+library(mvtnorm) # for rmvtnorm
+library(mvnmle)
+library(Hmisc) # For rMultinom function (easy sampling of class indicators C)
+library(coda)
+library(MCMCpack)
+library(matrixsampling)
+library(QRM)        # For rGumbel function
+library(nnet)       # To fit multinom function
+library(BayesLogit) # For rpg function
+
+#seed <- rnorm(1)
+#set.seed(seed)
+# No scientific notation
+options(scipen=999)
+options(warn=2)
+# Generate data
+k <- 4 # number of outcomes
+p <- 2 # number of predictors 
+h <- 3 # number of clusters
+n <- 1000 # number of observations (takes alot of data to estimate this many parameters)
+print(paste("Generating data with",n,"observations of dimension",k,
+            "with",p,"covariates and",h,"clusters"))
+
+N <- n*k # number of measurments 
+P <- matrix(rep(rep(1/h,h),each=n),n,h) # probability matrix governing cluster membership
+
+# Generate covariates
+X <- matrix(rnorm(n*p),nrow = n,ncol = p) # random predictors
+X[,1] <- 1
+t <- rtruncnorm(n,0,Inf,0,1) # trunc norm random effects
+Xstar <- cbind(X,t) # Combine X and t
+
+w1 <- rbinom(n,1,0.5) # binary predictor
+W <- cbind(1,w1) # design matrix
+v <- ncol(W) # number of multinomial predictors
+delta.true <- matrix(rtruncnorm(n = v*(h-1),
+                                a = -1, 
+                                b = 1,
+                                mean = 0),
+                     nrow = v,
+                     ncol = h-1) # true multinomial regression coefficients
+eta.true <- W %*% delta.true # true eta term for each observation
+U <- matrix(0,nrow = n,ncol = h) # make empty U matrix
[…]
